### 2023-09-18 paper/paper.qmd
Cross referenced first figure

---
title: "Elephant Seal Age JAE Intraspecific Variation"
format: html
editor: visual
---

```{r}
#| label: setup

library(tidyverse)
library(ggthemes)
sealdat <- read_csv(here::here("data/raw/128L pull 2023_06_29.csv"), 
                    show_col_types = FALSE) %>% 
  mutate(observed = if_else(observed == "B", "Breeder", "Non-breeder"))

```

## Introduction

## Methods

## Results

Pup sex shifted towards males as maternal age increased (@fig-repro_age).

```{r}
#| label: fig-repro_age
#| fig-cap: "Pup sex shifted towards males with increasing maternal age."

sealdat %>% 
  filter(pupsex %in% c("M", "F"), 
         age < 18) %>% 
  group_by(age, pupsex) %>%
  count(pupsex) %>% 
  pivot_wider(values_from = n, names_from = pupsex) %>% 
  mutate(num = M + F,
         percf = F / num,
         se = (percf * (1 - percf) / num)^0.5) %>% 
  ggplot(aes(x = age, y = percf, weight = num)) +
  geom_point() +
  geom_errorbar(aes(ymin = percf - se, ymax = percf + se), width = 0.2) +
  geom_hline(yintercept = 0.5, lty = 2) +
  stat_smooth(method = "glm", method.args = list(family = "binomial"), formula = y ~ x) +
  geom_text(aes(label = num, vjust = 1, y = 0.03)) +
  scale_y_continuous(labels = scales::percent, limits = c(0, NA)) +
  labs(x = "Maternal Age (Years)", y = "% Pups Female") +
  theme_few()

```

## Discussion

##

### 2023-09-18 scratch/128L 2023_06_25_Plots.r

@@ -1,0 +1,280 @@
+#FIGURE CHANGES TO MAKE
+#Create annual cycle figure
+
+#################################################################################
+
+setwd("C:/Users/roxan/Documents/Teaching/Current Courses/BIOE 128L/128L 2023/128L Projects")
+final=read.csv("128L pull 2023_06_29.csv")
+final$observed[final$observed=="B"]="Breeder"
+final$observed[final$observed==1]="Non-Breeder"
+
+library(tidyverse)
+library(ggthemes)
+library(emmeans)
+
+##### MOLT DURATION
+agg=final %>% group_by(age,observed) %>%
+  summarize(mean=mean(moltdur,na.rm=TRUE),
+            sd=sd(moltdur,na.rm=TRUE),
+            n=n(),
+            se=sd/sqrt(n))
+
+ggplot(data=agg,aes(x=age,y=mean,group=observed,col=observed))+
+  geom_point(aes(col=observed))+
+  geom_smooth(method="lm",mapping=aes(weight=n))+
+  geom_errorbar(aes(ymin=mean-se, ymax=mean+se,col=observed), width=.2)+
+  theme_few()+
+  ylab("Molt Haulout Duration (Mean +- SE)")+
+  geom_text(data=subset(agg,observed=="Breeder"),aes(label=n,vjust=1,
+                                                     col=observed),y=.03) + 
+  geom_text(data=subset(agg,observed=="Non-Breeder"),aes(label=n,vjust=-1,
+                                                         col=observed),y=.03) 
+require(lmerTest)
+#summary(lm(as.numeric(moltdur)~age+observed,data=final))
+summary(lmer(as.numeric(moltdur)~age+observed+(1|animalID)+(1|year),data=final))
+#summary(lm(as.numeric(moltdur)~age*observed,data=final))
+summary(lmer(as.numeric(moltdur)~age*observed+(1|animalID)+(1|year),data=final))
+
+#Molting Date
+agg=final %>% group_by(age,observed) %>%
+  subset(!is.na(lastobsmoltdoy)) %>% 
+  summarize(mean=mean(lastobsmoltdoy),
+            sd=sd(lastobsmoltdoy),
+            n=n(),
+            se=sd/sqrt(n))
+
+ggplot(data=agg,aes(x=age,y=mean,group=observed,col=observed))+
[…]

### 2023-09-18 paper/paper.qmd
Suppressed messages and code
Add citations

@@
 title: "Elephant Seal Age JAE Intraspecific Variation"
-format: html
+format: 
+  html: 
+    echo: false 
 editor: visual
+bibliography: references.bib
 ---
@@
 #| label: setup
+#| include: false
 
@@
 ## Introduction
+
+This paper is basically the same as another paper about polar bears (*Ursus maritimus*) [@naciri2022].
 

### 2023-09-19 paper/paper.qmd
Added two more reproductive success figures

@@
 library(ggthemes)
+library(gridExtra)
+
 sealdat <- read_csv(here::here("data/raw/128L pull 2023_06_29.csv"), 
@@
 
-Pup sex shifted towards males as maternal age increased (@fig-repro_age).
+### Changes to annual cycle timing with increased maternal age
+
+In order to create the timing figure, I need a csv with these columns: breeding arrival DOY, breeding departure DOY, molt arrival DOY, molt departure DOY.
+
+-   breeding arrival DOY = firstobsbreeddoy
+
+-   breeding departure DOY = firstobsbreeddoy + breeddur
+
+-   *molt arrival DOY = get from Roxanne*
+
+-   molt departure DOY = lastobsmoltdoy
+
+### Reproductive success and maternal age
+
+As maternal age increases, pup sex skews slightly towards males. Pup survival and recruitment are not significantly affected by maternal age (@fig-repro_age).
 
@@
 #| label: fig-repro_age
-#| fig-cap: "Pup sex shifted towards males with increasing maternal age."
+#| caption: "Pup sex skews slightly towards males with increasing maternal age. Pup survival decreases slightly, but survival and recruitment are not significantly affected by maternal age.
 
-sealdat %>% 
+#Pup sex shifts towards males with increasing maternal age.
+
+pup_sex <- sealdat %>% 
   filter(pupsex %in% c("M", "F"), 
@@
 
+#Pup survival doesn't change with increased maternal age.
+
+pup_survival <- sealdat %>% mutate(pup_surv = ifelse(pupseeneveragain > 0, 1, 0)) %>% 
+  drop_na(pup_surv) %>% 
+  group_by(age, pup_surv) %>%
+  count(pup_surv) %>% 
+  pivot_wider(names_from = pup_surv, values_from = n) %>% 
+  mutate(n = (`0`+`1`),
+         perc = `0` / n,
+         se = (perc * (1-perc) / n)^0.5) %>% 
+  filter(n > 5) %>% 
+  ggplot(aes(x = age, y = perc)) +
+  geom_point() +
+  geom_smooth(method = "glm", method.args = list(family = binomial), formula = y ~ x) +
+  geom_errorbar(aes(ymin = perc - se, ymax = perc + se), width = 0.2) +
+  geom_text(aes(label = n, 
+                vjust = -1), y = .5) +
+  labs(x = "Maternal Age (Years)", y = "Fraction Pups Not Seen Again") +
+  theme_few()
+
+#code that I haven't got to work yet that shows this by sex
+#ggplot(aes(x = age, y = perc, weight = n, col = pupsex)) +
+  # geom_text(pupsex == "M", aes(label = n, 
+  #                              vjust = -1, 
+  #                              col = pupsex), y = 0.5) + 
+  # geom_text(pupsex == "F", aes(label = n, 
+  #                              vjust = -1, 
+  #                              col = pupsex), y = 0.5) +
+
+#Pup recruitment doesn't change with maternal age
+
+pup_recruit <- sealdat %>% 
+  filter(pupsex == "F") %>% 
+  group_by(age, puprecruited) %>%
+  drop_na(puprecruited) %>% 
+  count(puprecruited) %>% 
+  pivot_wider(names_from = puprecruited, values_from = n) %>% 
+  mutate(n = (`0` + `1`),
+         perc = `0` / n,
+         se = (perc * (1 - perc) / n)^0.5) %>% 
+  filter(n > 5) %>% 
+  ggplot(aes(x = age, y = perc, weight = n)) +
+  geom_point() + 
+  xlim(2.5 , 13.5) +
+  geom_errorbar(aes(ymin = perc - se, ymax = perc + se), width = 0.2) +
+  stat_smooth(method = "glm", method.args = list(family = "binomial"), formula = y ~ x) +
+  labs(x = "Maternal Age (Years)", y = "Fraction Pups Never Breed") +
+  theme_few() +
+  geom_text(aes(label = n,
+                vjust = 1,
+                y = 1))
+
+grid.arrange(pup_sex, pup_survival, pup_recruit)
+
 ```

### 2023-09-19 paper/paper.qmd
Creating Fig 1A and 1B

--- 2023-09-18 paper/paper.qmd
@@
 library(ggthemes)
+library(cowplot)
+
 sealdat <- read_csv(here::here("data/raw/128L pull 2023_06_29.csv"), 
@@
 
-This paper is basically the same as another paper about polar bears (*Ursus maritimus*) [@naciri2022].
+This paper similar to another paper about polar bears (*Ursus maritimus*) [@naciri2022].
+
+Our hypotheses:
+
+1.  Elephant seal reproductive success declines with age.
+
+2.  Reproductive declines persist to the next generation (maternal effect senescence).
+
+3.  Seals optimize their annual cycle timing to compensate for reproductive declines.
 
@@
 
-Pup sex shifted towards males as maternal age increased (@fig-repro_age).
+### H1: Age distribution and breeding success
+
+```{r}
+#| label: fig-age_dist
+
+#Age distribution panel A = COUNTS, B = PERCENTAGES
+
+sealdat %>% 
+  group_by(age, observed) %>% 
+  count(age) %>% 
+  ggplot(aes(x = age, y = n, fill = observed)) + 
+  geom_col() + 
+  labs(x = "Number of individuals", y = "Age") + 
+  theme_few()
+
+#I want each of these bars to show the proportion of breeders and non breeders for each age
+
+```
+
+### H2: Maternal effect senescence
+
+As maternal age increases, pup sex skews slightly towards males.
+Pup survival declines slightly with maternal age, but pup recruitment is not significantly affected by maternal age (@fig-repro_age).
+Reproductive declines therefore do not persist to the next generation.
 
@@
 #| label: fig-repro_age
-#| fig-cap: "Pup sex shifted towards males with increasing maternal age."
+#| fig-cap: "Pup sex skews slightly towards males with increasing maternal age. Pup survival decreases slightly, but survival and recruitment are not significantly affected by maternal age."
+#| fig-width: 6
+#| fig-height: 6
 
-sealdat %>% 
+#Pup sex shifts towards males with increasing maternal age.
+
+pup_sex <- sealdat %>% 
   filter(pupsex %in% c("M", "F"), 
@@
   stat_smooth(method = "glm", method.args = list(family = "binomial"), formula = y ~ x) +
-  geom_text(aes(label = num, vjust = 1, y = 0.03)) +
   scale_y_continuous(labels = scales::percent, limits = c(0, NA)) +
@@
 
+#Pup survival doesn't change with increased maternal age.
+
+pup_surv_recruit <- sealdat %>% 
+  mutate(pup_surv = ifelse(pupseeneveragain > 0, 1, 0)) %>% 
+  filter(pupsex == "F") %>% 
+  group_by(age) %>%
+  summarize(n_surv = sum(!is.na(pup_surv)),
+            mean_surv = sum(pup_surv, na.rm = TRUE) / n_surv,  
+            se_surv = (mean_surv * (1 - mean_surv) / n_surv)^0.5,
+            n_recruit = sum(!is.na(puprecruited)), 
+            mean_recruit = sum(puprecruited, na.rm = TRUE) / n_recruit, 
+            se_recruit = (mean_recruit * (1 - mean_recruit) / n_recruit)^0.5)  %>%  
+  pivot_longer(cols = c(mean_surv, se_surv, mean_recruit, se_recruit), 
+               names_to = c("stat", "metric"), 
+               names_pattern = "(.*)_(.*)") %>% 
+  pivot_wider(names_from = "stat", 
+              values_from = "value") %>% 
+  mutate(low = mean - se, 
+         high = mean + se, 
+         age = age + ifelse(metric == "surv", -0.1, 0.1), 
+         metric = factor(metric, 
+                         levels = c("recruit", "surv"), 
+                         labels = c("Recruitment", "1st Year Survival"))) %>% 
+  filter(n_surv > 5) %>% 
+  ggplot(aes(x = age, y = mean, color = metric)) + 
+  geom_pointrange(aes(ymin = low, ymax = high)) +
+  geom_smooth(aes(weight = n_surv), 
+              method = "glm", 
+              method.args = list(family = binomial), 
+              formula = y ~ x) +
+  scale_y_continuous(labels = scales::percent, 
+                     limits = c(0, .5)) + 
+  labs(x = "Maternal Age (Years)", y = "% Pups") +
+  theme_few() + 
+  theme(legend.position = c(0.1,0.97), 
+        legend.justification = c(0,1), 
+        legend.title = element_blank(), 
+        legend.direction = "horizontal")
+
+plot_grid(pup_sex, pup_surv_recruit, ncol = 1, labels = "AUTO")
+
 ```
+
+### H3: Changes to annual cycle timing with increased maternal age
+
+In order to create the timing figure, I need a csv with these columns: breeding arrival DOY, breeding departure DOY, molt arrival DOY, molt departure DOY.
+
+-   breeding arrival DOY = firstobsbreeddoy
+
+-   breeding departure DOY = firstobsbreeddoy + breeddur
+
+-   molt arrival DOY = lastobsmoltdoy - moltdur
+
+-   molt departure DOY = lastobsmoltdoy
 

### 2023-09-20 paper/paper.qmd
Updating and making the colors consistent throughout the document

--- 2023-09-18 paper/paper.qmd
@@
 
-This paper is basically the same as another paper about polar bears (*Ursus maritimus*) [@naciri2022].
+Our hypotheses:
+
+1.  Elephant seal reproductive success declines with age.
+
+2.  Reproductive declines persist to the next generation (maternal effect senescence).
+
+3.  Seals optimize their annual cycle timing to compensate for reproductive declines.
+
+Note: This paper is similar to another paper about polar bears (*Ursus maritimus*) [@naciri2022].
 
@@
 
-Pup sex shifted towards males as maternal age increased (@fig-repro_age).
+### H1: Age distribution and breeding success
+
+```{r}
+#| label: fig-age_dist
+#| fig-cap: "A: Age structure of northern elephant seals at Año Nuevo. B: Percentage of each age class that were observed breeding."
+#| fig-width: 6
+#| fig-height: 8
+
+#Age distribution panel A
+
+age_dist <- sealdat %>% 
+  group_by(age) %>% 
+  count(age) %>% 
+  ggplot(aes(x = age, y = n)) + 
+  geom_col(fill = "#829BD1") + 
+  labs(x = "Number of individuals", y = "Age (Years)") + 
+  theme_few()
+
+#Panel B: percentage of each age class that were breeders or  nonbreeders
+
+breed_nonbreed <- sealdat %>% 
+  group_by(age) %>% 
+  summarize(n_observed = sum(!is.na(observed)), 
+            perc_breed = sum(observed == "Breeder") / n_observed, 
+            se = (perc_breed * (1 - perc_breed) / n_observed)^0.5) %>% 
+  ggplot(aes(x = age, y = perc_breed)) + 
+  stat_smooth(method = "lm", 
+              formula = y ~ poly(x , 2), 
+              se = TRUE, 
+              color = "#849ED5", 
+              fill = "#C8D3ED") +
+  geom_point() + 
+  geom_errorbar(aes(ymin = perc_breed - se, ymax = perc_breed + se), width = 0) +
+  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) + 
+  labs(x = "Maternal Age (Years)", y = "% Breeders") + 
+  theme_few()
+
+plot_grid(age_dist, breed_nonbreed, ncol = 1, labels = "AUTO")
+
+```
+
+### H2: Maternal effect senescence
+
+As maternal age increases, pup sex skews slightly towards males.
+Pup survival declines slightly with maternal age, but pup recruitment is not significantly affected by maternal age (@fig-repro_age).
+Reproductive declines therefore do not persist to the next generation.
 
@@
 #| label: fig-repro_age
-#| fig-cap: "Pup sex shifted towards males with increasing maternal age."
+#| fig-cap: "Pup sex skews slightly towards males with increasing maternal age. Pup survival decreases slightly, but survival and recruitment are not significantly affected by maternal age."
+#| fig-width: 6
+#| fig-height: 8
 
-sealdat %>% 
+#A: Pup sex shifts towards males with increasing maternal age.
+
+pup_sex <- sealdat %>% 
   filter(pupsex %in% c("M", "F"), 
@@
   ggplot(aes(x = age, y = percf, weight = num)) +
+  geom_hline(yintercept = 0.5, lty = 2) +
+  stat_smooth(method = "glm", 
+              method.args = list(family = "binomial"), 
+              formula = y ~ x, 
+              color = "#829BD1", 
+              fill = "#C8D3ED") +
   geom_point() +
-  geom_errorbar(aes(ymin = percf - se, ymax = percf + se), width = 0.2) +
-  geom_hline(yintercept = 0.5, lty = 2) +
-  stat_smooth(method = "glm", method.args = list(family = "binomial"), formula = y ~ x) +
-  geom_text(aes(label = num, vjust = 1, y = 0.03)) +
+  geom_errorbar(aes(ymin = percf - se, ymax = percf + se), width = 0) +
   scale_y_continuous(labels = scales::percent, limits = c(0, NA)) +
@@
   theme_few()
+
+#B Pup survival and recruitment doesn't change with increased maternal age.
+
+pup_surv_recruit <- sealdat %>% 
+  mutate(pup_surv = ifelse(pupseeneveragain > 0, 1, 0)) %>% 
+  filter(pupsex == "F") %>% 
+  group_by(age) %>%
+  summarize(n_surv = sum(!is.na(pup_surv)),
+            mean_surv = sum(pup_surv, na.rm = TRUE) / n_surv,  
+            se_surv = (mean_surv * (1 - mean_surv) / n_surv)^0.5,
+            n_recruit = sum(!is.na(puprecruited)), 
+            mean_recruit = sum(puprecruited, na.rm = TRUE) / n_recruit, 
+            se_recruit = (mean_recruit * (1 - mean_recruit) / n_recruit)^0.5)  %>%  
+  pivot_longer(cols = c(mean_surv, se_surv, mean_recruit, se_recruit), 
+               names_to = c("stat", "metric"), 
+               names_pattern = "(.*)_(.*)") %>% 
+  pivot_wider(names_from = "stat", 
+              values_from = "value") %>% 
+  mutate(low = mean - se, 
+         high = mean + se, 
+         age = age + ifelse(metric == "surv", -0.1, 0.1), 
+         metric = factor(metric, 
+                         levels = c("recruit", "surv"), 
+                         labels = c("Recruitment", "1st Year Survival"))) %>% 
+  filter(n_surv > 5) %>% 
+  ggplot(aes(x = age, y = mean, color = metric)) + 
+  geom_smooth(aes(weight = n_surv), 
+              method = "glm", 
+              method.args = list(family = binomial), 
+              formula = y ~ x, 
+              fill = "#C8D3ED") +
+  geom_pointrange(aes(ymin = low, ymax = high)) +
+  scale_color_manual(values = c("#829BD1", "#E9B668")) + 
+  scale_y_continuous(labels = scales::percent, limits = c(0, .5)) + 
+  labs(x = "Maternal Age (Years)", y = "% Pups") +
+  theme_few() + 
+  theme(legend.position = c(0.1,0.97), 
+        legend.justification = c(0,1), 
+        legend.title = element_blank(), 
+        legend.direction = "horizontal")
+
+plot_grid(pup_sex, pup_surv_recruit, ncol = 1, labels = "AUTO")
+
+```
+
+### H3: Changes to annual cycle timing with increased maternal age
+
+In order to create the timing figure, I need a csv with these columns: breeding arrival DOY, breeding departure DOY, molt arrival DOY, molt departure DOY.
+
+-   breeding arrival DOY = firstobsbreeddoy
+
+-   breeding departure DOY = firstobsbreeddoy + breeddur
+
+-   molt arrival DOY = lastobsmoltdoy - moltdur
+
+-   molt departure DOY = lastobsmoltdoy
+
+```{r}
+#| label: annual_timing
 

### 2023-09-20 scratch/128L 2023_06_25_Plots.r

@@ -3,8 +3,7 @@
 
 #################################################################################
 
-setwd("C:/Users/roxan/Documents/Teaching/Current Courses/BIOE 128L/128L 2023/128L Projects")
-final=read.csv("128L pull 2023_06_29.csv")
+final <- read_csv(here::here("data/raw/128L pull 2023_06_29.csv"))
 final$observed[final$observed=="B"]="Breeder"
 final$observed[final$observed==1]="Non-Breeder"
 

### 2023-09-20 paper/paper.qmd
Made the annual timing figure.

@@
 ### H1: Age distribution and breeding success
+
+We predicted that breeding success would decline with age. We found that breeding success increases to prime age, then decreases at older ages (@fig-age_dist).
 
@@
 
-As maternal age increases, pup sex skews slightly towards males.
-Pup survival declines slightly with maternal age, but pup recruitment is not significantly affected by maternal age (@fig-repro_age).
-Reproductive declines therefore do not persist to the next generation.
+We predicted that reproductive declines would carry over to the next generation.
+
+We found that as maternal age increases, pup sex skews slightly towards males.
+Pup survival declines slightly with maternal age, and pup recruitment is not significantly affected by maternal age (@fig-repro_age).
+
+We reject our hypothesis: reproductive declines do not persist to the next generation.
 
@@
 
-In order to create the timing figure, I need a csv with these columns: breeding arrival DOY, breeding departure DOY, molt arrival DOY, molt departure DOY.
+We predicted that seals would optimize the timing of their annual cycle.
 
--   breeding arrival DOY = firstobsbreeddoy
-
--   breeding departure DOY = firstobsbreeddoy + breeddur
-
--   molt arrival DOY = lastobsmoltdoy - moltdur
-
--   molt departure DOY = lastobsmoltdoy
+We found that as maternal age increases, there are slight adjustments to the timing of life history events.
+These events (breeding/molting arrival/departure) are typically shifted later for older animals (@fig-annual_timing).
 
 ```{r}
-#| label: annual_timing
+#| label: fig-annual_timing
+#| fig-cap: Timing of life history events across maternal age classes.
+
+sealdat %>% 
+  mutate(ba = firstobsbreeddoy, 
+         bd = firstobsbreeddoy + breeddur, 
+         ma = lastobsmoltdoy - moltdur, 
+         md = lastobsmoltdoy) %>% 
+  drop_na(age, ba, bd, ma, md) %>% 
+  group_by(age) %>% 
+  summarize(across(ba:md, mean)) %>% 
+  pivot_longer(cols = -age, 
+               names_to = "event", 
+               values_to = "doy") %>% 
+  group_by(age) %>% 
+  group_modify(\(data, keys) {
+    tibble(event = c("Long trip", "Breeding", "Short trip", "Molting", "longtrip2"),
+           doy = c(data$doy, 365) - c(0, data$doy))
+  }) %>% 
+  ungroup() %>% 
+  mutate(event = factor(event, levels = c("Long trip", 
+                                          "Breeding", 
+                                          "Short trip", 
+                                          "Molting", 
+                                          "longtrip2"))) %>% 
+  ggplot(aes(x = age, y = doy, fill = event)) + 
+  geom_area(position = position_stack(reverse = TRUE),
+            outline.type = "full",
+            color = "white") +
+  scale_fill_discrete(type = c("#829BD1", "#E9B668", "#7EC6CC", "#E091B9", "#829BD1"),
+                      breaks = c("Long trip", "Breeding", "Short trip", "Molting")) + 
+  labs(x = "Maternal Age (Years)", y = "Day of Year") + 
+  theme_few() + 
+  theme(legend.position = "right",
+        legend.title = element_blank())
 
@@
 ## Discussion
-
-##

### 2023-09-20 scratch/128L 2023_06_25_Plots.r

@@ -277,3 +277,33 @@ nd1=data.frame(age=3:21)
 nd1$pred=predict(f1,newdata=nd1,type="response")
 nd1$SE=predict(f1,newdata=nd1,type="response",se.fit=T)$se
 
+
+
+########################all models#########################
+
+require(lmerTest)
+
+#Molt duration model summaries
+moltdur_add <- summary(lmer(as.numeric(moltdur) ~ age + observed + (1 | animalID) + (1 | year),
+                            data = sealdat))
+moltdur_int <- summary(lmer(as.numeric(moltdur) ~ age * observed + (1 | animalID) + (1 | year),
+                            data = sealdat))
+
+#Molt departure date model summary
+moltdep <- summary(lmer(lastobsmoltdoy ~ age + observed + (1 | animalID) + (1 | year),
+                        data = sealdat))
+
+#Long trip duration model summary
+longtripdur_add <- summary(lmer(tripdur ~ age + observed + (1 | animalID) + (1 | year),
+                                data = sealdat))
+longtripdur_int <- lmer(tripdur ~ age * observed + (1 | animalID) + (1 | year),
+                        data = sealdat)
+emmeans::lstrends(longtripdur_int, "observed", var = "age", infer = T)
+emmeans(longtripdur_int, "observed", infer = T)
+longtripdur_int <- summary(longtripdur_int)
+
+#Breeding arrival date model summary
+brarr_add <- summary(lmer(firstobsbreeddoy ~ age + (1 | animalID) + (1 | year),
+                          data = sealdat))
+brarr_int <- summary(lmer(firstobsbreeddoy ~ age + observed + (1 | animalID) + (1 | year),
+                          data = sealdat))

### 2023-11-09 paper/paperv2.qmd
Catching up, MANY changes

---
title: "Elephant Seal Age JAE Intraspecific Variation"
format: 
  html: 
    echo: false
    fig-width: 4.5
    fig-height: 4.5
    embed-resources: true
    code-fold: true
    code-tools: true
    warning: false
editor: visual
bibliography: references.bib
---

```{r}
#| label: setup
#| include: false

library(ggthemes)
library(cowplot)
library(lme4)
library(lmerTest) # Loading this library will automatically append p-values to lmer models estimated using something called Satterthwaite's method
library(merTools)
library(tidyverse)

sealdat <- read_csv(here::here("data/raw/128L pull 2023_06_29.csv"), 
                    show_col_types = FALSE) %>% 
  mutate(observed = if_else(observed == "B", "Breeder", "Non-breeder"), 
         observed_int = if_else(observed == "Breeder", 1, 0), 
         pup_survived = pupseeneveragain > 0) %>% 
  filter(age > 3)

```

# Introduction

Our hypotheses:

1.  Elephant seal reproductive success declines with age.

2.  Reproductive declines persist to the next generation (maternal effect senescence).

3.  Seal phenology is shifted later for older animals and breeders.

**Note**: *This paper is similar to another paper about polar bears (*Ursus maritimus*)* [@naciri2022].

# Results

### H1: Age distribution and breeding status

We predicted that breeding success would decline with age. We found that breeding success increases to prime age, then decreases at older ages (@fig-age_dist).

**Note**: *statistics for increase in inter-birth interval?*

```{r}
#| label: fig-age_dist
#| fig-cap: "A: Age structure of northern elephant seals at Año Nuevo. B: Percentage of each age class that were observed breeding."
#| fig-width: 9

#Age distribution panel A

age_dist <- sealdat %>% 
  filter(age < 22) %>% 
  group_by(age) %>% 
  count(age) %>% 
  ggplot(aes(x = age, y = n)) + 
  geom_col(fill = "#829BD1") + 
  labs(x = "Number of individuals", y = "Age (Years)") + 
  theme_few()

#Panel B: percentage of each age class that were breeders or  nonbreeders

breed_nonbreed <- sealdat %>% 
  group_by(age) %>% 
  summarize(n_observed = sum(!is.na(observed)), 
            perc_breed = sum(observed == "Breeder") / n_observed, 
            se = (perc_breed * (1 - perc_breed) / n_observed)^0.5) %>% 
  filter(age < 22) %>%
  ggplot(aes(x = age, y = perc_breed)) + 
  stat_smooth(method = "lm", 
              formula = y ~ poly(x , 2), 
              se = TRUE, 
              color = "#849ED5", 
              fill = "#C8D3ED") +
  geom_point() + 
  geom_errorbar(aes(ymin = perc_breed - se, ymax = perc_breed + se), width = 0) +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) + 
  labs(x = "Maternal Age (Years)", y = "% Breeders") + 
  theme_few()

plot_grid(age_dist, breed_nonbreed, nrow = 1, labels = "AUTO")

```

Panel B shown with axis constrained below.

```{r}
#| label: fig-age_dist2
#| fig-cap: "B: Percentage of each age class that were observed breeding, with axes constrained."
#| fig-width: 9

#Panel B: percentage of each age class that were breeders or  nonbreeders

breed_nonbreed <- sealdat %>% 
  group_by(age) %>% 
  summarize(n_observed = sum(!is.na(observed)), 
            perc_breed = sum(observed == "Breeder") / n_observed, 
            se = (perc_breed * (1 - perc_breed) / n_observed)^0.5) %>% 
  filter(n_observed > 2) %>% 
  ggplot(aes(x = age, y = perc_breed)) + 
  stat_smooth(method = "lm", 
              formula = y ~ poly(x , 2), 
              se = TRUE, 
              color = "#849ED5", 
              fill = "#C8D3ED") +
  geom_point() + 
  geom_errorbar(aes(ymin = perc_breed - se, ymax = perc_breed + se), width = 0) +
  scale_y_continuous(labels = scales::percent, limits = c(.25, .75)) + 
  labs(x = "Maternal Age (Years)", y = "% Breeders") + 
  theme_few()

breed_nonbreed  
```

To test the quadratic fit of the model:

```{r}
summary(glmer(observed_int ~ poly(age, 2) + (1 | animalID) + (1 | year),
              data = sealdat, 
              family = "binomial"))
```

### H2: Maternal effect senescence

We predicted that reproductive declines would carry over to the next generation - that is, pups born to older mothers would have decreased survival and recruitment success.

We found that neither pup survival nor recruitment is significantly affected by maternal age (@fig-repro_age).

We reject our hypothesis: reproductive declines do not persist to the next generation, and there is no evidence of maternal effect senescence.

**To do:** *Add sample sizes to the plot because they are different than Figure 1.* *Do I need to do this if I have the sample sizes in the manuscript?*

**Note:** *Try to plot these quadraticlly??*

```{r}
#| label: fig-repro_age
#| fig-cap: "Pup survival decreases slightly, but survival and recruitment are not significantly affected by maternal age."
#| fig-width: 9

#B Pup survival and recruitment doesn't change with increased maternal age.

pup_surv_recruit <- sealdat %>% 
  mutate(pup_surv = ifelse(pupseeneveragain > 0, 1, 0)) %>% 
  filter(pupsex == "F", 
         pupyearborn < 2016) %>% 
  group_by(age) %>%
  summarize(n_surv = sum(!is.na(pup_surv)),
            mean_surv = sum(pup_surv, na.rm = TRUE) / n_surv,  
            se_surv = (mean_surv * (1 - mean_surv) / n_surv)^0.5,
            n_recruit = sum(!is.na(puprecruited)), 
            mean_recruit = sum(puprecruited, na.rm = TRUE) / n_recruit, 
            se_recruit = (mean_recruit * (1 - mean_recruit) / n_recruit)^0.5)  %>%  
  pivot_longer(cols = c(mean_surv, se_surv, mean_recruit, se_recruit), 
               names_to = c("stat", "metric"), 
               names_pattern = "(.*)_(.*)") %>% 
  pivot_wider(names_from = "stat", 
              values_from = "value") %>% 
  mutate(low = mean - se, 
         high = mean + se, 
         age = age + ifelse(metric == "surv", -0.1, 0.1), 
         metric = factor(metric, 
                         levels = c("recruit", "surv"), 
                         labels = c("Recruitment", "Survival"))) %>% 
  filter(n_surv > 5) %>% 
  ggplot(aes(x = age, y = mean, color = metric)) + 
  geom_pointrange(aes(ymin = low, ymax = high)) +
  scale_color_manual(values = c("#829BD1", "#E9B668")) + 
  scale_y_continuous(labels = scales::percent, limits = c(0, .5)) + 
  labs(x = "Maternal Age (Years)", y = "% Pups") +
  theme_few() + 
  theme(legend.position = c(0.1,0.97), 
        legend.justification = c(0,1), 
        legend.title = element_blank(), 
        legend.direction = "horizontal")

pup_surv_recruit
```

Results of the Pup Survival model, with individual and year as random effects:

```{r}
#Panel B Survival Model
surv_dat <- sealdat %>% filter(pupyearborn < 2016)
pup_survival_model <- glmer(pup_survived ~ age + (1 | animalID) + (1 | year),
                            data = surv_dat,
                            family = "binomial")
summary(pup_survival_model)

```

Results of the Pup Recruitment model, with individual and year as random effects:

**Note**: *Getting error "boundary (singular) fit" - possibly not enough data to fit this model? I.e., too few recruits. Also, need to apply 7 year filter to this one too????*

```{r}
#Panel B Recruitment Model

pup_recruit_model <- glmer(puprecruited ~ age + (1 | animalID) + (1 | year),
                           data = filter(sealdat, year > 1984), 
                           family = "binomial")
summary(pup_recruit_model)

```

The model of Survival \~ Maternal age was significant before we treated Year as a random effect. Was that because pups born earlier in the study had more time to be seen again (i.e., a sampling artifact?) If so, the plot of the random effect of year should have a negative slope.

```{r}
#| label: fig-surv_bias
#| fig-cap: "A: Plotting the intercept of the random effect of year shows that variation in ocean conditions is likely driving patterns in survival, rather than age of mother. B: 95% of seals that survived were resighted within ~seven years of their birth."
#| fig-width: 9

year_ranef <- ranef(pup_survival_model)$year %>% 
  as_tibble(rownames = "year") %>% 
  mutate(year = as.numeric(year)) %>% 
  rename(intercept = `(Intercept)`)

year_ranef_plot <- ggplot(year_ranef, aes(year, intercept)) + 
  geom_point() + 
  geom_smooth(method = "loess", 
              formula = y ~ x,
              color = "red") + 
  theme_classic()

survival_bias <- sealdat %>% 
  group_by(animalID) %>% 
  summarize(first_resight_yrs = min(age))

survival95 <- quantile(survival_bias$first_resight_yrs, 0.95)
#n for 95% quantile
sum(survival_bias$first_resight_yrs > 7)

survival_bias_plot <- ggplot(survival_bias, aes(x = first_resight_yrs)) +
  stat_ecdf() +
  geom_vline(xintercept = survival95, linetype = "dashed", color = "red") +
  scale_x_continuous(limits = c(0, 10)) + 
  theme_classic()

plot_grid(year_ranef_plot, survival_bias_plot, nrow = 1, labels = "AUTO")

```

There is a drop off in recent years that seems to support the idea that recent years have a sampling bias, but the earlier portion suggests that overall the variation is caused by ocean conditions and not sampling. See the Cohort paper!

If not maternal effect senescence, then perhaps a strategy shift?

### H3a: Changes in pup sex ratio with increased maternal age

We found that as maternal age increases, pup sex skews slightly towards males.

```{r}
#| label: fig-sex_ratio
#| fig-cap: "Pup sex skews slightly towards males with increasing maternal age, but it's not significant."
#| fig-width: 9
#A: Pup sex shifts towards males with increasing maternal age, but it's not significant.

pup_sex <- sealdat %>% 
  filter(pupsex %in% c("M", "F"), 
         age < 18) %>% 
  group_by(age, pupsex) %>%
  count(pupsex) %>% 
  pivot_wider(values_from = n, names_from = pupsex) %>% 
  mutate(num = M + F,
         percf = F / num,
         se = (percf * (1 - percf) / num)^0.5) %>% 
  ggplot(aes(x = age, y = percf, weight = num)) +
  geom_hline(yintercept = 0.5, lty = 2, color = "#829BD1") +
  geom_point() +
  geom_errorbar(aes(ymin = percf - se, ymax = percf + se), width = 0) +
  scale_y_continuous(labels = scales::percent, limits = c(0, NA)) +
  labs(x = "Maternal Age (Years)", y = "% Pups Female") +
  theme_few()

pup_sex
```

Results of the model for pup sex ratio:

```{r}
pup_sex_mod <- glmer(as.factor(pupsex) ~ age + (1 | animalID) + (1 | year),
                     data = filter(sealdat, 
                                   pupsex %in% c("M", "F"), 
                                   age < 18), 
                     family = "binomial")
summary(pup_sex_mod)
```

### H3b: Changes to annual cycle timing with increased maternal age

We predicted that seals would shift the timing of their annual cycle.

We found that as maternal age increases, there are slight adjustments to the timing of life history events. These events (breeding/molting arrival/departure) are typically shifted later for older animals (@fig-annual_timing).

```{r}

seal_phenology <- sealdat %>% 
  mutate(ba = firstobsbreeddoy, 
         bd = firstobsbreeddoy + breeddur, 
         ma = lastobsmoltdoy - moltdur, 
         md = lastobsmoltdoy) %>% 
  drop_na(age, ba, bd, ma, md)

phenology_n <- unique(seal_phenology$animalID))

timing_summary <- seal_phenology %>% 
  group_by(age, observed) %>% 
  summarize(across(ba:md, mean), 
            .groups = "drop") %>% 
  pivot_longer(cols = -c(age, observed),  
               names_to = "event", 
               values_to = "doy") %>% 
  group_by(age, observed) %>% 
  group_modify(\(data, keys) {
    tibble(event = c("Long trip", "Breeding", "Short trip", "Molting", "longtrip2"),
           doy = c(data$doy, 365) - c(0, data$doy))
  }) %>% 
  ungroup() %>% 
  mutate(event = factor(event, levels = c("Long trip", 
                                          "Breeding", 
                                          "Short trip", 
                                          "Molting", 
                                          "longtrip2")))
timing_summary_flipped <- timing_summary %>% 
  mutate(observed = ifelse(observed == "Breeder",
                           "Non-breeder",
                           "Breeder")) %>% 
  filter(event != "longtrip2")

```

**Note:** *Add in illustrator the colors between each phenological variable, so people can understand the timing. Fix the labels on panel B and write a caption.*

```{r}
#| label: fig-phenology
#| fig-height: 9
#| fig-cap: "Phenology caption"

phen_plot1 <- ggplot(timing_summary, aes(x = age, y = doy, fill = event)) + 
  geom_area(position = position_stack(reverse = TRUE),
            outline.type = "full",
            color = "white") +
  geom_line(data = timing_summary_flipped,
            position = position_stack(reverse = TRUE),
            color = "white",
            linetype = "dashed",
            alpha = 0.5) +
  scale_fill_discrete(type = c("#829BD1", "#E9B668", "#7EC6CC", "#E091B9", "#829BD1"),
                      breaks = c("Long trip", "Breeding", "Short trip", "Molting")) + 
  facet_grid(cols = vars(observed)) +
  labs(x = "Maternal Age (Years)", y = "Day of Year") + 
  theme_few() + 
  theme(legend.position = "bottom",
        legend.title = element_blank())

# Utility functions for model fitting and coefficient extraction
fit_model <- function(dat) {
  lmer(doy ~ age + observed + (1 | animalID) + (1 | year), data = dat)
}
extract_estimate <- function(models, param) {
  map_dbl(models, \(mod) summary(mod)$coefficients[param, "Estimate"])
}
extract_stderr <- function(models, param) {
  map_dbl(models, \(mod) summary(mod)$coefficients[param, "Std. Error"])
}
extract_pval <- function(models, param) {
  map_dbl(models, \(mod) summary(mod)$coefficients[param, "Pr(>|t|)"])
}

# This data frame contains a model for each phenological variable w.r.t. age
phenology_models <- seal_phenology %>% 
  pivot_longer(ba:md, names_to = "phenology", values_to = "doy") %>% 
  select(phenology, observed, doy, age, animalID, year) %>% 
  mutate(observed = factor(observed, levels = c("Non-breeder", "Breeder"))) %>% 
  nest(model_data = c(doy, age, observed, animalID, year)) %>% 
  mutate(model = map(model_data, fit_model),
         age_coef = extract_estimate(model, "age"),
         age_stderr = extract_stderr(model, "age"),
         age_pval = extract_pval(model, "age"),
         age_signif = age_pval * 8 < 0.05,
         breeder_coef = extract_estimate(model, "observedBreeder"),
         breeder_stderr = extract_stderr(model, "observedBreeder"),
         breeder_pval = extract_pval(model, "observedBreeder"),
         breeder_signif = breeder_pval * 8 < 0.05,)

phen_plot2 <- phenology_models %>% 
  select(-c(model_data, model)) %>% 
  pivot_longer(age_coef:breeder_stderr,
               names_to = c("param", "metric"),
               names_sep = "_",
               values_to = "value") %>% 
  pivot_wider(names_from = metric, values_from = value) %>% 
  ggplot(aes(x = coef, y = phenology)) +
  geom_pointrange(aes(xmin = coef - stderr, xmax = coef + stderr)) +
  expand_limits(x = 0) +
  facet_grid(cols = vars(param), scales = "free_x") +
  theme_classic()

plot_grid(phen_plot1, phen_plot2, nrow = 2, labels = "AUTO")

phenology_models %>% 
  select(-c(model_data, model)) %>% 
  pivot_longer(age_coef:breeder_signif, 
               names_to = c("param", "metric"),
               names_sep = "_",
               values_to = "value") %>% 
  pivot_wider(names_from = metric, values_from = value) %>% 
  mutate(coef = str_glue("{round(coef, 2)} +- {round(stderr, 2)}"),
         signif = str_glue("p = {round(pval, 4)} ({signif == 1})")) %>% 
  select(-c(stderr, pval)) %>% 
  pivot_wider(names_from = param, values_from = c(coef, signif)) %>% 
  select(phenology, ends_with("age"), ends_with("breeder")) %>% 
  knitr::kable()

```

## Discussion

### 2023-11-09 scratch/paternity.R

@@ -1,0 +1,53 @@
+# Is this the paper you do with Millie?
+
+seal_agent <- function(
+    # Parameterize mom's life history here
+    ) {
+  seal <- list(
+    age = 3,
+    sex = 0, # female
+    pups = 0,
+    grandpups = 0,
+    recruits = 0,
+    alive = TRUE
+  )
+  while (seal$alive) {
+    seal <- seal_agent_year(seal)
+  }
+  seal
+}
+
+seal_agent_year <- function(seal) {
+  survived <- rbinom(1, 1, some_probability(seal$age)) # from the age paper
+  bred <- rbinom(1, 1, some_probability(seal$age)) # from the age paper
+  if (bred == 1) {
+    pup = list(
+      # this is the strategy switching part
+      sex = rbinom(1, 1, some_probability(seal$age))
+    )
+    if (pup$sex == 0) { 
+      # if female
+      pup$recruited <- rbinom(1, 1, some_probability)
+      pup$lifetime_repro <- some_distribution() # parameterize using whole dataset?
+    } else {
+      # if male
+      pup$recruited <- one_in_nine_maybe() # ???
+      pup$lifetime_repro <- any_clue() # ???
+    }
+    seal$pups <- seal$pups + 1
+    seal$recruits <- seal$recruits + pup$recruited
+    seal$grandpups <- seal$grandpups + pup$lifetime_repro
+  }
+  if (survived) {
+    seal$age <- seal$age + 1
+  } else {
+    seal$alive <- FALSE
+  }
+  seal
+}
+
+##pup sex ratio change - taking on a riskier strategy as they get older. well established phenomenoon! 
+#seems to be a case of late-in-life change towards a riskier strategy, not maternal effect senescence
+#captial breeder? 
+#fitness benefits depend on the risk reward ratio for male vs female. we dont know this yet because we only track females
+#future work on paternity would allow us to quantitatively test optimal life history strategies in this model system

### 2023-11-09 paper/paperv2.qmd
Getting the doc ready to be a web page

@@
     fig-height: 4.5
-    embed-resources: true
     code-fold: true

### 2023-11-09 paper/paperv2.qmd
fixes to phenology figure

@@
 
-phenology_n <- unique(seal_phenology$animalID))
+phenology_n <- unique(seal_phenology$animalID)
 
@@
   geom_area(position = position_stack(reverse = TRUE),
-            outline.type = "full",
+            color = NA) +
+  geom_line(position = position_stack(reverse = TRUE),
             color = "white") +
@@
   labs(x = "Maternal Age (Years)", y = "Day of Year") + 
-  theme_few() + 
-  theme(legend.position = "bottom",
-        legend.title = element_blank())
+  theme_few()
 
@@
 
+phen_wrapped <- tibble(
+  event_begin = c(-Inf, 1:4),
+  event_end = c(1:4, Inf),
+  event = timing_summary$event[0:5]
+)
+
 phen_plot2 <- phenology_models %>% 
   select(-c(model_data, model)) %>% 
-  pivot_longer(age_coef:breeder_stderr,
+  pivot_longer(c(ends_with("coef"), ends_with("stderr"), ends_with("signif")),
                names_to = c("param", "metric"),
@@
   pivot_wider(names_from = metric, values_from = value) %>% 
+  mutate(signif = as.logical(signif)) %>% 
   ggplot(aes(x = coef, y = phenology)) +
-  geom_pointrange(aes(xmin = coef - stderr, xmax = coef + stderr)) +
+  # Prevent geom_rect() from establishing a continuous y-axis scale
+  geom_point(alpha = 0) +
+  geom_rect(aes(xmin = -Inf, xmax = Inf,
+                ymin = event_begin, ymax = event_end,
+                fill = event),
+            phen_wrapped,
+            inherit.aes = FALSE) +
+  geom_pointrange(aes(xmin = coef - stderr, 
+                      xmax = coef + stderr,
+                      color = signif)) +
+  scale_color_manual(values = c("FALSE" = "grey50", "TRUE" = "black"),
+                     guide = "none") +
+  scale_fill_discrete(type = c("#829BD1", "#E9B668", "#7EC6CC", "#E091B9", "#829BD1"),
+                      breaks = c("Long trip", "Breeding", "Short trip", "Molting")) + 
   expand_limits(x = 0) +
   facet_grid(cols = vars(param), scales = "free_x") +
-  theme_classic()
+  theme_few()
 
-plot_grid(phen_plot1, phen_plot2, nrow = 2, labels = "AUTO")
+phen_plot2
+
+library(patchwork)
+phen_plot1 /
+  phen_plot2 +
+  plot_annotation(tag_levels = "A") +
+  plot_layout(guides = "collect") &
+  theme(legend.position = "top",
+        legend.title = element_blank())
+
+plot_grid(phen_plot1, phen_plot2, nrow = 2, labels = "AUTO", align = "v")
 

### 2023-11-09 paper/paperv2.qmd
Fixing the plot labels.

@@
 library(tidyverse)
+library(patchwork)
 
@@
 seal_phenology <- sealdat %>% 
-  mutate(ba = firstobsbreeddoy, 
-         bd = firstobsbreeddoy + breeddur, 
-         ma = lastobsmoltdoy - moltdur, 
-         md = lastobsmoltdoy) %>% 
-  drop_na(age, ba, bd, ma, md)
+  mutate(BA = firstobsbreeddoy, 
+         BD = firstobsbreeddoy + breeddur, 
+         MA = lastobsmoltdoy - moltdur, 
+         MD = lastobsmoltdoy) %>% 
+  drop_na(age, BA, BD, MA, MD)
 
@@
   group_by(age, observed) %>% 
-  summarize(across(ba:md, mean), 
+  summarize(across(BA:MD, mean), 
             .groups = "drop") %>% 
@@
 phenology_models <- seal_phenology %>% 
-  pivot_longer(ba:md, names_to = "phenology", values_to = "doy") %>% 
+  pivot_longer(BA:MD, names_to = "phenology", values_to = "doy") %>% 
   select(phenology, observed, doy, age, animalID, year) %>% 
@@
   pivot_wider(names_from = metric, values_from = value) %>% 
-  mutate(signif = as.logical(signif)) %>% 
+  mutate(signif = as.logical(signif), 
+         param = factor(param, 
+                        labels = c("Maternal Age (Years)", 
+                                   "Breeding Status"))) %>% 
   ggplot(aes(x = coef, y = phenology)) +
@@
   facet_grid(cols = vars(param), scales = "free_x") +
+  xlab("LMM coefficeint") + 
+  ylab("Phenology") + 
   theme_few()
 
-phen_plot2
-
-library(patchwork)
 phen_plot1 /
@@
         legend.title = element_blank())
-
-plot_grid(phen_plot1, phen_plot2, nrow = 2, labels = "AUTO", align = "v")
 
@@
 ```
-
-## Discussion

### 2023-11-09 paper/paperv2.qmd
Phenology figure label tweaks

@@
 #| label: fig-phenology
-#| fig-height: 9
+#| fig-height: 7
 #| fig-cap: "Phenology caption"
@@
          param = factor(param, 
-                        labels = c("Maternal Age (Years)", 
+                        labels = c("Maternal Age", 
                                    "Breeding Status"))) %>% 

### 2023-11-13 paper/paperv2.qmd
Changing the styling of phenology table

@@
   select(phenology, ends_with("age"), ends_with("breeder")) %>% 
-  knitr::kable()
+  knitr::kable(col.names = c("Phenology", 
+                             "Age coef. (days/year)", 
+                             "Significance", 
+                             "Breeding status coef. (days)", 
+                             "Significance")) %>% 
+  kableExtra::kable_styling()
 

### 2023-11-13 paper/paperv2.qmd
modeling offspring metrics quadratically

@@
 We predicted that breeding success would decline with age. We found that breeding success increases to prime age, then decreases at older ages (@fig-age_dist).
-
-**Note**: *statistics for increase in inter-birth interval?*
 
@@
 
-**To do:** *Add sample sizes to the plot because they are different than Figure 1.* *Do I need to do this if I have the sample sizes in the manuscript?*
-
 **Note:** *Try to plot these quadraticlly??*
@@
 ```
+
+Modeling the above metrics as quadratic, just in case:
+
+```{r}
+surv_dat <- sealdat %>% filter(pupyearborn < 2016)
+pup_survival_model <- glmer(pup_survived ~ poly(age, 2) + (1 | animalID) + (1 | year),
+                            data = surv_dat,
+                            family = "binomial")
+summary(pup_survival_model)
+
+pup_recruit_model <- glmer(puprecruited ~ poly(age, 2) + (1 | animalID) + (1 | year),
+                           data = filter(sealdat, year > 1984), 
+                           family = "binomial")
+summary(pup_recruit_model)
+
+```
+
+Neither were significant as 2nd order polynomials.
 

### 2023-11-15 paper/paperv2.qmd
Fixing and saving figure 1

@@
   geom_errorbar(aes(ymin = perc_breed - se, ymax = perc_breed + se), width = 0) +
-  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) + 
+  scale_y_continuous(labels = scales::percent, limits = c(.25, .75)) + 
   labs(x = "Maternal Age (Years)", y = "% Breeders") + 
@@
 
-plot_grid(age_dist, breed_nonbreed, nrow = 1, labels = "AUTO")
+ggsave("outputs/agefig1.png", plot = plot_grid(age_dist, breed_nonbreed, nrow = 1, labels = "AUTO"), height = 4, width = 9)
 
@@
 pup_recruit_model <- glmer(puprecruited ~ age + (1 | animalID) + (1 | year),
-                           data = filter(sealdat, year > 1984), 
+                           data = filter(sealdat, year > 1984 & year < 2019), 
                            family = "binomial")
@@
 pup_recruit_model <- glmer(puprecruited ~ poly(age, 2) + (1 | animalID) + (1 | year),
-                           data = filter(sealdat, year > 1984), 
+                           data = filter(sealdat, year > 1984 & year < 2019), 
                            family = "binomial")
@@
   filter(pupsex %in% c("M", "F"), 
-         age < 18) %>% 
+         age < 20) %>% 
   group_by(age, pupsex) %>%

### 2023-11-15 paper/paperv2.qmd
Streamlining the html site

@@
 
-Panel B shown with axis constrained below.
-
-```{r}
-#| label: fig-age_dist2
-#| fig-cap: "B: Percentage of each age class that were observed breeding, with axes constrained."
-#| fig-width: 9
-
-#Panel B: percentage of each age class that were breeders or  nonbreeders
-
-breed_nonbreed <- sealdat %>% 
-  group_by(age) %>% 
-  summarize(n_observed = sum(!is.na(observed)), 
-            perc_breed = sum(observed == "Breeder") / n_observed, 
-            se = (perc_breed * (1 - perc_breed) / n_observed)^0.5) %>% 
-  filter(n_observed > 2) %>% 
-  ggplot(aes(x = age, y = perc_breed)) + 
-  stat_smooth(method = "lm", 
-              formula = y ~ poly(x , 2), 
-              se = TRUE, 
-              color = "#849ED5", 
-              fill = "#C8D3ED") +
-  geom_point() + 
-  geom_errorbar(aes(ymin = perc_breed - se, ymax = perc_breed + se), width = 0) +
-  scale_y_continuous(labels = scales::percent, limits = c(.25, .75)) + 
-  labs(x = "Maternal Age (Years)", y = "% Breeders") + 
-  theme_few()
-
-breed_nonbreed  
-```
-
 To test the quadratic fit of the model:
@@
 We reject our hypothesis: reproductive declines do not persist to the next generation, and there is no evidence of maternal effect senescence.
-
-**Note:** *Try to plot these quadraticlly??*
 
@@
 
-**Note**: *Getting error "boundary (singular) fit" - possibly not enough data to fit this model? I.e., too few recruits. Also, need to apply 7 year filter to this one too????*
-
 ```{r}
@@
 Was that because pups born earlier in the study had more time to be seen again (i.e., a sampling artifact?) If so, the plot of the random effect of year should have a negative slope.
+
+We also plotted survival bias to show that 95% of pups were resighted within 7 years. Both plots are supplemental.
 
@@
 
-If not maternal effect senescence, then perhaps a strategy shift?
-
 ### H3a: Changes in pup sex ratio with increased maternal age
 
-We found that as maternal age increases, pup sex skews slightly towards males.
+We found that as maternal age increases, pup sex skews slightly towards males - but it's not significant.
 
@@
   filter(pupsex %in% c("M", "F"), 
-         age < 20) %>% 
+         age < 19) %>% 
   group_by(age, pupsex) %>%
@@
                                    pupsex %in% c("M", "F"), 
-                                   age < 18), 
+                                   age < 19), 
                      family = "binomial")
@@
 
-**Note:** *Add in illustrator the colors between each phenological variable, so people can understand the timing. Fix the labels on panel B and write a caption.*
+The table is supplemental.
 
@@
 #| fig-height: 7
-#| fig-cap: "Phenology caption"
+#| fig-cap: "Successful breeding and increased maternal age both impact elephant seal annual cycles.
+A: Visualization of life history timing across ages, faceted by breeding status.
+White dashed lines on each panel represent the raw data from the other panel to facilitate comparisons between breeders and non-breeders.
+B: LMM coefficients for phenological parameters.
+The units for maternal age coefficients are in days per year.
+The coefficients for breeding status are in days.
+BA = Breeding arrival; BD = Breeding departure; MA = Molt arrival; MD = Molt departure.
+Black points are significant, gray points are not significant."
 

### 2023-11-24 paper/paperv2.qmd
Fixing the figures in the html
Fixing axis font sizes

@@
   geom_col(fill = "#829BD1") + 
-  labs(x = "Number of individuals", y = "Age (Years)") + 
-  theme_few()
+  labs(x = "Age (Years)", y = "Number of individuals") + 
+  theme_few() + 
+  theme(axis.title = element_text(size = 13))
 
@@
   scale_y_continuous(labels = scales::percent, limits = c(.25, .75)) + 
-  labs(x = "Maternal Age (Years)", y = "% Breeders") + 
-  theme_few()
+  labs(x = "Maternal Age (Years)", y = "% Breeders (Mean ± SE)") + 
+  theme_few() + 
+  theme(axis.title = element_text(size = 13))
 
-ggsave("outputs/agefig1.png", plot = plot_grid(age_dist, breed_nonbreed, nrow = 1, labels = "AUTO"), height = 4, width = 9)
+plot_grid(age_dist, breed_nonbreed, nrow = 1, labels = "AUTO")
+
+#ggsave("outputs/agefig1.png", plot = , height = 4, width = 9)
 
@@
         legend.title = element_blank(), 
-        legend.direction = "horizontal")
+        legend.direction = "horizontal", 
+        axis.title = element_text(size = 13)
+)
 
@@
   labs(x = "Maternal Age (Years)", y = "% Pups Female") +
-  theme_few()
+  theme_few() + 
+  theme(axis.title = element_text(size = 13))
 
@@
   labs(x = "Maternal Age (Years)", y = "Day of Year") + 
-  theme_few()
+  theme_few() + 
+  theme(axis.title = element_text(size = 13))
 
@@
   ylab("Phenology") + 
-  theme_few()
+  theme_few() + 
+  theme(axis.title = element_text(size = 13))
 
@@
         legend.title = element_blank())
+
+```
+
+Supplemental table
+
+```{r}
 

### 2023-11-24 scratch/phen_predict.R

@@ -1,0 +1,23 @@
+
+expand_grid(
+  age = 9:19,
+  observed = c("Breeder", "Non-breeder")
+) %>% 
+  mutate(ba = predict(phenology_models$model[[1]],
+                      newdata = .,
+                      re.form = NA),
+         bd = predict(phenology_models$model[[2]],
+                      newdata = .,
+                      re.form = NA), 
+         ma = predict(phenology_models$model[[3]],
+                      newdata = .,
+                      re.form = NA), 
+         md = predict(phenology_models$model[[4]],
+                      newdata = .,
+                      re.form = NA)) %>% 
+  mutate(short_trip = ma - bd, 
+         molt_dur = md - ma) %>% 
+  filter(observed == "Breeder", 
+         age %in% c(9,19)) %>% 
+  view()
+

### 2023-11-25 paper/paperv2.qmd
Figure updates

--- 2023-11-15 paper/paperv2.qmd
@@
 library(tidyverse)
+library(ggeffects)
 library(patchwork)
 
-sealdat <- read_csv(here::here("data/raw/128L pull 2023_06_29.csv"), 
+sealdat <- read_csv(here::here("data/raw/128L pull 2023_11_21.csv"), 
                     show_col_types = FALSE) %>% 
@@
 ```{r}
+#| label: age_breeding_mod
+
+breeding_mod_bad <- glmer(observed_int ~ poly(age, 2) + (1 | animalID) + (1 | year),
+                      data = filter(sealdat), 
+                      family = "binomial")
+
+count(sealdat, year) %>% 
+     left_join(as_tibble(ranef(breeding_mod_bad)$year, 
+                         rownames = "year") %>% 
+                   mutate(year = as.numeric(year)) %>% 
+                   rename(coef = `(Intercept)`),
+               by = "year") %>%
+     ggplot(aes(n, coef, color = year)) +
+     geom_point() +
+     scale_color_gradient(low = "blue", high = "red") +
+     theme_few()
+
+breeding_mod <- glmer(observed_int ~ poly(age, 2) + (1 | animalID) + (1 | year),
+                      data = filter(sealdat, year >= 1990), 
+                      family = "binomial")
+breed_pred <- ggpredict(breeding_mod, terms = c("age [all]"))
+summary(breeding_mod)
+
+```
+
+```{r}
 #| label: fig-age_dist
@@
 age_dist <- sealdat %>% 
-  filter(age < 22) %>% 
+  filter(age < 22,
+         year >= 1990) %>% 
   group_by(age) %>% 
@@
 breed_nonbreed <- sealdat %>% 
+  filter(year >= 1990) %>% 
   group_by(age) %>% 
@@
   ggplot(aes(x = age, y = perc_breed)) + 
-  stat_smooth(method = "lm", 
-              formula = y ~ poly(x , 2), 
-              se = TRUE, 
-              color = "#849ED5", 
-              fill = "#C8D3ED") +
+  geom_ribbon(aes(x, ymin = conf.low, ymax = conf.high),
+              breed_pred,
+              inherit.aes = FALSE,
+              fill = "#829BD1",
+              alpha = 0.25) +
+  geom_line(aes(x, predicted),
+            breed_pred,
+            inherit.aes = FALSE,
+            color = "#829BD1",
+            linewidth = 1) +
   geom_point() + 
   geom_errorbar(aes(ymin = perc_breed - se, ymax = perc_breed + se), width = 0) +
-  scale_y_continuous(labels = scales::percent, limits = c(.25, .75)) + 
-  labs(x = "Maternal Age (Years)", y = "% Breeders") + 
-  theme_few()
+  scale_y_continuous(labels = scales::percent, limits = c(0,1)) + 
+  labs(x = "Maternal Age (Years)", y = "% Breeders (Mean ± SE)") + 
+  theme_few() + 
+  theme(axis.title = element_text(size = 13))
 
-ggsave("outputs/agefig1.png", plot = plot_grid(age_dist, breed_nonbreed, nrow = 1, labels = "AUTO"), height = 4, width = 9)
+plot_grid(age_dist, breed_nonbreed, nrow = 1, labels = "AUTO")
+
+ggsave("outputs/agefig1.png", plot = , height = 4, width = 9)
 
@@
 
-To test the quadratic fit of the model:
+Population level prediction - the fact that the raw data falls above it comes from the fact that those age classes are correlated with years and individuals that were very successful.
 
 ```{r}
-summary(glmer(observed_int ~ poly(age, 2) + (1 | animalID) + (1 | year),
-              data = sealdat, 
-              family = "binomial"))
+#| label: fig-breeding_bias
+
+year_coef <- as_tibble(ranef(breeding_mod)$year, 
+                       rownames = "year") %>% 
+    mutate(year = as.numeric(year)) %>% 
+    rename(coef = `(Intercept)`)
+
+animal_coef <- as_tibble(ranef(breeding_mod)$animalID, 
+                         rownames = "animalID") %>% 
+    mutate(animalID = as.numeric(animalID)) %>% 
+    rename(coef = `(Intercept)`)
+
+year_bias <- sealdat %>% 
+  left_join(year_coef, by = "year") %>% 
+  mutate(age = factor(age)) %>% 
+  ggplot(aes(age, coef)) +
+  geom_boxplot() +
+  geom_hline(yintercept = 0, color = "#829BD1", linetype = "dashed") +
+  scale_x_discrete("Maternal Age (Years)", breaks = seq(5, 20, by = 5)) +
+  ylab("Random effect of year") +
+  theme_few()
+
+animalID_bias <- sealdat %>% 
+  left_join(animal_coef, by = "animalID") %>% 
+  mutate(age = factor(age)) %>% 
+  ggplot(aes(age, coef)) +
+  geom_boxplot() +
+  geom_hline(yintercept = 0, color = "#829BD1", linetype = "dashed") +
+  scale_x_discrete("Maternal Age (Years)", breaks = seq(5, 20, by = 5)) +  
+  ylab("Random effect of individual") +
+  theme_few()
+
+plot_grid(breed_nonbreed, year_bias, animalID_bias, nrow = 1)
+
 ```
@@
   scale_y_continuous(labels = scales::percent, limits = c(0, .5)) + 
-  labs(x = "Maternal Age (Years)", y = "% Pups") +
+  labs(x = "Maternal Age (Years)", y = "% Pups (Mean ± SE)") +
   theme_few() + 
@@
   scale_y_continuous(labels = scales::percent, limits = c(0, NA)) +
-  labs(x = "Maternal Age (Years)", y = "% Pups Female") +
-  theme_few()
+  labs(x = "Maternal Age (Years)", y = "% Female Pups (Mean ± SE)") +
+  theme_few() + 
+  theme(axis.title = element_text(size = 13))
 
@@
 We found that as maternal age increases, there are slight adjustments to the timing of life history events.
-These events (breeding/molting arrival/departure) are typically shifted later for older animals (@fig-annual_timing).
+These events (breeding/molting arrival/departure) are typically shifted later for older animals (@fig-phenology).
 
@@
 #| fig-height: 7
+#| fig-width: 7
 #| fig-cap: "Successful breeding and increased maternal age both impact elephant seal annual cycles.
@@
 The coefficients for breeding status are in days.
-BA = Breeding arrival; BD = Breeding departure; MA = Molt arrival; MD = Molt departure.
 Black points are significant, gray points are not significant."
@@
                       breaks = c("Long trip", "Breeding", "Short trip", "Molting")) + 
+  scale_x_continuous("Maternal Age (Years)", expand = expansion(mult = 0)) +
+  scale_y_continuous("Day of Year", expand = expansion(mult = 0)) +
   facet_grid(cols = vars(observed)) +
-  labs(x = "Maternal Age (Years)", y = "Day of Year") + 
-  theme_few()
+  theme_few() + 
+  theme(axis.title = element_text(size = 12),
+        panel.spacing = unit(1, "lines"))
 
@@
          param = factor(param, 
-                        labels = c("Maternal Age", 
-                                   "Breeding Status"))) %>% 
+                        labels = c("'Delay with maternal age (days year'^-1 * ')'", 
+                                   "'Delay with breeding (days)'")), 
+         phenology = factor(phenology, 
+                            labels = c("Breeding\narrrival", 
+                                       "Breeding\ndeparture", 
+                                       "Molt\narrrival", 
+                                       "Molt\ndeparture"))) %>% 
   ggplot(aes(x = coef, y = phenology)) +
@@
   expand_limits(x = 0) +
-  facet_grid(cols = vars(param), scales = "free_x") +
-  xlab("LMM coefficeint") + 
-  ylab("Phenology") + 
-  theme_few()
+  facet_grid(cols = vars(param), 
+             scales = "free_x", 
+             switch = "x",
+             labeller = label_parsed) +
+  theme_few() + 
+  theme(axis.title = element_blank(),
+        legend.direction = "horizontal",
+        legend.title = element_blank(),
+        strip.placement = "outside",
+        strip.text = element_text(size = 12, color = "black"),
+        panel.spacing = unit(1, "lines"))
 
-phen_plot1 /
-  phen_plot2 +
-  plot_annotation(tag_levels = "A") +
-  plot_layout(guides = "collect") &
-  theme(legend.position = "top",
-        legend.title = element_blank())
+plot_grid(
+  get_legend(phen_plot2),
+  phen_plot1 + theme(legend.position = "none"),
+  phen_plot2 + theme(legend.position = "none"),
+  align = "v",
+  axis = "lr",
+  labels = c("", "A", "B"),
+  ncol = 1,
+  rel_heights = c(0.1, 1, 1)
+)
+
+ggsave("outputs/agefig4.png", plot = , height = 9, width = 9)
+
+```
+
+Supplemental table
+
+```{r}
 

### 2023-11-27 paper/paperv2.qmd
adding labels to supplemental plot

@@
                by = "year") %>%
-     ggplot(aes(n, coef, color = year)) +
+     ggplot(aes(n, coef, color = year, label = year)) +
      geom_point() + 
+  geom_label() + 
      scale_color_gradient(low = "blue", high = "red") + 

### 2023-11-27 paper/paperv2.qmd
Updating to post-1994

@@
          pup_survived = pupseeneveragain > 0) %>% 
-  filter(age > 3)
+  filter(age > 3, 
+         year > 1994)
 
@@
 breeding_mod <- glmer(observed_int ~ poly(age, 2) + (1 | animalID) + (1 | year),
-                      data = filter(sealdat, year >= 1990), 
+                      data = filter(sealdat), 
                       family = "binomial")
@@
 age_dist <- sealdat %>% 
-  filter(age < 22,
-         year >= 1990) %>% 
+  filter(age < 22) %>% 
   group_by(age) %>% 
@@
 breed_nonbreed <- sealdat %>% 
-  filter(year >= 1990) %>% 
   group_by(age) %>% 

### 2023-11-27 scratch/phen_predict.R

@@ -1,6 +1,6 @@
 
 expand_grid(
-  age = 9:19,
+  age = 4:19,
   observed = c("Breeder", "Non-breeder")
 ) %>% 
   mutate(ba = predict(phenology_models$model[[1]],
@@ -18,6 +18,6 @@ expand_grid(
   mutate(short_trip = ma - bd, 
          molt_dur = md - ma) %>% 
   filter(observed == "Breeder", 
-         age %in% c(9,19)) %>% 
+         age %in% c(4,19)) %>% 
   view()
 

### 2023-12-01 00_clean_data.R
Data cleaning script from Roxanne

@@ -1,0 +1,179 @@
+# load packages
+library(RMySQL)
+library(lubridate)
+library(tidyverse)
+library(patchwork)
+library(marked)
+library(RMark)
+
+animal <- read_csv(here::here("data/raw/animalforallison.csv"))
+resight <- read_csv(here::here("data/raw/resightforallison.csv"))
+
+# specify which animals (pups) to select
+animals_to_select <- animal %>%
+  filter(
+    # from Año Nuevo
+    region == "AN",
+    # tagged as pup/wean
+    broad == "YOY",
+    # with a flipper tag
+    !is.na(origtag),
+    # with a correct date
+    !is.na(year(tagdate))
+  ) #both sex
+
+pups=subset(animals_to_select,momID!=0) #subset of pups where mom ID is known
+pups=pups[,c("animalID","tagsex","momID","tagdate")] #some columns from the pup dataset
+colnames(pups)=c("pupID","pupsex","animalID","tagdate") #rename the columns
+pups$year=year(pups$tagdate) #find year that the pup was tagged
+
+#get rid of pups with no tag sex (149 cases out of ~4,000)
+pups=pups %>% subset(pupsex=="M"|pupsex=="F")
+
+# final table for moms
+outall <- merge(# merge resight
+  resight,
+  # and animals to select (only tagdate, tagsex, pupID, and animalID columns)
+  animals_to_select %>% dplyr::select(tagdate,animalID,momID,tagsex),
+  # by animalID, which is the mom
+  by = "animalID") %>%
+  # create a new column yearborn
+  mutate(yearborn = year(tagdate),
+         yday=yday(date),
+         timeofyear=if_else(yday>274|yday<74,"Breeding", #between Dec 1 and Mar 15
+                            if_else(yday>=74&yday<182,"Molt","Other"))) %>%   #between Mar 15 and July 1
+  arrange(date)
+
+outall$calyear=year(outall$date) #make a new column for the calendar year of that observation
+
+#correct pup status for weird character values
+outall$withpupCor=NA
+outall$withpupCor[outall$withpup%in%c(as.character(1:8),"2+","3+",">8","1 or 2",">1")]=1
+outall$withpupCor[outall$withpup==0]=0
+
+#Encounter History
+g1=as.data.frame.matrix(with(outall,table(animalID,calyear))) #table : combinations of observation year and animal ID
[…]

### 2023-12-01 paper/paperv2_old.qmd
Updating the paper with Roxanne's feedback

--- 2023-11-27 paper/paperv2.qmd
@@
          pup_survived = pupseeneveragain > 0) %>% 
+  select(-c("tagsex", "season", "numbreed", "firstyearindicator", "indices")) %>% 
   filter(age > 3, 
@@
 
-We also plotted survival bias to show that 95% of pups were resighted within 7 years. Both plots are supplemental.
+We also plotted survival bias to show that 95% of pups were resighted within 10 years. Both plots are supplemental.
 
@@
 #n for 95% quantile
-sum(survival_bias$first_resight_yrs > 7)
+sum(survival_bias$first_resight_yrs > 10)
 

### 2023-12-01 paper/paperv3.qmd
Updating the paper with Roxanne's feedback

--- 2023-11-27 paper/paperv2.qmd
@@
 #| include: false  
-
 library(ggthemes)
@@
 
-sealdat <- read_csv(here::here("data/raw/128L pull 2023_11_21.csv"), 
+sealdat <- read_csv(here::here("data/raw/128L pull 2023_11_30.csv"), 
                     show_col_types = FALSE) %>% 
@@
   filter(age > 3, 
-         year > 1994)
-
+         year > 1987)
 ```
@@
 
+table(seal_phenology$age,seal_phenology$observed)
+
 ```

### 2023-12-04 paper/paperv3.qmd
paper v3 updates

@@
 #| label: age_breeding_mod
+#| #| fig-width: 9
 
@@
   geom_errorbar(aes(ymin = perc_breed - se, ymax = perc_breed + se), width = 0) +
-  scale_y_continuous(labels = scales::percent, limits = c(0,1)) + 
+  scale_y_continuous(labels = scales::percent, limits = c(.4, 1)) + 
   labs(x = "Maternal Age (Years)", y = "% Breeders (Mean ± SE)") + 
@@
 #| label: fig-breeding_bias
+#| fig-width: 9
 

### 2023-12-05 paper/paperv3.qmd
Start tackling the stats edits

@@
 summary(breeding_mod)
+#table s1, look at roxannes ecology letters paper supplemental 
+#Pull out the fixed effects table with 2 sfs, can paste in excel to adjust digits 
+#intercept, age (linear), age (quadratic) as row names 
+#in figure caption say "results from x analyis. The random effect variance of animal id is X (from results) and year was X. Apply to all the other stats. 
+#analyses fit to the age in which we have larger sample sizes fit also have a quadratic fit and don't change the story
 
@@
 
+pup_recruit_model <- glmer(puprecruited ~ poly(age, 2) + (1 | animalID) + (1 | year),
+                           data = filter(sealdat, year > 1984 & year < 2019), 
+                           family = "binomial")
+summary(pup_recruit_model)
+
 pup_surv_recruit <- sealdat %>% 
@@
   geom_pointrange(aes(ymin = low, ymax = high)) +
+  geom_ribbon(aes(x),
+              pup_recruit_model,
+              inherit.aes = FALSE,
+              fill = "#829BD1",
+              alpha = 0.25) + 
   scale_color_manual(values = c("#829BD1", "#E9B668")) + 
@@
 pup_surv_recruit
+
+#Why is recruitment higher than survival in some points? Add the quadratic model? Reverse order of legend in the figure
 ```
@@
 ```{r}
-surv_dat <- sealdat %>% filter(pupyearborn < 2016)
+surv_dat <- sealdat %>% filter(pupyearborn < 2016) %>% 
+  mutate(age10 = age/10)
+
 pup_survival_model <- glmer(pup_survived ~ poly(age, 2) + (1 | animalID) + (1 | year),
@@
 summary(pup_survival_model)
+#significant result, plot as quadratic on figure
+
+recruit_dat <- sealdat %>% mutate(age10 = age / 10)
 
 pup_recruit_model <- glmer(puprecruited ~ poly(age, 2) + (1 | animalID) + (1 | year),
-                           data = filter(sealdat, year > 1984 & year < 2019), 
-                           family = "binomial")
+                           data = filter(recruit_dat, year > 1984 & year < 2019), 
+                           family = "binomial", 
+                           control = glmerControl(optimizer = "bobyqa"))
 summary(pup_recruit_model)
+#divide age by 10 (ie rescale variables) - did not work
+#increase max iterations within glmer? 
+#change the optimizer? -> worked
+
+#survival but not recruitment success with age
+#figure with both models, with the recruitment one dashed or something
+
+#ranef variance goes in the table legend
 
@@
   scale_y_continuous(labels = scales::percent, limits = c(0, NA)) +
-  labs(x = "Maternal Age (Years)", y = "% Female Pups (Mean ± SE)") +
+  labs(x = "Maternal Age (Years)", y = "% Male Pups (Mean ± SE)") +
   theme_few() + 
@@
 pup_sex
+
+#take the p = 0.517 and divide by 2 because a priori. report in supplemental table that its reporting on 1 tailed (no absolute value of z, just z)
 ```
@@
 summary(pup_sex_mod)
+
+#quadratic
+
+pup_sex_quad <- glmer(as.factor(pupsex) ~ poly(age, 2) + (1 | animalID) + (1 | year),
+                            data = filter(sealdat, 
+                                   pupsex %in% c("M", "F"), 
+                                   age < 19),
+                            family = "binomial")
+summary(pup_sex_quad)
+
 ```
@@
 
+#Delete the non-breeders and add the dashed line as an average instead. 
+#put panels next to each other 
+#what do you do when breeding dates are in december? ifelse statement (if > 250), 200 - 365 (so use negative numbers if they are in december)
+#show xy plots of each phenological variable against age in the supplemental and summary stats for the model in the supplement too 
+#draw attention to the widths - 1) 2) 3) 
+#assess linearity from plots, but double check visually and see if quadratic might fit better (we werent sure if age patterns would be linear or not, so we tested quadratic analyses throughout)
+
 phen_plot1 <- ggplot(timing_summary, aes(x = age, y = doy, fill = event)) + 
@@
 fit_model <- function(dat) {
-  lmer(doy ~ age + observed + (1 | animalID) + (1 | year), data = dat)
+  lmer(doy ~ age + (1 | animalID) + (1 | year), data = dat)
 }
@@
   pivot_longer(BA:MD, names_to = "phenology", values_to = "doy") %>% 
-  select(phenology, observed, doy, age, animalID, year) %>% 
-  mutate(observed = factor(observed, levels = c("Non-breeder", "Breeder"))) %>% 
-  nest(model_data = c(doy, age, observed, animalID, year)) %>% 
+  filter(observed_int > 0) %>% 
+  select(phenology, doy, age, animalID, year) %>% 
+  nest(model_data = c(doy, age, animalID, year)) %>% 
   mutate(model = map(model_data, fit_model),
@@
          age_stderr = extract_stderr(model, "age"),
-         age_pval = extract_pval(model, "age"),
-         age_signif = age_pval * 8 < 0.05,
-         breeder_coef = extract_estimate(model, "observedBreeder"),
-         breeder_stderr = extract_stderr(model, "observedBreeder"),
-         breeder_pval = extract_pval(model, "observedBreeder"),
-         breeder_signif = breeder_pval * 8 < 0.05,)
+         age_pval = extract_pval(model, "age"))
+
+#  mutate(observed = factor(observed, levels = c("Non-breeder", "Breeder"))) %>% 
+         # age_signif = age_pval * 8 < 0.05,
+         # breeder_coef = extract_estimate(model, "observedBreeder"),
+         # breeder_stderr = extract_stderr(model, "observedBreeder"),
+         # breeder_pval = extract_pval(model, "observedBreeder"),
+         # breeder_signif = breeder_pval * 8 < 0.05,)
+
+#remove bonferroni correction
 

### 2023-12-05 scratch/phen_predict.R

@@ -16,7 +16,8 @@ expand_grid(
                       newdata = .,
                       re.form = NA)) %>% 
   mutate(short_trip = ma - bd, 
-         molt_dur = md - ma) %>% 
+         molt_dur = md - ma, 
+ #        long_trip = 365 - md ) %>% 
   filter(observed == "Breeder", 
          age %in% c(4,19)) %>% 
   view()

### 2023-12-05 paper/paperv3.qmd
model wont converge?

@@
 #| label: age_breeding_mod
-#| #| fig-width: 9
-
-breeding_mod_bad <- glmer(observed_int ~ poly(age, 2) + (1 | animalID) + (1 | year),
-                      data = filter(sealdat), 
-                      family = "binomial")
-
-count(sealdat, year) %>% 
-     left_join(as_tibble(ranef(breeding_mod_bad)$year, 
-                         rownames = "year") %>% 
-                   mutate(year = as.numeric(year)) %>% 
-                   rename(coef = `(Intercept)`),
-               by = "year") %>%
-     ggplot(aes(n, coef, color = year, label = year)) +
-     geom_point() + 
-  geom_label() + 
-     scale_color_gradient(low = "blue", high = "red") + 
-     theme_few()
+#| fig-width: 9
 
@@
                       family = "binomial")
+
 breed_pred <- ggpredict(breeding_mod, terms = c("age [all]"))
+
 summary(breeding_mod)
+
 #table s1, look at roxannes ecology letters paper supplemental 

### 2023-12-05 paper/paperv3.qmd
Figure 1 and 2 edits

@@
     warning: false
+    eval: false
 editor: visual
@@
 #| include: false  
+#| eval: true
 library(ggthemes)
@@
 #| label: age_breeding_mod
-#| fig-width: 9
+#| eval: true
 
@@
                       data = filter(sealdat), 
-                      family = "binomial")
+                      family = "binomial", 
+                      control = glmerControl(optimizer = "bobyqa"))
 
@@
 
-summary(breeding_mod)
+breed_mod_summary <- summary(breeding_mod)
+
+format_signif <- function(x, digits) {
+  formatC(signif(x, digits = digits), digits = digits, format = "fg", flag = "#")
+}
+format_pval <- function(p, digits) {
+  ifelse(p < 10^-digits,
+         sprintf("<%s", format(10^-digits, digits = 1, scientific = FALSE)),
+         sprintf(paste0("%0.", digits, "f"), p))
+}
+
+#p values are also rounded....
 
@@
 
+```
+
+```{r}
+#| tbl-cap: str_glue("The results from breeding probability analyis.
+The random effect variance of animal id is {VarCorr(breeding_mod, digits = 3)[1]} and year was VarCorr(breeding_mod, digits = 3)[2]".)
+#| eval: true
+
+breed_mod_summary$coefficients %>% 
+  as_tibble() %>% 
+  mutate(Parameter = c("Intercept", "Age (linear)", "Age (quadratic)"),
+         across(c(Estimate, `Std. Error`, `z value`), 
+                \(x) format_signif(x, digits = 2)),
+         `Pr(>|z|)` = format_pval(`Pr(>|z|)`, digits = 3)) %>%
+  relocate(Parameter) %>% 
+  knitr::kable()
 ```
@@
 
-pup_recruit_model <- glmer(puprecruited ~ poly(age, 2) + (1 | animalID) + (1 | year),
-                           data = filter(sealdat, year > 1984 & year < 2019), 
-                           family = "binomial")
-summary(pup_recruit_model)
+## NOTE: REMOVED RANDOM EFFECT OF INDIVIDUAL, OTHERWISE SINGULAR FIT (HYPOTHESIS: 
+## INSUFFICIENT SAMPLE SIZES PER INDIVIDUAL)
 
-pup_surv_recruit <- sealdat %>% 
-  mutate(pup_surv = ifelse(pupseeneveragain > 0, 1, 0)) %>% 
+surv_recruit_data <- sealdat %>% 
+  drop_na(pupID) %>% 
   filter(pupsex == "F",
          pupyearborn < 2016) %>% 
+  mutate(pup_survived_dbl = ifelse(pup_survived, 1, 0)) %>% 
   group_by(age) %>% 
-  summarize(n_surv = sum(!is.na(pup_surv)),
-            mean_surv = sum(pup_surv, na.rm = TRUE) / n_surv,  
-            se_surv = (mean_surv * (1 - mean_surv) / n_surv)^0.5,
-            n_recruit = sum(!is.na(puprecruited)), 
-            mean_recruit = sum(puprecruited, na.rm = TRUE) / n_recruit, 
-            se_recruit = (mean_recruit * (1 - mean_recruit) / n_recruit)^0.5)  %>%  
+  mutate(n_pups = n()) %>% 
+  ungroup() %>% 
+  filter(n_pups >= 5)
+pup_surv_model <- glmer(pup_survived_dbl ~ poly(age, 2) + (1 | year),
+                        data = surv_recruit_data, 
+                        family = "binomial")
+pup_surv_pred <- ggpredict(pup_surv_model, terms = c("age [all]")) %>% 
+  as_tibble() %>% 
+  transmute(age = x, 
+            metric = "surv", 
+            mean = predicted, 
+            mean_lwr = conf.low, 
+            mean_upr = conf.high)
+pup_recruit_model <- glmer(puprecruited ~ poly(age, 2) + (1 | year),
+                           data = surv_recruit_data, 
+                           family = "binomial")
+pup_recruit_pred <- ggpredict(pup_recruit_model, terms = c("age [all]")) %>% 
+  as_tibble() %>% 
+  transmute(age = x, 
+            metric = "recruit", 
+            mean = predicted, 
+            mean_lwr = conf.low, 
+            mean_upr = conf.high)
+surv_recruit_pred <- rbind(pup_surv_pred, pup_recruit_pred) %>% 
+  mutate(metric = factor(metric, 
+                         levels = c("surv", "recruit"), 
+                         labels = c("Survival", "Recruitment")))
+
+surv_recruit_summary <- surv_recruit_data %>% 
+  group_by(age) %>%
+  summarize(n_pups = n(),
+            n_surv = sum(pup_survived),
+            mean_surv = n_surv / n_pups,  
+            se_surv = (mean_surv * (1 - mean_surv) / n_pups)^0.5,
+            n_recruit = sum(puprecruited), 
+            mean_recruit = n_recruit / n_pups, 
+            se_recruit = (mean_recruit * (1 - mean_recruit) / n_pups)^0.5)  %>%  
   pivot_longer(cols = c(mean_surv, se_surv, mean_recruit, se_recruit), 
@@
          metric = factor(metric, 
-                         levels = c("recruit", "surv"), 
-                         labels = c("Recruitment", "Survival"))) %>% 
-  filter(n_surv > 5) %>% 
-  ggplot(aes(x = age, y = mean, color = metric)) + 
+                         levels = c("surv", "recruit"), 
+                         labels = c("Survival", "Recruitment")))
+
+surv_recruit_summary %>% 
+  ggplot(aes(x = age, y = mean, color = metric, fill = metric)) + 
+  geom_ribbon(aes(ymin = mean_lwr, ymax = mean_upr, alpha = metric),
+              surv_recruit_pred,
+              color = NA) +
   geom_pointrange(aes(ymin = low, ymax = high)) +
-  geom_ribbon(aes(x),
-              pup_recruit_model,
-              inherit.aes = FALSE,
-              fill = "#829BD1",
-              alpha = 0.25) + 
+  geom_line(aes(linetype = metric),
+            surv_recruit_pred) +
+  scale_y_continuous(labels = scales::percent) + 
   scale_color_manual(values = c("#829BD1", "#E9B668")) + 
-  scale_y_continuous(labels = scales::percent, limits = c(0, .5)) + 
+  scale_fill_manual(values = c("#829BD1", "#E9B668")) + 
+  scale_linetype_manual(values = c("solid", "dashed")) +
+  scale_alpha_manual(values = c(0.25, 0.1)) +
+  expand_limits(y = c(0, 0.5)) +
   labs(x = "Maternal Age (Years)", y = "% Pups (Mean ± SE)") +
@@
         legend.direction = "horizontal", 
-        axis.title = element_text(size = 13)
-)
-
-pup_surv_recruit
+        axis.title = element_text(size = 13))
 
@@
 ```
-
-Results of the Pup Survival model, with individual and year as random effects:
-
-```{r}
-#Panel B Survival Model
-surv_dat <- sealdat %>% filter(pupyearborn < 2016)
-pup_survival_model <- glmer(pup_survived ~ age + (1 | animalID) + (1 | year),
-                            data = surv_dat,
-                            family = "binomial")
-summary(pup_survival_model)
-
-```
-
-Results of the Pup Recruitment model, with individual and year as random effects:
-
-```{r}
-#Panel B Recruitment Model
-
-pup_recruit_model <- glmer(puprecruited ~ age + (1 | animalID) + (1 | year),
-                           data = filter(sealdat, year > 1984 & year < 2019), 
-                           family = "binomial")
-summary(pup_recruit_model)
-
-```
-
-Modeling the above metrics as quadratic, just in case:
-
-```{r}
-surv_dat <- sealdat %>% filter(pupyearborn < 2016) %>% 
-  mutate(age10 = age/10)
-
-pup_survival_model <- glmer(pup_survived ~ poly(age, 2) + (1 | animalID) + (1 | year),
-                            data = surv_dat,
-                            family = "binomial")
-summary(pup_survival_model)
-#significant result, plot as quadratic on figure
-
-recruit_dat <- sealdat %>% mutate(age10 = age / 10)
-
-pup_recruit_model <- glmer(puprecruited ~ poly(age, 2) + (1 | animalID) + (1 | year),
-                           data = filter(recruit_dat, year > 1984 & year < 2019), 
-                           family = "binomial", 
-                           control = glmerControl(optimizer = "bobyqa"))
-summary(pup_recruit_model)
-#divide age by 10 (ie rescale variables) - did not work
-#increase max iterations within glmer? 
-#change the optimizer? -> worked
-
-#survival but not recruitment success with age
-#figure with both models, with the recruitment one dashed or something
-
-#ranef variance goes in the table legend
-
-```
-
-Neither were significant as 2nd order polynomials.
 

### 2023-12-05 paper/paperv3.qmd
remove non-breeders from phenology plots

@@
          observed_int = if_else(observed == "Breeder", 1, 0), 
-         pup_survived = pupseeneveragain > 0) %>% 
+         pup_survived = pupseeneveragain > 0,) %>% 
   filter(age > 3, 
@@
 ```{r}
+#probably will need to take out animalID
 pup_sex_mod <- glmer(as.factor(pupsex) ~ age + (1 | animalID) + (1 | year),
@@
 seal_phenology <- sealdat %>% 
+  filter(observed == "Breeder") %>% 
   mutate(BA = firstobsbreeddoy, 
@@
 
+mean_phenology <- seal_phenology %>% 
+  summarize(across(BA:MD, mean), 
+            .groups = "drop") %>% 
+  pivot_longer(everything(),
+               names_to = "event", 
+               values_to = "doy")
+
 phenology_n <- unique(seal_phenology$animalID)
@@
 timing_summary <- seal_phenology %>% 
-  group_by(age, observed) %>% 
+  group_by(age) %>% 
   summarize(across(BA:MD, mean), 
             .groups = "drop") %>% 
-  pivot_longer(cols = -c(age, observed),  
+  pivot_longer(cols = -age,  
                names_to = "event", 
                values_to = "doy") %>% 
-  group_by(age, observed) %>% 
+  group_by(age) %>% 
   group_modify(\(data, keys) {
@@
                                           "longtrip2")))
-timing_summary_flipped <- timing_summary %>% 
-  mutate(observed = ifelse(observed == "Breeder",
-                           "Non-breeder",
-                           "Breeder")) %>% 
-  filter(event != "longtrip2")
 
@@
 #put panels next to each other 
-#what do you do when breeding dates are in december? ifelse statement (if > 250), 200 - 365 (so use negative numbers if they are in december)
+#what do you do when breeding dates are in december? ifelse statement (if > 250), 200 - 365 (so use negative numbers if they are in december) **does not happen! no worries!!**
 #show xy plots of each phenological variable against age in the supplemental and summary stats for the model in the supplement too 
@@
             color = "white") +
-  geom_line(data = timing_summary_flipped,
-            position = position_stack(reverse = TRUE),
+  geom_hline(aes(yintercept = doy),
+             mean_phenology,
              color = "white",
@@
   scale_y_continuous("Day of Year", expand = expansion(mult = 0)) +
-  facet_grid(cols = vars(observed)) +
   theme_few() + 
@@
   pivot_longer(BA:MD, names_to = "phenology", values_to = "doy") %>% 
-  filter(observed_int > 0) %>% 
   select(phenology, doy, age, animalID, year) %>% 
@@
          age_stderr = extract_stderr(model, "age"),
-         age_pval = extract_pval(model, "age"))
-
-#  mutate(observed = factor(observed, levels = c("Non-breeder", "Breeder"))) %>% 
-         # age_signif = age_pval * 8 < 0.05,
-         # breeder_coef = extract_estimate(model, "observedBreeder"),
-         # breeder_stderr = extract_stderr(model, "observedBreeder"),
-         # breeder_pval = extract_pval(model, "observedBreeder"),
-         # breeder_signif = breeder_pval * 8 < 0.05,)
-
-#remove bonferroni correction
+         age_pval = extract_pval(model, "age"),
+         age_signif = age_pval < 0.05)
 
@@
   mutate(signif = as.logical(signif), 
-         param = factor(param, 
-                        labels = c("'Delay with maternal age (days year'^-1 * ')'", 
-                                   "'Delay with breeding (days)'")), 
+         # param = factor(param, 
+         #                labels = "'Delay with maternal age (days year'^-1 * ')'"), 
          phenology = factor(phenology, 
@@
                       breaks = c("Long trip", "Breeding", "Short trip", "Molting")) + 
+  labs(x = expression('Delay with maternal age (days year'^-1 * ')')) +
   expand_limits(x = 0) +
-  facet_grid(cols = vars(param), 
-             scales = "free_x", 
-             switch = "x",
-             labeller = label_parsed) +
   theme_few() + 
-  theme(axis.title = element_blank(),
+  theme(axis.title.y = element_blank(),
         legend.direction = "horizontal",
-        legend.title = element_blank(),
-        strip.placement = "outside",
-        strip.text = element_text(size = 12, color = "black"),
-        panel.spacing = unit(1, "lines"))
+        legend.title = element_blank())
 
@@
   get_legend(phen_plot2),
+  plot_grid(
     phen_plot1 + theme(legend.position = "none"),
     phen_plot2 + theme(legend.position = "none"),
-  align = "v",
-  axis = "lr",
-  labels = c("", "A", "B"),
+    align = "h",
+    axis = "tb",
+    labels = c("A", "B"),
+    nrow = 1
+  ),
   ncol = 1,
-  rel_heights = c(0.1, 1, 1)
+  rel_heights = c(0.1, 1)
 )

### 2023-12-06 paper/paperv3.qmd
fixing pup sex figure

@@
     warning: false
-    eval: false
 editor: visual
@@
 #| include: false  
-#| eval: true
 library(ggthemes)
@@
 #| label: age_breeding_mod
-#| eval: true
 
@@
 
+#change rightmost value to say p value 
+#add an extra sig fig to all columns
 #p values are also rounded....
@@
          across(c(Estimate, `Std. Error`, `z value`), 
-                \(x) format_signif(x, digits = 2)),
-         `Pr(>|z|)` = format_pval(`Pr(>|z|)`, digits = 3)) %>%
+                \(x) format_signif(x, digits = 3)),
+         P_value = format_pval(`Pr(>|z|)`, digits = 2)) %>%
+  select(-`Pr(>|z|)`) %>% 
   relocate(Parameter) %>% 
@@
   expand_limits(y = c(0, 0.5)) +
-  labs(x = "Maternal Age (Years)", y = "% Pups (Mean ± SE)") +
+  labs(x = "Maternal Age (Years)", y = "Pup Survival or Recruitment/n") +
   theme_few() + 
@@
 #Why is recruitment higher than survival in some points? Add the quadratic model? Reverse order of legend in the figure
+#caption - error bars show SE. Points are mean values for maternal age. Make sure that language in the caption matches both the model and the data 
 ```
@@
 
-pup_sex <- sealdat %>% 
+pup_sex_data <- sealdat %>% 
   filter(pupsex %in% c("M", "F"), 
          age < 19) %>% 
+  mutate(is_male = ifelse(pupsex == "M", 1, 0))
+
+#linear
+pup_sex_mod <- glmer(is_male ~ age + (1 | animalID) + (1 | year),
+                     data = pup_sex_data, 
+                     family = "binomial")
+summary(pup_sex_mod)
+
+# We fit a quadratic equation, but only the linear term was significant, so the linear model is the one that is plotted.
+
+pup_sex_pred <- ggpredict(pup_sex_mod, terms = c("age [all]")) %>% 
+  as_tibble() %>% 
+  transmute(age = x, 
+            percm = predicted, 
+            percm_lwr = conf.low, 
+            percm_upr = conf.high)
+
+pup_sex_data %>% 
   group_by(age, pupsex) %>%
-  count(pupsex) %>% 
+  summarize(n = n(), .groups = "drop") %>% 
   pivot_wider(values_from = n, names_from = pupsex) %>% 
   mutate(num = M + F,
-         percf = F / num,
-         se = (percf * (1 - percf) / num)^0.5) %>% 
-  ggplot(aes(x = age, y = percf, weight = num)) +
-  geom_hline(yintercept = 0.5, lty = 2, color = "#829BD1") +
-  geom_point() +
-  geom_errorbar(aes(ymin = percf - se, ymax = percf + se), width = 0) +
-  scale_y_continuous(labels = scales::percent, limits = c(0, NA)) +
+         percm = M / num,
+         se = (percm * (1 - percm) / num)^0.5) %>% 
+  ggplot(aes(x = age, y = percm)) +
+  geom_hline(yintercept = 0.5, color = "gray75", linetype = "dashed") +
+  geom_ribbon(aes(ymin = percm_lwr, ymax = percm_upr), 
+              pup_sex_pred,
+              fill = "#829BD1",
+              alpha = 0.25) + 
+  geom_line(data = pup_sex_pred, 
+            color = "#829BD1", 
+            linewidth = 1) + 
+  geom_pointrange(aes(ymin = percm - se, ymax = percm + se), 
+                  size = 0.25) +
+  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
   labs(x = "Maternal Age (Years)", y = "% Male Pups (Mean ± SE)") +
@@
 
-pup_sex
-
 #take the p = 0.517 and divide by 2 because a priori. report in supplemental table that its reporting on 1 tailed (no absolute value of z, just z)
-```
-
-Results of the model for pup sex ratio:
-
-```{r}
-#probably will need to take out animalID
-pup_sex_mod <- glmer(as.factor(pupsex) ~ age + (1 | animalID) + (1 | year),
-                     data = filter(sealdat, 
-                                   pupsex %in% c("M", "F"), 
-                                   age < 19), 
-                     family = "binomial")
-summary(pup_sex_mod)
-
-#quadratic
-
-pup_sex_quad <- glmer(as.factor(pupsex) ~ poly(age, 2) + (1 | animalID) + (1 | year),
-                            data = filter(sealdat, 
-                                   pupsex %in% c("M", "F"), 
-                                   age < 19),
-                            family = "binomial")
-summary(pup_sex_quad)
-
 ```
@@
 
+#average the averages - global average of age-specific averages 
 ```

### 2023-12-06 paper/paperv3.qmd
fixing the average line on phenology plot

@@
 #| fig-cap: "Pup survival decreases slightly, but survival and recruitment are not significantly affected by maternal age."
-#| fig-width: 9
+#| fig-width: 4.5
 
@@
   expand_limits(y = c(0, 0.5)) +
-  labs(x = "Maternal Age (Years)", y = "Pup Survival or Recruitment/n") +
+  labs(x = "Maternal Age (Years)", y = "Pup Survival or Recruitment") +
   theme_few() + 
@@
 
-year_ranef <- ranef(pup_survival_model)$year %>% 
+year_ranef <- ranef(pup_surv_model)$year %>% 
   as_tibble(rownames = "year") %>% 
@@
 #| fig-cap: "Pup sex skews slightly towards males with increasing maternal age, but it's not significant."
-#| fig-width: 9
+#| fig-width: 4.5
 #A: Pup sex shifts towards males with increasing maternal age, but it's not significant.
@@
 mean_phenology <- seal_phenology %>% 
-  summarize(across(BA:MD, mean), 
-            .groups = "drop") %>% 
+  group_by(age) %>% 
+  summarize(across(BA:MD, mean)) %>% 
+  summarize(across(BA:MD, mean)) %>% 
   pivot_longer(everything(),
@@
 
-The table is supplemental.
-
 ```{r}
 #| label: fig-phenology
-#| fig-height: 7
+#| fig-height: 4.5
 #| fig-width: 7
@@
 
-ggsave("outputs/agefig4.png", plot = , height = 9, width = 9)
+# table(seal_phenology$age,seal_phenology$observed)
 
-table(seal_phenology$age,seal_phenology$observed)
-
-#average the averages - global average of age-specific averages 
+#white dashed lines average the averages - global average of age-specific averages 
 ```
-
-Supplemental table
-
-```{r}
-
-phenology_models %>% 
-  select(-c(model_data, model)) %>% 
-  pivot_longer(age_coef:breeder_signif, 
-               names_to = c("param", "metric"),
-               names_sep = "_",
-               values_to = "value") %>% 
-  pivot_wider(names_from = metric, values_from = value) %>% 
-  mutate(coef = str_glue("{round(coef, 2)} +- {round(stderr, 2)}"),
-         signif = str_glue("p = {round(pval, 4)} ({signif == 1})")) %>% 
-  select(-c(stderr, pval)) %>% 
-  pivot_wider(names_from = param, values_from = c(coef, signif)) %>% 
-  select(phenology, ends_with("age"), ends_with("breeder")) %>% 
-  knitr::kable(col.names = c("Phenology", 
-                             "Age coef. (days/year)", 
-                             "Significance", 
-                             "Breeding status coef. (days)", 
-                             "Significance")) %>% 
-  kableExtra::kable_styling()
-
-```

### 2023-12-06 paper/paperv3.qmd
update with new feedback from Roxanne

@@
 
-sealdat <- read_csv(here::here("data/raw/128L pull 2023_11_30.csv"), 
+sealdat <- read_csv(here::here("data/raw/128L pull 2023_12_05.csv"), 
                     show_col_types = FALSE) %>% 
@@
          observed_int = if_else(observed == "Breeder", 1, 0), 
-         pup_survived = pupseeneveragain > 0,) %>% 
+         pup_survived = pupseeneveragain > 0) %>% 
   filter(age > 3, 
@@
 #| tbl-cap: str_glue("The results from breeding probability analyis.
-The random effect variance of animal id is {VarCorr(breeding_mod, digits = 3)[1]} and year was VarCorr(breeding_mod, digits = 3)[2]".)
+The random effect variance of animal id is {VarCorr(breeding_mod, digits = 3)[1]} and year was VarCorr(breeding_mod, digits = 3)[2].")
 #| eval: true
@@
   filter(n_pups >= 5)
-pup_surv_model <- glmer(pup_survived_dbl ~ poly(age, 2) + (1 | year),
+pup_surv_model <- glmer(pup_survived_dbl ~ poly(age, 2) + (1 | animalID) + (1 | year),
                         data = surv_recruit_data, 
@@
             mean_upr = conf.high)
-pup_recruit_model <- glmer(puprecruited ~ poly(age, 2) + (1 | year),
+pup_recruit_model <- glmer(puprecruited ~ poly(age, 2) + (1 | animalID) + (1 | year),
                            data = surv_recruit_data, 

### 2023-12-06 00_clean_data.R

@@ -93,7 +93,6 @@ dateobsmolt=outall %>%
 dateobs=left_join(dateobsmolt,dateobsbreed,
                   by=c("season","animalID")) %>%
   #only last observed molt less than july 1 ish and only first breeding < march 16 ish
-  mutate(firstobsbreed=if_else(yday(firstobsbreed)>70,NA,firstobsbreed)) %>%
   mutate(firstobsbreeddoy=yday(as.Date(firstobsbreed))) %>%
   mutate(firstobsbreeddoy=if_else(firstobsbreeddoy>300,firstobsbreeddoy-365,firstobsbreeddoy)) %>%
   mutate(lastobsmolt=if_else(yday(lastobsmolt)>185,NA,lastobsmolt)) %>%
@@ -175,5 +174,5 @@ for(i in rows){ #for each row of those rows
 #subset to females only here (moms) for the main dataset
 final = final %>% filter(tagsex=="F")
 
-write.csv(final,here::here("data/raw/128L pull 2023_11_30.csv"),
+write.csv(final,here::here("data/raw/128L pull 2023_12_05.csv"),
           row.names=FALSE)

### 2023-12-07 paper/paperv3.qmd
Moved supplemental table to end

@@
 ```{r}
-#| tbl-cap: str_glue("The results from breeding probability analyis.
-The random effect variance of animal id is {VarCorr(breeding_mod, digits = 3)[1]} and year was VarCorr(breeding_mod, digits = 3)[2].")
-#| eval: true
-
-breed_mod_summary$coefficients %>% 
-  as_tibble() %>% 
-  mutate(Parameter = c("Intercept", "Age (linear)", "Age (quadratic)"),
-         across(c(Estimate, `Std. Error`, `z value`), 
-                \(x) format_signif(x, digits = 3)),
-         P_value = format_pval(`Pr(>|z|)`, digits = 2)) %>%
-  select(-`Pr(>|z|)`) %>% 
-  relocate(Parameter) %>% 
-  knitr::kable()
-```
-
-```{r}
 #| label: fig-age_dist
@@
   geom_errorbar(aes(ymin = perc_breed - se, ymax = perc_breed + se), width = 0) +
-  scale_y_continuous(labels = scales::percent, limits = c(.4, 1)) + 
+  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) + 
   labs(x = "Maternal Age (Years)", y = "% Breeders (Mean ± SE)") + 
@@
 
-```{r}
-#| label: fig-breeding_bias
-#| fig-width: 9
-
-year_coef <- as_tibble(ranef(breeding_mod)$year, 
-                       rownames = "year") %>% 
-    mutate(year = as.numeric(year)) %>% 
-    rename(coef = `(Intercept)`)
-
-animal_coef <- as_tibble(ranef(breeding_mod)$animalID, 
-                         rownames = "animalID") %>% 
-    mutate(animalID = as.numeric(animalID)) %>% 
-    rename(coef = `(Intercept)`)
-
-year_bias <- sealdat %>% 
-  left_join(year_coef, by = "year") %>% 
-  mutate(age = factor(age)) %>% 
-  ggplot(aes(age, coef)) +
-  geom_boxplot() +
-  geom_hline(yintercept = 0, color = "#829BD1", linetype = "dashed") +
-  scale_x_discrete("Maternal Age (Years)", breaks = seq(5, 20, by = 5)) +
-  ylab("Random effect of year") +
-  theme_few()
-
-animalID_bias <- sealdat %>% 
-  left_join(animal_coef, by = "animalID") %>% 
-  mutate(age = factor(age)) %>% 
-  ggplot(aes(age, coef)) +
-  geom_boxplot() +
-  geom_hline(yintercept = 0, color = "#829BD1", linetype = "dashed") +
-  scale_x_discrete("Maternal Age (Years)", breaks = seq(5, 20, by = 5)) +  
-  ylab("Random effect of individual") +
-  theme_few()
-
-plot_grid(breed_nonbreed, year_bias, animalID_bias, nrow = 1)
-
-```
-
 ### H2: Maternal effect senescence
@@
 #| label: fig-repro_age
-#| fig-cap: "Pup survival decreases slightly, but survival and recruitment are not significantly affected by maternal age."
+#| fig-cap: "Pup survival and recruitment are not significantly affected by maternal age. Points are mean values for maternal age. Error bars show SE."
 #| fig-width: 4.5
@@
 #B Pup survival and recruitment doesn't change with increased maternal age.
-
-## NOTE: REMOVED RANDOM EFFECT OF INDIVIDUAL, OTHERWISE SINGULAR FIT (HYPOTHESIS: 
-## INSUFFICIENT SAMPLE SIZES PER INDIVIDUAL)
 
@@
   scale_fill_manual(values = c("#829BD1", "#E9B668")) + 
-  scale_linetype_manual(values = c("solid", "dashed")) +
+  scale_linetype_manual(values = c("dashed", "dashed")) +
   scale_alpha_manual(values = c(0.25, 0.1)) +
@@
 ```
+
+### Supplemental Material
+
+Supplemental Table 1: Statistics for breeding success model.
+
+```{r}
+#| label: tbl-breed-cap
+
+tbl_breed_cap <- str_glue("The results from breeding probability analyis.
+The random effect variance of animal id is {format_signif(VarCorr(breeding_mod)$animalID, 3)} and of year was {format_signif(VarCorr(breeding_mod)$year, 3)}.")
+
+```
+
+```{r}
+#| label: tbl-breed-success
+#| tbl-cap: !expr "tbl_breed_cap"
+
+breed_mod_summary$coefficients %>% 
+  as_tibble() %>% 
+  mutate(Parameter = c("Intercept", "Age (linear)", "Age (quadratic)"),
+         across(c(Estimate, `Std. Error`, `z value`), 
+                \(x) format_signif(x, digits = 3)),
+         `p value` = format_pval(`Pr(>|z|)`, digits = 2)) %>%
+  select(-`Pr(>|z|)`) %>% 
+  relocate(Parameter) %>% 
+  knitr::kable()
+```

### 2023-12-07 paper/paperv3.qmd
Add details for pup sex

@@
 #| label: fig-sex_ratio
-#| fig-cap: "Pup sex skews slightly towards males with increasing maternal age, but it's not significant."
+#| fig-cap: !expr "str_glue('Offspring sex ratio skews towards males with increasing maternal age (p = {format_pval(summary(pup_sex_mod)$coefficients[\"age\", \"Pr(>|z|)\"] / 2, 3)}).
+Mothers aged 12, 13, and 17 gave birth to significantly more male than female offspring (red stars).')"
 #| fig-width: 4.5
-#A: Pup sex shifts towards males with increasing maternal age, but it's not significant.
 
@@
   mutate(is_male = ifelse(pupsex == "M", 1, 0))
+
+pup_sex_summary <- pup_sex_data %>% 
+  group_by(age, pupsex) %>%
+  summarize(n = n(), .groups = "drop") %>% 
+  pivot_wider(values_from = n, names_from = pupsex) %>% 
+  mutate(num = M + F,
+         percm = M / num,
+         se = (percm * (1 - percm) / num)^0.5,
+         pval = map2_dbl(M, num, \(x, n) binom.test(x, n, alternative = "greater")$p.value))
+
+signif_male <- filter(pup_sex_summary, pval < 0.05)
 
@@
 
-pup_sex_data %>% 
-  group_by(age, pupsex) %>%
-  summarize(n = n(), .groups = "drop") %>% 
-  pivot_wider(values_from = n, names_from = pupsex) %>% 
-  mutate(num = M + F,
-         percm = M / num,
-         se = (percm * (1 - percm) / num)^0.5) %>% 
-  ggplot(aes(x = age, y = percm)) +
+ggplot(pup_sex_summary, aes(x = age, y = percm)) +
   geom_hline(yintercept = 0.5, color = "gray75", linetype = "dashed") +
@@
                   size = 0.25) +
+  geom_point(aes(y = percm + se + 0.05),
+             signif_male,
+             shape = 8,
+             color = "firebrick") +
   scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
@@
 ```
+
+Supplemental table for sex ratios
+
+```{r}
+#| label: tbl-pup-sex
+#| tbl-cap: "Offspring sex ratio skewed towards males with increasing maternal age. Mothers age 12, 13, and 17 gave birth to significantly more male than female pups (bold)."
+
+pup_sex_tbl <- pup_sex_summary %>% 
+  transmute(
+    Age = age,
+    `Male pups` = M,
+    `Total pups` = num,
+    `% male` = scales::label_percent(accuracy = 0.1)(percm),
+    `p value` = format_pval(pval, 3)
+  )
+
+is_signif <- which(pup_sex_tbl$`p value` < "0.05")
+
+knitr::kable(pup_sex_tbl) %>% 
+  kableExtra::kable_styling() %>% 
+  kableExtra::row_spec(is_signif, bold = TRUE)
+```

### 2023-12-07 paper/paperv3.qmd
Move functions to R/. Constrain fig 1 y axis

@@
 library(patchwork)
+source(here::here("R/format.R"))
 
@@
 breed_mod_summary <- summary(breeding_mod)
-
-format_signif <- function(x, digits) {
-  formatC(signif(x, digits = digits), digits = digits, format = "fg", flag = "#")
-}
-format_pval <- function(p, digits) {
-  ifelse(p < 10^-digits,
-         sprintf("<%s", format(10^-digits, digits = 1, scientific = FALSE)),
-         sprintf(paste0("%0.", digits, "f"), p))
-}
 
@@
   geom_errorbar(aes(ymin = perc_breed - se, ymax = perc_breed + se), width = 0) +
-  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) + 
+  scale_y_continuous(labels = scales::percent, limits = c(0.5, 1)) + 
   labs(x = "Maternal Age (Years)", y = "% Breeders (Mean ± SE)") + 

### 2023-12-07 R/format.R

@@ -1,0 +1,9 @@
+format_signif <- function(x, digits) {
+  formatC(signif(x, digits = digits), digits = digits, format = "fg", flag = "#")
+}
+
+format_pval <- function(p, digits) {
+  ifelse(p < 10^-digits,
+         sprintf("<%s", format(10^-digits, digits = 1, scientific = FALSE)),
+         sprintf(paste0("%0.", digits, "f"), p))
+}

### 2023-12-07 paper/paperv3.qmd
Adding supplemental figures and tables

@@
 
-We found that as maternal age increases, pup sex skews slightly towards males - but it's not significant.
+We found that as maternal age increases, pup sex skews slightly towards males.
 
@@
 
-tbl_breed_cap <- str_glue("The results from breeding probability analyis.
+tbl_breed_cap <- str_glue("The results from breeding probability analysis.
 The random effect variance of animal id is {format_signif(VarCorr(breeding_mod)$animalID, 3)} and of year was {format_signif(VarCorr(breeding_mod)$year, 3)}.")
@@
 
-Supplemental table for sex ratios
+Supplemental table for offspring survival
+
+```{r}
+surv_mod_summary <- summary(pup_surv_model)
+surv_mod_kable <- surv_mod_summary$coefficients %>% 
+  as_tibble() %>% 
+  mutate(Parameter = c("Intercept", "Age (linear)", "Age (quadratic)"),
+         across(c(Estimate, `Std. Error`, `z value`), 
+                \(x) format_signif(x, digits = 3)),
+         `p value` = format_pval(`Pr(>|z|)`, digits = 2),
+         Response = "Survival") %>%
+  select(-`Pr(>|z|)`) %>% 
+  relocate(Response, Parameter)
+
+recruit_mod_summary <- summary(pup_recruit_model)
+recruit_mod_kable <- recruit_mod_summary$coefficients %>% 
+  as_tibble() %>% 
+  mutate(Parameter = c("Intercept", "Age (linear)", "Age (quadratic)"),
+         across(c(Estimate, `Std. Error`, `z value`), 
+                \(x) format_signif(x, digits = 3)),
+         `p value` = format_pval(`Pr(>|z|)`, digits = 2),
+         Response = "Recruitment") %>%
+  select(-`Pr(>|z|)`) %>% 
+  relocate(Response, Parameter)
+
+rbind(surv_mod_kable, recruit_mod_kable) %>% 
+  knitr::kable() %>% 
+  kableExtra::kable_styling() %>% 
+  kableExtra::collapse_rows(columns = 1)
+```
+
+Supplemental table for pup sex ratio model.
+
+```{r}
+pupsex_mod_summary <- summary(pup_sex_mod)
+pupsex_mod_summary$coefficients %>% 
+  as_tibble() %>% 
+  mutate(Parameter = c("Intercept", "Age (linear)"),
+         across(c(Estimate, `Std. Error`, `z value`), 
+                \(x) format_signif(x, digits = 3)),
+         `p value` = format_pval(`Pr(>|z|)`/2, digits = 3)) %>%
+  select(-`Pr(>|z|)`) %>% 
+  relocate(Parameter) %>% 
+  knitr::kable()
+
+# We fit a quadratic equation, but only the linear term was significant, so the linear model is the one that is plotted.
+take the p = 0.517 and divide by 2 because a priori.
+report in supplemental table that its reporting on 1 tailed (no absolute value of z, just z)
+```
+
+Supplemental table for male biased sex ratios at older ages.
 
@@
 ```
+
+Supplemental figure for phenology models
+
+```{r}
+
+#NOTE: Both BA and BD have some points that are really late (above 300 doy) that need fixing
+BA_age <- ggplot(seal_phenology, aes(age, BA)) + 
+  geom_point() + 
+  scale_y_continuous(limits = c(0, 150)) + 
+  theme_classic() + 
+  labs(y = "Breeding arrival (DOY)", x = "Maternal age (Years)")
+
+BD_age <- ggplot(seal_phenology, aes(age, BD)) + 
+  geom_point() + 
+  scale_y_continuous(limits = c(0, 150)) + 
+  theme_classic() + 
+  labs(y = "Breeding departure (DOY)", x = "Maternal age (Years)")
+
+MA_age <- ggplot(seal_phenology, aes(age, MA)) + 
+  geom_point() + 
+  geom_smooth(method = "lm", color = "#829BD1") + 
+  scale_y_continuous(limits = c(50, 200)) + 
+  theme_classic() + 
+  labs(y = "Molt arrival (DOY)", x = "Maternal age (Years)")
+
+MD_age <- ggplot(seal_phenology, aes(age, MD)) + 
+  geom_point() + 
+  geom_smooth(method = "lm", color = "#829BD1") + 
+  scale_y_continuous(limits = c(50, 200)) + 
+  theme_classic() + 
+  labs(y = "Molt departure (DOY)", x = "Maternal age (Years)")
+
+plot_grid(BA_age, BD_age, MA_age, MD_age, nrow = 2, ncol = 2)
+```
+
+Supplemental table for the phenology models.
+
+```{r}
+BA_summary <- phenology_models$model[1]
+ <- BA_summary$coefficients %>% 
+  as_tibble() %>% 
+  mutate(Parameter = c("Intercept", "Age (linear)", "Age (quadratic)"),
+         across(c(Estimate, `Std. Error`, `z value`), 
+                \(x) format_signif(x, digits = 3)),
+         `p value` = format_pval(`Pr(>|z|)`, digits = 2),
+         Response = "Survival") %>%
+  select(-`Pr(>|z|)`) %>% 
+  relocate(Response, Parameter)
+
+recruit_mod_summary <- summary(pup_recruit_model)
+recruit_mod_kable <- recruit_mod_summary$coefficients %>% 
+  as_tibble() %>% 
+  mutate(Parameter = c("Intercept", "Age (linear)", "Age (quadratic)"),
+         across(c(Estimate, `Std. Error`, `z value`), 
+                \(x) format_signif(x, digits = 3)),
+         `p value` = format_pval(`Pr(>|z|)`, digits = 2),
+         Response = "Recruitment") %>%
+  select(-`Pr(>|z|)`) %>% 
+  relocate(Response, Parameter)
+
+rbind(surv_mod_kable, recruit_mod_kable) %>% 
+  knitr::kable() %>% 
+  kableExtra::kable_styling() %>% 
+  kableExtra::collapse_rows(columns = 1)
+```

### 2023-12-07 paper/paperv3.qmd
Finishing supplemental tables

@@
 ```{r}
-BA_summary <- phenology_models$model[1]
- <- BA_summary$coefficients %>% 
+ba_data <- phenology_models$model_data[[1]]
+ba_mod <- lmer(doy ~ age + (1 | animalID) + (1 | year), data = ba_data)
+ba_summary <- summary(ba_mod)
+ba_ktable <- ba_summary$coefficients %>% 
   as_tibble() %>% 
-  mutate(Parameter = c("Intercept", "Age (linear)", "Age (quadratic)"),
-         across(c(Estimate, `Std. Error`, `z value`), 
+  mutate(Parameter = c("Intercept", "Age (linear)"),
+         across(c(Estimate, `Std. Error`, `df`, `t value`), 
                 \(x) format_signif(x, digits = 3)),
-         `p value` = format_pval(`Pr(>|z|)`, digits = 2),
-         Response = "Survival") %>%
-  select(-`Pr(>|z|)`) %>% 
+         `p value` = format_pval(`Pr(>|t|)`, digits = 2),
+         Response = "Breeding arrival") %>%
+  select(-`Pr(>|t|)`) %>% 
   relocate(Response, Parameter)
 
-recruit_mod_summary <- summary(pup_recruit_model)
-recruit_mod_kable <- recruit_mod_summary$coefficients %>% 
+bd_data <- phenology_models$model_data[[2]]
+bd_mod <- lmer(doy ~ age + (1 | animalID) + (1 | year), data = ba_data)
+bd_summary <- summary(ba_mod)
+bd_ktable <- bd_summary$coefficients %>% 
   as_tibble() %>% 
-  mutate(Parameter = c("Intercept", "Age (linear)", "Age (quadratic)"),
-         across(c(Estimate, `Std. Error`, `z value`), 
+  mutate(Parameter = c("Intercept", "Age (linear)"),
+         across(c(Estimate, `Std. Error`, `df`, `t value`), 
                 \(x) format_signif(x, digits = 3)),
-         `p value` = format_pval(`Pr(>|z|)`, digits = 2),
-         Response = "Recruitment") %>%
-  select(-`Pr(>|z|)`) %>% 
+         `p value` = format_pval(`Pr(>|t|)`, digits = 2),
+         Response = "Breeding departure") %>%
+  select(-`Pr(>|t|)`) %>% 
   relocate(Response, Parameter)
 
-rbind(surv_mod_kable, recruit_mod_kable) %>% 
+ma_data <- phenology_models$model_data[[3]]
+ma_mod <- lmer(doy ~ age + (1 | animalID) + (1 | year), data = ma_data)
+ma_summary <- summary(ma_mod)
+ma_ktable <- ma_summary$coefficients %>% 
+  as_tibble() %>% 
+  mutate(Parameter = c("Intercept", "Age (linear)"),
+         across(c(Estimate, `Std. Error`, `df`, `t value`), 
+                \(x) format_signif(x, digits = 3)),
+         `p value` = format_pval(`Pr(>|t|)`, digits = 2),
+         Response = "Molt arrival") %>%
+  select(-`Pr(>|t|)`) %>% 
+  relocate(Response, Parameter)
+
+md_data <- phenology_models$model_data[[4]]
+md_mod <- lmer(doy ~ age + (1 | animalID) + (1 | year), data = md_data)
+md_summary <- summary(md_mod)
+md_ktable <- md_summary$coefficients %>% 
+  as_tibble() %>% 
+  mutate(Parameter = c("Intercept", "Age (linear)"),
+         across(c(Estimate, `Std. Error`, `df`, `t value`), 
+                \(x) format_signif(x, digits = 3)),
+         `p value` = format_pval(`Pr(>|t|)`, digits = 2),
+         Response = "Molt departure") %>%
+  select(-`Pr(>|t|)`) %>% 
+  relocate(Response, Parameter)
+
+rbind(ba_ktable, bd_ktable, ma_ktable, md_ktable) %>% 
   knitr::kable() %>% 

### 2023-12-07 paper/paperv3.qmd
Little tweaks to the figure names and small edits to the text. Also updating the index to go to v3 of the paper

@@
 
-2.  Reproductive declines persist to the next generation (maternal effect senescence).
+2.  Reproductive declines persist to the next generation (maternal effect senescence). That is, offspring survival and recruitment are decreased.
 
-3.  Seal phenology is shifted later for older animals and breeders.
+3.  Offspring sex ratios shift towards males with old age.
 
-**Note**: *This paper is similar to another paper about polar bears (*Ursus maritimus*)* [@naciri2022].
+4.  Phenology is shifted later for older animals and breeders.
 
@@
 
-The model of Survival \~ Maternal age was significant before we treated Year as a random effect.
-Was that because pups born earlier in the study had more time to be seen again (i.e., a sampling artifact?) If so, the plot of the random effect of year should have a negative slope.
-
-We also plotted survival bias to show that 95% of pups were resighted within 7 years. Both plots are supplemental.
-
-```{r}
-#| label: fig-surv_bias
-#| fig-cap: "A: Plotting the intercept of the random effect of year shows that variation in ocean conditions is likely driving patterns in survival, rather than age of mother.
-B: 95% of seals that survived were resighted within ~seven years of their birth."
-#| fig-width: 9
-
-year_ranef <- ranef(pup_surv_model)$year %>% 
-  as_tibble(rownames = "year") %>% 
-  mutate(year = as.numeric(year)) %>% 
-  rename(intercept = `(Intercept)`)
-
-year_ranef_plot <- ggplot(year_ranef, aes(year, intercept)) + 
-  geom_point() + 
-  geom_smooth(method = "loess", 
-              formula = y ~ x,
-              color = "red") + 
-  theme_classic()
-
-survival_bias <- sealdat %>% 
-  group_by(animalID) %>% 
-  summarize(first_resight_yrs = min(age))
-
-survival95 <- quantile(survival_bias$first_resight_yrs, 0.95)
-#n for 95% quantile
-sum(survival_bias$first_resight_yrs > 7)
-
-survival_bias_plot <- ggplot(survival_bias, aes(x = first_resight_yrs)) +
-  stat_ecdf() +
-  geom_vline(xintercept = survival95, linetype = "dashed", color = "red") +
-  scale_x_continuous(limits = c(0, 10)) + 
-  theme_classic()
-
-plot_grid(year_ranef_plot, survival_bias_plot, nrow = 1, labels = "AUTO")
-
-```
-
-There is a drop off in recent years that seems to support the idea that recent years have a sampling bias, but the earlier portion suggests that overall the variation is caused by ocean conditions and not sampling.
-See the Cohort paper!
-
 ### H3a: Changes in pup sex ratio with increased maternal age
 
-We found that as maternal age increases, pup sex skews slightly towards males.
+We found that as maternal age increases, pup sex skews slightly towards males (@fig-sex_ratio).
 
@@
                      family = "binomial")
-summary(pup_sex_mod)
 
@@
 
+Supplemental figure for survival bias:
+
+```{r}
+#| label: fig-surv_bias
+#| fig-cap: "A: Plotting the intercept of the random effect of year shows that variation in ocean conditions is likely driving patterns in survival, rather than age of mother.
+B: 95% of seals that survived were resighted within ~seven years of their birth."
+#| fig-width: 9
+
+year_ranef <- ranef(pup_surv_model)$year %>% 
+  as_tibble(rownames = "year") %>% 
+  mutate(year = as.numeric(year)) %>% 
+  rename(intercept = `(Intercept)`)
+
+year_ranef_plot <- ggplot(year_ranef, aes(year, intercept)) + 
+  geom_point() + 
+  geom_smooth(method = "loess", 
+              formula = y ~ x,
+              color = "red") + 
+  theme_classic()
+
+survival_bias <- sealdat %>% 
+  group_by(animalID) %>% 
+  summarize(first_resight_yrs = min(age))
+
+survival95 <- quantile(survival_bias$first_resight_yrs, 0.95)
+#n for 95% quantile
+sum(survival_bias$first_resight_yrs > 7)
+
+survival_bias_plot <- ggplot(survival_bias, aes(x = first_resight_yrs)) +
+  stat_ecdf() +
+  geom_vline(xintercept = survival95, linetype = "dashed", color = "red") +
+  scale_x_continuous(limits = c(0, 10)) + 
+  theme_classic()
+
+plot_grid(year_ranef_plot, survival_bias_plot, nrow = 1, labels = "AUTO")
+```
+
 Supplemental table for pup sex ratio model.

### 2023-12-07 paper/paperv3.qmd
Adding sample sizes and fixing the initial data filter

@@
          year > 1987)
+
+# 17 observations where pup recruited but did not survive
+problem_children <- filter(sealdat, puprecruited == 1 & !pup_survived)
+
+sealdat <- anti_join(sealdat, problem_children, by = c("animalID", "pupID"))
 ```
@@
 
+n = `r n_distinct(sealdat$animalID)` distinct animals
+
 ### H1: Age distribution and breeding status
@@
 breeding_mod <- glmer(observed_int ~ poly(age, 2) + (1 | animalID) + (1 | year),
-                      data = filter(sealdat), 
+                      data = sealdat, 
                       family = "binomial", 
@@
 
+Sample size n = `r nrow(sealdat)`.
+
 ### H2: Maternal effect senescence
@@
 #B Pup survival and recruitment doesn't change with increased maternal age.
+
+##NOTE: Need to take out the instances where recruitment > survival
 
@@
 
+Sample size n = `r nrow(surv_recruit_data)`.
+
 ### H3a: Changes in pup sex ratio with increased maternal age
@@
 ```
+
+Sample size n = `r nrow(pup_sex_data)`.
 
@@
 ```
+
+Sample size n = `r nrow(seal_phenology)`.
 

### 2023-12-07 scratch/phen_predict.R

@@ -1,8 +1,6 @@
 
 expand_grid(
-  age = 4:19,
-  observed = c("Breeder", "Non-breeder")
-) %>% 
+  age = 9:19) %>% 
   mutate(ba = predict(phenology_models$model[[1]],
                       newdata = .,
                       re.form = NA),
@@ -16,9 +14,7 @@ expand_grid(
                       newdata = .,
                       re.form = NA)) %>% 
   mutate(short_trip = ma - bd, 
-         molt_dur = md - ma, 
- #        long_trip = 365 - md ) %>% 
-  filter(observed == "Breeder", 
-         age %in% c(4,19)) %>% 
+         molt_dur = md - ma) %>%  
+  filter(age %in% c(9,19)) %>% 
   view()
 

### 2023-12-08 paper/paperv3.qmd
Fix quadratic fits, add year lines to fig 1b

@@
   filter(age > 3, 
-         year > 1987)
+         year > 1987) %>% 
+  mutate(across(c(animalID, year), factor),
+         age10 = age / 10)
 
@@
 
-breeding_mod <- glmer(observed_int ~ poly(age, 2) + (1 | animalID) + (1 | year),
+breeding_mod <- glmer(
+  observed_int ~ poly(age10, 2, raw = TRUE) + (1 | animalID) + (1 | year),
   data = sealdat, 
   family = "binomial",
-                      control = glmerControl(optimizer = "bobyqa"))
+  control = glmerControl(optimizer="bobyqa")
+)
 
-breed_pred <- ggpredict(breeding_mod, terms = c("age [all]"))
+breed_pred <- ggpredict(breeding_mod, terms = "age10 [all]") %>% 
+  as_tibble() %>% 
+  mutate(age = x * 10)
+
+ranef_year <- ranef(breeding_mod)$year %>% 
+  as_tibble(rownames = "year") %>% 
+  arrange(`(Intercept)`) %>% 
+  slice(1, floor(nrow(.) * c(0.25, 0.5, 0.75, 1))) %>% 
+  expand_grid(age10 = unique(sealdat$age10)) %>% 
+  mutate(predicted = predict(breeding_mod, newdata = ., re.form = ~ (1 | year), type = "response"),
+         age = age10 * 10)
 
 breed_mod_summary <- summary(breeding_mod)
+
+breed_mod_summary
 
@@
   ggplot(aes(x = age, y = perc_breed)) + 
-  geom_ribbon(aes(x, ymin = conf.low, ymax = conf.high),
+  geom_ribbon(aes(age, ymin = conf.low, ymax = conf.high),
               breed_pred,
@@
               alpha = 0.25) +
-  geom_line(aes(x, predicted),
+  geom_line(aes(age, predicted, group = year),
+            ranef_year,
+            inherit.aes = FALSE,
+            alpha = 0.2) +
+  geom_text(aes(x = 22.1, predicted, label = year),
+            filter(ranef_year, age == 22),
+            inherit.aes = FALSE,
+            hjust = 0,
+            alpha = 0.5,
+            size = 3) +
+  geom_line(aes(age, predicted),
             breed_pred,
@@
   scale_y_continuous(labels = scales::percent, limits = c(0.5, 1)) + 
+  expand_limits(x = 23) +
   labs(x = "Maternal Age (Years)", y = "% Breeders (Mean ± SE)") + 
@@
 ##NOTE: Need to take out the instances where recruitment > survival
+#what does ggpredict do with random effects? 
+#thin gray lines for the different year effects? show the interannual variation (5 that include extreme ones, just something to illustrate the variation of year)
 
@@
   filter(n_pups >= 5)
-pup_surv_model <- glmer(pup_survived_dbl ~ poly(age, 2) + (1 | animalID) + (1 | year),
+pup_surv_model <- glmer(
+  pup_survived_dbl ~ poly(age10, 2, raw = TRUE) + (1 | animalID) + (1 | year),
   data = surv_recruit_data, 
-                        family = "binomial")
-pup_surv_pred <- ggpredict(pup_surv_model, terms = c("age [all]")) %>% 
+  family = "binomial"
+)
+summary(pup_surv_model)
+pup_surv_pred <- ggpredict(pup_surv_model, terms = c("age10 [all]")) %>% 
   as_tibble() %>% 
-  transmute(age = x, 
+  transmute(age = x * 10, 
             metric = "surv", 
@@
             mean_upr = conf.high)
-pup_recruit_model <- glmer(puprecruited ~ poly(age, 2) + (1 | animalID) + (1 | year),
+pup_recruit_model <- glmer(
+  puprecruited ~ poly(age10, 2, raw = TRUE) + (1 | animalID) + (1 | year),
   data = surv_recruit_data, 
-                           family = "binomial")
-pup_recruit_pred <- ggpredict(pup_recruit_model, terms = c("age [all]")) %>% 
+  family = "binomial"
+)
+summary(pup_recruit_model)
+pup_recruit_pred <- ggpredict(pup_recruit_model, terms = c("age10 [all]")) %>% 
   as_tibble() %>% 
-  transmute(age = x, 
+  transmute(age = x * 10, 
             metric = "recruit", 
@@
                      family = "binomial")
+summary(pup_sex_mod)
 
@@
   filter(observed == "Breeder") %>% 
-  mutate(BA = firstobsbreeddoy, 
-         BD = firstobsbreeddoy + breeddur, 
+  mutate(BA = ifelse(firstobsbreeddoy > 200, firstobsbreeddoy - 365, firstobsbreeddoy), 
+         BD = BA + breeddur, 
          MA = lastobsmoltdoy - moltdur, 
@@
 
-#NOTE: Both BA and BD have some points that are really late (above 300 doy) that need fixing
-BA_age <- ggplot(seal_phenology, aes(age, BA)) + 
-  geom_point() + 
-  scale_y_continuous(limits = c(0, 150)) + 
-  theme_classic() + 
-  labs(y = "Breeding arrival (DOY)", x = "Maternal age (Years)")
-
-BD_age <- ggplot(seal_phenology, aes(age, BD)) + 
-  geom_point() + 
-  scale_y_continuous(limits = c(0, 150)) + 
-  theme_classic() + 
-  labs(y = "Breeding departure (DOY)", x = "Maternal age (Years)")
-
-MA_age <- ggplot(seal_phenology, aes(age, MA)) + 
-  geom_point() + 
-  geom_smooth(method = "lm", color = "#829BD1") + 
-  scale_y_continuous(limits = c(50, 200)) + 
-  theme_classic() + 
-  labs(y = "Molt arrival (DOY)", x = "Maternal age (Years)")
-
-MD_age <- ggplot(seal_phenology, aes(age, MD)) + 
-  geom_point() + 
-  geom_smooth(method = "lm", color = "#829BD1") + 
-  scale_y_continuous(limits = c(50, 200)) + 
-  theme_classic() + 
-  labs(y = "Molt departure (DOY)", x = "Maternal age (Years)")
-
-plot_grid(BA_age, BD_age, MA_age, MD_age, nrow = 2, ncol = 2)
+#NOTE: Both BA and BD have some points that are really late (above 300 doy) that need fixing. Maybe make boxplots also..?
+seal_phenology %>% 
+  rename(`Breeding\narrival` = BA,
+         `Breeding\ndeparture` = BD,
+         `Molt\narrival` = MA,
+         `Molt\ndeparture` = MD) %>% 
+  pivot_longer(`Breeding\narrival`:`Molt\ndeparture`, 
+               names_to = "event", 
+               values_to = "DOY") %>% 
+ggplot(aes(y = DOY)) +
+  geom_boxplot(aes(factor(age))) +
+  geom_smooth(aes(as.integer(factor(age))), method = "lm", formula = y ~ x, se = TRUE) +
+  facet_wrap(~ event, nrow = 2, scales = "free_y") +
+  labs(x = "Age (years)", y = "Day of year") +
+  theme_classic()
 ```

### 2023-12-08 paper/paperv3.qmd
Updating supplemental tables

@@
 #| fig-cap: "Pup survival and recruitment are not significantly affected by maternal age. Points are mean values for maternal age. Error bars show SE."
-#| fig-width: 4.5
+#| fig-width: 5.5
 
@@
 
+```{r}
+#| label: tbl-survrecruit-cap
+
+surv_recruit_cap <- str_glue("The results from offspring survival and recruitment analyses.
+In the offspring survival model, the random effect variance of animal id is {format_signif(VarCorr(pup_surv_model)$animalID, 3)} and of year was {format_signif(VarCorr(pup_surv_model)$year, 3)}.
+In the offspring recruitment model, the random effect variance of animal id is {format_signif(VarCorr(pup_recruit_model)$animalID, 3)} and of year was {format_signif(VarCorr(pup_recruit_model)$year, 3)}")
+
+```
+
 Supplemental table for offspring survival
@@
 ```{r}
+#| tbl-cap: !expr "surv_recruit_cap"
+#| label: tbl-surv_recruit
+
 surv_mod_summary <- summary(pup_surv_model)
@@
   kableExtra::collapse_rows(columns = 1)
-```
 
-Supplemental figure for survival bias:
-
-```{r}
-#| label: fig-surv_bias
-#| fig-cap: "A: Plotting the intercept of the random effect of year shows that variation in ocean conditions is likely driving patterns in survival, rather than age of mother.
-B: 95% of seals that survived were resighted within ~seven years of their birth."
-#| fig-width: 9
-
-year_ranef <- ranef(pup_surv_model)$year %>% 
-  as_tibble(rownames = "year") %>% 
-  mutate(year = as.numeric(year)) %>% 
-  rename(intercept = `(Intercept)`)
-
-year_ranef_plot <- ggplot(year_ranef, aes(year, intercept)) + 
-  geom_point() + 
-  geom_smooth(method = "loess", 
-              formula = y ~ x,
-              color = "red") + 
-  theme_classic()
-
-survival_bias <- sealdat %>% 
-  group_by(animalID) %>% 
-  summarize(first_resight_yrs = min(age))
-
-survival95 <- quantile(survival_bias$first_resight_yrs, 0.95)
-#n for 95% quantile
-sum(survival_bias$first_resight_yrs > 7)
-
-survival_bias_plot <- ggplot(survival_bias, aes(x = first_resight_yrs)) +
-  stat_ecdf() +
-  geom_vline(xintercept = survival95, linetype = "dashed", color = "red") +
-  scale_x_continuous(limits = c(0, 10)) + 
-  theme_classic()
-
-plot_grid(year_ranef_plot, survival_bias_plot, nrow = 1, labels = "AUTO")
 ```
@@
 ```{r}
+#| label: tbl-pupsex-cap
+
+pupsex_cap <- str_glue("The results from offspring sex ratio analyses.
+In the model, the random effect variance of animal id is {format_signif(VarCorr(pup_sex_mod)$animalID, 3)} and of year was {format_signif(VarCorr(pup_sex_mod)$year, 3)}.
+```
+
+```{r}
+#| label: tbl-pupsex
+#| tbl-cap: !expr "pupsex_cap"
+
 pupsex_mod_summary <- summary(pup_sex_mod)

### 2023-12-08 00_clean_data.R

@@ -54,6 +54,7 @@ outall$withpupCor[outall$withpup==0]=0
 #Encounter History
 g1=as.data.frame.matrix(with(outall,table(animalID,calyear))) #table : combinations of observation year and animal ID
 #presence/absence with only encounter history (no breeding info, and no multiple sightings per year)
+gg=g1
 ############################################ NEW***
 g1[g1<=3]=0 #seen less than or equal to 3 times means "missing" = 0
 g1[g1!=0]=1 #seen more than 3 times is "seen" = 1
@@ -160,7 +161,7 @@ rows=rownames(subset(final,!is.na(pupID)))
 
 for(i in rows){ #for each row of those rows
   #make encounter history subset of g1 where the row names of g1 are equal to the pup ID
-  enc=subset(g1,row.names(g1)==final$pupID[rownames(final)==i])
+  enc=subset(gg,row.names(gg)==final$pupID[rownames(final)==i])
   if(nrow(enc)>0&final$firstyearindicator[rownames(final)==i]!=2024){
     final[i,"pupseeneveragain"]=sum(enc[,final$indices[rownames(final)==i]:ncol(g1)])
     encb=subset(g1b,row.names(g1b)==final$pupID[rownames(final)==i])

### 2023-12-08 scratch/phen_predict.R

@@ -1,6 +1,6 @@
 
 expand_grid(
-  age = 9:19) %>% 
+  age = 6:16) %>% 
   mutate(ba = predict(phenology_models$model[[1]],
                       newdata = .,
                       re.form = NA),
@@ -15,6 +15,6 @@ expand_grid(
                       re.form = NA)) %>% 
   mutate(short_trip = ma - bd, 
          molt_dur = md - ma) %>%  
-  filter(age %in% c(9,19)) %>% 
+  filter(age %in% c(6,16)) %>% 
   view()
 

### 2023-12-08 paper/paperv3.qmd
Fixing the recruitment problem seals

@@
          age10 = age / 10)
-
-# 17 observations where pup recruited but did not survive
-problem_children <- filter(sealdat, puprecruited == 1 & !pup_survived)
-
-sealdat <- anti_join(sealdat, problem_children, by = c("animalID", "pupID"))
 ```
@@
 pupsex_cap <- str_glue("The results from offspring sex ratio analyses.
-In the model, the random effect variance of animal id is {format_signif(VarCorr(pup_sex_mod)$animalID, 3)} and of year was {format_signif(VarCorr(pup_sex_mod)$year, 3)}.
+In the model, the random effect variance of animal id is {format_signif(VarCorr(pup_sex_mod)$animalID, 3)} and of year was {format_signif(VarCorr(pup_sex_mod)$year, 3)}.")
 ```

### 2023-12-08 paper/paperv3.qmd
Getting the p value for quadratic pup sex model

@@
 
-### H3a: Changes in pup sex ratio with increased maternal age
+### H3: Changes in pup sex ratio with increased maternal age
 
@@
 
+pup_sex_quad <- glmer(
+  is_male ~ poly(age10, 2, raw = TRUE) + (1 | animalID) + (1 | year),
+  data = pup_sex_data, 
+  family = "binomial"
+  )
+
 pup_sex_pred <- ggpredict(pup_sex_mod, terms = c("age [all]")) %>% 
@@
 
-### H3b: Changes to annual cycle timing with increased maternal age
+### H4: Changes to annual cycle timing with increased maternal age
 
@@
 pupsex_mod_summary <- summary(pup_sex_mod)
+pupsex_quad_summary <- summary(pup_sex_quad)
 pupsex_mod_summary$coefficients %>% 
@@
 ```{r}
+#| fig-width: 8
 

### 2024-04-25 paper/paperv3.qmd
Making a figure to illustrate detection probability

@@
          age10 = age / 10)
+
+resight <- read.csv(here::here("data/raw/fullresights.csv"))
+
 ```
@@
 
-n = `r n_distinct(sealdat$animalID)` distinct animals
+First, address detection probability.
+We do resights near-daily during breeding and molting seasons, so while we may miss an individual on a given day, it is unlikely that we miss a marked individual throughout the entire breeding or molting season.
+
+```{r}
+#| label: detection-prob
+
+resight <- resight %>% 
+  dplyr::select(animalID, resightID, season, date, timeofyear, calyear) %>% 
+  mutate(animalID = as.factor(animalID))
+
+full <- semi_join(resight, sealdat, "animalID") %>% group_by(animalID, calyear) %>% count(calyear)
+
+ggplot(full, aes(calyear, n)) + 
+  geom_point()
+
+```
+
+#
+
+n = `r n_distinct(sealdat$animalID)` distinct animals.
 

### 2024-04-25 00_clean_data.R

@@ -6,8 +6,8 @@ library(patchwork)
 library(marked)
 library(RMark)
 
-animal <- read_csv(here::here("data/raw/animalforallison.csv"))
-resight <- read_csv(here::here("data/raw/resightforallison.csv"))
+animal <- read_csv(here::here("data/raw/animal.csv"))
+resight <- read_csv(here::here("data/raw/resight.csv"))
 
 # specify which animals (pups) to select
 animals_to_select <- animal %>%
@@ -46,10 +46,14 @@ outall <- merge(# merge resight
 
 outall$calyear=year(outall$date) #make a new column for the calendar year of that observation
 
+
 #correct pup status for weird character values
 outall$withpupCor=NA
 outall$withpupCor[outall$withpup%in%c(as.character(1:8),"2+","3+",">8","1 or 2",">1")]=1
 outall$withpupCor[outall$withpup==0]=0
+
+write.csv(outall,here::here("data/raw/fullresights.csv"),
+          row.names=FALSE)
 
 #Encounter History
 g1=as.data.frame.matrix(with(outall,table(animalID,calyear))) #table : combinations of observation year and animal ID

### 2024-04-25 paper/paperv3.qmd
fixing the semi join

@@
 
-full <- semi_join(resight, sealdat, "animalID") %>% group_by(animalID, calyear) %>% count(calyear)
+full <- semi_join(resight, sealdat, "animalID") %>% 
+  group_by(animalID, calyear) %>% 
+  count(calyear) %>% 
+  mutate(year = as.numeric(calyear)) %>% 
+  filter(year > 1987)
 
-ggplot(full, aes(calyear, n)) + 
+ggplot(full, aes(year, n)) + 
   geom_point()

### 2024-05-01 paper/paperv3.qmd
Adding sections to check out recapture rate

--- 2024-04-25 paper/paperv3.qmd
@@
 
-resight <- read.csv(here::here("data/raw/fullresights.csv"))
+resight <- read.csv(here::here("data/raw/fullresights.csv")) %>% 
+  dplyr::select(animalID, resightID, season, date, timeofyear, calyear, yday) %>% 
+  mutate(animalID = as.factor(animalID))
 
@@
 
-resight <- resight %>% 
-  dplyr::select(animalID, resightID, season, date, timeofyear, calyear) %>% 
-  mutate(animalID = as.factor(animalID))
+full <- semi_join(resight, sealdat, "animalID") %>% 
+  mutate(fiscalyear = ifelse(yday > 200, calyear + 1, calyear)) %>% 
+  group_by(animalID, fiscalyear) %>% 
+  count(fiscalyear) %>% 
+  mutate(year = as.numeric(fiscalyear)) %>% 
+  filter(year > 1987)
 
-full <- semi_join(resight, sealdat, "animalID") %>% group_by(animalID, calyear) %>% count(calyear)
+full %>% 
+  group_by(year) %>% 
+  summarize(across(n, list(median = median, 
+                           q25 = \(x) quantile(x, 0.25), 
+                           q75 = \(x) quantile(x, 0.75)))) %>% 
+  ggplot(aes(year, n_median)) +
+  geom_ribbon(aes(ymin = n_q25, ymax = n_q75), fill = "grey80") +
+  geom_line() +
+  geom_hline(yintercept = 3, linetype = "dashed", color = "firebrick") +
+  labs(y = "Resights per year") + 
+  theme_classic()
 
-ggplot(full, aes(calyear, n)) + 
+#some animals have multiple rows because they have multiple pups associated with them...
+sealdat %>% 
+  group_by(animalID, year) %>% 
+  mutate(n = n()) %>% 
+  filter(n > 2)
+
+skips <- sealdat %>% 
+  distinct(animalID, year) %>% 
+  group_by(animalID) %>% 
+  summarize(n_obs = n(),
+            obs_window = max(year) - min(year) + 1,
+            skips = obs_window - n_obs) %>% 
+  group_by(obs_window, skips) %>% 
+  summarize(ind = n(), .groups = "drop")
+
+resights_per_year <- full %>% 
+  group_by(year) %>% 
+  summarize(median_resights = median(n))
+
+skips_per_year <- skips
+
+skip_frac <- sealdat %>% 
+  distinct(animalID, year) %>% 
+  mutate(year = as.numeric(as.character(year))) %>% 
+  group_by(animalID) %>% 
+  summarize(firstobs = min(year),
+            lastobs = max(year)) %>% 
+  mutate(n = lastobs - firstobs + 1) %>% 
+  uncount(n) %>% 
+  group_by(animalID) %>% 
+  mutate(year = firstobs + seq(n()) - 1) %>% 
+  ungroup() %>% 
+  mutate(year = factor(year)) %>% 
+  left_join(transmute(sealdat, animalID, year, skip = 0), by = c("animalID", "year")) %>% 
+  mutate(skip = ifelse(is.na(skip), 1, 0)) %>% 
+  group_by(year) %>% 
+  summarize(skip_frac = sum(skip) / n())
+
+resights_per_year %>% 
+  mutate(year = factor(year)) %>% 
+  left_join(skip_frac, by = "year") %>% 
+  mutate(year = as.numeric(as.character(year))) %>% 
+  ggplot(aes(median_resights, skip_frac)) +
+  geom_point(aes(fill = year), shape = 21, color = "grey20", size = 2) +
+  ggrepel::geom_label_repel(aes(label = year), max.overlaps = 8) +
+  scale_y_continuous("Skip fraction", labels = scales::percent) +
+  scale_fill_distiller(palette = "PuRd") +
+  labs(x = "Median resights") +
+  theme_classic()
+
+#mark recapture model is necessary? could have bred at another colony. Many skips!!
+
+ggplot(skips, aes(obs_window, skips, size = ind)) +
   geom_point()

### 2024-05-01 scratch/markrecapture.R
Mark recapture initiated!

@@ -1,0 +1,51 @@
+library(tidyverse)
+
+sealdat <- read_csv(here::here("data/raw/128L pull 2023_12_05.csv")) %>% 
+  filter(age > 3)
+
+# int<lower=0> nind;
+# int<lower=0> n_occasions;
+# array[nind, n_occasions] int<lower=1, upper=3> y;                   
+
+animalIDs <- sort(unique(sealdat$animalID))
+years <- sort(unique(sealdat$year))
+n_indiv <- length(animalIDs)
+n_occ <- length(years)
+
+sealobs <- matrix(0, nrow = n_indiv, ncol = n_occ)
+
+for (i in seq(n_indiv)) {
+  for (j in seq(n_occ)) {
+    animalID <- animalIDs[i]
+    year <- years[j]
+    obs <- sealdat$observed[sealdat$animalID == animalID & sealdat$year == year]
+    #workaround for some seals that have multiple observations in a year
+    if (length(obs) > 1)
+      obs <- "B"
+    if (length(obs) == 0)
+      sealobs[i, j] <- 3
+    else if (obs == "B")
+      sealobs[i, j] <- 1 
+    else 
+      sealobs[i, j] <- 2
+  }
+}
+
+# Seal resight map
+image(years, seq_along(animalIDs), t(sealobs))
+
+# Fit the dang model
+markrecapdat <- list(
+  nind = n_indiv, 
+  n_occasions = n_occ, 
+  y = sealobs
+)
+markrecapmod <- rstan::stan("scratch/multistate.stan", 
+                            data = markrecapdat, 
+                            chains = 1, iter = 200)
+
+# Look at posterior distribution
+library(posterior)
+markrecapdraws <- as_draws_df(markrecapmod) %>% 
+  subset_draws(variable = names(markrecapmod)[1:10])
+summarize_draws(markrecapdraws)

### 2024-05-01 scratch/markrecapture.R
adding time varying detection

@@ -1,4 +1,6 @@
+library(posterior)
 library(tidyverse)
+library(rstan)
 
 sealdat <- read_csv(here::here("data/raw/128L pull 2023_12_05.csv")) %>% 
   filter(age > 3)
@@ -34,18 +36,36 @@ for (i in seq(n_indiv)) {
 # Seal resight map
 image(years, seq_along(animalIDs), t(sealobs))
 
+# Basic model -------------------------------------------------------------
+
 # Fit the dang model
 markrecapdat <- list(
   nind = n_indiv, 
   n_occasions = n_occ, 
   y = sealobs
 )
-markrecapmod <- rstan::stan("scratch/multistate.stan", 
-                            data = markrecapdat, 
-                            chains = 1, iter = 200)
+markrecapmod <- stan("scratch/multistate.stan", 
+                     data = markrecapdat, 
+                     chains = 4, iter = 2000, cores = 4)
 
 # Look at posterior distribution
-library(posterior)
 markrecapdraws <- as_draws_df(markrecapmod) %>% 
   subset_draws(variable = names(markrecapmod)[1:10])
 summarize_draws(markrecapdraws)
+
+# Time varying detection  -------------------------------------------------
+
+# Fit the dang model
+markrecapdat <- list(
+  nind = n_indiv, 
+  n_occasions = n_occ, 
+  y = sealobs
+)
+markrecapmod <- stan("scratch/multistate2.stan", 
+                     data = markrecapdat, 
+                     chains = 4, iter = 2000, cores = 4)
+
+# Look at posterior distribution
+markrecapdraws <- as_draws_df(markrecapmod)
+summarize_draws(markrecapdraws)
+

### 2024-05-02 cluster/test.R
test for hummingbird

@@ -1,0 +1,40 @@
+library(posterior)
+library(rstan)
+library(tidyverse)
+set.seed(1234)
+
+bern.stan <-
+"
+data {
+  int<lower=0> N;               // number of trials
+  int<lower=0, upper=1> y[N];   // success on trial n
+}
+
+parameters {
+  real<lower=0, upper=1> theta; // chance of success
+}
+
+model {
+  theta ~ uniform(0, 1);        // prior
+  y ~ bernoulli(theta);         // likelihood
+}
+"
+
+theta <- 0.30
+N <- 20
+y <- rbinom(N, 1, theta)
+print(y)
+
+fit <- stan(model_code = bern.stan,
+          data = list(y = y, N = N),
+            iter = 5000)
+
+print(fit, probs = c(0.1, 0.9))
+
+theta_draws <- as_draws_df(fit)
+plotpostre <- ggplot(theta_draws, aes(theta)) +
+geom_density(color = "cornflowerblue", linewidth = 1.5) +
+    theme_classic()
+ggsave("test.png", plotpostre)
+
+print("DONE")

### 2024-05-02 scratch/markrecapture.R
Adding fixed effect for age

@@ -3,7 +3,8 @@ library(tidyverse)
 library(rstan)
 
 sealdat <- read_csv(here::here("data/raw/128L pull 2023_12_05.csv")) %>% 
-  filter(age > 3)
+  filter(age > 3,
+         year < 2023)
 
 # int<lower=0> nind;
 # int<lower=0> n_occasions;
@@ -13,6 +14,10 @@ animalIDs <- sort(unique(sealdat$animalID))
 years <- sort(unique(sealdat$year))
 n_indiv <- length(animalIDs)
 n_occ <- length(years)
+birth_years <- sealdat %>% 
+  arrange(animalID) %>%
+  distinct(animalID, yearborn) %>%
+  pull(yearborn) - min(years)
 
 sealobs <- matrix(0, nrow = n_indiv, ncol = n_occ)
 
@@ -33,20 +38,38 @@ for (i in seq(n_indiv)) {
   }
 }
 
+rownames(sealobs) <- animalIDs
+colnames(sealobs) <- years
+
 # Seal resight map
 image(years, seq_along(animalIDs), t(sealobs))
+
+test <- FALSE
+n_indiv_test <- 100
+animalIDs_test <- sample(animalIDs, n_indiv_test)
 
 # Basic model -------------------------------------------------------------
 
 # Fit the dang model
-markrecapdat <- list(
-  nind = n_indiv, 
-  n_occasions = n_occ, 
-  y = sealobs
-)
-markrecapmod <- stan("scratch/multistate.stan", 
-                     data = markrecapdat, 
-                     chains = 4, iter = 2000, cores = 4)
+if (test) {
+  markrecapdat <- list(
+    nind = n_indiv_test, 
+    n_occasions = n_occ, 
+    y = sealobs[animalIDs %in% animalIDs_test, ]
+  )
+  markrecapmod <- stan("scratch/multistate.stan", 
+                       data = markrecapdat, 
+                       chains = 1, iter = 200, cores = 1)
+} else {
+  markrecapdat <- list(
+    nind = n_indiv, 
+    n_occasions = n_occ, 
+    y = sealobs
+  )
+  markrecapmod <- stan("scratch/multistate.stan", 
+                       data = markrecapdat, 
+                       chains = 4, iter = 2000, cores = 4)
+}
 
 # Look at posterior distribution
 markrecapdraws <- as_draws_df(markrecapmod) %>% 
@@ -56,16 +79,138 @@ summarize_draws(markrecapdraws)
[…]

### 2024-05-02 cluster/markrecapture.R
slurm script for mark recap models

@@ -2,6 +2,15 @@ library(posterior)
 library(tidyverse)
 library(rstan)
 
+args <- commandArgs(trailingOnly = TRUE)
+if (args[1] == "test") {
+  test <- TRUE
+} else if (args[1] == "full") {
+  test <- FALSE
+} else {
+  stop(sprintf("command line argument must be 'test' or 'full', not %s", args[1]))
+}
+
 sealdat <- read_csv(here::here("data/raw/128L pull 2023_12_05.csv")) %>% 
   filter(age > 3,
          year < 2023)
@@ -44,7 +53,7 @@ colnames(sealobs) <- years
 # Seal resight map
 image(years, seq_along(animalIDs), t(sealobs))
 
-test <- FALSE
+# Test parameters
 n_indiv_test <- 100
 animalIDs_test <- sample(animalIDs, n_indiv_test)
 

### 2024-05-02 cluster/markrecapture.R
Making sure that plots will be saved

@@ -57,6 +57,10 @@ image(years, seq_along(animalIDs), t(sealobs))
 n_indiv_test <- 100
 animalIDs_test <- sample(animalIDs, n_indiv_test)
 
+#Setting up directory to save plots
+direct_name <- here::here(file.path("cluster", Sys.time() %>% format("%Y%m%d%H%M")))
+dir.create(direct_name)
+
 # Basic model -------------------------------------------------------------
 
 # Fit the dang model
@@ -66,7 +70,7 @@ if (test) {
     n_occasions = n_occ, 
     y = sealobs[animalIDs %in% animalIDs_test, ]
   )
-  markrecapmod <- stan("scratch/multistate.stan", 
+  markrecapmod <- stan("cluster/multistate.stan", 
                        data = markrecapdat, 
                        chains = 1, iter = 200, cores = 1)
 } else {
@@ -75,7 +79,7 @@ if (test) {
     n_occasions = n_occ, 
     y = sealobs
   )
-  markrecapmod <- stan("scratch/multistate.stan", 
+  markrecapmod <- stan("cluster/multistate.stan", 
                        data = markrecapdat, 
                        chains = 4, iter = 2000, cores = 4)
 }
@@ -94,7 +98,7 @@ if (test) {
     n_occasions = n_occ, 
     y = sealobs[animalIDs %in% animalIDs_test, ]
   )
-  markrecapmod2 <- stan("scratch/multistate2.stan", 
+  markrecapmod2 <- stan("cluster/multistate2.stan", 
                         data = markrecapdat2, 
                         chains = 1, iter = 200, cores = 1)
 } else {
@@ -103,7 +107,7 @@ if (test) {
     n_occasions = n_occ, 
     y = sealobs
   )
-  markrecapmod2 <- stan("scratch/multistate2.stan", 
+  markrecapmod2 <- stan("cluster/multistate2.stan", 
                         data = markrecapdat2, 
                         chains = 4, iter = 2000, cores = 4)
 }
@@ -111,7 +115,7 @@ if (test) {
 # Look at posterior distribution
 markrecapdraws2 <- as_draws_df(markrecapmod2)
 # Detection probability for breeders and non-breeders
-markrecapdraws2 %>% 
+p1 <- markrecapdraws2 %>% 
   select(starts_with(c("pA[", "pB["))) %>% 
   pivot_longer(everything(), 
                names_to = "year", 
[…]

### 2024-05-02 cluster/markrecapture.R
Adding .gitignore for outputs

@@ -58,8 +58,9 @@ n_indiv_test <- 100
 animalIDs_test <- sample(animalIDs, n_indiv_test)
 
 #Setting up directory to save plots
-direct_name <- here::here(file.path("cluster", Sys.time() %>% format("%Y%m%d%H%M")))
-dir.create(direct_name)
+output_dir <- paste0("output", format(Sys.time(), "%Y%m%d%H%M"))
+output_path <- here::here(file.path("cluster", output_dir))
+dir.create(output_path)
 
 # Basic model -------------------------------------------------------------
 
@@ -137,7 +138,7 @@ p1 <- markrecapdraws2 %>%
                                 `Non-breeder` = "cornflowerblue")) +
   labs(x = "Year", y = "Detection probability") +
   theme_classic()
-ggsave(file.path(direct_name, "DetectProbTimeVary.png"), p1)
+ggsave(file.path(output_path, "DetectProbTimeVary.png"), p1)
 
 
 # Survival for breeders and non-breeders
@@ -152,7 +153,7 @@ p2 <- markrecapdraws2 %>%
   scale_color_manual(values = c(Breeder = "firebrick", 
                                 `Non-breeder` = "cornflowerblue")) +
   theme_classic()
-ggsave(file.path(direct_name, "SurvivalTimeVary.png"), p2)
+ggsave(file.path(output_path, "SurvivalTimeVary.png"), p2)
 
 # Add fixed effect of year on survival and repro --------------------------
 
@@ -206,7 +207,7 @@ p3 <- markrecapdraws3 %>%
   labs(x = "Year", y = "Detection probability") +
   theme_classic() +
   theme(legend.position = "none")
-ggsave(file.path(direct_name, "DetectProbFixedEffect.png"), p3)
+ggsave(file.path(output_path, "DetectProbFixedEffect.png"), p3)
 
 # Survival by age
 inv_logit <- function(x) exp(x) / (1 + exp(x))
@@ -232,4 +233,4 @@ p4 <- markrecapdraws3 %>%
   theme_classic() +
   theme(legend.position = "none")
 
-ggsave(file.path(direct_name, "SurvivalFixedEffect.png"), p4)
+ggsave(file.path(output_path, "SurvivalFixedEffect.png"), p4)

### 2024-05-21 paper/paperv3.qmd
Trying segmented. No GLMM support tho :(

@@
 library(patchwork)
+library(segmented)
 source(here::here("R/format.R"))
@@
          year > 1987) %>% 
-  mutate(across(c(animalID, year), factor),
+  mutate(animalID = factor(animalID),
+         year_fct = factor(year),
          age10 = age / 10)
@@
 resight <- read.csv(here::here("data/raw/fullresights.csv")) %>% 
-  dplyr::select(animalID, resightID, season, date, timeofyear, calyear, yday) %>% 
+  select(animalID, resightID, season, date, timeofyear, calyear, yday) %>% 
   mutate(animalID = as.factor(animalID))
@@
   distinct(animalID, year) %>% 
-  mutate(year = as.numeric(as.character(year))) %>% 
   group_by(animalID) %>% 
@@
   ungroup() %>% 
-  mutate(year = factor(year)) %>% 
-  left_join(transmute(sealdat, animalID, year, skip = 0), by = c("animalID", "year")) %>% 
+  left_join(transmute(sealdat, animalID, year, skip = 0), 
+            by = c("animalID", "year")) %>% 
   mutate(skip = ifelse(is.na(skip), 1, 0)) %>% 
@@
 resights_per_year %>% 
-  mutate(year = factor(year)) %>% 
   left_join(skip_frac, by = "year") %>% 
-  mutate(year = as.numeric(as.character(year))) %>% 
   ggplot(aes(median_resights, skip_frac)) +
@@
 breeding_mod <- glmer(
-  observed_int ~ poly(age10, 2, raw = TRUE) + (1 | animalID) + (1 | year),
+  observed_int ~ poly(age10, 2, raw = TRUE) + (1 | animalID) + (1 | year_fct),
   data = sealdat, 
@@
 
-ranef_year <- ranef(breeding_mod)$year %>% 
-  as_tibble(rownames = "year") %>% 
+ranef_year <- ranef(breeding_mod)$year_fct %>% 
+  as_tibble(rownames = "year_fct") %>% 
   arrange(`(Intercept)`) %>% 
@@
   expand_grid(age10 = unique(sealdat$age10)) %>% 
-  mutate(predicted = predict(breeding_mod, newdata = ., re.form = ~ (1 | year), type = "response"),
+  mutate(year_fct = factor(year_fct, levels = levels(sealdat$year_fct))) %>% 
+  mutate(predicted = predict(breeding_mod, 
+                             newdata = ., 
+                             re.form = ~ (1 | year_fct), type = "response"),
          age = age10 * 10)
@@
 #analyses fit to the age in which we have larger sample sizes fit also have a quadratic fit and don't change the story
+
+```
+
+Trying segmented analysis. Note: `segmented` does NOT support mixed effect GLMs!
+
+```{r}
+#| label: age_breeding_seg
+
+breeding_mod2 <- glm(
+  observed_int ~ age,
+  data = sealdat, 
+  family = "binomial"
+)
+breed_nseg <- lapply(
+  list(15, c(9, 15)),
+  \(.psi) {
+    segmented(breeding_mod2, seg.Z = ~age, psi = .psi)
+  }
+)
+AIC(breeding_mod2, breed_nseg[[1]], breed_nseg[[2]]) %>% 
+  as_tibble() %>% 
+  mutate(model = paste(1:3, "segments")) %>% 
+  transmute(model, AIC, dAIC = AIC - min(AIC)) %>% 
+  arrange(dAIC)
+
+breeding_best <- breed_nseg[[1]]
+breeding_pred <- tibble(age = seq(4, 22, length.out = 100)) %>% 
+  predict(breeding_best, 
+          newdata = ., 
+          type = "response", 
+          interval = "confidence") %>% 
+  as_tibble() %>% 
+  set_names(c("bred", "bred_lwr", "bred_upr")) %>% 
+  mutate(age = seq(4, 22, length.out = 100))
+breed_nonbreed <- sealdat %>% 
+  group_by(age) %>% 
+  summarize(n_observed = sum(!is.na(observed)), 
+            perc_breed = sum(observed == "Breeder") / n_observed, 
+            se = (perc_breed * (1 - perc_breed) / n_observed)^0.5) %>% 
+  filter(age < 22)
+
+ggplot(breeding_pred, aes(age)) +
+  geom_ribbon(aes(ymin = bred_lwr, ymax = bred_upr),
+              fill = "grey90") + 
+  geom_line(aes(y = bred)) +
+  geom_point(aes(y = perc_breed), breed_nonbreed) + 
+  geom_errorbar(aes(ymin = perc_breed - se, ymax = perc_breed + se), 
+                breed_nonbreed,
+                width = 0) +
+  theme_classic()
 
@@
               alpha = 0.25) +
-  geom_line(aes(age, predicted, group = year),
+  geom_line(aes(age, predicted, group = year_fct),
             ranef_year,
@@
             alpha = 0.2) +
-  geom_text(aes(x = 22.1, predicted, label = year),
+  geom_text(aes(x = 22.1, predicted, label = year_fct),
             filter(ranef_year, age == 22),
@@
   geom_point() + 
-  geom_errorbar(aes(ymin = perc_breed - se, ymax = perc_breed + se), width = 0) +
+  geom_errorbar(aes(ymin = perc_breed - se, ymax = perc_breed + se), 
+                width = 0) +
   scale_y_continuous(labels = scales::percent, limits = c(0.5, 1)) + 

### 2024-05-22 paper/paperv3.qmd
Tried filling "gaps", but will use mark-recap instead

@@
 library(ggeffects)
+library(ggridges)
 library(patchwork)
-library(segmented)
 source(here::here("R/format.R"))
@@
          year_fct = factor(year),
+         age10 = age / 10) %>% 
+  group_by(animalID) %>% 
+  mutate(longevity = ifelse(max(year) < 2020,
+                            max(age),
+                            NA))
+
+all_years <- sealdat %>% 
+  group_by(animalID, yearborn, longevity) %>% 
+  summarize(age4 = yearborn[1] + 4,
+            first_obs = min(year),
+            last_obs = max(year),
+            .groups = "keep") %>% 
+  reframe(year = seq(age4, last_obs)) %>% 
+  filter(between(year, 1987, 2020)) %>% 
+  left_join(select(sealdat, animalID, year, observed),
+            by = c("animalID", "year")) %>% 
+  mutate(observed2 = factor(ifelse(is.na(observed), 
+                                   "Non-breeder", 
+                                   observed),
+                            levels = c("Non-breeder", "Breeder")),
+         year_fct = factor(year),
+         age = year - yearborn,
          age10 = age / 10)
@@
 
-#
-
 n = `r n_distinct(sealdat$animalID)` distinct animals.
@@
 
-We predicted that breeding success would decline with age. We found that breeding success increases to prime age, then decreases at older ages (@fig-age_dist).
+~~We predicted that breeding success would decline with age. We found that breeding success increases to prime age, then decreases at older ages (@fig-age_dist).~~
+
+```{r}
+#| label: breeding-mod 
+#| fig-width: 9
+#| fig-height: 6
+
+breed_age_mod <- glmer(
+  observed2 ~ age10 + (1 | animalID) + (1 | year_fct),
+  data = all_years, 
+  family = "binomial",
+  control = glmerControl(optimizer="bobyqa")
+)
+
+year_ranef <- ranef(breed_age_mod)$year_fct %>% 
+  as_tibble(rownames = "year") %>% 
+  mutate(year = as.numeric(year)) %>% 
+  rename(ranef = 2) %>% 
+  arrange(ranef)
+
+year_ranef2 <- slice(year_ranef, 
+                     round(seq(1, nrow(year_ranef), length.out = 5)))
+
+indiv_ranef <- ranef(breed_age_mod)$animalID %>% 
+  as_tibble(rownames = "animalID") %>% 
+  mutate(animalID = factor(animalID)) %>% 
+  rename(ranef = 2) %>% 
+  arrange(ranef)
+
+indiv_ranef2 <- slice(indiv_ranef, 
+                      round(seq(1, nrow(indiv_ranef), length.out = 5)))
+
+breed_age_pred_year <- ggpredict(breed_age_mod, 
+                                 terms = c("age10 [all]", "year_fct"), 
+                                 type = "random",
+                                 interval = "confidence") %>% 
+  as_tibble() %>% 
+  mutate(age = x * 10,
+         year_fct = group,
+         year = as.numeric(as.character(group))) %>% 
+  semi_join(year_ranef2, by = "year") %>% 
+  filter(age <= 20)
+
+breed_age_pred_indiv <- ggpredict(breed_age_mod, 
+                                  terms = c("age10 [all]", "animalID"), 
+                                  type = "random",
+                                  interval = "confidence") %>% 
+  as_tibble() %>% 
+  mutate(age = x * 10,
+         animalID = group) %>% 
+  semi_join(indiv_ranef2, by = "animalID") %>% 
+  filter(age <= 20)
+
+breed_age_pred <- ggpredict(breed_age_mod, 
+                                 terms = "age10 [all]", 
+                                 type = "fixed",
+                                 interval = "confidence") %>% 
+  as_tibble() %>% 
+  mutate(age = x * 10) %>% 
+  filter(age <= 20)
+
+breed_nonbreed <- all_years %>% 
+  group_by(age) %>% 
+  summarize(n_observed = sum(!is.na(observed2)), 
+            perc_breed = sum(observed2 == "Breeder") / n_observed, 
+            se = (perc_breed * (1 - perc_breed) / n_observed)^0.5) %>% 
+  filter(age <= 20)
+
+pred_year_plot <- ggplot(breed_age_pred_year, aes(age)) +
+  # Prediction across years (fixed effect only)
+  geom_ribbon(aes(ymin = conf.low, ymax = conf.high),
+              breed_age_pred,
+              fill = "grey", alpha = 0.5) +
+  geom_line(aes(y = predicted), 
+            breed_age_pred,
+            color = "firebrick", linewidth = 1.5) +
+  # Predictions for specific years (random effect)
+  geom_line(aes(y = predicted, group = year_fct), 
+            color = "cornflowerblue", linewidth = 1) +
+  # Raw data
+  geom_pointrange(aes(y = perc_breed, 
+                      ymin = perc_breed - se, 
+                      ymax = perc_breed + se),
+                  breed_nonbreed) + 
+  # Year labels
+  geom_text(aes(y = predicted, label = year_fct), 
+            filter(breed_age_pred_year, age == 20),
+            x = 20.2,
+            hjust = 0,
+            size = 3) +
+  scale_x_continuous("Age", limits = c(NA, 21)) +
+  scale_y_continuous("Breeding %", 
+                     limits = c(0, 1), 
+                     labels = scales::percent) +
+  theme_classic()
+
+pred_indiv_plot <- ggplot(breed_age_pred_indiv, aes(age)) +
+  # Prediction across years (fixed effect only)
+  geom_ribbon(aes(ymin = conf.low, ymax = conf.high),
+              breed_age_pred,
+              fill = "grey", alpha = 0.5) +
+  geom_line(aes(y = predicted), 
+            breed_age_pred,
+            color = "firebrick", linewidth = 1.5) +
+  # Predictions for specific years (random effect)
+  geom_line(aes(y = predicted, group = animalID), 
+            color = "cornflowerblue", linewidth = 1) +
+  # Raw data
+  geom_pointrange(aes(y = perc_breed, 
+                      ymin = perc_breed - se, 
+                      ymax = perc_breed + se),
+                  breed_nonbreed) + 
+  # Individual labels
+  geom_text(aes(y = predicted, label = animalID), 
+            filter(breed_age_pred_indiv, age == 20),
+            x = 20.2,
+            hjust = 0,
+            size = 3) +
+  scale_x_continuous("Age", limits = c(NA, 21)) +
+  scale_y_continuous("Breeding %", 
+                     limits = c(0, 1), 
+                     labels = scales::percent) +
+  theme_classic()
+
+pred_year_plot | pred_indiv_plot
+
+year_plot <- all_years %>% 
+  filter(age <= 20) %>% 
+  left_join(year_ranef, by = "year") %>% 
+  group_by(age) %>% 
+  summarize(across(ranef, list(mean = mean, sd = sd))) %>% 
+  ggplot(aes(age, ranef_mean)) +
+  geom_pointrange(aes(ymin = ranef_mean - ranef_sd,
+                      ymax = ranef_mean + ranef_sd)) +
+  geom_hline(yintercept = 0, color = "firebrick", linewidth = 1.5) +
+  labs(x = "Age", y = "Year ranef") +
+  theme_classic()
+
+indiv_plot <- all_years %>% 
+  filter(age <= 20) %>% 
+  left_join(indiv_ranef, by = "animalID") %>% 
+  group_by(age) %>% 
+  summarize(across(ranef, list(mean = mean, sd = sd))) %>% 
+  ggplot(aes(age, ranef_mean)) +
+  geom_pointrange(aes(ymin = ranef_mean - ranef_sd,
+                      ymax = ranef_mean + ranef_sd)) +
+  geom_hline(yintercept = 0, color = "firebrick", linewidth = 1.5) +
+  labs(x = "Age", y = "Indiv ranef") +
+  theme_classic()
+
+year_plot | indiv_plot
+
+```
+
+```{r}
+#| label: age_breeding_b4after
+
+senescence_age <- 12
+
+breeding_mod_b4 <- glmer(
+  observed2 ~ age10 + (1 | animalID) + (1 | year_fct),
+  data = filter(all_years, age <= senescence_age), 
+  family = "binomial",
+  control = glmerControl(optimizer="bobyqa")
+)
+
+breeding_mod_after <- glmer(
+  observed2 ~ age10 + (1 | animalID) + (1 | year_fct),
+  data = filter(all_years, age > senescence_age), 
+  family = "binomial",
+  control = glmerControl(optimizer="bobyqa")
+)
+
+summary(breeding_mod_b4)
+summary(breeding_mod_after)
+
+```
+
+```{r}
+#| label: fig-age_breeding
+#| fig-cap: "A: Age structure of northern elephant seals at Año Nuevo. B: Percentage of each age class that were observed breeding."
+#| fig-width: 9
+
+#Age distribution panel A
+
+age_dist <- all_years %>% 
+  filter(age < 22) %>% 
+  group_by(age) %>% 
+  count(age) %>% 
+  ggplot(aes(x = age, y = n)) + 
+  geom_col(fill = "#829BD1") + 
+  labs(x = "Age (Years)", y = "Number of individuals") + 
+  theme_few() + 
+  theme(axis.title = element_text(size = 13))
+
+#Panel B: percentage of each age class that were breeders or  nonbreeders
+
+breed_pred <- rbind(
+  ggpredict(breeding_mod_b4,
+            terms = "age10 [all]") %>% 
+    as_tibble(),
+  ggpredict(breeding_mod_after,
+            terms = "age10 [all]") %>% 
+    as_tibble()
+) %>% 
+    mutate(age = x * 10,
+           age_class = ifelse(age <= senescence_age,
+                              "Before",
+                              "After"))
+
+breed_nonbreed <- all_years %>% 
+  group_by(age) %>% 
+  summarize(n_observed = sum(!is.na(observed2)), 
+            perc_breed = sum(observed2 == "Breeder") / n_observed, 
+            se = (perc_breed * (1 - perc_breed) / n_observed)^0.5) %>% 
+  filter(age < 22) %>%
+  ggplot(aes(x = age, y = perc_breed)) + 
+  geom_ribbon(aes(age, ymin = conf.low, ymax = conf.high, group = age_class),
+              breed_pred,
+              inherit.aes = FALSE,
+              fill = "#829BD1",
+              alpha = 0.25) +
+  geom_line(aes(age, predicted, group = age_class),
+            breed_pred,
+            inherit.aes = FALSE,
+            color = "#829BD1",
+            linewidth = 1) +
+  geom_point() + 
+  geom_errorbar(aes(ymin = perc_breed - se, ymax = perc_breed + se), 
+                width = 0) +
+  scale_y_continuous(labels = scales::percent, limits = c(0.4, 1)) +
+  expand_limits(x = 23) +
+  labs(x = "Maternal Age (Years)", y = "% Breeders (Mean ± SE)") + 
+  theme_few() + 
+  theme(axis.title = element_text(size = 13))
+
+plot_grid(age_dist, breed_nonbreed, nrow = 1, labels = "AUTO")
+
+ggsave("outputs/agefig1.png", plot = , height = 4, width = 9)
+
+```
 
@@
 
-Trying segmented analysis. Note: `segmented` does NOT support mixed effect GLMs!
-
-```{r}
-#| label: age_breeding_seg
-
-breeding_mod2 <- glm(
-  observed_int ~ age,
-  data = sealdat, 
-  family = "binomial"
-)
-breed_nseg <- lapply(
-  list(15, c(9, 15)),
-  \(.psi) {
-    segmented(breeding_mod2, seg.Z = ~age, psi = .psi)
-  }
-)
-AIC(breeding_mod2, breed_nseg[[1]], breed_nseg[[2]]) %>% 
-  as_tibble() %>% 
-  mutate(model = paste(1:3, "segments")) %>% 
-  transmute(model, AIC, dAIC = AIC - min(AIC)) %>% 
-  arrange(dAIC)
-
-breeding_best <- breed_nseg[[1]]
-breeding_pred <- tibble(age = seq(4, 22, length.out = 100)) %>% 
-  predict(breeding_best, 
-          newdata = ., 
-          type = "response", 
-          interval = "confidence") %>% 
-  as_tibble() %>% 
-  set_names(c("bred", "bred_lwr", "bred_upr")) %>% 
-  mutate(age = seq(4, 22, length.out = 100))
-breed_nonbreed <- sealdat %>% 
-  group_by(age) %>% 
-  summarize(n_observed = sum(!is.na(observed)), 
-            perc_breed = sum(observed == "Breeder") / n_observed, 
-            se = (perc_breed * (1 - perc_breed) / n_observed)^0.5) %>% 
-  filter(age < 22)
-
-ggplot(breeding_pred, aes(age)) +
-  geom_ribbon(aes(ymin = bred_lwr, ymax = bred_upr),
-              fill = "grey90") + 
-  geom_line(aes(y = bred)) +
-  geom_point(aes(y = perc_breed), breed_nonbreed) + 
-  geom_errorbar(aes(ymin = perc_breed - se, ymax = perc_breed + se), 
-                breed_nonbreed,
-                width = 0) +
-  theme_classic()
-
-```
-
 ```{r}
@@
 
-age_dist <- sealdat %>% 
+age_dist <- all_years %>% 
   filter(age < 22) %>% 

### 2024-05-22 paper/paperv3.qmd
p(breed) split regression in 2

--- 2024-05-21 paper/paperv3.qmd
@@
 library(ggeffects)
+library(ggridges)
 library(patchwork)
-library(segmented)
 source(here::here("R/format.R"))
+
+age_senesce <- 12
 
@@
          year_fct = factor(year),
-         age10 = age / 10)
+         age10 = age / 10) %>% 
+  group_by(animalID) %>% 
+  mutate(longevity = ifelse(max(year) < 2020,
+                            max(age),
+                            NA),
+         longevity10 = longevity / 10) %>% 
+  ungroup()
 
@@
 
-We predicted that breeding success would decline with age. We found that breeding success increases to prime age, then decreases at older ages (@fig-age_dist).
+~~We predicted that breeding success would decline with age. We found that breeding success increases to prime age, then decreases at older ages (@fig-age_dist).~~
+
+```{r}
+#| label: breeding-mod 
+#| fig-width: 9
+#| fig-height: 6
+
+breed_age_mod1 <- glmer(
+    observed_int ~ age10 + (1 | animalID) + (1 | year_fct),
+    data = filter(sealdat, age < age_senesce),
+    family = "binomial",
+    control = glmerControl(optimizer="bobyqa")
+)
+breed_age_mod2 <- glmer(
+    observed_int ~ age10 + (1 | animalID) + (1 | year_fct),
+    data = filter(sealdat, age >= age_senesce),
+    family = "binomial",
+    control = glmerControl(optimizer="bobyqa")
+)
+
+breed_pred <- rbind(
+  ggpredict(
+    breed_age_mod1,
+    terms = c("age10 [all]"),
+    type = "fixed",
+    interval = "confidence"
+  ) %>% 
+    as_tibble() %>% 
+    mutate(age = x * 10,
+           age_class = "young"),
+  ggpredict(
+    breed_age_mod2,
+    terms = c("age10 [all]"),
+    type = "fixed",
+    interval = "confidence"
+  ) %>% 
+    as_tibble() %>% 
+    mutate(age = x * 10,
+           age_class = "old")
+) %>% 
+  filter(age <= 20)
+
+get_ranefs <- function(model, r, reduce = TRUE) {
+  result <- ranef(model)[[r]] %>% 
+    as_tibble(rownames = r) %>% 
+    rename(ranef = 2) %>% 
+    arrange(ranef)
+
+  if (reduce)
+    slice(result, round(seq(1, nrow(result), length.out = 5)))
+  else 
+    result
+}
+
+year_ranef1 <- get_ranefs(breed_age_mod1, "year_fct")
+year_ranef2 <- get_ranefs(breed_age_mod2, "year_fct")
+indiv_ranef1 <- get_ranefs(breed_age_mod1, "animalID")
+indiv_ranef2 <- get_ranefs(breed_age_mod2, "animalID")
+
+get_ranef_pred <- function(model, r, keep) {
+  ggpredict(model,
+            terms = c("age10 [all]", r), 
+                             type = "random",
+                             interval = "confidence") %>% 
+  as_tibble() %>% 
+  mutate(age = x * 10,
+         "{r}" := group) %>% 
+  semi_join(keep, by = r) %>% 
+  filter(age <= 20)
+}
+
+breed_pred_year <- rbind(
+  get_ranef_pred(breed_age_mod1, "year_fct", year_ranef1) %>% 
+    mutate(age_class = "young"),
+  get_ranef_pred(breed_age_mod2, "year_fct", year_ranef2) %>% 
+    mutate(age_class = "old")
+) %>% 
+  filter(age <= 20)
+
+breed_pred_indiv <- rbind(
+  get_ranef_pred(breed_age_mod1, "animalID", indiv_ranef1) %>% 
+    mutate(age_class = "young"),
+  get_ranef_pred(breed_age_mod2, "animalID", indiv_ranef2) %>% 
+    mutate(age_class = "old")
+) %>% 
+  filter(age <= 20)
+
+breed_nonbreed <- sealdat %>% 
+  group_by(age) %>% 
+  summarize(n_observed = n(), 
+            perc_breed = sum(observed_int) / n_observed, 
+            se = (perc_breed * (1 - perc_breed) / n_observed)^0.5) %>% 
+  filter(age <= 20)
+
+ggplot(breed_pred_year, aes(age)) +
+  # Prediction across years (fixed effect only)
+  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, 
+                  group = age_class),
+              breed_pred,
+              fill = "grey", alpha = 0.5) +
+  geom_line(aes(y = predicted, color = age_class), 
+            breed_pred,
+            linewidth = 1.5) +
+  # Predictions for specific years (random effect)
+  geom_line(aes(y = predicted, group = interaction(year_fct, age_class)),
+            linewidth = 0.75, alpha = 0.5) +
+  # Raw data
+  geom_pointrange(aes(y = perc_breed, 
+                      ymin = perc_breed - se, 
+                      ymax = perc_breed + se),
+                  breed_nonbreed) + 
+  # Year labels
+  geom_text(aes(y = predicted, label = year_fct),
+            filter(breed_pred_year, age == 20),
+            x = 20.2,
+            hjust = 0,
+            size = 3) +
+  geom_text(aes(y = predicted, label = year_fct),
+            filter(breed_pred_year, age == 4),
+            x = 3.8,
+            hjust = 1,
+            size = 3) +
+  scale_x_continuous("Age", limits = c(3.5, 20.5)) +
+  scale_y_continuous("Breeding %", 
+                     limits = c(NA, 1), 
+                     labels = scales::percent) +
+  scale_color_manual(values = c("cornflowerblue", "firebrick")) +
+  theme_classic() +
+  theme(legend.position = "none")
+
+```
+
+```{r}
+#| label: age_breeding_b4after
+
+senescence_age <- 12
+
+breeding_mod_b4 <- glmer(
+  observed2 ~ age10 + (1 | animalID) + (1 | year_fct),
+  data = filter(all_years, age <= senescence_age), 
+  family = "binomial",
+  control = glmerControl(optimizer="bobyqa")
+)
+
+breeding_mod_after <- glmer(
+  observed2 ~ age10 + (1 | animalID) + (1 | year_fct),
+  data = filter(all_years, age > senescence_age), 
+  family = "binomial",
+  control = glmerControl(optimizer="bobyqa")
+)
+
+summary(breeding_mod_b4)
+summary(breeding_mod_after)
+
+```
+
+```{r}
+#| label: fig-age_breeding
+#| fig-cap: "A: Age structure of northern elephant seals at Año Nuevo. B: Percentage of each age class that were observed breeding."
+#| fig-width: 9
+
+#Age distribution panel A
+
+age_dist <- all_years %>% 
+  filter(age < 22) %>% 
+  group_by(age) %>% 
+  count(age) %>% 
+  ggplot(aes(x = age, y = n)) + 
+  geom_col(fill = "#829BD1") + 
+  labs(x = "Age (Years)", y = "Number of individuals") + 
+  theme_few() + 
+  theme(axis.title = element_text(size = 13))
+
+#Panel B: percentage of each age class that were breeders or  nonbreeders
+
+breed_pred <- rbind(
+  ggpredict(breeding_mod_b4,
+            terms = "age10 [all]") %>% 
+    as_tibble(),
+  ggpredict(breeding_mod_after,
+            terms = "age10 [all]") %>% 
+    as_tibble()
+) %>% 
+    mutate(age = x * 10,
+           age_class = ifelse(age <= senescence_age,
+                              "Before",
+                              "After"))
+
+breed_nonbreed <- all_years %>% 
+  group_by(age) %>% 
+  summarize(n_observed = sum(!is.na(observed2)), 
+            perc_breed = sum(observed2 == "Breeder") / n_observed, 
+            se = (perc_breed * (1 - perc_breed) / n_observed)^0.5) %>% 
+  filter(age < 22) %>%
+  ggplot(aes(x = age, y = perc_breed)) + 
+  geom_ribbon(aes(age, ymin = conf.low, ymax = conf.high, group = age_class),
+              breed_pred,
+              inherit.aes = FALSE,
+              fill = "#829BD1",
+              alpha = 0.25) +
+  geom_line(aes(age, predicted, group = age_class),
+            breed_pred,
+            inherit.aes = FALSE,
+            color = "#829BD1",
+            linewidth = 1) +
+  geom_point() + 
+  geom_errorbar(aes(ymin = perc_breed - se, ymax = perc_breed + se), 
+                width = 0) +
+  scale_y_continuous(labels = scales::percent, limits = c(0.4, 1)) +
+  expand_limits(x = 23) +
+  labs(x = "Maternal Age (Years)", y = "% Breeders (Mean ± SE)") + 
+  theme_few() + 
+  theme(axis.title = element_text(size = 13))
+
+plot_grid(age_dist, breed_nonbreed, nrow = 1, labels = "AUTO")
+
+ggsave("outputs/agefig1.png", plot = , height = 4, width = 9)
+
+```
 

### 2024-05-23 paper/paperv3.qmd
clean up reproductive senescence model

--- 2024-05-21 paper/paperv3.qmd
@@
 
-First, address detection probability.
-We do resights near-daily during breeding and molting seasons, so while we may miss an individual on a given day, it is unlikely that we miss a marked individual throughout the entire breeding or molting season.
-
-```{r}
-#| label: detection-prob
-
-full <- semi_join(resight, sealdat, "animalID") %>% 
-  mutate(fiscalyear = ifelse(yday > 200, calyear + 1, calyear)) %>% 
-  group_by(animalID, fiscalyear) %>% 
-  count(fiscalyear) %>% 
-  mutate(year = as.numeric(fiscalyear)) %>% 
-  filter(year > 1987)
-
-full %>% 
-  group_by(year) %>% 
-  summarize(across(n, list(median = median, 
-                           q25 = \(x) quantile(x, 0.25), 
-                           q75 = \(x) quantile(x, 0.75)))) %>% 
-  ggplot(aes(year, n_median)) +
-  geom_ribbon(aes(ymin = n_q25, ymax = n_q75), fill = "grey80") +
-  geom_line() +
-  geom_hline(yintercept = 3, linetype = "dashed", color = "firebrick") +
-  labs(y = "Resights per year") + 
-  theme_classic()
-
-#some animals have multiple rows because they have multiple pups associated with them...
-sealdat %>% 
-  group_by(animalID, year) %>% 
-  mutate(n = n()) %>% 
-  filter(n > 2)
-
-skips <- sealdat %>% 
-  distinct(animalID, year) %>% 
-  group_by(animalID) %>% 
-  summarize(n_obs = n(),
-            obs_window = max(year) - min(year) + 1,
-            skips = obs_window - n_obs) %>% 
-  group_by(obs_window, skips) %>% 
-  summarize(ind = n(), .groups = "drop")
-
-resights_per_year <- full %>% 
-  group_by(year) %>% 
-  summarize(median_resights = median(n))
-
-skips_per_year <- skips
-
-skip_frac <- sealdat %>% 
-  distinct(animalID, year) %>% 
-  group_by(animalID) %>% 
-  summarize(firstobs = min(year),
-            lastobs = max(year)) %>% 
-  mutate(n = lastobs - firstobs + 1) %>% 
-  uncount(n) %>% 
-  group_by(animalID) %>% 
-  mutate(year = firstobs + seq(n()) - 1) %>% 
-  ungroup() %>% 
-  left_join(transmute(sealdat, animalID, year, skip = 0), 
-            by = c("animalID", "year")) %>% 
-  mutate(skip = ifelse(is.na(skip), 1, 0)) %>% 
-  group_by(year) %>% 
-  summarize(skip_frac = sum(skip) / n())
-
-resights_per_year %>% 
-  left_join(skip_frac, by = "year") %>% 
-  ggplot(aes(median_resights, skip_frac)) +
-  geom_point(aes(fill = year), shape = 21, color = "grey20", size = 2) +
-  ggrepel::geom_label_repel(aes(label = year), max.overlaps = 8) +
-  scale_y_continuous("Skip fraction", labels = scales::percent) +
-  scale_fill_distiller(palette = "PuRd") +
-  labs(x = "Median resights") +
-  theme_classic()
-
-#mark recapture model is necessary? could have bred at another colony. Many skips!!
-
-ggplot(skips, aes(obs_window, skips, size = ind)) +
-  geom_point()
-
-```
-
-#
+First, address detection probability. DO THIS IN MARK.
 
@@
 
-We predicted that breeding success would decline with age. We found that breeding success increases to prime age, then decreases at older ages (@fig-age_dist).
+~~We predicted that breeding success would decline with age. We found that breeding success increases to prime age, then decreases at older ages (@fig-age_dist).~~
 
 ```{r}
-#| label: age_breeding_mod
+#| label: breeding-mod
 
-breeding_mod <- glmer(
-  observed_int ~ poly(age10, 2, raw = TRUE) + (1 | animalID) + (1 | year_fct),
-  data = sealdat, 
+# Fit separate GLMMs to the data subsetted before and after onset of senescence
+# We a priori chose age 12 for that (age_senesce)
+# Pre-senescence model
+breed_age_mod1 <- glmer(
+    observed_int ~ age10 + (1 | animalID) + (1 | year_fct),
+    # Here's the age filter
+    data = filter(sealdat, age < age_senesce),
+    family = "binomial",
+    control = glmerControl(optimizer="bobyqa")
+)
+# Post-senescence model
+breed_age_mod2 <- glmer(
+    observed_int ~ age10 + (1 | animalID) + (1 | year_fct),
+    data = filter(sealdat, age >= age_senesce),
     family = "binomial",
@@
 
-breed_pred <- ggpredict(breeding_mod, terms = "age10 [all]") %>% 
-  as_tibble() %>% 
-  mutate(age = x * 10)
-
-ranef_year <- ranef(breeding_mod)$year_fct %>% 
-  as_tibble(rownames = "year_fct") %>% 
-  arrange(`(Intercept)`) %>% 
-  slice(1, floor(nrow(.) * c(0.25, 0.5, 0.75, 1))) %>% 
-  expand_grid(age10 = unique(sealdat$age10)) %>% 
-  mutate(year_fct = factor(year_fct, levels = levels(sealdat$year_fct))) %>% 
-  mutate(predicted = predict(breeding_mod, 
-                             newdata = ., 
-                             re.form = ~ (1 | year_fct), type = "response"),
-         age = age10 * 10)
-
-breed_mod_summary <- summary(breeding_mod)
-
-breed_mod_summary
-
-#change rightmost value to say p value 
-#add an extra sig fig to all columns
-#p values are also rounded....
-
-#table s1, look at roxannes ecology letters paper supplemental 
-#Pull out the fixed effects table with 2 sfs, can paste in excel to adjust digits 
-#intercept, age (linear), age (quadratic) as row names 
-#in figure caption say "results from x analyis. The random effect variance of animal id is X (from results) and year was X. Apply to all the other stats. 
-#analyses fit to the age in which we have larger sample sizes fit also have a quadratic fit and don't change the story
-
-```
-
-Trying segmented analysis. Note: `segmented` does NOT support mixed effect GLMs!
-
-```{r}
-#| label: age_breeding_seg
-
-breeding_mod2 <- glm(
-  observed_int ~ age,
-  data = sealdat, 
-  family = "binomial"
-)
-breed_nseg <- lapply(
-  list(15, c(9, 15)),
-  \(.psi) {
-    segmented(breeding_mod2, seg.Z = ~age, psi = .psi)
-  }
-)
-AIC(breeding_mod2, breed_nseg[[1]], breed_nseg[[2]]) %>% 
-  as_tibble() %>% 
-  mutate(model = paste(1:3, "segments")) %>% 
-  transmute(model, AIC, dAIC = AIC - min(AIC)) %>% 
-  arrange(dAIC)
-
-breeding_best <- breed_nseg[[1]]
-breeding_pred <- tibble(age = seq(4, 22, length.out = 100)) %>% 
-  predict(breeding_best, 
-          newdata = ., 
-          type = "response", 
-          interval = "confidence") %>% 
-  as_tibble() %>% 
-  set_names(c("bred", "bred_lwr", "bred_upr")) %>% 
-  mutate(age = seq(4, 22, length.out = 100))
-breed_nonbreed <- sealdat %>% 
-  group_by(age) %>% 
-  summarize(n_observed = sum(!is.na(observed)), 
-            perc_breed = sum(observed == "Breeder") / n_observed, 
-            se = (perc_breed * (1 - perc_breed) / n_observed)^0.5) %>% 
-  filter(age < 22)
-
-ggplot(breeding_pred, aes(age)) +
-  geom_ribbon(aes(ymin = bred_lwr, ymax = bred_upr),
-              fill = "grey90") + 
-  geom_line(aes(y = bred)) +
-  geom_point(aes(y = perc_breed), breed_nonbreed) + 
-  geom_errorbar(aes(ymin = perc_breed - se, ymax = perc_breed + se), 
-                breed_nonbreed,
-                width = 0) +
-  theme_classic()
+tibble(
+  model = c("Pre-senescence", "Post-senescence"),
+  age_coef = c(fixef(breed_age_mod1)["age10"],
+               fixef(breed_age_mod2)["age10"]),
+  age_coef_se = c(summary(breed_age_mod1)$coef["age10", "Std. Error"],
+                  summary(breed_age_mod2)$coef["age10", "Std. Error"]),
+  `p(coef<0)` = c(pnorm(summary(breed_age_mod1)$coef["age10", "z value"]),
+                  pnorm(summary(breed_age_mod2)$coef["age10", "z value"]))
+) %>% 
+  mutate(across(where(is.numeric), \(x) signif(x, 3)))
 
@@
 ```{r}
-#| label: fig-age_dist
-#| fig-cap: "A: Age structure of northern elephant seals at Año Nuevo. B: Percentage of each age class that were observed breeding."
+#| label: fig-breeding-mod 
 #| fig-width: 9
+#| fig-height: 6
 
-#Age distribution panel A
+# Age distribution panel A
 
@@
   ggplot(aes(x = age, y = n)) + 
-  geom_col(fill = "#829BD1") + 
+  geom_col(fill = "grey50") + 
   labs(x = "Age (Years)", y = "Number of individuals") + 
@@
 
-#Panel B: percentage of each age class that were breeders or  nonbreeders
+# Create POPULATION-LEVEL (i.e. without rand effects) predictions for both
+# models and row bind them together
+breed_pred <- rbind(
+  # Pre-senescence
+  ggpredict(
+    breed_age_mod1,
+    terms = c("age10 [all]"),
+    type = "fixed",
+    interval = "confidence"
+  ) %>% 
+    as_tibble() %>% 
+    mutate(age = x * 10,
+           # Add an age_class variable to facilitate plotting
+           age_class = "young"),
+  # Post-senescence
+  ggpredict(
+    breed_age_mod2,
+    terms = c("age10 [all]"),
+    type = "fixed",
+    interval = "confidence"
+  ) %>% 
+    as_tibble() %>% 
+    mutate(age = x * 10,
+           age_class = "old")
+) %>% 
+  filter(age <= 20)
 
+# Now create rand effect predictions
+# This utility function pulls rand effect *coefs* from a model
+get_ranefs <- function(model, r, reduce = TRUE) {
+  result <- ranef(model)[[r]] %>% 
+    as_tibble(rownames = r) %>% 
+    rename(ranef = 2) %>% 
+    arrange(ranef)
+
+  if (reduce)
+    slice(result, round(seq(1, nrow(result), length.out = 5)))
+  else 
+    result
+}
+
+# Get year and individual random effeccts from pre- and post-senescence models
+year_ranef1 <- get_ranefs(breed_age_mod1, "year_fct")
+year_ranef2 <- get_ranefs(breed_age_mod2, "year_fct")
+indiv_ranef1 <- get_ranefs(breed_age_mod1, "animalID")
+indiv_ranef2 <- get_ranefs(breed_age_mod2, "animalID")
+
+# This util fun pulls rand effect *predictions* from a model
+get_ranef_pred <- function(model, r, keep) {
+  ggpredict(model,
+            terms = c("age10 [all]", r), 
+                             type = "random",
+                             interval = "confidence") %>% 
+  as_tibble() %>% 
+  mutate(age = x * 10,
+         "{r}" := group) %>% 
+  semi_join(keep, by = r) %>% 
+  filter(age <= 20)
+}
+
+# Predictions based on 5 different *years* (min, q1, median, q3, max ranef
+# coef)
+breed_pred_year <- rbind(
+  # Pre-senescence
+  get_ranef_pred(breed_age_mod1, "year_fct", year_ranef1) %>% 
+    mutate(age_class = "young"),
+  # Post-senescence
+  get_ranef_pred(breed_age_mod2, "year_fct", year_ranef2) %>% 
+    mutate(age_class = "old")
+) %>% 
+  filter(age <= 20)
+
+# As above, but 5 different *individuals*
+breed_pred_indiv <- rbind(
+  get_ranef_pred(breed_age_mod1, "animalID", indiv_ranef1) %>% 
+    mutate(age_class = "young"),
+  get_ranef_pred(breed_age_mod2, "animalID", indiv_ranef2) %>% 
+    mutate(age_class = "old")
+) %>% 
+  filter(age <= 20)
+
+# The actual raw data
 breed_nonbreed <- sealdat %>% 
   group_by(age) %>% 
-  summarize(n_observed = sum(!is.na(observed)), 
-            perc_breed = sum(observed == "Breeder") / n_observed, 
+  summarize(n_observed = n(), 
+            perc_breed = sum(observed_int) / n_observed, 
             se = (perc_breed * (1 - perc_breed) / n_observed)^0.5) %>% 
-  filter(age < 22) %>%
-  ggplot(aes(x = age, y = perc_breed)) + 
-  geom_ribbon(aes(age, ymin = conf.low, ymax = conf.high),
-              breed_pred,
-              inherit.aes = FALSE,
-              fill = "#829BD1",
-              alpha = 0.25) +
-  geom_line(aes(age, predicted, group = year_fct),
-            ranef_year,
-            inherit.aes = FALSE,
-            alpha = 0.2) +
-  geom_text(aes(x = 22.1, predicted, label = year_fct),
-            filter(ranef_year, age == 22),
-            inherit.aes = FALSE,
+  filter(age <= 20)
+
+# Figure with raw data and model predictions
+breed_fig <- ggplot(breed_pred, aes(x = age)) +
+  # Prediction across years (fixed effect only)
+  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, 
+                  group = age_class),
+              fill = "grey", alpha = 0.5) +
+  geom_line(aes(y = predicted, color = age_class), 
+            linewidth = 1.5) +
+  # Predictions for specific years (random effect)
+  geom_line(aes(y = predicted, group = interaction(year_fct, age_class)),
+            breed_pred_year,
+            linewidth = 0.75, alpha = 0.5) +
+  # Raw data
+  geom_pointrange(aes(y = perc_breed, 
+                      ymin = perc_breed - se, 
+                      ymax = perc_breed + se),
+                  breed_nonbreed) + 
+  # Year labels
+  geom_text(aes(y = predicted, label = year_fct),
+            filter(breed_pred_year, age == 20),
+            x = 20.2,
             hjust = 0,
-            alpha = 0.5,
             size = 3) +
-  geom_line(aes(age, predicted),
-            breed_pred,
-            inherit.aes = FALSE,
-            color = "#829BD1",
-            linewidth = 1) +
-  geom_point() + 
-  geom_errorbar(aes(ymin = perc_breed - se, ymax = perc_breed + se), 
-                width = 0) +
-  scale_y_continuous(labels = scales::percent, limits = c(0.5, 1)) + 
-  expand_limits(x = 23) +
-  labs(x = "Maternal Age (Years)", y = "% Breeders (Mean ± SE)") + 
+  geom_text(aes(y = predicted, label = year_fct),
+            filter(breed_pred_year, age == 4),
+            x = 3.8,
+            hjust = 1,
+            size = 3) +
+  # Add visual indicator for senescence
+  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
+  scale_x_continuous("Age", limits = c(3.5, 20.5)) +
+  scale_y_continuous("Breeding %", 
+                     limits = c(NA, 1), 
+                     labels = scales::percent) +
+  scale_color_manual(values = c("firebrick", "cornflowerblue")) +
   theme_few() +
-  theme(axis.title = element_text(size = 13))
+  theme(legend.position = "none",
+        axis.title = element_text(size = 13))
 
-plot_grid(age_dist, breed_nonbreed, nrow = 1, labels = "AUTO")
-
-ggsave("outputs/agefig1.png", plot = , height = 4, width = 9)
+age_dist | breed_fig
 
 ```
-
-Population level prediction - the fact that the raw data falls above it comes from the fact that those age classes are correlated with years and individuals that were very successful.
 

### 2024-05-23 paper/paperv3.qmd
pup surv/recr pre-/post-senescence

@@
 
-We found that neither pup survival nor recruitment is significantly affected by maternal age (@fig-repro_age).
+We found that neither pup survival nor recruitment is significantly affected by maternal age (@fig-repro-age).
 
@@
 ```{r}
-#| label: fig-repro_age
+#| label: repro-age
+
+##NOTE: Need to take out the instances where recruitment > survival
+#what does ggpredict do with random effects? 
+#thin gray lines for the different year effects? show the interannual variation (5 that include extreme ones, just something to illustrate the variation of year)
+
+# Isolate pup survival/recruitment 
+surv_recruit_data <- sealdat %>% 
+  drop_na(pupID) %>% 
+  # Only pups born 2016 or earlier (otherwise survival biiased)
+  filter(pupyearborn <= 2016) %>% 
+  mutate(pup_survived_int = ifelse(pup_survived, 1L, 0L)) %>% 
+  # Only keep maternal age classes with 5+ births, which drops moms 19+
+  group_by(age) %>% 
+  mutate(n_pups = n()) %>% 
+  ungroup() %>%
+  filter(n_pups >= 5)
+
+# Fit pup *survival* models pre- and post-senescence
+pup_surv_model1 <- glmer(
+  pup_survived_int ~ age10 + (1 | animalID) + (1 | year),
+  data = filter(surv_recruit_data, age < age_senesce),
+  family = "binomial"
+)
+pup_surv_model2 <- glmer(
+  pup_survived_int ~ age10 + (1 | animalID) + (1 | year),
+  data = filter(surv_recruit_data, age >= age_senesce),
+  family = "binomial"
+)
+
+# Fit pup *recruitment* models pre- and post-senescence
+pup_recr_model1 <- glmer(
+  puprecruited ~ age10 + (1 | animalID) + (1 | year),
+  # For recruitment, ONLY female pups
+  data = filter(surv_recruit_data, 
+                age < age_senesce,
+                pupsex == "F"),
+  family = "binomial"
+)
+pup_recr_model2 <- glmer(
+  puprecruited ~ age10 + (1 | animalID) + (1 | year),
+  data = filter(surv_recruit_data, 
+                age >= age_senesce,
+                pupsex == "F"),
+  family = "binomial",
+  control = glmerControl(optimizer="bobyqa")
+)
+
+# Summary table
+tibble(
+  response = rep(c("Survival", "Recruitment"), each = 2),
+  n = map_dbl(
+    list(pup_surv_model1@frame, 
+         pup_surv_model2@frame, 
+         pup_recr_model1@frame, 
+         pup_recr_model2@frame),
+    nrow
+  ),
+  model = rep(c("Pre-senescence", "Post-senescence"), 2),
+  age_coef = c(fixef(pup_surv_model1)["age10"],
+               fixef(pup_surv_model2)["age10"],
+               fixef(pup_recr_model1)["age10"],
+               fixef(pup_recr_model2)["age10"]),
+  age_coef_se = c(summary(pup_surv_model1)$coef["age10", "Std. Error"],
+                  summary(pup_surv_model2)$coef["age10", "Std. Error"],
+                  summary(pup_recr_model1)$coef["age10", "Std. Error"],
+                  summary(pup_recr_model2)$coef["age10", "Std. Error"]),
+  `p(coef<0)` = c(pnorm(summary(pup_surv_model1)$coef["age10", "z value"]),
+                  pnorm(summary(pup_surv_model2)$coef["age10", "z value"]),
+                  pnorm(summary(pup_recr_model1)$coef["age10", "z value"]),
+                  pnorm(summary(pup_recr_model2)$coef["age10", "z value"]))
+) %>% 
+  mutate(across(where(is.numeric), \(x) signif(x, 3)))
+
+```
+
+```{r}
+#| label: fig-repro-age
 #| fig-cap: "Pup survival and recruitment are not significantly affected by maternal age. Points are mean values for maternal age. Error bars show SE."
@@
 
-##NOTE: Need to take out the instances where recruitment > survival
-#what does ggpredict do with random effects? 
-#thin gray lines for the different year effects? show the interannual variation (5 that include extreme ones, just something to illustrate the variation of year)
+# Pup survival 
+surv_pred <- rbind(
+  # Pre-senescence
+  ggpredict(
+    pup_surv_model1,
+    terms = c("age10 [all]"),
+    type = "fixed",
+    interval = "confidence"
+  ) %>% 
+    as_tibble() %>% 
+    mutate(age = x * 10,
+           age_class = "young"),
+  # Post-senescence
+  ggpredict(
+    pup_surv_model2,
+    terms = c("age10 [all]"),
+    type = "fixed",
+    interval = "confidence"
+  ) %>% 
+    as_tibble() %>% 
+    mutate(age = x * 10,
+           age_class = "old")
+)
 
-surv_recruit_data <- sealdat %>% 
-  drop_na(pupID) %>% 
-  filter(pupsex == "F",
-         pupyearborn < 2016) %>% 
-  mutate(pup_survived_dbl = ifelse(pup_survived, 1, 0)) %>% 
-  group_by(age) %>% 
-  mutate(n_pups = n()) %>% 
-  ungroup() %>% 
-  filter(n_pups >= 5)
-pup_surv_model <- glmer(
-  pup_survived_dbl ~ poly(age10, 2, raw = TRUE) + (1 | animalID) + (1 | year),
-  data = surv_recruit_data, 
-  family = "binomial"
-)
-summary(pup_surv_model)
-pup_surv_pred <- ggpredict(pup_surv_model, terms = c("age10 [all]")) %>% 
-  as_tibble() %>% 
-  transmute(age = x * 10, 
-            metric = "surv", 
-            mean = predicted, 
-            mean_lwr = conf.low, 
-            mean_upr = conf.high)
-pup_recruit_model <- glmer(
-  puprecruited ~ poly(age10, 2, raw = TRUE) + (1 | animalID) + (1 | year),
-  data = surv_recruit_data, 
-  family = "binomial"
-)
-summary(pup_recruit_model)
-pup_recruit_pred <- ggpredict(pup_recruit_model, terms = c("age10 [all]")) %>% 
-  as_tibble() %>% 
-  transmute(age = x * 10, 
-            metric = "recruit", 
-            mean = predicted, 
-            mean_lwr = conf.low, 
-            mean_upr = conf.high)
-surv_recruit_pred <- rbind(pup_surv_pred, pup_recruit_pred) %>% 
-  mutate(metric = factor(metric, 
-                         levels = c("surv", "recruit"), 
-                         labels = c("Survival", "Recruitment")))
+repro_raw <- surv_recruit_data %>% 
+  mutate(pupsex = factor(pupsex, labels = c("Female", "Male"))) %>% 
+  group_by(age, pupsex) %>% 
+  summarize(n_born = n(), 
+            n_surv = sum(pup_survived_int),
+            perc_surv = n_surv / n_born, 
+            se_surv = (perc_surv * (1 - perc_surv) / n_born)^0.5,
+            n_recr = sum(puprecruited),
+            perc_recr = n_recr / n_born,
+            se_recr = (perc_recr * (1 - perc_recr) / n_born)^0.5,
+            .groups = "drop")
 
-surv_recruit_summary <- surv_recruit_data %>% 
-  group_by(age) %>%
-  summarize(n_pups = n(),
-            n_surv = sum(pup_survived),
-            mean_surv = n_surv / n_pups,  
-            se_surv = (mean_surv * (1 - mean_surv) / n_pups)^0.5,
-            n_recruit = sum(puprecruited), 
-            mean_recruit = n_recruit / n_pups, 
-            se_recruit = (mean_recruit * (1 - mean_recruit) / n_pups)^0.5)  %>%  
-  pivot_longer(cols = c(mean_surv, se_surv, mean_recruit, se_recruit), 
-               names_to = c("stat", "metric"), 
-               names_pattern = "(.*)_(.*)") %>% 
-  pivot_wider(names_from = "stat", 
-              values_from = "value") %>% 
-  mutate(low = mean - se, 
-         high = mean + se, 
-         age = age + ifelse(metric == "surv", -0.1, 0.1), 
-         metric = factor(metric, 
-                         levels = c("surv", "recruit"), 
-                         labels = c("Survival", "Recruitment")))
-
-surv_recruit_summary %>% 
-  ggplot(aes(x = age, y = mean, color = metric, fill = metric)) + 
-  geom_ribbon(aes(ymin = mean_lwr, ymax = mean_upr, alpha = metric),
-              surv_recruit_pred,
-              color = NA) +
-  geom_pointrange(aes(ymin = low, ymax = high)) +
-  geom_line(aes(linetype = metric),
-            surv_recruit_pred) +
-  scale_y_continuous(labels = scales::percent) + 
-  scale_color_manual(values = c("#829BD1", "#E9B668")) + 
-  scale_fill_manual(values = c("#829BD1", "#E9B668")) + 
-  scale_linetype_manual(values = c("dashed", "dashed")) +
-  scale_alpha_manual(values = c(0.25, 0.1)) +
-  expand_limits(y = c(0, 0.5)) +
-  labs(x = "Maternal Age (Years)", y = "Pup Survival or Recruitment") +
+ggplot(surv_pred, aes(x = age)) +
+  # Survival by age (fixed effect only)
+  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, 
+                  group = age_class),
+              fill = "grey", alpha = 0.5) +
+  geom_line(aes(y = predicted, color = age_class), 
+            linewidth = 1.5) +
+  # Raw survival and recruitment data
+  geom_pointrange(aes(y = perc_surv,
+                      ymin = perc_surv - se_surv,
+                      ymax = perc_surv + se_surv),
+                  repro_raw) +
+  geom_pointrange(aes(y = perc_recr,
+                      ymin = perc_recr - se_recr,
+                      ymax = perc_recr + se_recr),
+                  shape = 21,
+                  filter(repro_raw, pupsex == "Female")) +
+  # Add visual indicator for senescence
+  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
+  scale_x_continuous("Age", limits = c(3.5, 18.5)) +
+  scale_y_continuous("Pup survival/recruitment", 
+                     limits = c(0, 0.65), 
+                     labels = scales::percent) +
+  scale_color_manual(values = c("firebrick", "cornflowerblue")) +
+  facet_wrap(~pupsex) +
   theme_few() +
-  theme(legend.position = c(0.1,0.97), 
-        legend.justification = c(0,1), 
-        legend.title = element_blank(), 
-        legend.direction = "horizontal", 
+  theme(legend.position = "none",
         axis.title = element_text(size = 13))
 
-#Why is recruitment higher than survival in some points? Add the quadratic model? Reverse order of legend in the figure
-#caption - error bars show SE. Points are mean values for maternal age. Make sure that language in the caption matches both the model and the data 
 ```

### 2024-05-23 paper/paperv3.qmd
add coef plot for breeding

@@
 
+### Raw data
+
+Resighting plot (image) and age structure bar plot
+
+```{r}
+#| label: fig-raw-data
+
+age_dist <- sealdat %>% 
+  filter(age < 22) %>% 
+  group_by(age) %>% 
+  count(age) %>% 
+  ggplot(aes(x = age, y = n)) + 
+  geom_col(fill = "grey50") + 
+  labs(x = "Age (Years)", y = "Number of individuals") + 
+  theme_few() + 
+  theme(axis.title = element_text(size = 13))
+
+age_dist
+
+```
+
 ### H1: Age distribution and breeding status
@@
 
-tibble(
+breed_summ <- tibble(
   model = c("Pre-senescence", "Post-senescence"),
@@
                   pnorm(summary(breed_age_mod2)$coef["age10", "z value"]))
-) %>% 
-  mutate(across(where(is.numeric), \(x) signif(x, 3)))
+)
+
+breed_summ %>% 
+  mutate(across(where(is.numeric), \(x) signif(x, 3))) %>%
+  set_names(c("Age class", 
+              "Age", 
+              "se(Age)",
+              "p(Age) < 0")) %>% 
+  knitr::kable()
 
@@
 #| fig-height: 6
-
-# Age distribution panel A
-
-age_dist <- sealdat %>% 
-  filter(age < 22) %>% 
-  group_by(age) %>% 
-  count(age) %>% 
-  ggplot(aes(x = age, y = n)) + 
-  geom_col(fill = "grey50") + 
-  labs(x = "Age (Years)", y = "Number of individuals") + 
-  theme_few() + 
-  theme(axis.title = element_text(size = 13))
 
@@
 
-age_dist | breed_fig
+# Coefficient panel
+breed_coef_fig <- breed_summ %>% 
+  mutate(model = fct_rev(model)) %>% 
+  ggplot(aes(color = model)) + 
+  geom_pointrange(aes(ymin = age_coef - age_coef_se,
+                      y = age_coef,
+                      ymax = age_coef + age_coef_se,
+                      x = model)) +
+  geom_hline(yintercept = 0, linetype = "dashed") +
+  scale_color_manual(values = c("cornflowerblue", "firebrick")) +
+  labs(y = "Age coefficient") +
+  theme_few() +
+  theme(legend.position = "none",
+        axis.title.x = element_blank())
+
+(breed_fig | breed_coef_fig) + plot_layout(widths = c(3, 1))
 
@@
 
-```{r}
+```{r, code = FALSE}
 #| label: tbl-breed-cap
@@
 
-```{r}
+```{r, code = FALSE}
 #| label: tbl-breed-success
-#| tbl-cap: !expr "tbl_breed_cap"
+tbl-cap: !expr "tbl_breed_cap"
 
@@
 
-```{r}
+```{r, code = FALSE}
 #| label: tbl-survrecruit-cap
@@
 
-```{r}
-#| tbl-cap: !expr "surv_recruit_cap"
+```{r, code = FALSE}
 #| label: tbl-surv_recruit
+tbl-cap: !expr "surv_recruit_cap"
 
@@
 
-```{r}
+```{r, code = FALSE}
 #| label: tbl-pupsex-cap
@@
 
-```{r}
+```{r, code = FALSE}
 #| label: tbl-pupsex
-#| tbl-cap: !expr "pupsex_cap"
+tbl-cap: !expr "pupsex_cap"
 
@@
 
-```{r}
+```{r, code = FALSE}
 #| label: tbl-pup-sex
@@
 
-```{r}
+```{r, code = FALSE}
 #| fig-width: 8
@@
 
-```{r}
+```{r, code = FALSE}
 ba_data <- phenology_models$model_data[[1]]

### 2024-05-23 paper/paperv3.qmd
raw data and surv/recr figs

@@
     echo: false
-    fig-width: 4.5
-    fig-height: 4.5
+    fig-width: 5
+    fig-height: 5
     code-fold: true
@@
 source(here::here("R/format.R"))
+theme_set(theme_few(base_size = 13))
 
@@
 
+sighting_img <- sealdat %>% 
+  group_by(animalID) %>% 
+  summarize(yearborn = yearborn[1],
+            first_obs = min(year),
+            last_obs = max(year),
+            .groups = "keep") %>% 
+  reframe(yearborn = yearborn,
+          year = yearborn:last_obs,
+          age = year - yearborn,
+          prerecruit = year < first_obs) %>% 
+  filter(year > 1987) %>% 
+  left_join(sealdat %>% 
+              group_by(animalID, year) %>% 
+              summarize(observed_int = max(observed_int),
+                        .groups = "drop"),
+            by = c("animalID", "year")) %>% 
+  mutate(
+    observed = case_when(
+      age == 0 ~ "Born",
+      prerecruit ~ "Pre-recruit",
+      observed_int == 0 ~ "Non-breeder",
+      observed_int == 1 ~ "Breeder",
+      is.na(observed_int) ~ "Unobserved"
+    ),
+    observed = factor(observed, levels = c("Born",
+                                           "Pre-recruit",
+                                           "Non-breeder",
+                                           "Breeder",
+                                           "Unobserved"))
+  ) %>% 
+  group_by(animalID) %>% 
+  mutate(longevity = max(age),
+         yearborn = na.omit(yearborn)[1]) %>% 
+  ungroup() %>% 
+  arrange(yearborn, longevity, animalID) %>% 
+  mutate(animalID = factor(animalID, levels = unique(animalID))) %>% 
+  ggplot(aes(year, animalID)) +
+  geom_raster(aes(fill = observed)) +
+  scale_fill_viridis_d(direction = -1) +
+  guides(fill = guide_legend(position = "inside", 
+                             direction = "horizontal",
+                             ncol = 1,
+                             byrow = TRUE)) +
+  labs(y ="Individual") +
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.x = element_blank(),
+        legend.title = element_blank(),
+        legend.position.inside = c(0.01, 0.99),
+        legend.justification = c(0, 1),
+        legend.background = element_blank(),
+        legend.key.size = unit(4, "mm"),
+        legend.text = element_text(size = 8))
+
 age_dist <- sealdat %>% 
@@
   labs(x = "Age (Years)", y = "Number of individuals") + 
-  theme_few() + 
-  theme(axis.title = element_text(size = 13))
+  theme()
 
-age_dist
+(sighting_img | age_dist) + plot_layout(widths = c(2, 1))
 
@@
 #| label: fig-breeding-mod 
-#| fig-width: 9
-#| fig-height: 6
+#| fig-width: 7.5
 
@@
             hjust = 0,
-            size = 3) +
+            size = 2.5) +
   geom_text(aes(y = predicted, label = year_fct),
@@
             hjust = 1,
-            size = 3) +
+            size = 2.5) +
   # Add visual indicator for senescence
@@
   scale_x_continuous("Age", limits = c(3.5, 20.5)) +
-  scale_y_continuous("Breeding %", 
+  scale_y_continuous("Adult female breeding", 
                      limits = c(NA, 1), 
@@
 breed_coef_fig <- breed_summ %>% 
-  mutate(model = fct_rev(model)) %>% 
+  mutate(model = str_extract(model, ".*-"),
+         model = fct_rev(model)) %>% 
   ggplot(aes(color = model)) + 
@@
   scale_color_manual(values = c("cornflowerblue", "firebrick")) +
-  labs(y = "Age coefficient") +
+  labs(x = "Senescence",
+       y = "Age coefficient") +
   theme_few() +
-  theme(legend.position = "none",
-        axis.title.x = element_blank())
+  theme(legend.position = "none")
 
-(breed_fig | breed_coef_fig) + plot_layout(widths = c(3, 1))
+(breed_fig | breed_coef_fig) + 
+  plot_layout(widths = c(3, 1)) +
+  plot_annotation(tag_levels = "A")
 
@@
 # Fit pup *survival* models pre- and post-senescence
+# Commented-out formulae vary age10 coef by pupsex, end up with similar
+# results. Can add to supplemental?
 pup_surv_model1 <- glmer(
+  # pup_survived_int ~ age10 : pupsex + (1 | animalID) + (1 | year),
   pup_survived_int ~ age10 + (1 | animalID) + (1 | year),
@@
 pup_surv_model2 <- glmer(
+  # pup_survived_int ~ age10 : pupsex + (1 | animalID) + (1 | year),
   pup_survived_int ~ age10 + (1 | animalID) + (1 | year),
@@
 # Summary table
-tibble(
+surv_recr_summ <- tibble(
   response = rep(c("Survival", "Recruitment"), each = 2),
@@
                   pnorm(summary(pup_recr_model2)$coef["age10", "z value"]))
-) %>% 
-  mutate(across(where(is.numeric), \(x) signif(x, 3)))
+)
+
+surv_recr_summ %>% 
+  mutate(across(where(is.numeric), \(x) signif(x, 3))) %>% 
+  knitr::kable()
 
@@
 #| label: fig-repro-age
-#| fig-cap: "Pup survival and recruitment are not significantly affected by maternal age. Points are mean values for maternal age. Error bars show SE."
+#| fig-cap: "Offspring survival decreases as mothers senesce.
+**A** Survival of female and male offspring (filled points) is not negatively impacted by maternal age (blue line) prior to age 12 (vertical dashed line).
+After age 12, offspring survival declines with maternal age (red line), i.e., maternal effect senescence.
+Hollow points represent female offspring recruitment, which did not significantly decline with maternal age after age 12.
+**B** The coefficient of age w.r.t.
+offspring survival (solid points) was not significantly negative before age 12 (blue), but was significantly negative after (red).
+The coefficient of age w.r.t.
+female offspring recruitment (hollow points) was not significant in either maternal age class.
+Lines represent standard error of the coefficient, which is ommitted for the post-age 12 offspring recruitment coefficent (hollow red point) because the standard error was too large (14.8) to visually represent without hiding other patterns.
+Numbers at top indicate sample size in each model."
 #| fig-width: 5.5
@@
 repro_raw <- surv_recruit_data %>% 
-  mutate(pupsex = factor(pupsex, labels = c("Female", "Male"))) %>% 
-  group_by(age, pupsex) %>% 
+  # Uncomment to group by female and male offspring
+  # mutate(
+  #   pupsex = factor(pupsex, 
+  #                   labels = c("Female offspring", "Male offspring"))
+  # ) %>% 
+  # group_by(age, pupsex) %>% 
+  group_by(age) %>% 
   summarize(n_born = n(), 
@@
             se_surv = (perc_surv * (1 - perc_surv) / n_born)^0.5,
-            n_recr = sum(puprecruited),
-            perc_recr = n_recr / n_born,
-            se_recr = (perc_recr * (1 - perc_recr) / n_born)^0.5,
+            f_born = sum(pupsex == "F"),
+            f_recr = sum(puprecruited & pupsex == "F"),
+            perc_recr = f_recr / f_born,
+            se_recr = (perc_recr * (1 - perc_recr) / f_born)^0.5,
             .groups = "drop")
 
-ggplot(surv_pred, aes(x = age)) +
+surv_recr_fig <- ggplot(surv_pred, aes(x = age)) +
   # Survival by age (fixed effect only)
@@
                       ymax = perc_recr + se_recr),
-                  shape = 21,
-                  filter(repro_raw, pupsex == "Female")) +
+                  repro_raw,
+                  shape = 21) +
   # Add visual indicator for senescence
@@
   scale_x_continuous("Age", limits = c(3.5, 18.5)) +
-  scale_y_continuous("Pup survival/recruitment", 
+  scale_y_continuous("Offspring survival/recruitment", 
                      limits = c(0, 0.65), 
@@
   scale_color_manual(values = c("firebrick", "cornflowerblue")) +
-  facet_wrap(~pupsex) +
+  # Uncomment to facet by offspring sex
+  # facet_wrap(~pupsex) +
   theme_few() +
@@
         axis.title = element_text(size = 13))
+
+# Survival/recruitment figure. In the caption, mention the post-senescence
+# recruitment std err was omitted because it was 14.8, so it swamps all the
+# other std errs
+surv_recr_coef_fig <- surv_recr_summ %>% 
+  mutate(response = fct_rev(str_sub(response, 1, 4)),
+         model = fct_rev(str_extract(model, ".*-")),
+         # Remove post-senescence recruitment std err
+         age_coef_se = c(age_coef_se[1:3], 0)) %>% 
+  ggplot(aes(x = response, ymax = 1.1)) +
+  geom_pointrange(aes(ymin = age_coef - age_coef_se, 
+                      y = age_coef,
+                      ymax = age_coef + age_coef_se,
+                      color = model,
+                      shape = response),
+                  position = position_dodge(width = 0.9)) +
+  geom_text(aes(label = n, group = model), 
+            y = 0.9, 
+            size = 3,
+            angle = 30,
+            hjust = 0,
+            position = position_dodge(width = 1)) +
+  geom_hline(yintercept = 0, linetype = "dashed") +
+  scale_color_manual(values = c("cornflowerblue", "firebrick")) +
+  scale_shape_manual(values = c(16, 1)) +
+  labs(x = "Response", y = "Age coefficient") +
+  expand_limits(y = 1.1) +
+  theme(legend.position = "none")
+
+(surv_recr_fig | surv_recr_coef_fig) + 
+  plot_layout(widths = c(3, 1)) +
+  plot_annotation(tag_levels = "A")
 

### 2024-05-24 paper/paperv3.qmd
sex ratio revisions

@@
 ```{r}
-#| label: fig-sex_ratio
-#| fig-cap: !expr "str_glue('Offspring sex ratio skews towards males with increasing maternal age (p = {format_pval(summary(pup_sex_mod)$coefficients[\"age\", \"Pr(>|z|)\"] / 2, 3)}).
-Mothers aged 12, 13, and 17 gave birth to significantly more male than female offspring (red stars).')"
-#| fig-width: 4.5
+#| label: sex-ratio
 
@@
 
-pup_sex_summary <- pup_sex_data %>% 
-  group_by(age, pupsex) %>%
-  summarize(n = n(), .groups = "drop") %>% 
-  pivot_wider(values_from = n, names_from = pupsex) %>% 
-  mutate(num = M + F,
-         percm = M / num,
-         se = (percm * (1 - percm) / num)^0.5,
-         pval = map2_dbl(M, num, \(x, n) binom.test(x, n, alternative = "greater")$p.value))
-
-signif_male <- filter(pup_sex_summary, pval < 0.05)
-
-#linear
-pup_sex_mod <- glmer(is_male ~ age + (1 | animalID) + (1 | year),
-                     data = pup_sex_data, 
-                     family = "binomial")
-summary(pup_sex_mod)
-
-# We fit a quadratic equation, but only the linear term was significant, so the linear model is the one that is plotted.
-
-pup_sex_quad <- glmer(
-  is_male ~ poly(age10, 2, raw = TRUE) + (1 | animalID) + (1 | year),
-  data = pup_sex_data, 
+sex_ratio_mod1 <- glmer(
+  is_male ~ age10 + (1 | animalID) + (1 | year),
+  filter(pup_sex_data, age < age_senesce),
+  family = "binomial"
+)
+sex_ratio_mod2 <- glmer(
+  is_male ~ age10 + (1 | animalID) + (1 | year),
+  filter(pup_sex_data, age >= age_senesce),
   family = "binomial"
@@
 
-pup_sex_pred <- ggpredict(pup_sex_mod, terms = c("age [all]")) %>% 
+sex_ratio_summ <- tibble(
+  model = c("Pre-senescence", "Post-senescence"),
+  age_coef = c(fixef(sex_ratio_mod1)["age10"],
+               fixef(sex_ratio_mod2)["age10"]),
+  age_coef_se = c(summary(sex_ratio_mod1)$coef["age10", "Std. Error"],
+                  summary(sex_ratio_mod2)$coef["age10", "Std. Error"]),
+  `p(coef!=0)` = c(summary(sex_ratio_mod1)$coef["age10", "Pr(>|z|)"],
+                   summary(sex_ratio_mod2)$coef["age10", "Pr(>|z|)"])
+)
+
+sex_ratio_summ %>% 
+  mutate(across(where(is.numeric), \(x) signif(x, 3))) %>%
+  set_names(c("Age class", 
+              "Age", 
+              "se(Age)",
+              "p(Age) != 0")) %>% 
+  knitr::kable()
+
+```
+
+```{r}
+#| label: fig-sex-ratio
+#| fig-cap: "The effect of age on offspring sex ratio was not significantly different for pre- and post-senescence adults.
+**A** Raw data (points with error bars, indicating mean and standard error of the proportion) and model predictions (lines with 95% confidence intervals).
+**B** Estimate and standard error of age coefficients in pre- and post-senescence models."
+#| fig-width: 5.5
+
+# Model predictions
+sex_ratio_pred <- rbind(
+  # Pre-senescence
+  ggpredict(
+    sex_ratio_mod1,
+    terms = c("age10 [all]"),
+    type = "fixed",
+    interval = "confidence"
+  ) %>% 
     as_tibble() %>% 
-  transmute(age = x, 
-            percm = predicted, 
-            percm_lwr = conf.low, 
-            percm_upr = conf.high)
+    mutate(age = x * 10,
+           # Add an age_class variable to facilitate plotting
+           age_class = "young"),
+  # Post-senescence
+  ggpredict(
+    sex_ratio_mod2,
+    terms = c("age10 [all]"),
+    type = "fixed",
+    interval = "confidence"
+  ) %>% 
+    as_tibble() %>% 
+    mutate(age = x * 10,
+           age_class = "old")
+)
 
-ggplot(pup_sex_summary, aes(x = age, y = percm)) +
-  geom_hline(yintercept = 0.5, color = "gray75", linetype = "dashed") +
-  geom_ribbon(aes(ymin = percm_lwr, ymax = percm_upr), 
-              pup_sex_pred,
-              fill = "#829BD1",
-              alpha = 0.25) + 
-  geom_line(data = pup_sex_pred, 
-            color = "#829BD1", 
-            linewidth = 1) + 
-  geom_pointrange(aes(ymin = percm - se, ymax = percm + se), 
-                  size = 0.25) +
-  geom_point(aes(y = percm + se + 0.05),
-             signif_male,
-             shape = 8,
-             color = "firebrick") +
-  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
-  labs(x = "Maternal Age (Years)", y = "% Male Pups (Mean ± SE)") +
+# Sex ratio raw data
+sex_ratio <- pup_sex_data %>% 
+  group_by(age) %>% 
+  summarize(n_observed = n(), 
+            perc_male = sum(is_male) / n_observed, 
+            se = (perc_male * (1 - perc_male) / n_observed)^0.5)
+
+# Figure with raw data and model predictions
+sex_ratio_fig <- ggplot(sex_ratio_pred, aes(x = age)) +
+  # Prediction across years (fixed effect only)
+  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, 
+                  group = age_class),
+              fill = "grey", alpha = 0.5) +
+  geom_line(aes(y = predicted, color = age_class), 
+            linewidth = 1.5) +
+  # Raw data
+  geom_pointrange(aes(y = perc_male, 
+                      ymin = perc_male - se, 
+                      ymax = perc_male + se),
+                  sex_ratio) + 
+  # Add visual indicator for senescence
+  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
+  scale_y_continuous("Male offspring", 
+                     labels = scales::percent) +
+  scale_color_manual(values = c("firebrick", "cornflowerblue")) +
+  labs(x = "Age") +
   theme_few() +
-  theme(axis.title = element_text(size = 13))
+  theme(legend.position = "none",
+        axis.title = element_text(size = 13))
 
-#take the p = 0.517 and divide by 2 because a priori. report in supplemental table that its reporting on 1 tailed (no absolute value of z, just z)
+# Coefficient panel
+sex_ratio_coef_fig <- sex_ratio_summ %>% 
+  mutate(model = str_extract(model, ".*-"),
+         model = fct_rev(model)) %>% 
+  ggplot(aes(color = model)) + 
+  geom_pointrange(aes(ymin = age_coef - age_coef_se,
+                      y = age_coef,
+                      ymax = age_coef + age_coef_se,
+                      x = model)) +
+  geom_hline(yintercept = 0, linetype = "dashed") +
+  scale_color_manual(values = c("cornflowerblue", "firebrick")) +
+  labs(x = "Senescence",
+       y = "Age coefficient") +
+  theme_few() +
+  theme(legend.position = "none")
+
+(sex_ratio_fig | sex_ratio_coef_fig) + 
+  plot_layout(widths = c(3, 1)) +
+  plot_annotation(tag_levels = "A")
 ```
@@
 
-### H4: Changes to annual cycle timing with increased maternal age
-
-We predicted that seals would shift the timing of their annual cycle.
-
-We found that as maternal age increases, there are slight adjustments to the timing of life history events.
-These events (breeding/molting arrival/departure) are typically shifted later for older animals (@fig-phenology).
+There appears to be a step change biasing offspring sex ratio towards males post-senescence.
+We simulated the sex ratio of 1,000,000 offspring born to mothers pre- and post-senesecence using the coefficient estimates and standard errors of the fitted sex ratio models.
+The mean offspring sex ratio increases from 51.3% male before maternal age 12 to 61.2% male after maternal age 12.
+However, the uncertainty in post-senescence sex ratio is considerable, and only 59.3% of the contrasts of simulated post- and pre-senescence sex ratios are biased towards male offspring.
+I.e., a simulated post-senescent adult female was more likely to have a male offspring than a simulated pre-senescent adult female in 59.3% of simulations.
 
 ```{r}
+#| label: fig-sim-sex-ratio
+#| fig-cap: "Simulated sex ratios of offspring born to pre- and post-senescence adults.
+**A** Comparison of simulated sex ratios by pre- (blue) and post- (red) senescence adults.
+The mean (vertical lines) sex ratio of post-senescence adults is greater than pre-senescence adults.
+However the distributions overlap extensively.
+**B** The contrast of post-senescence sex ratio relative to pre-senescence is greater than 0 in only 59.3% of simulations."
+#| fig-width: 5.5
 
-seal_phenology <- sealdat %>% 
-  filter(observed == "Breeder") %>% 
-  mutate(BA = ifelse(firstobsbreeddoy > 200, firstobsbreeddoy - 365, firstobsbreeddoy), 
-         BD = BA + breeddur, 
-         MA = lastobsmoltdoy - moltdur, 
-         MD = lastobsmoltdoy) %>% 
-  drop_na(age, BA, BD, MA, MD)
+# pre_sex_ratio <- tibble(age10 = seq(4, age_senesce - 1)) %>% 
+inv_logit <- \(x) exp(x) / (1 + exp(x))
+n_sim <- 1e6
+sim_sex_ratio <- rbind(
+  tibble(intercept_mean = summary(sex_ratio_mod1)$coef[1, 1],
+         intercept_sd = summary(sex_ratio_mod1)$coef[1, 2],
+         slope_mean = summary(sex_ratio_mod1)$coef[2, 1],
+         slope_sd = summary(sex_ratio_mod1)$coef[2, 2],
+         age10 = sample(seq(4, age_senesce - 1) / 10, n_sim, TRUE),
+         intercept = rnorm(n_sim, intercept_mean, intercept_sd),
+         slope = rnorm(n_sim, slope_mean, slope_sd),
+         logit_p = intercept + slope * age10,
+         p = inv_logit(logit_p),
+         age_class = "Pre-senescence"),
+  tibble(intercept_mean = summary(sex_ratio_mod2)$coef[1, 1],
+         intercept_sd = summary(sex_ratio_mod2)$coef[1, 2],
+         slope_mean = summary(sex_ratio_mod2)$coef[2, 1],
+         slope_sd = summary(sex_ratio_mod2)$coef[2, 2],
+         age10 = sample(seq(age_senesce, 18) / 10, n_sim, TRUE),
+         intercept = rnorm(n_sim, intercept_mean, intercept_sd),
+         slope = rnorm(n_sim, slope_mean, slope_sd),
+         logit_p = intercept + slope * age10,
+         p = inv_logit(logit_p),
+         age_class = "Post-senescence")
+) %>% 
+  mutate(age_class = fct_rev(age_class))
 
-mean_phenology <- seal_phenology %>% 
-  group_by(age) %>% 
-  summarize(across(BA:MD, mean)) %>% 
-  summarize(across(BA:MD, mean)) %>% 
-  pivot_longer(everything(),
-               names_to = "event", 
-               values_to = "doy")
+pre_logit_p <- with(sim_sex_ratio, logit_p[age_class == "Pre-senescence"])
+post_logit_p <- with(sim_sex_ratio, logit_p[age_class == "Post-senescence"])
+sex_ratio_contrast <- tibble(
+  contrast_logit_p = post_logit_p - pre_logit_p
+)
 
-phenology_n <- unique(seal_phenology$animalID)
+sim_sex_ratio_fig <- ggplot(sim_sex_ratio, aes(p)) +
+  geom_density(aes(fill = age_class), alpha = 0.5) +
+  geom_density(aes(color = age_class)) +
+  geom_vline(aes(xintercept = p, color = age_class),
+             sim_sex_ratio %>% 
+               group_by(age_class) %>% 
+               summarize(p = inv_logit(mean(logit_p))),
+             linewidth = 1.5) +
+  scale_x_continuous("Male offspring", 
+                     labels = scales::percent) +
+  scale_color_manual(values = c("cornflowerblue", "firebrick")) +
+  scale_fill_manual(values = c("cornflowerblue", "firebrick")) +
+  theme(legend.position = "none")
 
-timing_summary <- seal_phenology %>% 
-  group_by(age) %>% 
-  summarize(across(BA:MD, mean), 
-            .groups = "drop") %>% 
-  pivot_longer(cols = -age,  
-               names_to = "event", 
-               values_to = "doy") %>% 
-  group_by(age) %>% 
-  group_modify(\(data, keys) {
-    tibble(event = c("Long trip", "Breeding", "Short trip", "Molting", "longtrip2"),
-           doy = c(data$doy, 365) - c(0, data$doy))
-  }) %>% 
-  ungroup() %>% 
-  mutate(event = factor(event, levels = c("Long trip", 
-                                          "Breeding", 
-                                          "Short trip", 
-                                          "Molting", 
-                                          "longtrip2")))
+sim_sr_contrast_fig <- ggplot(sex_ratio_contrast, aes(contrast_logit_p)) +
+  geom_density(fill = "grey90") +
+  geom_vline(xintercept = 0, linetype = "dashed") +
+  labs(x = "Contrast of logit(p)")
+
+(sim_sex_ratio_fig | sim_sr_contrast_fig) +
+  plot_annotation(tag_levels = "A")
 
 ```
-
-```{r}
-#| label: fig-phenology
-#| fig-height: 4.5
-#| fig-width: 7
-#| fig-cap: "Successful breeding and increased maternal age both impact elephant seal annual cycles.
-A: Visualization of life history timing across ages, faceted by breeding status.
-White dashed lines on each panel represent the raw data from the other panel to facilitate comparisons between breeders and non-breeders.
-B: LMM coefficients for phenological parameters.
-The units for maternal age coefficients are in days per year.
-The coefficients for breeding status are in days.
-Black points are significant, gray points are not significant."
-
-#Delete the non-breeders and add the dashed line as an average instead. 
-#put panels next to each other 
-#what do you do when breeding dates are in december? ifelse statement (if > 250), 200 - 365 (so use negative numbers if they are in december) **does not happen! no worries!!**
-#show xy plots of each phenological variable against age in the supplemental and summary stats for the model in the supplement too 
-#draw attention to the widths - 1) 2) 3) 
-#assess linearity from plots, but double check visually and see if quadratic might fit better (we werent sure if age patterns would be linear or not, so we tested quadratic analyses throughout)
-
-phen_plot1 <- ggplot(timing_summary, aes(x = age, y = doy, fill = event)) + 
-  geom_area(position = position_stack(reverse = TRUE),
-            color = NA) +
-  geom_line(position = position_stack(reverse = TRUE),
-            color = "white") +
-  geom_hline(aes(yintercept = doy),
-             mean_phenology,
-             color = "white",
-             linetype = "dashed",
-             alpha = 0.5) +
-  scale_fill_discrete(type = c("#829BD1", "#E9B668", "#7EC6CC", "#E091B9", "#829BD1"),
-                      breaks = c("Long trip", "Breeding", "Short trip", "Molting")) + 
-  scale_x_continuous("Maternal Age (Years)", expand = expansion(mult = 0)) +
-  scale_y_continuous("Day of Year", expand = expansion(mult = 0)) +
-  theme_few() + 
-  theme(axis.title = element_text(size = 12),
-        panel.spacing = unit(1, "lines"))
-
-# Utility functions for model fitting and coefficient extraction
-fit_model <- function(dat) {
-  lmer(doy ~ age + (1 | animalID) + (1 | year), data = dat)
-}
-extract_estimate <- function(models, param) {
-  map_dbl(models, \(mod) summary(mod)$coefficients[param, "Estimate"])
-}
-extract_stderr <- function(models, param) {
-  map_dbl(models, \(mod) summary(mod)$coefficients[param, "Std. Error"])
-}
-extract_pval <- function(models, param) {
-  map_dbl(models, \(mod) summary(mod)$coefficients[param, "Pr(>|t|)"])
-}
-
-# This data frame contains a model for each phenological variable w.r.t. age
-phenology_models <- seal_phenology %>% 
-  pivot_longer(BA:MD, names_to = "phenology", values_to = "doy") %>% 
-  select(phenology, doy, age, animalID, year) %>% 
-  nest(model_data = c(doy, age, animalID, year)) %>% 
-  mutate(model = map(model_data, fit_model),
-         age_coef = extract_estimate(model, "age"),
-         age_stderr = extract_stderr(model, "age"),
-         age_pval = extract_pval(model, "age"),
-         age_signif = age_pval < 0.05)
-
-phen_wrapped <- tibble(
-  event_begin = c(-Inf, 1:4),
-  event_end = c(1:4, Inf),
-  event = timing_summary$event[0:5]
-)
-
-phen_plot2 <- phenology_models %>% 
-  select(-c(model_data, model)) %>% 
-  pivot_longer(c(ends_with("coef"), ends_with("stderr"), ends_with("signif")),
-               names_to = c("param", "metric"),
-               names_sep = "_",
-               values_to = "value") %>% 
-  pivot_wider(names_from = metric, values_from = value) %>% 
-  mutate(signif = as.logical(signif), 
-         # param = factor(param, 
-         #                labels = "'Delay with maternal age (days year'^-1 * ')'"), 
-         phenology = factor(phenology, 
-                            labels = c("Breeding\narrrival", 
-                                       "Breeding\ndeparture", 
-                                       "Molt\narrrival", 
-                                       "Molt\ndeparture"))) %>% 
-  ggplot(aes(x = coef, y = phenology)) +
-  # Prevent geom_rect() from establishing a continuous y-axis scale
-  geom_point(alpha = 0) +
-  geom_rect(aes(xmin = -Inf, xmax = Inf,
-                ymin = event_begin, ymax = event_end,
-                fill = event),
-            phen_wrapped,
-            inherit.aes = FALSE) +
-  geom_pointrange(aes(xmin = coef - stderr, 
-                      xmax = coef + stderr,
-                      color = signif)) +
-  scale_color_manual(values = c("FALSE" = "grey50", "TRUE" = "black"),
-                     guide = "none") +
-  scale_fill_discrete(type = c("#829BD1", "#E9B668", "#7EC6CC", "#E091B9", "#829BD1"),
-                      breaks = c("Long trip", "Breeding", "Short trip", "Molting")) + 
-  labs(x = expression('Delay with maternal age (days year'^-1 * ')')) +
-  expand_limits(x = 0) +
-  theme_few() + 
-  theme(axis.title.y = element_blank(),
-        legend.direction = "horizontal",
-        legend.title = element_blank())
-
-plot_grid(
-  get_legend(phen_plot2),
-  plot_grid(
-    phen_plot1 + theme(legend.position = "none"),
-    phen_plot2 + theme(legend.position = "none"),
-    align = "h",
-    axis = "tb",
-    labels = c("A", "B"),
-    nrow = 1
-  ),
-  ncol = 1,
-  rel_heights = c(0.1, 1)
-)
-
-# table(seal_phenology$age,seal_phenology$observed)
-
-#white dashed lines average the averages - global average of age-specific averages 
-```
-
-Sample size n = `r nrow(seal_phenology)`.
 

### 2024-05-24 cluster/sealmark.R
got mark to run on a docker image!

@@ -1,0 +1,75 @@
+library(RMark)
+library(tidyverse)
+
+sealdat <- read_csv("data/raw/128L pull 2023_12_05.csv", 
+                    show_col_types = FALSE) %>% 
+  mutate(observed = if_else(observed == "B", "Breeder", "Non-breeder"), 
+         observed_int = if_else(observed == "Breeder", 1, 0), 
+         pup_survived = pupseeneveragain > 0) %>% 
+  filter(age > 3, 
+         year > 1987) %>% 
+  mutate(animalID = factor(animalID),
+         year_fct = factor(year),
+         age10 = age / 10) %>% 
+  group_by(animalID) %>% 
+  mutate(longevity = ifelse(max(year) < 2020,
+                            max(age),
+                            NA),
+         longevity10 = longevity / 10) %>% 
+  ungroup()
+
+sealhist <- sealdat %>% 
+  select(animalID, year, observed) %>% 
+  mutate(observed = substr(observed, 1, 1)) %>% 
+  group_by(animalID, year) %>% 
+  summarize(observed = ifelse(any(observed == "B"), "B", "N"),
+            .groups = "drop") %>% 
+  pivot_wider(names_from = year, 
+              values_from = observed,
+              values_fill = "0") %>% 
+  pivot_longer(-animalID, names_to = "year", values_to = "observation") %>%
+  group_by(animalID) %>% 
+  summarize(ch = paste(observation, collapse = "")) %>% 
+  count(ch, name = "freq")
+
+seal_proc <- process.data(sealhist, model = "Multistrata")
+seal_ddl <- make.design.data(seal_proc)
+
+S.stratum <- list(formula =~ stratum)
+S.stratumxtime <- list(formula =~ stratum * time)
+
+p.stratum <- list(formula =~ stratum)
+p.stratumxtime <- list(formula =~ stratum * time)
+p.stratumxtime.fixed <- list(formula =~ stratum * time,
+                             fixed = list(time = 4, value = 1))
+
+Psi.s <- list(formula =~ -1 + stratum:tostratum)
+
+model.list <- create.model.list("Multistrata")
+
+model.list <- rbind(model.list,
+                    c(S = "S.stratumxtime", 
+                      p = "p.stratumxtime",
+                      Psi = "Psi.s"))
+
+ms_mod <- mark.wrapper(model.list,
+                       data = seal_proc,
[…]

### 2024-05-29 paper/paperv3.qmd
Fit seg mdl w dummy var, move to supp. Replace w poly2 mdl in main.

@@
 
-age_senesce <- 12
+age_senesce <- 11
 
@@
          year_fct = factor(year),
-         age10 = age / 10) %>% 
+         age10 = age / 10,
+         age_cat = factor(age >= age_senesce, 
+                          labels = c("Young", "Old"))) %>% 
   group_by(animalID) %>% 
@@
 
-# Fit separate GLMMs to the data subsetted before and after onset of senescence
-# We a priori chose age 12 for that (age_senesce)
-# Pre-senescence model
-breed_age_mod1 <- glmer(
-    observed_int ~ age10 + (1 | animalID) + (1 | year_fct),
-    # Here's the age filter
-    data = filter(sealdat, age < age_senesce),
+breed_age_mod <- glmer(
+  observed_int ~ poly(age10, 2, raw = TRUE) + (1 | animalID) + (1 | year_fct),
+  sealdat,
   family = "binomial",
-    control = glmerControl(optimizer="bobyqa")
-)
-# Post-senescence model
-breed_age_mod2 <- glmer(
-    observed_int ~ age10 + (1 | animalID) + (1 | year_fct),
-    data = filter(sealdat, age >= age_senesce),
-    family = "binomial",
-    control = glmerControl(optimizer="bobyqa")
+  control = glmerControl(optimizer = "bobyqa")
 )
 
-breed_summ <- tibble(
-  model = c("Pre-senescence", "Post-senescence"),
-  age_coef = c(fixef(breed_age_mod1)["age10"],
-               fixef(breed_age_mod2)["age10"]),
-  age_coef_se = c(summary(breed_age_mod1)$coef["age10", "Std. Error"],
-                  summary(breed_age_mod2)$coef["age10", "Std. Error"]),
-  `p(coef<0)` = c(pnorm(summary(breed_age_mod1)$coef["age10", "z value"]),
-                  pnorm(summary(breed_age_mod2)$coef["age10", "z value"]))
-)
-
-breed_summ %>% 
-  mutate(across(where(is.numeric), \(x) signif(x, 3))) %>%
-  set_names(c("Age class", 
-              "Age", 
-              "se(Age)",
-              "p(Age) < 0")) %>% 
-  knitr::kable()
+summary(breed_age_mod)
 
@@
 
-# Create POPULATION-LEVEL (i.e. without rand effects) predictions for both
-# models and row bind them together
-breed_pred <- rbind(
-  # Pre-senescence
-  ggpredict(
-    breed_age_mod1,
-    terms = c("age10 [all]"),
-    type = "fixed",
-    interval = "confidence"
-  ) %>% 
-    as_tibble() %>% 
-    mutate(age = x * 10,
-           # Add an age_class variable to facilitate plotting
-           age_class = "young"),
-  # Post-senescence
-  ggpredict(
-    breed_age_mod2,
-    terms = c("age10 [all]"),
-    type = "fixed",
-    interval = "confidence"
-  ) %>% 
-    as_tibble() %>% 
-    mutate(age = x * 10,
-           age_class = "old")
-) %>% 
-  filter(age <= 20)
-
-# Now create rand effect predictions
-# This utility function pulls rand effect *coefs* from a model
-get_ranefs <- function(model, r, reduce = TRUE) {
-  result <- ranef(model)[[r]] %>% 
-    as_tibble(rownames = r) %>% 
-    rename(ranef = 2) %>% 
-    arrange(ranef)
-
-  if (reduce)
-    slice(result, round(seq(1, nrow(result), length.out = 5)))
-  else 
-    result
-}
-
-# Get year and individual random effeccts from pre- and post-senescence models
-year_ranef1 <- get_ranefs(breed_age_mod1, "year_fct")
-year_ranef2 <- get_ranefs(breed_age_mod2, "year_fct")
-indiv_ranef1 <- get_ranefs(breed_age_mod1, "animalID")
-indiv_ranef2 <- get_ranefs(breed_age_mod2, "animalID")
-
-# This util fun pulls rand effect *predictions* from a model
-get_ranef_pred <- function(model, r, keep) {
-  ggpredict(model,
-            terms = c("age10 [all]", r), 
-                             type = "random",
-                             interval = "confidence") %>% 
-  as_tibble() %>% 
-  mutate(age = x * 10,
-         "{r}" := group) %>% 
-  semi_join(keep, by = r) %>% 
-  filter(age <= 20)
-}
-
-# Predictions based on 5 different *years* (min, q1, median, q3, max ranef
-# coef)
-breed_pred_year <- rbind(
-  # Pre-senescence
-  get_ranef_pred(breed_age_mod1, "year_fct", year_ranef1) %>% 
-    mutate(age_class = "young"),
-  # Post-senescence
-  get_ranef_pred(breed_age_mod2, "year_fct", year_ranef2) %>% 
-    mutate(age_class = "old")
-) %>% 
-  filter(age <= 20)
-
-# As above, but 5 different *individuals*
-breed_pred_indiv <- rbind(
-  get_ranef_pred(breed_age_mod1, "animalID", indiv_ranef1) %>% 
-    mutate(age_class = "young"),
-  get_ranef_pred(breed_age_mod2, "animalID", indiv_ranef2) %>% 
-    mutate(age_class = "old")
-) %>% 
-  filter(age <= 20)
-
-# The actual raw data
+# Summarize the raw data
 breed_nonbreed <- sealdat %>% 
@@
 
-# Figure with raw data and model predictions
-breed_fig <- ggplot(breed_pred, aes(x = age)) +
-  # Prediction across years (fixed effect only)
-  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, 
-                  group = age_class),
-              fill = "grey", alpha = 0.5) +
-  geom_line(aes(y = predicted, color = age_class), 
-            linewidth = 1.5) +
-  # Predictions for specific years (random effect)
-  geom_line(aes(y = predicted, group = interaction(year_fct, age_class)),
-            breed_pred_year,
-            linewidth = 0.75, alpha = 0.5) +
-  # Raw data
+# Create POPULATION-LEVEL (i.e. without rand effects) predictions
+breed_pred_pop <- ggpredict(
+  breed_age_mod, 
+  terms = "age10 [all]",
+  type = "fixed",
+  interval = "confidence"
+) %>% 
+  as_tibble() %>% 
+  mutate(age = x * 10) %>% 
+  filter(age <= 20)
+
+# Create WEIGHTED-AVERAGE predictions using random effects
+seal_years <- count(sealdat, year_fct)
+breed_pred_wgt <- ggpredict(
+  breed_age_mod,
+  terms = c("age10 [all]", "year_fct [all]"),
+  type = "random",
+  interval = "confidence"
+) %>% 
+  as_tibble() %>% 
+  mutate(age = x * 10,
+         year_fct = group) %>% 
+  filter(age <= 20) %>% 
+  left_join(seal_years, by = "year_fct") %>% 
+  group_by(age) %>% 
+  summarize(predicted = weighted.mean(predicted, n))
+
+ggplot(breed_pred_pop, aes(age, predicted)) +
+  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), 
+              alpha = 0.2) +
+  geom_line(linewidth = 1.2, color = "grey50") +
+  geom_line(data = breed_pred_wgt, 
+            color = "cornflowerblue", 
+            linewidth = 1.2) +
   geom_pointrange(aes(y = perc_breed, 
@@
                   breed_nonbreed) +
-  # Year labels
-  geom_text(aes(y = predicted, label = year_fct),
-            filter(breed_pred_year, age == 20),
-            x = 20.2,
-            hjust = 0,
-            size = 2.5) +
-  geom_text(aes(y = predicted, label = year_fct),
-            filter(breed_pred_year, age == 4),
-            x = 3.8,
-            hjust = 1,
-            size = 2.5) +
-  # Add visual indicator for senescence
-  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
-  scale_x_continuous("Age", limits = c(3.5, 20.5)) +
-  scale_y_continuous("Adult female breeding", 
-                     limits = c(NA, 1), 
-                     labels = scales::percent) +
-  scale_color_manual(values = c("firebrick", "cornflowerblue")) +
-  theme_few() +
-  theme(legend.position = "none",
-        axis.title = element_text(size = 13))
-
-# Coefficient panel
-breed_coef_fig <- breed_summ %>% 
-  mutate(model = str_extract(model, ".*-"),
-         model = fct_rev(model)) %>% 
-  ggplot(aes(color = model)) + 
-  geom_pointrange(aes(ymin = age_coef - age_coef_se,
-                      y = age_coef,
-                      ymax = age_coef + age_coef_se,
-                      x = model)) +
-  geom_hline(yintercept = 0, linetype = "dashed") +
-  scale_color_manual(values = c("cornflowerblue", "firebrick")) +
-  labs(x = "Senescence",
-       y = "Age coefficient") +
-  theme_few() +
+  scale_y_continuous("Breeding", labels = scales::percent) +
+  labs(x = "Age (years)") +
   theme(legend.position = "none")
-
-(breed_fig | breed_coef_fig) + 
-  plot_layout(widths = c(3, 1)) +
-  plot_annotation(tag_levels = "A")
 
@@
 
+```{r}
+#| label: seg-breeding-mod
+
+# Fit GLMMs to the data with binary variable indicating before and after onset # of senescence. We a priori chose age 11 for senescence (age_senesce).
+breed_age_seg <- glmer(
+  observed_int ~ age10 * age_cat + (1 | animalID) + (1 | year_fct),
+  sealdat,
+  family = "binomial",
+  control = glmerControl(optimizer = "bobyqa")
+)
+
+summary(breed_age_seg)
+```
+
+```{r}
+#| label: fig-seg-breeding-mod
+
+breed_pred_seg <- ggpredict(breed_age_seg, 
+                            terms = c("age10 [all]", "age_cat")) %>% 
+  as_tibble() %>% 
+  mutate(age = x * 10,
+         age_cat = factor(group, levels = c("Young", "Old"))) %>% 
+  filter((age_cat == "Young" & age < age_senesce) |
+           (age_cat == "Old" & age >= age_senesce))
+
+ggplot(breed_pred_seg, aes(age, predicted)) +
+  geom_ribbon(aes(fill = age_cat, ymin = conf.low, ymax = conf.high), 
+              alpha = 0.2) +
+  geom_line(aes(color = age_cat)) +
+  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
+  scale_color_manual(values = c("cornflowerblue", "firebrick")) +
+  scale_fill_manual(values = c("cornflowerblue", "firebrick")) +
+  scale_y_continuous("Breeding", labels = scales::percent) +
+  labs(x = "Age (years)") +
+  theme(legend.position = "none")
+
+```
+
 Supplemental Table 1: Statistics for breeding success model.

### 2024-05-29 cluster/sealmark.R

@@ -1,75 +1,51 @@
-library(RMark)
+library(marked)
 library(tidyverse)
 
 sealdat <- read_csv("data/raw/128L pull 2023_12_05.csv", 
                     show_col_types = FALSE) %>% 
-  mutate(observed = if_else(observed == "B", "Breeder", "Non-breeder"), 
-         observed_int = if_else(observed == "Breeder", 1, 0), 
-         pup_survived = pupseeneveragain > 0) %>% 
-  filter(age > 3, 
+  filter(between(age, 3, 18), 
          year > 1987) %>% 
-  mutate(animalID = factor(animalID),
-         year_fct = factor(year),
-         age10 = age / 10) %>% 
-  group_by(animalID) %>% 
-  mutate(longevity = ifelse(max(year) < 2020,
-                            max(age),
-                            NA),
-         longevity10 = longevity / 10) %>% 
-  ungroup()
+  mutate(observed = 1)
 
 sealhist <- sealdat %>% 
-  select(animalID, year, observed) %>% 
-  mutate(observed = substr(observed, 1, 1)) %>% 
   group_by(animalID, year) %>% 
-  summarize(observed = ifelse(any(observed == "B"), "B", "N"),
+  summarize(observed = 1,
             .groups = "drop") %>% 
   pivot_wider(names_from = year, 
               values_from = observed,
-              values_fill = "0") %>% 
-  pivot_longer(-animalID, names_to = "year", values_to = "observation") %>%
+              values_fill = 0) %>% 
+  pivot_longer(-animalID, 
+               names_to = "year", 
+               values_to = "observation") %>%
+  arrange(animalID, year) %>% 
   group_by(animalID) %>% 
-  summarize(ch = paste(observation, collapse = "")) %>% 
-  count(ch, name = "freq")
+  summarize(ch = paste(observation, collapse = "")) 
 
-seal_proc <- process.data(sealhist, model = "Multistrata")
+seal_proc <- process.data(sealhist)
 seal_ddl <- make.design.data(seal_proc)
 
-S.stratum <- list(formula =~ stratum)
-S.stratumxtime <- list(formula =~ stratum * time)
+fit.models <- function() {
+  Phi.age <- list(formula = ~age)
+  p.age <- list(formula = ~age)
+  cml <- create.model.list(c("Phi","p"))
+  crm.wrapper(cml, 
+              data = seal_proc, 
[…]

### 2024-05-29 scratch/markrecapbase.R

@@ -2,6 +2,8 @@ library(posterior)
 library(tidyverse)
 library(rstan)
 
+inv_logit <- \(x) 1 / (1 + exp(-x))
+
 sealdat <- read_csv("data/raw/128L pull 2023_12_05.csv") %>% 
   filter(age > 3,
          year < 2023,
@@ -44,9 +46,19 @@ image(years, seq_along(animalIDs), t(obs))
 n_ind_test <- 100
 animalIDs_test <- sample(animalIDs, n_ind_test)
 
+# Read command line
+args <- commandArgs(trailingOnly = TRUE)
+if (args[1] == "test") {
+  test <- TRUE
+} else if (args[1] == "full") {
+  test <- FALSE
+} else {
+  stop(sprintf("First argument must be 'test' or 'full', not %s", args[1]))
+}
+
 # Setting up directory to save plots
-# output_path <- args[2]
-# dir.create(output_path)
+output_path <- args[2]
+dir.create(output_path)
 
 # Fit basic model ---------------------------------------------------------
 if (test) {
@@ -66,16 +78,16 @@ if (test) {
     obs = obs,
     age = age
   )
-  markrecapmod <- stan("scratch/sealaging.stan",
+  markrecapmod0 <- stan("scratch/sealaging.stan",
                        data = markrecapdat,
                        chains = 4, iter = 8000, cores = 4)
 }
 
 markrecapdraws <- as_draws_df(markrecapmod)
-summarize_draws(markrecapdraws) %>% view()
+summarize_draws(markrecapdraws)
 stan_trace(markrecapmod, 
-           pars = c("pb_maturity", "ps[1]", "ps[2]", "pd[1]", "pd[2]", 
-                    "pb[1]", "pb[2]"))
+           pars = c("pb_maturity", "lps[1]", "lps[2]", "lpb[1]", "lpb[2]",
+                    "lpd_mu[1]", "lpd_mu[2]"))
 
 # Fit time-varying detection model ----------------------------------------
 if (test) {
@@ -85,7 +97,7 @@ if (test) {
     obs = obs[animalIDs %in% animalIDs_test, ],
     age = age[animalIDs %in% animalIDs_test, ]
   )
-  markrecapmod <- stan("scratch/sealaging2.stan",
+  markrecapmod <- stan("scratch/sealaging3.stan",
                        data = markrecapdat,
                        chains = 1, iter = 200, cores = 1)
 } else {
[…]

### 2024-05-29 paper/paperv3.qmd
Update pup surv/recr

@@
 #| label: fig-raw-data
+#| fig-width: 9
 
@@
 
-summary(breed_age_mod)
+summary(breed_age_mod)$coefficients
 
@@
 #| label: fig-breeding-mod
-#| fig-width: 7.5
 
@@
 ##NOTE: Need to take out the instances where recruitment > survival
-#what does ggpredict do with random effects? 
-#thin gray lines for the different year effects? show the interannual variation (5 that include extreme ones, just something to illustrate the variation of year)
 
 # Isolate pup survival/recruitment 
-surv_recruit_data <- sealdat %>% 
+surv_recr_data <- sealdat %>% 
   drop_na(pupID) %>% 
-  # Only pups born 2016 or earlier (otherwise survival biiased)
+  # Only pups born 2016 or earlier (otherwise survival biased)
   filter(pupyearborn <= 2016) %>% 
-  mutate(pup_survived_int = ifelse(pup_survived, 1L, 0L)) %>% 
+  mutate(pup_surv_int = as.integer(pup_survived),
+         pup_recr_int = as.integer(puprecruited)) %>% 
   # Only keep maternal age classes with 5+ births, which drops moms 19+
@@
 
-# Fit pup *survival* models pre- and post-senescence
-# Commented-out formulae vary age10 coef by pupsex, end up with similar
-# results. Can add to supplemental?
-pup_surv_model1 <- glmer(
-  # pup_survived_int ~ age10 : pupsex + (1 | animalID) + (1 | year),
-  pup_survived_int ~ age10 + (1 | animalID) + (1 | year),
-  data = filter(surv_recruit_data, age < age_senesce),
-  family = "binomial"
-)
-pup_surv_model2 <- glmer(
-  # pup_survived_int ~ age10 : pupsex + (1 | animalID) + (1 | year),
-  pup_survived_int ~ age10 + (1 | animalID) + (1 | year),
-  data = filter(surv_recruit_data, age >= age_senesce),
-  family = "binomial"
+# Fit pup *survival* model
+pup_surv_mod <- glmer(
+  pup_surv_int ~ poly(age10, 2, raw = TRUE) + (1 | animalID) + (1 | year_fct),
+  surv_recr_data,
+  family = "binomial",
+  control = glmerControl(optimizer = "bobyqa")
 )
 
-# Fit pup *recruitment* models pre- and post-senescence
-pup_recr_model1 <- glmer(
-  puprecruited ~ age10 + (1 | animalID) + (1 | year),
-  # For recruitment, ONLY female pups
-  data = filter(surv_recruit_data, 
-                age < age_senesce,
-                pupsex == "F"),
-  family = "binomial"
-)
-pup_recr_model2 <- glmer(
-  puprecruited ~ age10 + (1 | animalID) + (1 | year),
-  data = filter(surv_recruit_data, 
-                age >= age_senesce,
-                pupsex == "F"),
+# Fit pup *recruitment* model (female offspring only)
+pup_recr_mod <- glmer(
+  pup_recr_int ~ poly(age10, 2, raw = TRUE) + (1 | animalID) + (1 | year_fct),
+  filter(surv_recr_data, pupsex == "F"),
   family = "binomial",
-  control = glmerControl(optimizer="bobyqa")
+  control = glmerControl(optimizer = "bobyqa")
 )
 
-# Summary table
-surv_recr_summ <- tibble(
-  response = rep(c("Survival", "Recruitment"), each = 2),
-  n = map_dbl(
-    list(pup_surv_model1@frame, 
-         pup_surv_model2@frame, 
-         pup_recr_model1@frame, 
-         pup_recr_model2@frame),
-    nrow
-  ),
-  model = rep(c("Pre-senescence", "Post-senescence"), 2),
-  age_coef = c(fixef(pup_surv_model1)["age10"],
-               fixef(pup_surv_model2)["age10"],
-               fixef(pup_recr_model1)["age10"],
-               fixef(pup_recr_model2)["age10"]),
-  age_coef_se = c(summary(pup_surv_model1)$coef["age10", "Std. Error"],
-                  summary(pup_surv_model2)$coef["age10", "Std. Error"],
-                  summary(pup_recr_model1)$coef["age10", "Std. Error"],
-                  summary(pup_recr_model2)$coef["age10", "Std. Error"]),
-  `p(coef<0)` = c(pnorm(summary(pup_surv_model1)$coef["age10", "z value"]),
-                  pnorm(summary(pup_surv_model2)$coef["age10", "z value"]),
-                  pnorm(summary(pup_recr_model1)$coef["age10", "z value"]),
-                  pnorm(summary(pup_recr_model2)$coef["age10", "z value"]))
-)
-
-surv_recr_summ %>% 
-  mutate(across(where(is.numeric), \(x) signif(x, 3))) %>% 
-  knitr::kable()
+"Survival"
+summary(pup_surv_mod)$coefficients
+"Recruitment"
+summary(pup_recr_mod)$coefficients
 
@@
 #| label: fig-repro-age
-#| fig-cap: "Offspring survival decreases as mothers senesce.
-**A** Survival of female and male offspring (filled points) is not negatively impacted by maternal age (blue line) prior to age 12 (vertical dashed line).
-After age 12, offspring survival declines with maternal age (red line), i.e., maternal effect senescence.
-Hollow points represent female offspring recruitment, which did not significantly decline with maternal age after age 12.
-**B** The coefficient of age w.r.t.
-offspring survival (solid points) was not significantly negative before age 12 (blue), but was significantly negative after (red).
-The coefficient of age w.r.t.
-female offspring recruitment (hollow points) was not significant in either maternal age class.
-Lines represent standard error of the coefficient, which is ommitted for the post-age 12 offspring recruitment coefficent (hollow red point) because the standard error was too large (14.8) to visually represent without hiding other patterns.
-Numbers at top indicate sample size in each model."
-#| fig-width: 5.5
+#| fig-width: 9
 
-#B Pup survival and recruitment doesn't change with increased maternal age.
+# Summarize the raw data
+surv_recr_summ <- surv_recr_data %>% 
+  group_by(age) %>% 
+  summarize(n_pups = n(), 
+            n_fpups = sum(pupsex == "F"),
+            perc_surv = sum(pup_surv_int) / n_pups, 
+            se_surv = (perc_surv * (1 - perc_surv) / n_pups)^0.5,
+            perc_recr = sum(pup_recr_int) / n_fpups, 
+            se_recr = (perc_recr * (1 - perc_recr) / n_fpups)^0.5) %>% 
+  filter(age <= 20)
 
-# Pup survival 
-surv_pred <- rbind(
-  # Pre-senescence
-  ggpredict(
-    pup_surv_model1,
-    terms = c("age10 [all]"),
+# Create pup survival POPULATION-LEVEL (i.e. without rand effects) predictions
+surv_pred_pop <- ggpredict(
+  pup_surv_mod, 
+  terms = "age10 [all]",
   type = "fixed",
@@
   as_tibble() %>% 
+  mutate(age = x * 10) %>% 
+  filter(age <= 20)
+
+# Create pup survival WEIGHTED-AVERAGE predictions using random effects
+seal_years <- count(sealdat, year_fct)
+surv_pred_wgt <- ggpredict(
+  pup_surv_mod,
+  terms = c("age10 [all]", "year_fct [all]"),
+  type = "random",
+  interval = "confidence"
+) %>% 
+  as_tibble() %>% 
   mutate(age = x * 10,
-           age_class = "young"),
-  # Post-senescence
-  ggpredict(
-    pup_surv_model2,
-    terms = c("age10 [all]"),
+         year_fct = group) %>% 
+  filter(age <= 20) %>% 
+  left_join(seal_years, by = "year_fct") %>% 
+  group_by(age) %>% 
+  summarize(predicted = weighted.mean(predicted, n))
+
+fig_surv <- ggplot(surv_pred_pop, aes(age, predicted)) +
+  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), 
+              alpha = 0.2) +
+  geom_line(linewidth = 1.2, color = "grey50") +
+  geom_line(data = surv_pred_wgt, 
+            color = "cornflowerblue", 
+            linewidth = 1.2) +
+  geom_pointrange(aes(y = perc_surv, 
+                      ymin = perc_surv - se_surv, 
+                      ymax = perc_surv + se_surv),
+                  surv_recr_summ) +
+  scale_y_continuous("Offspring survival", 
+                     labels = scales::percent,
+                     limits = c(0, 0.62)) +
+  labs(x = "Age (years)") +
+  theme(legend.position = "none")
+
+# Create pup recruit POPULATION-LEVEL (i.e. without rand effects) predictions
+recr_pred_pop <- ggpredict(
+  pup_recr_mod, 
+  terms = "age10 [all]",
   type = "fixed",
@@
   as_tibble() %>% 
+  mutate(age = x * 10) %>% 
+  filter(age <= 20)
+
+# Create pup recruit WEIGHTED-AVERAGE predictions using random effects
+seal_years <- count(sealdat, year_fct)
+recr_pred_wgt <- ggpredict(
+  pup_recr_mod,
+  terms = c("age10 [all]", "year_fct [all]"),
+  type = "random",
+  interval = "confidence"
+) %>% 
+  as_tibble() %>% 
   mutate(age = x * 10,
-           age_class = "old")
-)
+         year_fct = group) %>% 
+  filter(age <= 20) %>% 
+  left_join(seal_years, by = "year_fct") %>% 
+  group_by(age) %>% 
+  summarize(predicted = weighted.mean(predicted, n))
 
-repro_raw <- surv_recruit_data %>% 
-  # Uncomment to group by female and male offspring
-  # mutate(
-  #   pupsex = factor(pupsex, 
-  #                   labels = c("Female offspring", "Male offspring"))
-  # ) %>% 
-  # group_by(age, pupsex) %>% 
-  group_by(age) %>% 
-  summarize(n_born = n(), 
-            n_surv = sum(pup_survived_int),
-            perc_surv = n_surv / n_born, 
-            se_surv = (perc_surv * (1 - perc_surv) / n_born)^0.5,
-            f_born = sum(pupsex == "F"),
-            f_recr = sum(puprecruited & pupsex == "F"),
-            perc_recr = f_recr / f_born,
-            se_recr = (perc_recr * (1 - perc_recr) / f_born)^0.5,
-            .groups = "drop")
-
-surv_recr_fig <- ggplot(surv_pred, aes(x = age)) +
-  # Survival by age (fixed effect only)
-  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, 
-                  group = age_class),
-              fill = "grey", alpha = 0.5) +
-  geom_line(aes(y = predicted, color = age_class), 
-            linewidth = 1.5) +
-  # Raw survival and recruitment data
-  geom_pointrange(aes(y = perc_surv,
-                      ymin = perc_surv - se_surv,
-                      ymax = perc_surv + se_surv),
-                  repro_raw) +
+fig_recr <- ggplot(recr_pred_pop, aes(age, predicted)) +
+  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), 
+              alpha = 0.2) +
+  geom_line(linewidth = 1.2, color = "grey50") +
+  geom_line(data = recr_pred_wgt, 
+            color = "cornflowerblue", 
+            linewidth = 1.2) +
   geom_pointrange(aes(y = perc_recr, 
@@
                       ymax = perc_recr + se_recr),
-                  repro_raw,
-                  shape = 21) +
-  # Add visual indicator for senescence
-  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
-  scale_x_continuous("Age", limits = c(3.5, 18.5)) +
-  scale_y_continuous("Offspring survival/recruitment", 
-                     limits = c(0, 0.65), 
-                     labels = scales::percent) +
-  scale_color_manual(values = c("firebrick", "cornflowerblue")) +
-  # Uncomment to facet by offspring sex
-  # facet_wrap(~pupsex) +
-  theme_few() +
-  theme(legend.position = "none",
-        axis.title = element_text(size = 13))
-
-# Survival/recruitment figure. In the caption, mention the post-senescence
-# recruitment std err was omitted because it was 14.8, so it swamps all the
-# other std errs
-surv_recr_coef_fig <- surv_recr_summ %>% 
-  mutate(response = fct_rev(str_sub(response, 1, 4)),
-         model = fct_rev(str_extract(model, ".*-")),
-         # Remove post-senescence recruitment std err
-         age_coef_se = c(age_coef_se[1:3], 0)) %>% 
-  ggplot(aes(x = response, ymax = 1.1)) +
-  geom_pointrange(aes(ymin = age_coef - age_coef_se, 
-                      y = age_coef,
-                      ymax = age_coef + age_coef_se,
-                      color = model,
-                      shape = response),
-                  position = position_dodge(width = 0.9)) +
-  geom_text(aes(label = n, group = model), 
-            y = 0.9, 
-            size = 3,
-            angle = 30,
-            hjust = 0,
-            position = position_dodge(width = 1)) +
-  geom_hline(yintercept = 0, linetype = "dashed") +
-  scale_color_manual(values = c("cornflowerblue", "firebrick")) +
-  scale_shape_manual(values = c(16, 1)) +
-  labs(x = "Response", y = "Age coefficient") +
-  expand_limits(y = 1.1) +
+                  surv_recr_summ) +
+  scale_y_continuous("Offspring recruitment", 
+                     labels = scales::percent,
+                     limits = c(0, 0.62)) +
+  labs(x = "Age (years)") +
   theme(legend.position = "none")
 
-(surv_recr_fig | surv_recr_coef_fig) + 
-  plot_layout(widths = c(3, 1)) +
-  plot_annotation(tag_levels = "A")
+(fig_surv | fig_recr) + plot_annotation(tag_levels = "A")
 
@@
 
-Sample size n = `r nrow(surv_recruit_data)`.
+Sample size n = `r nrow(surv_recr_data)`.
 
@@
               alpha = 0.2) +
-  geom_line(aes(color = age_cat)) +
+  geom_line(aes(color = age_cat), linewidth = 1.2) +
+  geom_pointrange(aes(y = perc_breed, 
+                      ymin = perc_breed - se, 
+                      ymax = perc_breed + se),
+                  breed_nonbreed) +
   geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
@@
   theme(legend.position = "none")
+
+```
+
+```{r}
+#| label: seg-repro-mod
+
+# Fit pup *survival* model
+pup_surv_seg <- glmer(
+  pup_surv_int ~ age10 * age_cat + (1 | animalID) + (1 | year),
+  surv_recr_data,
+  family = "binomial",
+  control = glmerControl(optimizer = "bobyqa")
+)
+
+# Fit pup *recruitment* model
+pup_recr_seg <- glmer(
+  pup_recr_int ~ age10 * age_cat + (1 | animalID) + (1 | year),
+  filter(surv_recr_data, pupsex == "F"),
+  family = "binomial",
+  control = glmerControl(optimizer = "bobyqa")
+)
+
+"Survival"
+summary(pup_surv_seg)$coefficients
+"Recruitment"
+summary(pup_recr_seg)$coefficients
+
+```
+
+```{r}
+#| label: fig-seg-repro-mod
+#| fig-width: 9
+
+surv_pred_seg <- ggpredict(pup_surv_seg, 
+                           terms = c("age10 [all]", "age_cat")) %>% 
+  as_tibble() %>% 
+  mutate(age = x * 10,
+         age_cat = factor(group, levels = c("Young", "Old"))) %>% 
+  filter((age_cat == "Young" & age < age_senesce) |
+           (age_cat == "Old" & age >= age_senesce))
+
+fig_surv <- ggplot(surv_pred_seg, aes(age, predicted)) +
+  geom_ribbon(aes(fill = age_cat, ymin = conf.low, ymax = conf.high), 
+              alpha = 0.2) +
+  geom_line(aes(color = age_cat), linewidth = 1.2) +
+  geom_pointrange(aes(y = perc_surv, 
+                      ymin = perc_surv - se_surv, 
+                      ymax = perc_surv + se_surv),
+                  surv_recr_summ) +
+  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
+  scale_color_manual(values = c("cornflowerblue", "firebrick")) +
+  scale_fill_manual(values = c("cornflowerblue", "firebrick")) +
+  scale_y_continuous("Offspring survival", labels = scales::percent) +
+  labs(x = "Age (years)") +
+  theme(legend.position = "none")
+
+recr_pred_seg <- ggpredict(pup_recr_seg, 
+                           terms = c("age10 [all]", "age_cat")) %>% 
+  as_tibble() %>% 
+  mutate(age = x * 10,
+         age_cat = factor(group, levels = c("Young", "Old"))) %>% 
+  filter((age_cat == "Young" & age < age_senesce) |
+           (age_cat == "Old" & age >= age_senesce))
+
+fig_recr <- ggplot(recr_pred_seg, aes(age, predicted)) +
+  geom_ribbon(aes(fill = age_cat, ymin = conf.low, ymax = conf.high), 
+              alpha = 0.2) +
+  geom_line(aes(color = age_cat), linewidth = 1.2) +
+  geom_pointrange(aes(y = perc_recr, 
+                      ymin = perc_recr - se_recr, 
+                      ymax = perc_recr + se_recr),
+                  surv_recr_summ) +
+  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
+  scale_color_manual(values = c("cornflowerblue", "firebrick")) +
+  scale_fill_manual(values = c("cornflowerblue", "firebrick")) +
+  scale_y_continuous("Offspring recruitment", labels = scales::percent) +
+  labs(x = "Age (years)") +
+  theme(legend.position = "none")
+
+(fig_surv | fig_recr) + plot_annotation(tag_levels = "A")
 

### 2024-05-29 paper/paperv3.qmd
finish sex ratio

@@
 # Create pup survival WEIGHTED-AVERAGE predictions using random effects
-seal_years <- count(sealdat, year_fct)
+seal_years <- count(surv_recr_data, year_fct)
 surv_pred_wgt <- ggpredict(
@@
 # Create pup recruit WEIGHTED-AVERAGE predictions using random effects
-seal_years <- count(sealdat, year_fct)
+seal_years <- count(filter(surv_recr_data, pupsex == "F"), year_fct)
 recr_pred_wgt <- ggpredict(
@@
          age < 19) %>% 
-  mutate(is_male = ifelse(pupsex == "M", 1, 0))
+  mutate(is_male = as.integer(pupsex == "M"))
 
-sex_ratio_mod1 <- glmer(
-  is_male ~ age10 + (1 | animalID) + (1 | year),
-  filter(pup_sex_data, age < age_senesce),
-  family = "binomial"
-)
-sex_ratio_mod2 <- glmer(
-  is_male ~ age10 + (1 | animalID) + (1 | year),
-  filter(pup_sex_data, age >= age_senesce),
+sex_ratio_mod <- glmer(
+  is_male ~ poly(age10, 2, raw = TRUE) + (1 | animalID) + (1 | year_fct),
+  pup_sex_data,
   family = "binomial"
@@
 
-sex_ratio_summ <- tibble(
-  model = c("Pre-senescence", "Post-senescence"),
-  age_coef = c(fixef(sex_ratio_mod1)["age10"],
-               fixef(sex_ratio_mod2)["age10"]),
-  age_coef_se = c(summary(sex_ratio_mod1)$coef["age10", "Std. Error"],
-                  summary(sex_ratio_mod2)$coef["age10", "Std. Error"]),
-  `p(coef!=0)` = c(summary(sex_ratio_mod1)$coef["age10", "Pr(>|z|)"],
-                   summary(sex_ratio_mod2)$coef["age10", "Pr(>|z|)"])
-)
-
-sex_ratio_summ %>% 
-  mutate(across(where(is.numeric), \(x) signif(x, 3))) %>%
-  set_names(c("Age class", 
-              "Age", 
-              "se(Age)",
-              "p(Age) != 0")) %>% 
-  knitr::kable()
+summary(sex_ratio_mod)$coefficients
 
@@
 #| label: fig-sex-ratio
-#| fig-cap: "The effect of age on offspring sex ratio was not significantly different for pre- and post-senescence adults.
-**A** Raw data (points with error bars, indicating mean and standard error of the proportion) and model predictions (lines with 95% confidence intervals).
-**B** Estimate and standard error of age coefficients in pre- and post-senescence models."
-#| fig-width: 5.5
 
-# Model predictions
-sex_ratio_pred <- rbind(
-  # Pre-senescence
-  ggpredict(
-    sex_ratio_mod1,
-    terms = c("age10 [all]"),
+# Summarize the raw data
+sexrat_summ <- pup_sex_data %>% 
+  group_by(age) %>% 
+  summarize(n_observed = n(), 
+            perc_male = mean(is_male), 
+            se = (perc_male * (1 - perc_male) / n_observed)^0.5) %>% 
+  filter(age <= 20)
+
+# Create POPULATION-LEVEL (i.e. without rand effects) predictions
+sexrat_pred_pop <- ggpredict(
+  sex_ratio_mod, 
+  terms = "age10 [all]",
   type = "fixed",
@@
   as_tibble() %>% 
-    mutate(age = x * 10,
-           # Add an age_class variable to facilitate plotting
-           age_class = "young"),
-  # Post-senescence
-  ggpredict(
-    sex_ratio_mod2,
-    terms = c("age10 [all]"),
-    type = "fixed",
+  mutate(age = x * 10) %>% 
+  filter(age <= 20)
+
+# Create WEIGHTED-AVERAGE predictions using random effects
+seal_years <- count(pup_sex_data, year_fct)
+sexrat_pred_wgt <- ggpredict(
+  sex_ratio_mod,
+  terms = c("age10 [all]", "year_fct [all]"),
+  type = "random",
   interval = "confidence"
@@
   mutate(age = x * 10,
-           age_class = "old")
-)
+         year_fct = group) %>% 
+  filter(age <= 20) %>% 
+  left_join(seal_years, by = "year_fct") %>% 
+  group_by(age) %>% 
+  summarize(pred_sd = sd(predicted),
+            predicted = weighted.mean(predicted, n))
 
-# Sex ratio raw data
-sex_ratio <- pup_sex_data %>% 
-  group_by(age) %>% 
-  summarize(n_observed = n(), 
-            perc_male = sum(is_male) / n_observed, 
-            se = (perc_male * (1 - perc_male) / n_observed)^0.5)
-
-# Figure with raw data and model predictions
-sex_ratio_fig <- ggplot(sex_ratio_pred, aes(x = age)) +
-  # Prediction across years (fixed effect only)
-  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, 
-                  group = age_class),
-              fill = "grey", alpha = 0.5) +
-  geom_line(aes(y = predicted, color = age_class), 
-            linewidth = 1.5) +
-  # Raw data
+ggplot(sexrat_pred_pop, aes(age, predicted)) +
+  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), 
+              alpha = 0.2) +
+  geom_line(linewidth = 1.2, color = "grey50") +
   geom_pointrange(aes(y = perc_male, 
@@
                       ymax = perc_male + se),
-                  sex_ratio) + 
-  # Add visual indicator for senescence
-  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
-  scale_y_continuous("Male offspring", 
-                     labels = scales::percent) +
-  scale_color_manual(values = c("firebrick", "cornflowerblue")) +
-  labs(x = "Age") +
-  theme_few() +
-  theme(legend.position = "none",
-        axis.title = element_text(size = 13))
-
-# Coefficient panel
-sex_ratio_coef_fig <- sex_ratio_summ %>% 
-  mutate(model = str_extract(model, ".*-"),
-         model = fct_rev(model)) %>% 
-  ggplot(aes(color = model)) + 
-  geom_pointrange(aes(ymin = age_coef - age_coef_se,
-                      y = age_coef,
-                      ymax = age_coef + age_coef_se,
-                      x = model)) +
-  geom_hline(yintercept = 0, linetype = "dashed") +
-  scale_color_manual(values = c("cornflowerblue", "firebrick")) +
-  labs(x = "Senescence",
-       y = "Age coefficient") +
-  theme_few() +
+                  sexrat_summ) +
+  scale_y_continuous("Male offspring", labels = scales::percent) +
+  labs(x = "Age (years)") +
   theme(legend.position = "none")
-
-(sex_ratio_fig | sex_ratio_coef_fig) + 
-  plot_layout(widths = c(3, 1)) +
-  plot_annotation(tag_levels = "A")
 ```
@@
 Sample size n = `r nrow(pup_sex_data)`.
-
-There appears to be a step change biasing offspring sex ratio towards males post-senescence.
-We simulated the sex ratio of 1,000,000 offspring born to mothers pre- and post-senesecence using the coefficient estimates and standard errors of the fitted sex ratio models.
-The mean offspring sex ratio increases from 51.3% male before maternal age 12 to 61.2% male after maternal age 12.
-However, the uncertainty in post-senescence sex ratio is considerable, and only 59.3% of the contrasts of simulated post- and pre-senescence sex ratios are biased towards male offspring.
-I.e., a simulated post-senescent adult female was more likely to have a male offspring than a simulated pre-senescent adult female in 59.3% of simulations.
-
-```{r}
-#| label: fig-sim-sex-ratio
-#| fig-cap: "Simulated sex ratios of offspring born to pre- and post-senescence adults.
-**A** Comparison of simulated sex ratios by pre- (blue) and post- (red) senescence adults.
-The mean (vertical lines) sex ratio of post-senescence adults is greater than pre-senescence adults.
-However the distributions overlap extensively.
-**B** The contrast of post-senescence sex ratio relative to pre-senescence is greater than 0 in only 59.3% of simulations."
-#| fig-width: 5.5
-
-# pre_sex_ratio <- tibble(age10 = seq(4, age_senesce - 1)) %>% 
-inv_logit <- \(x) exp(x) / (1 + exp(x))
-n_sim <- 1e6
-sim_sex_ratio <- rbind(
-  tibble(intercept_mean = summary(sex_ratio_mod1)$coef[1, 1],
-         intercept_sd = summary(sex_ratio_mod1)$coef[1, 2],
-         slope_mean = summary(sex_ratio_mod1)$coef[2, 1],
-         slope_sd = summary(sex_ratio_mod1)$coef[2, 2],
-         age10 = sample(seq(4, age_senesce - 1) / 10, n_sim, TRUE),
-         intercept = rnorm(n_sim, intercept_mean, intercept_sd),
-         slope = rnorm(n_sim, slope_mean, slope_sd),
-         logit_p = intercept + slope * age10,
-         p = inv_logit(logit_p),
-         age_class = "Pre-senescence"),
-  tibble(intercept_mean = summary(sex_ratio_mod2)$coef[1, 1],
-         intercept_sd = summary(sex_ratio_mod2)$coef[1, 2],
-         slope_mean = summary(sex_ratio_mod2)$coef[2, 1],
-         slope_sd = summary(sex_ratio_mod2)$coef[2, 2],
-         age10 = sample(seq(age_senesce, 18) / 10, n_sim, TRUE),
-         intercept = rnorm(n_sim, intercept_mean, intercept_sd),
-         slope = rnorm(n_sim, slope_mean, slope_sd),
-         logit_p = intercept + slope * age10,
-         p = inv_logit(logit_p),
-         age_class = "Post-senescence")
-) %>% 
-  mutate(age_class = fct_rev(age_class))
-
-pre_logit_p <- with(sim_sex_ratio, logit_p[age_class == "Pre-senescence"])
-post_logit_p <- with(sim_sex_ratio, logit_p[age_class == "Post-senescence"])
-sex_ratio_contrast <- tibble(
-  contrast_logit_p = post_logit_p - pre_logit_p
-)
-
-sim_sex_ratio_fig <- ggplot(sim_sex_ratio, aes(p)) +
-  geom_density(aes(fill = age_class), alpha = 0.5) +
-  geom_density(aes(color = age_class)) +
-  geom_vline(aes(xintercept = p, color = age_class),
-             sim_sex_ratio %>% 
-               group_by(age_class) %>% 
-               summarize(p = inv_logit(mean(logit_p))),
-             linewidth = 1.5) +
-  scale_x_continuous("Male offspring", 
-                     labels = scales::percent) +
-  scale_color_manual(values = c("cornflowerblue", "firebrick")) +
-  scale_fill_manual(values = c("cornflowerblue", "firebrick")) +
-  theme(legend.position = "none")
-
-sim_sr_contrast_fig <- ggplot(sex_ratio_contrast, aes(contrast_logit_p)) +
-  geom_density(fill = "grey90") +
-  geom_vline(xintercept = 0, linetype = "dashed") +
-  labs(x = "Contrast of logit(p)")
-
-(sim_sex_ratio_fig | sim_sr_contrast_fig) +
-  plot_annotation(tag_levels = "A")
-
-```
+In the above figure, the population-level and weighted-average would lie directly on top of each other, so we've excluded the weighted average.
+This is because the random effect of year has very little impact on offspring sex ratio, which makes sense biologically.
+It's easy to imagine scenarios where reproduction would vary year-to-year, but harder to think of a mechanism linking year and offspring sex ratio.
 
@@
 ```{r}
-#| label: fig-seg-repro-mod
+#| label: fig-seg-repro
 #| fig-width: 9
@@
 (fig_surv | fig_recr) + plot_annotation(tag_levels = "A")
+
+```
+
+```{r}
+#| label: seg-sexrat-mod
+
+# Fit GLMMs to the data with binary variable indicating before and after onset 
+# of senescence. We a priori chose age 11 for senescence (age_senesce).
+sexrat_age_seg <- glmer(
+  is_male ~ age10 * age_cat + (1 | animalID) + (1 | year_fct),
+  pup_sex_data,
+  family = "binomial",
+  control = glmerControl(optimizer = "bobyqa")
+)
+
+summary(sexrat_age_seg)$coefficients
+```
+
+```{r}
+#| label: fig-seg-sexrat
+
+sexrat_pred_seg <- ggpredict(sexrat_age_seg, 
+                             terms = c("age10 [all]", "age_cat")) %>% 
+  as_tibble() %>% 
+  mutate(age = x * 10,
+         age_cat = factor(group, levels = c("Young", "Old"))) %>% 
+  filter((age_cat == "Young" & age < age_senesce) |
+           (age_cat == "Old" & age >= age_senesce))
+
+ggplot(sexrat_pred_seg, aes(age, predicted)) +
+  geom_ribbon(aes(fill = age_cat, ymin = conf.low, ymax = conf.high), 
+              alpha = 0.2) +
+  geom_line(aes(color = age_cat), linewidth = 1.2) +
+  geom_pointrange(aes(y = perc_male, 
+                      ymin = perc_male - se, 
+                      ymax = perc_male + se),
+                  sexrat_summ) +
+  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
+  scale_color_manual(values = c("cornflowerblue", "firebrick")) +
+  scale_fill_manual(values = c("cornflowerblue", "firebrick")) +
+  scale_y_continuous("Breeding", labels = scales::percent) +
+  labs(x = "Age (years)") +
+  theme(legend.position = "none")
 

### 2024-06-07 paper/paperv3.qmd

@@
 
-### H1: Age distribution and breeding status
+### H1: Reproductive senescence
 
@@
 
+# Fit GLMMs to the data with binary variable indicating before and after onset # of senescence. We a priori chose age 11 for senescence (age_senesce).
 breed_age_mod <- glmer(
-  observed_int ~ poly(age10, 2, raw = TRUE) + (1 | animalID) + (1 | year_fct),
+  observed_int ~ age10 * age_cat + (1 | animalID) + (1 | year_fct),
   sealdat,
@@
 #| label: fig-breeding-mod
+#| fig-cap: "Breeding rates increased until age 10, then decreased from age 11 on.
+Black points represent mean and standard error of the observed breeding rates.
+The solid lines and shaded areas represent the mean and 95% confidence interval of the population-level model predictions.
+The observed data are unequally distributed across years, so the dashed lines indicate the predicted breeding rates averaged over years (i.e., including the random effect of year), weighted by the number of seals observed in each year."
 
@@
 # Create POPULATION-LEVEL (i.e. without rand effects) predictions
-breed_pred_pop <- ggpredict(
-  breed_age_mod, 
-  terms = "age10 [all]",
-  type = "fixed",
-  interval = "confidence"
-) %>% 
+breed_pred_pop <- ggpredict(breed_age_mod, 
+                            terms = c("age10 [all]", "age_cat")) %>% 
   as_tibble() %>% 
-  mutate(age = x * 10) %>% 
-  filter(age <= 20)
+  mutate(age = x * 10,
+         age_cat = factor(group, levels = c("Young", "Old"))) %>% 
+  filter((age_cat == "Young" & age < age_senesce) |
+           (age_cat == "Old" & age >= age_senesce),
+         age <= 20)
 
-# Create WEIGHTED-AVERAGE predictions using random effects
+# Create WEIGHTED-AVERAGE predictions using years' random effects
 seal_years <- count(sealdat, year_fct)
@@
   breed_age_mod,
-  terms = c("age10 [all]", "year_fct [all]"),
+  terms = c("age10 [all]", "age_cat", "year_fct [all]"),
   type = "random",
@@
   mutate(age = x * 10,
-         year_fct = group) %>% 
-  filter(age <= 20) %>% 
+         age_cat = group,
+         year_fct = facet) %>% 
+  filter((age_cat == "Young" & age < age_senesce) |
+           (age_cat == "Old" & age >= age_senesce),
+         age <= 20) %>% 
   left_join(seal_years, by = "year_fct") %>% 
-  group_by(age) %>% 
-  summarize(predicted = weighted.mean(predicted, n))
+  group_by(age, age_cat) %>% 
+  summarize(predicted = weighted.mean(predicted, n),
+            .groups = "drop")
 
 ggplot(breed_pred_pop, aes(age, predicted)) +
-  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), 
+  geom_ribbon(aes(fill = age_cat, ymin = conf.low, ymax = conf.high), 
               alpha = 0.2) +
-  geom_line(linewidth = 1.2, color = "grey50") +
-  geom_line(data = breed_pred_wgt, 
-            color = "cornflowerblue", 
-            linewidth = 1.2) +
+  geom_line(aes(color = age_cat), linewidth = 1.2) +
+  geom_line(aes(color = age_cat),
+            breed_pred_wgt, 
+            linewidth = 1.2,
+            linetype = "dashed") +
   geom_pointrange(aes(y = perc_breed, 
@@
                   breed_nonbreed) +
+  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
   scale_y_continuous("Breeding", labels = scales::percent) +
+  scale_color_manual(values = c("cornflowerblue", "firebrick")) +
+  scale_fill_manual(values = c("cornflowerblue", "firebrick")) +
   labs(x = "Age (years)") +
@@
 
-##NOTE: Need to take out the instances where recruitment > survival
-
 # Isolate pup survival/recruitment 
@@
   mutate(pup_surv_int = as.integer(pup_survived),
-         pup_recr_int = as.integer(puprecruited)) %>% 
-  # Only keep maternal age classes with 5+ births, which drops moms 19+
-  group_by(age) %>% 
-  mutate(n_pups = n()) %>% 
-  ungroup() %>%
-  filter(n_pups >= 5)
+         pup_recr_int = as.integer(puprecruited))
 
@@
 pup_surv_mod <- glmer(
-  pup_surv_int ~ poly(age10, 2, raw = TRUE) + (1 | animalID) + (1 | year_fct),
+  pup_surv_int ~ age10 * age_cat + (1 | animalID) + (1 | year_fct),
   surv_recr_data,
@@
 pup_recr_mod <- glmer(
-  pup_recr_int ~ poly(age10, 2, raw = TRUE) + (1 | animalID) + (1 | year_fct),
+  # NOTE: no random effect for individual bc singular fit
+  # In survival model, 93 individuals (15%) had >3 pups (male+female pups)
+  # In recruit model, only 17 individuals (4%) had >3 pups (female pups only)
+  pup_recr_int ~ age10 * age_cat + (1 | year_fct),
   filter(surv_recr_data, pupsex == "F"),
@@
   pup_surv_mod, 
-  terms = "age10 [all]",
+  terms = c("age10 [all]", "age_cat"),
   type = "fixed",
@@
   as_tibble() %>% 
-  mutate(age = x * 10) %>% 
-  filter(age <= 20)
-
-# Create pup survival WEIGHTED-AVERAGE predictions using random effects
-seal_years <- count(surv_recr_data, year_fct)
-surv_pred_wgt <- ggpredict(
-  pup_surv_mod,
-  terms = c("age10 [all]", "year_fct [all]"),
-  type = "random",
-  interval = "confidence"
-) %>% 
-  as_tibble() %>% 
   mutate(age = x * 10,
-         year_fct = group) %>% 
-  filter(age <= 20) %>% 
-  left_join(seal_years, by = "year_fct") %>% 
-  group_by(age) %>% 
-  summarize(predicted = weighted.mean(predicted, n))
+         age_cat = group) %>% 
+  filter((age_cat == "Young" & age < age_senesce) |
+           (age_cat == "Old" & age >= age_senesce),
+         age <= 20)
 
 fig_surv <- ggplot(surv_pred_pop, aes(age, predicted)) +
-  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), 
+  geom_ribbon(aes(fill = age_cat, ymin = conf.low, ymax = conf.high), 
               alpha = 0.2) +
-  geom_line(linewidth = 1.2, color = "grey50") +
-  geom_line(data = surv_pred_wgt, 
-            color = "cornflowerblue", 
-            linewidth = 1.2) +
+  geom_line(aes(color = age_cat), linewidth = 1.2) +
   geom_pointrange(aes(y = perc_surv, 
@@
                   surv_recr_summ) +
+  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
   scale_y_continuous("Offspring survival", 
@@
                      limits = c(0, 0.62)) +
+  scale_color_manual(values = c("cornflowerblue", "firebrick")) +
+  scale_fill_manual(values = c("cornflowerblue", "firebrick")) +
   labs(x = "Age (years)") +
@@
   pup_recr_mod, 
-  terms = "age10 [all]",
+  terms = c("age10 [all]", "age_cat"),
   type = "fixed",
@@
   as_tibble() %>% 
-  mutate(age = x * 10) %>% 
-  filter(age <= 20)
-
-# Create pup recruit WEIGHTED-AVERAGE predictions using random effects
-seal_years <- count(filter(surv_recr_data, pupsex == "F"), year_fct)
-recr_pred_wgt <- ggpredict(
-  pup_recr_mod,
-  terms = c("age10 [all]", "year_fct [all]"),
-  type = "random",
-  interval = "confidence"
-) %>% 
-  as_tibble() %>% 
   mutate(age = x * 10,
-         year_fct = group) %>% 
-  filter(age <= 20) %>% 
-  left_join(seal_years, by = "year_fct") %>% 
-  group_by(age) %>% 
-  summarize(predicted = weighted.mean(predicted, n))
+         age_cat = group) %>% 
+  filter((age_cat == "Young" & age < age_senesce) |
+           (age_cat == "Old" & age >= age_senesce),
+         age <= 20)
 
 fig_recr <- ggplot(recr_pred_pop, aes(age, predicted)) +
-  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), 
+  geom_ribbon(aes(fill = age_cat, ymin = conf.low, ymax = conf.high), 
               alpha = 0.2) +
-  geom_line(linewidth = 1.2, color = "grey50") +
-  geom_line(data = recr_pred_wgt, 
-            color = "cornflowerblue", 
-            linewidth = 1.2) +
+  geom_line(aes(color = age_cat), linewidth = 1.2) +
   geom_pointrange(aes(y = perc_recr, 
@@
                   surv_recr_summ) +
+  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
   scale_y_continuous("Offspring recruitment", 
@@
 
-sex_ratio_mod <- glmer(
-  is_male ~ poly(age10, 2, raw = TRUE) + (1 | animalID) + (1 | year_fct),
+sexrat_age_mod <- glmer(
+  is_male ~ age10 * age_cat + (1 | animalID) + (1 | year_fct),
   pup_sex_data,
@@
 
-summary(sex_ratio_mod)$coefficients
+summary(sexrat_age_mod)$coefficients
 
@@
 # Create POPULATION-LEVEL (i.e. without rand effects) predictions
-sexrat_pred_pop <- ggpredict(
-  sex_ratio_mod, 
-  terms = "age10 [all]",
-  type = "fixed",
-  interval = "confidence"
-) %>% 
-  as_tibble() %>% 
-  mutate(age = x * 10) %>% 
-  filter(age <= 20)
-
-# Create WEIGHTED-AVERAGE predictions using random effects
-seal_years <- count(pup_sex_data, year_fct)
-sexrat_pred_wgt <- ggpredict(
-  sex_ratio_mod,
-  terms = c("age10 [all]", "year_fct [all]"),
-  type = "random",
-  interval = "confidence"
-) %>% 
+sexrat_pred <- ggpredict(sexrat_age_mod, 
+                         terms = c("age10 [all]", "age_cat")) %>% 
   as_tibble() %>% 
   mutate(age = x * 10,
-         year_fct = group) %>% 
-  filter(age <= 20) %>% 
-  left_join(seal_years, by = "year_fct") %>% 
-  group_by(age) %>% 
-  summarize(pred_sd = sd(predicted),
-            predicted = weighted.mean(predicted, n))
+         age_cat = factor(group, levels = c("Young", "Old"))) %>% 
+  filter((age_cat == "Young" & age < age_senesce) |
+           (age_cat == "Old" & age >= age_senesce))
 
-ggplot(sexrat_pred_pop, aes(age, predicted)) +
-  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), 
+ggplot(sexrat_pred, aes(age, predicted)) +
+  geom_ribbon(aes(fill = age_cat, ymin = conf.low, ymax = conf.high), 
               alpha = 0.2) +
-  geom_line(linewidth = 1.2, color = "grey50") +
+  geom_line(aes(color = age_cat), linewidth = 1.2) +
   geom_pointrange(aes(y = perc_male, 
@@
                   sexrat_summ) +
+  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
+  scale_color_manual(values = c("cornflowerblue", "firebrick")) +
+  scale_fill_manual(values = c("cornflowerblue", "firebrick")) +
   scale_y_continuous("Male offspring", labels = scales::percent) +
@@
 
-### Supplemental Material
+## Supplemental Material
+
+### Models' sensitivity to age of senescence
 
 ```{r}
-#| label: seg-breeding-mod
+#| label: mod-sens
 
-# Fit GLMMs to the data with binary variable indicating before and after onset # of senescence. We a priori chose age 11 for senescence (age_senesce).
-breed_age_seg <- glmer(
+fit_models <- function(a) {
+  age_senesce <- a
+  sealdat <- sealdat %>% 
+    mutate(age_cat = factor(age >= age_senesce, 
+                            labels = c("Young", "Old")))
+
+  # H1
+  breed_age_mod <- glmer(
     observed_int ~ age10 * age_cat + (1 | animalID) + (1 | year_fct),
@@
 
-summary(breed_age_seg)
-```
+  # H2
+  surv_recr_data <- sealdat %>% 
+    drop_na(pupID) %>% 
+    filter(pupyearborn <= 2016) %>% 
+    mutate(pup_surv_int = as.integer(pup_survived),
+           pup_recr_int = as.integer(puprecruited))
 
-```{r}
-#| label: fig-seg-breeding-mod
-
-breed_pred_seg <- ggpredict(breed_age_seg, 
-                            terms = c("age10 [all]", "age_cat")) %>% 
-  as_tibble() %>% 
-  mutate(age = x * 10,
-         age_cat = factor(group, levels = c("Young", "Old"))) %>% 
-  filter((age_cat == "Young" & age < age_senesce) |
-           (age_cat == "Old" & age >= age_senesce))
-
-ggplot(breed_pred_seg, aes(age, predicted)) +
-  geom_ribbon(aes(fill = age_cat, ymin = conf.low, ymax = conf.high), 
-              alpha = 0.2) +
-  geom_line(aes(color = age_cat), linewidth = 1.2) +
-  geom_pointrange(aes(y = perc_breed, 
-                      ymin = perc_breed - se, 
-                      ymax = perc_breed + se),
-                  breed_nonbreed) +
-  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
-  scale_color_manual(values = c("cornflowerblue", "firebrick")) +
-  scale_fill_manual(values = c("cornflowerblue", "firebrick")) +
-  scale_y_continuous("Breeding", labels = scales::percent) +
-  labs(x = "Age (years)") +
-  theme(legend.position = "none")
-
-```
-
-```{r}
-#| label: seg-repro-mod
-
-# Fit pup *survival* model
-pup_surv_seg <- glmer(
-  pup_surv_int ~ age10 * age_cat + (1 | animalID) + (1 | year),
+  pup_surv_mod <- glmer(
+    pup_surv_int ~ age10 * age_cat + (1 | animalID) + (1 | year_fct),
     surv_recr_data,
@@
 
-# Fit pup *recruitment* model
-pup_recr_seg <- glmer(
-  pup_recr_int ~ age10 * age_cat + (1 | animalID) + (1 | year),
+  pup_recr_mod <- glmer(
+    pup_recr_int ~ age10 * age_cat + (1 | year_fct),
     filter(surv_recr_data, pupsex == "F"),
@@
 
-"Survival"
-summary(pup_surv_seg)$coefficients
-"Recruitment"
-summary(pup_recr_seg)$coefficients
+  # H3
+  pup_sex_data <- sealdat %>% 
+    filter(pupsex %in% c("M", "F"), 
+           age < 19) %>% 
+    mutate(is_male = as.integer(pupsex == "M"))
+
+  sexrat_age_mod <- glmer(
+    is_male ~ age10 * age_cat + (1 | animalID) + (1 | year_fct),
+    pup_sex_data,
+    family = "binomial"
+  )
+
+  list(
+    h1 = summary(breed_age_mod)$coefficients,
+    h2_surv = summary(pup_surv_mod)$coefficients,
+    h2_recr = summary(pup_recr_mod)$coefficients,
+    h3 = summary(sexrat_age_mod)$coefficients
+  )
+}
+
+# Fit models with cut-off at ages 8-13
+cutoff <- 8:13
+mod_sens <- map(cutoff, fit_models)
+
+# Are hypotheses supported by the model?
+extract_p <- \(h) map_dbl(mod_sens, \(m) m[[h]]["age10:age_catOld", "Pr(>|z|)"])
+p_tbl <- tibble(age_cutoff = cutoff, 
+                "H1 (breeding)" = extract_p("h1"),
+                "H2a (survival)" = extract_p("h2_surv"),
+                "H2b (recruitment)" = extract_p("h2_recr"),
+                "H3 (sex ratio)" = extract_p("h3"))
+
+knitr::kable(p_tbl)
 
@@
 ```{r}
-#| label: fig-seg-repro
-#| fig-width: 9
+#| label: fig-mod-sens
+#| fig-cap: "Model sensitivity to onset of senescence.
+Age at onset of senescence on the x-axis, p-value for the coefficient of the interaction (age category = post-senescence) x (response variable) on the y-axis.
+Statistical significance (p = 0.05) indicated by dashed line.
+Hypotheses H1 (reproductive senescence), H2b (maternal-effect senescence; offspring recruitment), and H3 (offspring sex ratio) were insensitive to choice of age for onset of senescence.
+Hypothesis H2a (maternal-effect senescence; offspring survival) was significantly supported for onset of senescence age 11+.
+At age 10, the p-vale for the coefficient of the interaction (age category = post-senescence) x (offspring survival) was 0.051."
 
-surv_pred_seg <- ggpredict(pup_surv_seg, 
-                           terms = c("age10 [all]", "age_cat")) %>% 
-  as_tibble() %>% 
-  mutate(age = x * 10,
-         age_cat = factor(group, levels = c("Young", "Old"))) %>% 
-  filter((age_cat == "Young" & age < age_senesce) |
-           (age_cat == "Old" & age >= age_senesce))
-
-fig_surv <- ggplot(surv_pred_seg, aes(age, predicted)) +
-  geom_ribbon(aes(fill = age_cat, ymin = conf.low, ymax = conf.high), 
-              alpha = 0.2) +
-  geom_line(aes(color = age_cat), linewidth = 1.2) +
-  geom_pointrange(aes(y = perc_surv, 
-                      ymin = perc_surv - se_surv, 
-                      ymax = perc_surv + se_surv),
-                  surv_recr_summ) +
-  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
-  scale_color_manual(values = c("cornflowerblue", "firebrick")) +
-  scale_fill_manual(values = c("cornflowerblue", "firebrick")) +
-  scale_y_continuous("Offspring survival", labels = scales::percent) +
-  labs(x = "Age (years)") +
-  theme(legend.position = "none")
-
-recr_pred_seg <- ggpredict(pup_recr_seg, 
-                           terms = c("age10 [all]", "age_cat")) %>% 
-  as_tibble() %>% 
-  mutate(age = x * 10,
-         age_cat = factor(group, levels = c("Young", "Old"))) %>% 
-  filter((age_cat == "Young" & age < age_senesce) |
-           (age_cat == "Old" & age >= age_senesce))
-
-fig_recr <- ggplot(recr_pred_seg, aes(age, predicted)) +
-  geom_ribbon(aes(fill = age_cat, ymin = conf.low, ymax = conf.high), 
-              alpha = 0.2) +
-  geom_line(aes(color = age_cat), linewidth = 1.2) +
-  geom_pointrange(aes(y = perc_recr, 
-                      ymin = perc_recr - se_recr, 
-                      ymax = perc_recr + se_recr),
-                  surv_recr_summ) +
-  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
-  scale_color_manual(values = c("cornflowerblue", "firebrick")) +
-  scale_fill_manual(values = c("cornflowerblue", "firebrick")) +
-  scale_y_continuous("Offspring recruitment", labels = scales::percent) +
-  labs(x = "Age (years)") +
-  theme(legend.position = "none")
-
-(fig_surv | fig_recr) + plot_annotation(tag_levels = "A")
+p_tbl %>% 
+  pivot_longer(-age_cutoff, names_to = "hypothesis", values_to = "p") %>% 
+  ggplot(aes(age_cutoff, p)) +
+  geom_point() +
+  geom_hline(yintercept = 0.05, linetype = "dashed") +
+  scale_y_log10(labels = scales::number) +
+  labs(x = "Senescence onset (years)") +
+  facet_wrap(~hypothesis)
 
@@
 
-```{r}
-#| label: seg-sexrat-mod
-
-# Fit GLMMs to the data with binary variable indicating before and after onset 
-# of senescence. We a priori chose age 11 for senescence (age_senesce).
-sexrat_age_seg <- glmer(
-  is_male ~ age10 * age_cat + (1 | animalID) + (1 | year_fct),
-  pup_sex_data,
-  family = "binomial",
-  control = glmerControl(optimizer = "bobyqa")
-)
-
-summary(sexrat_age_seg)$coefficients
-```
+### Haul-out duration by reproductive state
 
 ```{r}
-#| label: fig-seg-sexrat
+#| label: fig-haul-dur
 
-sexrat_pred_seg <- ggpredict(sexrat_age_seg, 
-                             terms = c("age10 [all]", "age_cat")) %>% 
-  as_tibble() %>% 
-  mutate(age = x * 10,
-         age_cat = factor(group, levels = c("Young", "Old"))) %>% 
-  filter((age_cat == "Young" & age < age_senesce) |
-           (age_cat == "Old" & age >= age_senesce))
-
-ggplot(sexrat_pred_seg, aes(age, predicted)) +
-  geom_ribbon(aes(fill = age_cat, ymin = conf.low, ymax = conf.high), 
-              alpha = 0.2) +
-  geom_line(aes(color = age_cat), linewidth = 1.2) +
-  geom_pointrange(aes(y = perc_male, 
-                      ymin = perc_male - se, 
-                      ymax = perc_male + se),
-                  sexrat_summ) +
-  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
-  scale_color_manual(values = c("cornflowerblue", "firebrick")) +
-  scale_fill_manual(values = c("cornflowerblue", "firebrick")) +
-  scale_y_continuous("Breeding", labels = scales::percent) +
-  labs(x = "Age (years)") +
+haul_dur <- drop_na(sealdat, breeddur)
+haul_summ <- haul_dur %>% 
+  group_by(observed) %>% 
+  summarize(breeddur_10 = quantile(breeddur, 0.1),
+            breeddur_lbl = str_glue("10% < {breeddur_10} days"))
+ggplot(haul_dur, aes(breeddur)) +
+  geom_density(aes(fill = observed), alpha = 0.5) +
+  geom_density(aes(color = observed)) +
+  geom_vline(aes(color = observed, xintercept = breeddur_10), 
+             haul_summ) +
+  geom_text(aes(x = breeddur_10 + 0.5, label = breeddur_lbl), 
+            haul_summ, 
+            y = 0.09,
+            hjust = 0) +
+  scale_x_continuous("Haul-out duration (days)", 
+                     limits = c(0, 50)) +
+  scale_color_manual(values = c("darkgreen", "darkorchid")) +
+  scale_fill_manual(values = c("darkgreen", "darkorchid")) +
   theme(legend.position = "none")
@@
 ```
-
-Supplemental Table 1: Statistics for breeding success model.
-
-```{r, code = FALSE}
-#| label: tbl-breed-cap
-
-tbl_breed_cap <- str_glue("The results from breeding probability analysis.
-The random effect variance of animal id is {format_signif(VarCorr(breeding_mod)$animalID, 3)} and of year was {format_signif(VarCorr(breeding_mod)$year, 3)}.")
-
-```
-
-```{r, code = FALSE}
-#| label: tbl-breed-success
-tbl-cap: !expr "tbl_breed_cap"
-
-breed_mod_summary$coefficients %>% 
-  as_tibble() %>% 
-  mutate(Parameter = c("Intercept", "Age (linear)", "Age (quadratic)"),
-         across(c(Estimate, `Std. Error`, `z value`), 
-                \(x) format_signif(x, digits = 3)),
-         `p value` = format_pval(`Pr(>|z|)`, digits = 2)) %>%
-  select(-`Pr(>|z|)`) %>% 
-  relocate(Parameter) %>% 
-  knitr::kable()
-```
-
-```{r, code = FALSE}
-#| label: tbl-survrecruit-cap
-
-surv_recruit_cap <- str_glue("The results from offspring survival and recruitment analyses.
-In the offspring survival model, the random effect variance of animal id is {format_signif(VarCorr(pup_surv_model)$animalID, 3)} and of year was {format_signif(VarCorr(pup_surv_model)$year, 3)}.
-In the offspring recruitment model, the random effect variance of animal id is {format_signif(VarCorr(pup_recruit_model)$animalID, 3)} and of year was {format_signif(VarCorr(pup_recruit_model)$year, 3)}")
-
-```
-
-Supplemental table for offspring survival
-
-```{r, code = FALSE}
-#| label: tbl-surv_recruit
-tbl-cap: !expr "surv_recruit_cap"
-
-surv_mod_summary <- summary(pup_surv_model)
-surv_mod_kable <- surv_mod_summary$coefficients %>% 
-  as_tibble() %>% 
-  mutate(Parameter = c("Intercept", "Age (linear)", "Age (quadratic)"),
-         across(c(Estimate, `Std. Error`, `z value`), 
-                \(x) format_signif(x, digits = 3)),
-         `p value` = format_pval(`Pr(>|z|)`, digits = 2),
-         Response = "Survival") %>%
-  select(-`Pr(>|z|)`) %>% 
-  relocate(Response, Parameter)
-
-recruit_mod_summary <- summary(pup_recruit_model)
-recruit_mod_kable <- recruit_mod_summary$coefficients %>% 
-  as_tibble() %>% 
-  mutate(Parameter = c("Intercept", "Age (linear)", "Age (quadratic)"),
-         across(c(Estimate, `Std. Error`, `z value`), 
-                \(x) format_signif(x, digits = 3)),
-         `p value` = format_pval(`Pr(>|z|)`, digits = 2),
-         Response = "Recruitment") %>%
-  select(-`Pr(>|z|)`) %>% 
-  relocate(Response, Parameter)
-
-rbind(surv_mod_kable, recruit_mod_kable) %>% 
-  knitr::kable() %>% 
-  kableExtra::kable_styling() %>% 
-  kableExtra::collapse_rows(columns = 1)
-
-```
-
-Supplemental table for pup sex ratio model.
-
-```{r, code = FALSE}
-#| label: tbl-pupsex-cap
-
-pupsex_cap <- str_glue("The results from offspring sex ratio analyses.
-In the model, the random effect variance of animal id is {format_signif(VarCorr(pup_sex_mod)$animalID, 3)} and of year was {format_signif(VarCorr(pup_sex_mod)$year, 3)}.")
-```
-
-```{r, code = FALSE}
-#| label: tbl-pupsex
-tbl-cap: !expr "pupsex_cap"
-
-pupsex_mod_summary <- summary(pup_sex_mod)
-pupsex_quad_summary <- summary(pup_sex_quad)
-pupsex_mod_summary$coefficients %>% 
-  as_tibble() %>% 
-  mutate(Parameter = c("Intercept", "Age (linear)"),
-         across(c(Estimate, `Std. Error`, `z value`), 
-                \(x) format_signif(x, digits = 3)),
-         `p value` = format_pval(`Pr(>|z|)`/2, digits = 3)) %>%
-  select(-`Pr(>|z|)`) %>% 
-  relocate(Parameter) %>% 
-  knitr::kable()
-
-# We fit a quadratic equation, but only the linear term was significant, so the linear model is the one that is plotted.
-take the p = 0.517 and divide by 2 because a priori.
-report in supplemental table that its reporting on 1 tailed (no absolute value of z, just z)
-```
-
-Supplemental table for male biased sex ratios at older ages.
-
-```{r, code = FALSE}
-#| label: tbl-pup-sex
-#| tbl-cap: "Offspring sex ratio skewed towards males with increasing maternal age. Mothers age 12, 13, and 17 gave birth to significantly more male than female pups (bold)."
-
-pup_sex_tbl <- pup_sex_summary %>% 
-  transmute(
-    Age = age,
-    `Male pups` = M,
-    `Total pups` = num,
-    `% male` = scales::label_percent(accuracy = 0.1)(percm),
-    `p value` = format_pval(pval, 3)
-  )
-
-is_signif <- which(pup_sex_tbl$`p value` < "0.05")
-
-knitr::kable(pup_sex_tbl) %>% 
-  kableExtra::kable_styling() %>% 
-  kableExtra::row_spec(is_signif, bold = TRUE)
-```
-
-Supplemental figure for phenology models
-
-```{r, code = FALSE}
-#| fig-width: 8
-
-#NOTE: Both BA and BD have some points that are really late (above 300 doy) that need fixing. Maybe make boxplots also..?
-seal_phenology %>% 
-  rename(`Breeding\narrival` = BA,
-         `Breeding\ndeparture` = BD,
-         `Molt\narrival` = MA,
-         `Molt\ndeparture` = MD) %>% 
-  pivot_longer(`Breeding\narrival`:`Molt\ndeparture`, 
-               names_to = "event", 
-               values_to = "DOY") %>% 
-ggplot(aes(y = DOY)) +
-  geom_boxplot(aes(factor(age))) +
-  geom_smooth(aes(as.integer(factor(age))), method = "lm", formula = y ~ x, se = TRUE) +
-  facet_wrap(~ event, nrow = 2, scales = "free_y") +
-  labs(x = "Age (years)", y = "Day of year") +
-  theme_classic()
-```
-
-Supplemental table for the phenology models.
-
-```{r, code = FALSE}
-ba_data <- phenology_models$model_data[[1]]
-ba_mod <- lmer(doy ~ age + (1 | animalID) + (1 | year), data = ba_data)
-ba_summary <- summary(ba_mod)
-ba_ktable <- ba_summary$coefficients %>% 
-  as_tibble() %>% 
-  mutate(Parameter = c("Intercept", "Age (linear)"),
-         across(c(Estimate, `Std. Error`, `df`, `t value`), 
-                \(x) format_signif(x, digits = 3)),
-         `p value` = format_pval(`Pr(>|t|)`, digits = 2),
-         Response = "Breeding arrival") %>%
-  select(-`Pr(>|t|)`) %>% 
-  relocate(Response, Parameter)
-
-bd_data <- phenology_models$model_data[[2]]
-bd_mod <- lmer(doy ~ age + (1 | animalID) + (1 | year), data = ba_data)
-bd_summary <- summary(ba_mod)
-bd_ktable <- bd_summary$coefficients %>% 
-  as_tibble() %>% 
-  mutate(Parameter = c("Intercept", "Age (linear)"),
-         across(c(Estimate, `Std. Error`, `df`, `t value`), 
-                \(x) format_signif(x, digits = 3)),
-         `p value` = format_pval(`Pr(>|t|)`, digits = 2),
-         Response = "Breeding departure") %>%
-  select(-`Pr(>|t|)`) %>% 
-  relocate(Response, Parameter)
-
-ma_data <- phenology_models$model_data[[3]]
-ma_mod <- lmer(doy ~ age + (1 | animalID) + (1 | year), data = ma_data)
-ma_summary <- summary(ma_mod)
-ma_ktable <- ma_summary$coefficients %>% 
-  as_tibble() %>% 
-  mutate(Parameter = c("Intercept", "Age (linear)"),
-         across(c(Estimate, `Std. Error`, `df`, `t value`), 
-                \(x) format_signif(x, digits = 3)),
-         `p value` = format_pval(`Pr(>|t|)`, digits = 2),
-         Response = "Molt arrival") %>%
-  select(-`Pr(>|t|)`) %>% 
-  relocate(Response, Parameter)
-
-md_data <- phenology_models$model_data[[4]]
-md_mod <- lmer(doy ~ age + (1 | animalID) + (1 | year), data = md_data)
-md_summary <- summary(md_mod)
-md_ktable <- md_summary$coefficients %>% 
-  as_tibble() %>% 
-  mutate(Parameter = c("Intercept", "Age (linear)"),
-         across(c(Estimate, `Std. Error`, `df`, `t value`), 
-                \(x) format_signif(x, digits = 3)),
-         `p value` = format_pval(`Pr(>|t|)`, digits = 2),
-         Response = "Molt departure") %>%
-  select(-`Pr(>|t|)`) %>% 
-  relocate(Response, Parameter)
-
-rbind(ba_ktable, bd_ktable, ma_ktable, md_ktable) %>% 
-  knitr::kable() %>% 
-  kableExtra::kable_styling() %>% 
-  kableExtra::collapse_rows(columns = 1)
-```

### 2024-06-07 paper/paperv3.qmd
phenology

@@
 
+### Phenology
+
+Does the timing of breeding and molting suggest a mechanistic explanation for observed patterns in reproductive and maternal-effect senescence?
+
+We fit a generalized linear mixed-effects model with the duration of a phase (molting or breeding) as the response variable and age, interacting with phase, as the predictor, using year as a random effect.
+We attempted to fit a model with random effects for both year and individual, but that failed to converge.
+We assumed the effect of year on phase duration (molting or breeding) was greater than the effect of individual, so we retained year as the random effect.
+
+$$
+\begin{align}
+d &\sim NB(\mu, k) \\
+\mu &= exp(\eta) \\
+\eta &= \beta_0 + \beta_p * a + Zy
+\end{align}
+$$
+
+Where: $d$ is the duration of the phase in days, $NB$ is the negative binomial distribution with location $\mu$ and dispersion $k$, $\eta$ is the linear predictor, $\beta_0$ is the intercept, $\beta_p$ is the coefficient for age ($a$) by phase (i.e., for $p$ in "molting" or "breeding"), $y$ is the year, and $Z$ is the design matrix relating $y$ to $\eta$.
+
+```{r}
+#| label: pheno-model
+
+seal_phenology <- sealdat %>% 
+  mutate(BA = firstobsbreeddoy, 
+         BD = firstobsbreeddoy + breeddur, 
+         MA = lastobsmoltdoy - moltdur, 
+         MD = lastobsmoltdoy) %>% 
+  drop_na(age, BA, BD, MA, MD)
+
+pheno_model <- seal_phenology %>% 
+  transmute(animalID, year, age10 = age / 10, moltdur, breeddur) %>% 
+  pivot_longer(moltdur:breeddur, 
+               names_to = "phase", 
+               values_to = "dur_days") %>% 
+  glmer.nb(dur_days ~ age10:phase + (1 | year), 
+           data = .)
+
+summary(pheno_model)
+
+```
+
+```{r}
+#| label: fig-pheno
+#| fig-cap: "Molting gets shorter but breeding doesn't.
+A: Observed annual time allocation averaged over individuals.
+B: Raw data for breeding and molting durations (mean ± se) and predictions (mean and 95% CI) from negative binomial weighted by number of seals obersved per year."
+#| fig-width: 9
+
+phenology_n <- unique(seal_phenology$animalID)
+
+timing_summary <- seal_phenology %>% 
+  group_by(age) %>% 
+  summarize(across(BA:MD, mean)) %>% 
+  pivot_longer(cols = -age,  
+               names_to = "event", 
+               values_to = "doy") %>% 
+  group_by(age) %>% 
+  group_modify(\(data, keys) {
+    tibble(event = c("Foraging", "Breeding", "Post-breeding", "Molting", "foraging2"),
+           doy = c(data$doy, 365) - c(0, data$doy))
+  }) %>% 
+  ungroup() %>% 
+  mutate(event = factor(event, levels = c("Foraging", 
+                                          "Breeding", 
+                                          "Post-breeding", 
+                                          "Molting", 
+                                          "foraging2")))
+
+annual_plot <- ggplot(timing_summary, aes(x = age, y = doy, fill = event)) + 
+  geom_area(position = position_stack(reverse = TRUE),
+            color = NA) +
+  geom_line(position = position_stack(reverse = TRUE),
+            color = "white") +
+  scale_fill_discrete(
+    type = c("#829BD1", "#D05794", "#829BD1", "#E09A2D", "#829BD1"),
+    breaks = c("Foraging", "Breeding", "Molting")
+  ) + 
+  scale_x_continuous("Maternal age (years)", expand = expansion(mult = 0)) +
+  scale_y_continuous("Day of year", expand = expansion(mult = 0)) +
+  theme(panel.spacing = unit(1, "lines"),
+        legend.position = "bottom",
+        legend.title = element_blank(),
+        legend.text = element_text(size = 8),
+        legend.key.size = unit(3, "mm"))
+
+n_phenology <- seal_phenology %>% 
+  count(year)
+
+pheno_raw <- seal_phenology %>% 
+  select(breeddur, moltdur, age) %>% 
+  pivot_longer(c(breeddur, moltdur), 
+               names_to = "phase", 
+               values_to = "dur_days") %>% 
+  group_by(age, phase) %>% 
+  summarize(n = n(),
+            dur_days_se = sd(dur_days) / sqrt(n),
+            dur_days = mean(dur_days),
+            dur_days_lwr = dur_days - dur_days_se,
+            dur_days_upr = dur_days + dur_days_se,
+            .groups = "drop")
+
+duration_plot <- ggpredict(
+  pheno_model, 
+  terms = c("age10 [all]", "phase", "year [all]")
+) %>% 
+  as_tibble() %>% 
+  transmute(age = x * 10,
+            dur_days = predicted,
+            dur_days_lwr = conf.low,
+            dur_days_upr = conf.high,
+            phase = group,
+            year = as.numeric(as.character(facet))) %>% 
+  left_join(n_phenology, by = "year") %>% 
+  group_by(age, phase) %>% 
+  summarize(across(starts_with("dur_days"), \(x) weighted.mean(x, n)),
+            .groups = "drop") %>% 
+  ggplot(aes(age)) +
+  geom_ribbon(aes(ymin = dur_days_lwr, ymax = dur_days_upr, fill = phase),
+              alpha = 0.25) +
+  geom_pointrange(aes(y = dur_days, 
+                      ymin = dur_days_lwr, 
+                      ymax = dur_days_upr,
+                      color = phase),
+                  pheno_raw,
+                  position = position_dodge(width = 0.5)) +
+  geom_line(aes(y = dur_days, color = phase)) +
+  annotate("text", 12, 14, label = "Molting", color = "#E09A2D", hjust = 0) +
+  annotate("text", 12, 32, label = "Breeding", color = "#D05794", hjust = 0) +
+  scale_fill_manual(values = c("#D05794", "#E09A2D")) +
+  scale_color_manual(values = c("#D05794", "#E09A2D")) +
+  labs(x = "Maternal age (years)",
+       y = "Phase duration (days)") +
+  theme(legend.position = "none")
+
+(annual_plot | duration_plot) + plot_annotation(tag_levels = "A")
+
+```
+
 ## Supplemental Material

### 2024-06-07 paper/paperv3.qmd
from max's pull request

@@
         legend.title = element_blank(),
-        legend.position.inside = c(0.01, 0.99),
+        # legend.position.inside = c(0.01, 0.99),
         legend.justification = c(0, 1),

### 2024-06-13 paper/paperv3.qmd
Updating color schemes

@@
 
-(sighting_img | age_dist) + plot_layout(widths = c(2, 1))
+(sighting_img | age_dist) + plot_layout(widths = c(2, 1)) + plot_annotation(tag_levels = "A")
 
@@
   scale_y_continuous("Breeding", labels = scales::percent) +
-  scale_color_manual(values = c("cornflowerblue", "firebrick")) +
-  scale_fill_manual(values = c("cornflowerblue", "firebrick")) +
+  scale_color_manual(values = c("#7fbc41", "#de77ae")) +
+  scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
   labs(x = "Age (years)") +
@@
                      limits = c(0, 0.62)) +
-  scale_color_manual(values = c("cornflowerblue", "firebrick")) +
-  scale_fill_manual(values = c("cornflowerblue", "firebrick")) +
+  scale_color_manual(values = c("#7fbc41", "#de77ae")) +
+  scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
   labs(x = "Age (years)") +
@@
   geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
-  scale_color_manual(values = c("cornflowerblue", "firebrick")) +
-  scale_fill_manual(values = c("cornflowerblue", "firebrick")) +
+  scale_color_manual(values = c("#7fbc41", "#de77ae")) +
+  scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
   scale_y_continuous("Male offspring", labels = scales::percent) +
@@
     type = c("#829BD1", "#D05794", "#829BD1", "#E09A2D", "#829BD1"),
-    breaks = c("Foraging", "Breeding", "Molting")
+    breaks = c("Foraging", "Breeding Haulout", "Molt Haulout")
   ) + 
@@
 
-(annual_plot | duration_plot) + plot_annotation(tag_levels = "A")
+(annual_plot | duration_plot) + plot_annotation(tag_levels = "A") + plot_layout(widths = c(1, 2))
 

### 2024-06-13 paper/paperv3.qmd
adding the number of observations to the text
Fixing the text, titles, sample size calculations

@@
 ---
-title: "Elephant Seal Age JAE Intraspecific Variation"
+title: "Reproduction frequency and offspring survival decline in elephant seals past prime age"
 format: 
@@
 
-2.  Reproductive declines persist to the next generation (maternal effect senescence). That is, offspring survival and recruitment are decreased.
+2.  Reproductive declines persist to the next generation (maternal effect senescence). That is, offspring survival and reproduction decrease with maternal age.
 
@@
 
-4.  Phenology is shifted later for older animals and breeders.
+4.  Phenology (breeding haulout and molt haulout) is shifted later for older animals and breeders.
 
@@
 
-First, address detection probability. DO THIS IN MARK.
-
-n = `r n_distinct(sealdat$animalID)` distinct animals.
+n = `r n_distinct(sealdat$animalID)` distinct animals, with `r n_distinct(sealdat)` total observations.
 
 ### Raw data
-
-Resighting plot (image) and age structure bar plot
 
@@
 #| fig-width: 9
+#| fig-cap: "A: Raw longitudinal data for 1,202 known-age female elephant seals and their observations from birth, to recruitment, to death.
+B: Histogram of the number of seals in each age class in the dataset."
 
@@
 
-~~We predicted that breeding success would decline with age. We found that breeding success increases to prime age, then decreases at older ages (@fig-age_dist).~~
+We predicted that breeding success would decline with age.
+Based on the literature, we chose age `r age_senesce` to be the end of prime age and indicate the onset of senescence.
+If breeding success declined with age, we would expect to see a negative difference between the slopes of the two models (pre and post senescence) (@fig-breeding-mod).
 
@@
 
-Sample size n = `r nrow(sealdat)`.
+Sample size n = `r nrow(sealdat)` observations of `r n_distinct(sealdat$animalID)` individuals.
 
@@
 
-We predicted that reproductive declines would carry over to the next generation - that is, pups born to older mothers would have decreased survival and recruitment success.
-
-We found that neither pup survival nor recruitment is significantly affected by maternal age (@fig-repro-age).
-
-We reject our hypothesis: reproductive declines do not persist to the next generation, and there is no evidence of maternal effect senescence.
+We predicted that reproductive declines would carry over to the next generation - that is, pups born to older mothers would have decreased survival (for female and male pups) and reproductive success (for female pups).
+Pup survival significantly decreased with maternal age, but pup reproduction did not (@fig-repro-age).
 
@@
 
-Sample size n = `r nrow(surv_recr_data)`.
+Sample size n = `r nrow(surv_recr_data)` observations of `r n_distinct(surv_recr_data$animalID)` individuals.
 
@@
 
-We found that as maternal age increases, pup sex skews slightly towards males (@fig-sex_ratio).
+We found that as maternal age increases, pup sex skews slightly towards males, but it is not significant (@fig-sex_ratio).
 
@@
 #| label: fig-sex-ratio
+#| fig-cap: "Offspring sex ratio skews towards males with increasing maternal age, but the result is not significant (p = 0.6).
+Black points show the mean and solid black lines show the standard error of the observed sex ratio."
 
@@
 
-Sample size n = `r nrow(pup_sex_data)`.
-In the above figure, the population-level and weighted-average would lie directly on top of each other, so we've excluded the weighted average.
+Sample size n = `r nrow(pup_sex_data)` observations of `r n_distinct(pup_sex_data$animalID)` individuals.
+
+**Note**: In the above figure, the population-level and weighted-average would lie directly on top of each other, so we've excluded the weighted average.
 This is because the random effect of year has very little impact on offspring sex ratio, which makes sense biologically.
@@
   group_modify(\(data, keys) {
-    tibble(event = c("Foraging", "Breeding", "Post-breeding", "Molting", "foraging2"),
+    tibble(event = c("Foraging", "Breeding haulout", "Post-breeding", "Molt haulout", "foraging2"),
            doy = c(data$doy, 365) - c(0, data$doy))
@@
   mutate(event = factor(event, levels = c("Foraging", 
-                                          "Breeding", 
+                                          "Breeding haulout", 
                                           "Post-breeding", 
-                                          "Molting", 
+                                          "Molt haulout", 
                                           "foraging2")))
@@
     type = c("#829BD1", "#D05794", "#829BD1", "#E09A2D", "#829BD1"),
-    breaks = c("Foraging", "Breeding Haulout", "Molt Haulout")
+    breaks = c("Foraging", "Breeding haulout", "Molt haulout")
   ) + 
@@
 
-(annual_plot | duration_plot) + plot_annotation(tag_levels = "A") + plot_layout(widths = c(1, 2))
+(annual_plot | duration_plot) + plot_annotation(tag_levels = "A") + plot_layout(widths = c(1.25, 2))
 

### 2024-06-18 paper/paperv3.qmd
updating phenology n()
Updating sample sizes

@@
 
+n_recruit <- surv_recr_data %>% 
+  filter(pupsex == "F")
+
 # Summarize the raw data
@@
 
-Sample size n = `r nrow(surv_recr_data)` observations of `r n_distinct(surv_recr_data$animalID)` individuals.
+Sample size for offspring survival was n = `r nrow(surv_recr_data)` observations of `r n_distinct(surv_recr_data$animalID)` individuals.
+Sample size for offspring recruitment was n = `r nrow(n_recruit)` observations of `r n_distinct(n_recruit$animalID)` female individuals.
 
@@
 
+Sample size n = `r nrow(seal_phenology)` observations of `r n_distinct(phenology_n)` individuals.
+
 ## Supplemental Material

### 2024-06-18 paper/paperv3.qmd
selective disappearance h1

--- 2024-06-13 paper/paperv3.qmd
@@
                             NA),
-         longevity10 = longevity / 10) %>% 
+         longevity10 = longevity / 10,
+         recruit_age = ifelse(max(year) < 2020 && any(observed_int == 1),
+                              min(age[observed_int == 1]),
+                              NA)) %>% 
   ungroup()
@@
 ```
+
+## Selective appearance and disappearance
+
+```{r}
+#| label: selective_values
+
+recruit_by_6 <- mean(sealdat$recruit_age <= 6, na.rm = TRUE)
+
+```
+
+Age-dependent life history shifts can result from within- or between-animal changes [@vandepol2006].
+We hypothesized age-dependent changes in reproduction were due to senescence, i.e.
+within-animal changes.
+However, it is possible that selective appearance or disappearance of animals in the population can create the same pattern.
+For example, if animals the breed more frequently have lower survival, then older animals will have lower reproductive traits not because of senescence but rather because only infrequent breeders survive to older ages (i.e.
+selective disappearance).
+We tested whether our results were influenced by selective disappearance, following the methods @vandepol2006, by adding a term for longevity to our models.
+We did not test for selective appearance because delayed sexual maturity was very rare in this population.
+`r scales::label_percent(accuracy = 0.1)(recruit_by_6)` of seals recruited at ages 4, 5, or 6.
+
+### H1: Reproductive senescence
+
+```{r}
+#| label: sel-dis-h1
+
+breed_sel_dis <- glmer(
+  observed_int ~ age10 * age_cat + longevity + (1 | animalID) + (1 | year_fct),
+  sealdat,
+  family = "binomial",
+  control = glmerControl(optimizer = "bobyqa")
+)
+
+summary(breed_sel_dis)$coefficients
+
+longevity_est <- summary(breed_sel_dis)$coefficients["longevity", "Estimate"]
+longevity_se <- summary(breed_sel_dis)$coefficients["longevity", "Std. Error"]
+longevity_coef <- sprintf(
+  "%0.5f (95%% CI: %0.5f - %0.5f)", 
+  longevity_est,
+  longevity_est - 1.96 * longevity_se,
+  longevity_est + 1.96 * longevity_se
+)
+
+longevity_p <- sprintf(
+  "%0.3f", 
+  summary(breed_sel_dis)$coefficients["longevity", "Pr(>|z|)"]
+)
+
+```
+
+Selective disappearance did not explain the observed pattern of declining breeding probabilities with age.
+The coefficient for longevity was estimated at `r longevity_coef`, which was not significantly different than 0 ($p$ = `r longevity_p`).
+The predicted breeding probabilities for seals with longevity of 5, 10, and 17 years are shown in @fig-sel-dis-h1.
+
+```{r}
+#| label: fig-sel-dis-h1
+#| fig-cap: Longevity did not explain changes in breeding probabilities with age.
+Predicted breeding probabilities for seals with longevity of 5, 10, and 17 years shown with solid, dotted, and dashed lines, respectively.
+95% confidence intervals shown in shaded regions.
+
+# Create POPULATION-LEVEL (i.e. without rand effects) predictions
+breed_pred_pop <- ggpredict(
+  breed_sel_dis, 
+  terms = c("age10 [all]", "age_cat", "longevity [5,10,17]")
+) %>% 
+  as_tibble() %>% 
+  mutate(age = x * 10,
+         age_cat = factor(group, levels = c("Young", "Old")),
+         longevity = facet,
+         longevity_fct = factor(longevity)) %>% 
+  filter((age_cat == "Young" & age < age_senesce) |
+           (age_cat == "Old" & age >= age_senesce),
+         age <= 20)
+
+ggplot(breed_pred_pop, aes(age, predicted)) +
+  geom_ribbon(aes(fill = age_cat, 
+                  group = interaction(age_cat, longevity_fct),
+                  ymin = conf.low,
+                  ymax = conf.high),
+              alpha = 0.25) +
+  geom_line(aes(color = age_cat, 
+                linetype = longevity_fct)) +
+  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
+  scale_y_continuous("Breeding", 
+                     labels = scales::percent) +
+  scale_color_manual(values = c("#7fbc41", "#de77ae")) +
+  scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
+  labs(x = "Age (years)") +
+  theme(legend.position = "none")
+
+```

### 2024-06-18 paper/paperv3.qmd
selective disappearance h2

@@
 
-## Supplemental Material
+# Supplemental Material
 
-### Models' sensitivity to age of senescence
+## Models' sensitivity to age of senescence
 
@@
 
-### Haul-out duration by reproductive state
+## Haul-out duration by reproductive state
 
@@
 selective disappearance).
-We tested whether our results were influenced by selective disappearance, following the methods @vandepol2006, by adding a term for longevity to our models.
+We tested whether our significant results (reproductive senescence, maternal effect senescence in pup survival) were influenced by selective disappearance, following the methods @vandepol2006, by adding a term for longevity to our models.
 We did not test for selective appearance because delayed sexual maturity was very rare in this population.
@@
 
-summary(breed_sel_dis)$coefficients
-
-longevity_est <- summary(breed_sel_dis)$coefficients["longevity", "Estimate"]
-longevity_se <- summary(breed_sel_dis)$coefficients["longevity", "Std. Error"]
-longevity_coef <- sprintf(
+breed_long_est <- summary(breed_sel_dis)$coefficients["longevity", "Estimate"]
+breed_long_se <- summary(breed_sel_dis)$coefficients["longevity", "Std. Error"]
+breed_long_coef <- sprintf(
   "%0.5f (95%% CI: %0.5f - %0.5f)", 
-  longevity_est,
-  longevity_est - 1.96 * longevity_se,
-  longevity_est + 1.96 * longevity_se
+  breed_long_est,
+  breed_long_est - 1.96 * breed_long_se,
+  breed_long_est + 1.96 * breed_long_se
 )
 
-longevity_p <- sprintf(
+breed_long_p <- sprintf(
   "%0.3f", 
@@
 
+summary(breed_sel_dis)$coefficients
+
 ```
@@
 Selective disappearance did not explain the observed pattern of declining breeding probabilities with age.
-The coefficient for longevity was estimated at `r longevity_coef`, which was not significantly different than 0 ($p$ = `r longevity_p`).
+The coefficient for longevity was estimated at `r breed_long_coef`, which was not significantly different than 0 ($p$ = `r breed_long_p`).
 The predicted breeding probabilities for seals with longevity of 5, 10, and 17 years are shown in @fig-sel-dis-h1.
@@
                   ymax = conf.high),
-              alpha = 0.25) +
+              alpha = 0.2) +
   geom_line(aes(color = age_cat, 
@@
 ```
+
+### H2: Maternal effect senescence
+
+```{r}
+#| label: sel-dis-h2
+
+# Fit pup *survival* model
+surv_sel_dis <- glmer(
+  pup_surv_int ~ age10 * age_cat + longevity + (1 | animalID) + (1 | year_fct),
+  surv_recr_data,
+  family = "binomial",
+  control = glmerControl(optimizer = "bobyqa")
+)
+
+surv_long_est <- summary(surv_sel_dis)$coefficients["longevity", "Estimate"]
+surv_long_se <- summary(surv_sel_dis)$coefficients["longevity", "Std. Error"]
+surv_long_coef <- sprintf(
+  "%0.5f (95%% CI: %0.5f - %0.5f)", 
+  surv_long_est,
+  surv_long_est - 1.96 * surv_long_se,
+  surv_long_est + 1.96 * surv_long_se
+)
+
+surv_long_p <- sprintf(
+  "%0.3f", 
+  summary(surv_sel_dis)$coefficients["longevity", "Pr(>|z|)"]
+)
+
+summary(surv_sel_dis)$coefficients
+
+```
+
+Selective disappearance did not explain the observed pattern of declining offspring survival with maternal age.
+The coefficient for longevity was estimated at `r surv_long_coef`, which was not significantly different than 0 ($p$ = `r surv_long_p`).
+The predicted breeding probabilities for seals with longevity of 5, 10, and 17 years are shown in @fig-sel-dis-h2.
+
+```{r}
+#| label: fig-sel-dis-h2
+#| fig-cap: Longevity did not explain changes in offspring survival or recruitment with maternal age.
+Predicted offspring survival and recruitment probabilities for seals with longevity of 5, 10, and 17 years shown with solid, dotted, and dashed lines, respectively.
+95% confidence intervals shown in shaded regions.
+
+surv_pred_long <- ggpredict(
+  surv_sel_dis, 
+  terms = c("age10 [all]", "age_cat", "longevity [5,10,17]"),
+  type = "fixed",
+  interval = "confidence"
+) %>% 
+  as_tibble() %>% 
+  mutate(age = x * 10,
+         age_cat = group,
+         longevity = facet,
+         longevity_fct = factor(longevity)) %>% 
+  filter((age_cat == "Young" & age < age_senesce) |
+           (age_cat == "Old" & age >= age_senesce),
+         age <= 20)
+
+ggplot(surv_pred_long, aes(age, predicted)) +
+  geom_ribbon(aes(fill = age_cat, 
+                  group = interaction(age_cat, longevity_fct),
+                  ymin = conf.low, 
+                  ymax = conf.high), 
+              alpha = 0.2) +
+  geom_line(aes(color = age_cat, linetype = longevity_fct)) +
+  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
+  scale_y_continuous("Offspring survival", 
+                     labels = scales::percent,
+                     limits = c(0, 0.62)) +
+  scale_color_manual(values = c("#7fbc41", "#de77ae")) +
+  scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
+  labs(x = "Age (years)") +
+  theme(legend.position = "none")
+
+```

### 2024-06-18 paper/paperv3.qmd
saving output of breeding success fig

--- 2024-06-18 paper/paperv3.qmd
@@
                             NA),
-         longevity10 = longevity / 10,
-         recruit_age = ifelse(max(year) < 2020 && any(observed_int == 1),
-                              min(age[observed_int == 1]),
-                              NA)) %>% 
+         longevity10 = longevity / 10) %>% 
   ungroup()
@@
 
+ggsave("outputs/breeding_success.svg", height = 6, width = 10, units = "in")
 ```
@@
 #| fig-width: 9
+
+n_recruit <- surv_recr_data %>% 
+  filter(pupsex == "F")
 
@@
 ```
-
-## Selective appearance and disappearance
-
-```{r}
-#| label: selective_values
-
-recruit_by_6 <- mean(sealdat$recruit_age <= 6, na.rm = TRUE)
-
-```
-
-Age-dependent life history shifts can result from within- or between-animal changes [@vandepol2006].
-We hypothesized age-dependent changes in reproduction were due to senescence, i.e.
-within-animal changes.
-However, it is possible that selective appearance or disappearance of animals in the population can create the same pattern.
-For example, if animals the breed more frequently have lower survival, then older animals will have lower reproductive traits not because of senescence but rather because only infrequent breeders survive to older ages (i.e.
-selective disappearance).
-We tested whether our results were influenced by selective disappearance, following the methods @vandepol2006, by adding a term for longevity to our models.
-We did not test for selective appearance because delayed sexual maturity was very rare in this population.
-`r scales::label_percent(accuracy = 0.1)(recruit_by_6)` of seals recruited at ages 4, 5, or 6.
-
-### H1: Reproductive senescence
-
-```{r}
-#| label: sel-dis-h1
-
-breed_sel_dis <- glmer(
-  observed_int ~ age10 * age_cat + longevity + (1 | animalID) + (1 | year_fct),
-  sealdat,
-  family = "binomial",
-  control = glmerControl(optimizer = "bobyqa")
-)
-
-summary(breed_sel_dis)$coefficients
-
-longevity_est <- summary(breed_sel_dis)$coefficients["longevity", "Estimate"]
-longevity_se <- summary(breed_sel_dis)$coefficients["longevity", "Std. Error"]
-longevity_coef <- sprintf(
-  "%0.5f (95%% CI: %0.5f - %0.5f)", 
-  longevity_est,
-  longevity_est - 1.96 * longevity_se,
-  longevity_est + 1.96 * longevity_se
-)
-
-longevity_p <- sprintf(
-  "%0.3f", 
-  summary(breed_sel_dis)$coefficients["longevity", "Pr(>|z|)"]
-)
-
-```
-
-Selective disappearance did not explain the observed pattern of declining breeding probabilities with age.
-The coefficient for longevity was estimated at `r longevity_coef`, which was not significantly different than 0 ($p$ = `r longevity_p`).
-The predicted breeding probabilities for seals with longevity of 5, 10, and 17 years are shown in @fig-sel-dis-h1.
-
-```{r}
-#| label: fig-sel-dis-h1
-#| fig-cap: Longevity did not explain changes in breeding probabilities with age.
-Predicted breeding probabilities for seals with longevity of 5, 10, and 17 years shown with solid, dotted, and dashed lines, respectively.
-95% confidence intervals shown in shaded regions.
-
-# Create POPULATION-LEVEL (i.e. without rand effects) predictions
-breed_pred_pop <- ggpredict(
-  breed_sel_dis, 
-  terms = c("age10 [all]", "age_cat", "longevity [5,10,17]")
-) %>% 
-  as_tibble() %>% 
-  mutate(age = x * 10,
-         age_cat = factor(group, levels = c("Young", "Old")),
-         longevity = facet,
-         longevity_fct = factor(longevity)) %>% 
-  filter((age_cat == "Young" & age < age_senesce) |
-           (age_cat == "Old" & age >= age_senesce),
-         age <= 20)
-
-ggplot(breed_pred_pop, aes(age, predicted)) +
-  geom_ribbon(aes(fill = age_cat, 
-                  group = interaction(age_cat, longevity_fct),
-                  ymin = conf.low,
-                  ymax = conf.high),
-              alpha = 0.25) +
-  geom_line(aes(color = age_cat, 
-                linetype = longevity_fct)) +
-  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
-  scale_y_continuous("Breeding", 
-                     labels = scales::percent) +
-  scale_color_manual(values = c("#7fbc41", "#de77ae")) +
-  scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
-  labs(x = "Age (years)") +
-  theme(legend.position = "none")
-
-```

### 2024-06-21 paper/paperv3.qmd
outputs for figures

--- 2024-06-18 paper/paperv3.qmd
@@
 breed_age_mod <- glmer(
-  observed_int ~ age10 * age_cat + (1 | animalID) + (1 | year_fct),
+  observed_int ~ age_cat + age10 : age_cat + (1 | animalID) + (1 | year_fct),
   sealdat,

### 2024-06-21 paper/paperv3.qmd
updating plots with dashed lines

@@
               alpha = 0.2) +
-  geom_line(aes(color = age_cat), linewidth = 1.2) +
+  geom_line(aes(color = age_cat), linewidth = 1.2, linetype = "dashed") +
   geom_pointrange(aes(y = perc_recr, 
@@
               alpha = 0.2) +
-  geom_line(aes(color = age_cat), linewidth = 1.2) +
+  geom_line(aes(color = age_cat), linewidth = 1.2, linetype = "dashed") +
   geom_pointrange(aes(y = perc_male, 

### 2024-06-21 paper/paperv3.qmd
more dashed line updates

@@
                   position = position_dodge(width = 0.5)) +
-  geom_line(aes(y = dur_days, color = phase)) +
+  geom_line(aes(y = dur_days, color = phase, linetype = phase)) +
   annotate("text", 12, 14, label = "Molting", color = "#E09A2D", hjust = 0) +
@@
   scale_color_manual(values = c("#D05794", "#E09A2D")) +
+  scale_linetype_manual(values = c("dashed", "solid")) +
   labs(x = "Maternal age (years)",

### 2024-06-22 paper/paperv3.qmd
bayesian segmented regression works!

--- 2024-06-21 paper/paperv3.qmd
@@
          year_fct = factor(year),
-         age10 = age / 10,
+         age10 = (age - age_senesce) / 10,
          age_cat = factor(age >= age_senesce, 
@@
 breed_age_mod <- glmer(
-  observed_int ~ age_cat + age10 : age_cat + (1 | animalID) + (1 | year_fct),
+  observed_int ~ age10 : age_cat + (1 | animalID) + (1 | year_fct),
   sealdat,
@@
   as_tibble() %>% 
-  mutate(age = x * 10,
+  mutate(age = x * 10 + age_senesce,
          age_cat = factor(group, levels = c("Young", "Old"))) %>% 
@@
   as_tibble() %>% 
-  mutate(age = x * 10,
+  mutate(age = x * 10 + age_senesce,
          age_cat = group,
@@
 
-ggsave("outputs/breeding_success.svg", height = 6, width = 10, units = "in")
 ```
@@
 pup_surv_mod <- glmer(
-  pup_surv_int ~ age10 * age_cat + (1 | animalID) + (1 | year_fct),
+  pup_surv_int ~ age10 : age_cat + (1 | animalID) + (1 | year_fct),
   surv_recr_data,
@@
   # In recruit model, only 17 individuals (4%) had >3 pups (female pups only)
-  pup_recr_int ~ age10 * age_cat + (1 | year_fct),
+  pup_recr_int ~ age10 : age_cat + (1 | year_fct),
   filter(surv_recr_data, pupsex == "F"),
@@
   as_tibble() %>% 
-  mutate(age = x * 10,
+  mutate(age = x * 10 + age_senesce,
          age_cat = group) %>% 
@@
 sexrat_age_mod <- glmer(
-  is_male ~ age10 * age_cat + (1 | animalID) + (1 | year_fct),
+  is_male ~ age10 : age_cat + (1 | animalID) + (1 | year_fct),
   pup_sex_data,

### 2024-06-22 cluster/markrecapture.R

@@ -1,247 +1,0 @@
-library(posterior)
-library(tidyverse)
-library(rstan)
-
-args <- commandArgs(trailingOnly = TRUE)
-if (args[1] == "test") {
-  test <- TRUE
-} else if (args[1] == "full") {
-  test <- FALSE
-} else {
-  stop(sprintf("command line argument must be 'test' or 'full', not %s", args[1]))
-}
-
-sealdat <- read_csv("data/raw/128L pull 2023_12_05.csv") %>% 
-  filter(age > 3,
-         year < 2023)
-
-# int<lower=0> nind;
-# int<lower=0> n_occasions;
-# array[nind, n_occasions] int<lower=1, upper=3> y;                   
-
-animalIDs <- sort(unique(sealdat$animalID))
-years <- sort(unique(sealdat$year))
-n_indiv <- length(animalIDs)
-n_occ <- length(years)
-birth_years <- sealdat %>% 
-  arrange(animalID) %>%
-  distinct(animalID, yearborn) %>%
-  pull(yearborn) - min(years)
-
-sealobs <- matrix(0, nrow = n_indiv, ncol = n_occ)
-
-for (i in seq(n_indiv)) {
-  for (j in seq(n_occ)) {
-    animalID <- animalIDs[i]
-    year <- years[j]
-    obs <- sealdat$observed[sealdat$animalID == animalID & sealdat$year == year]
-    #workaround for some seals that have multiple observations in a year
-    if (length(obs) > 1)
-      obs <- "B"
-    if (length(obs) == 0)
-      sealobs[i, j] <- 3
-    else if (obs == "B")
-      sealobs[i, j] <- 1 
-    else 
-      sealobs[i, j] <- 2
-  }
-}
-
-rownames(sealobs) <- animalIDs
-colnames(sealobs) <- years
-
-# Seal resight map
-image(years, seq_along(animalIDs), t(sealobs))
-
-# Test parameters
-n_indiv_test <- 100
-animalIDs_test <- sample(animalIDs, n_indiv_test)
-
-# Setting up directory to save plots
-output_path <- args[2]
-dir.create(output_path)
-
-# Basic model -------------------------------------------------------------
-
-print("BEGINNING BASIC MODEL")
-
-# Fit the dang model
-if (test) {
-  markrecapdat <- list(
-    nind = n_indiv_test, 
-    n_occasions = n_occ, 
-    y = sealobs[animalIDs %in% animalIDs_test, ]
-  )
[…]

### 2024-06-22 cluster/setup.sh

@@ -1,15 +1,0 @@
-# You should have already cloned the reo
-# git clone https://github.com/FlukeAndFeather/sealaging
-
-# Set up conda environment
-module load miniconda3.9
-conda create -n sealaging r-essentials r-base
-
-# Install packages
-conda activate sealaging
-conda install r-posterior
-conda install r-rstan
-
-# Clone repo
-echo Final step: transfer data files from your computer.
-echo scp -r \<path-to-data\> $USER@hb.ucsc.edu:sealaging/data

### 2024-06-22 scratch/128L 2023_06_25_Plots.r

@@ -1,309 +1,0 @@
-#FIGURE CHANGES TO MAKE
-#Create annual cycle figure
-
-#################################################################################
-
-final <- read_csv(here::here("data/raw/128L pull 2023_06_29.csv"))
-final$observed[final$observed=="B"]="Breeder"
-final$observed[final$observed==1]="Non-Breeder"
-
-library(tidyverse)
-library(ggthemes)
-library(emmeans)
-
-##### MOLT DURATION
-agg=final %>% group_by(age,observed) %>%
-  summarize(mean=mean(moltdur,na.rm=TRUE),
-            sd=sd(moltdur,na.rm=TRUE),
-            n=n(),
-            se=sd/sqrt(n))
-
-ggplot(data=agg,aes(x=age,y=mean,group=observed,col=observed))+
-  geom_point(aes(col=observed))+
-  geom_smooth(method="lm",mapping=aes(weight=n))+
-  geom_errorbar(aes(ymin=mean-se, ymax=mean+se,col=observed), width=.2)+
-  theme_few()+
-  ylab("Molt Haulout Duration (Mean +- SE)")+
-  geom_text(data=subset(agg,observed=="Breeder"),aes(label=n,vjust=1,
-                                                     col=observed),y=.03) + 
-  geom_text(data=subset(agg,observed=="Non-Breeder"),aes(label=n,vjust=-1,
-                                                         col=observed),y=.03) 
-require(lmerTest)
-#summary(lm(as.numeric(moltdur)~age+observed,data=final))
-summary(lmer(as.numeric(moltdur)~age+observed+(1|animalID)+(1|year),data=final))
-#summary(lm(as.numeric(moltdur)~age*observed,data=final))
-summary(lmer(as.numeric(moltdur)~age*observed+(1|animalID)+(1|year),data=final))
-
-#Molting Date
-agg=final %>% group_by(age,observed) %>%
-  subset(!is.na(lastobsmoltdoy)) %>% 
-  summarize(mean=mean(lastobsmoltdoy),
-            sd=sd(lastobsmoltdoy),
-            n=n(),
-            se=sd/sqrt(n))
-
-ggplot(data=agg,aes(x=age,y=mean,group=observed,col=observed))+
-  geom_jitter(data=final,aes(x=age,y=lastobsmoltdoy),alpha=0.02)+
-  geom_point(size=3)+
[…]

### 2024-06-22 scratch/changepoint.R

@@ -1,0 +1,45 @@
+library(brms)
+library(tidyverse)
+
+sealdat <- read_csv(here::here("data/raw/128L pull 2023_12_05.csv"), 
+                    show_col_types = FALSE) %>% 
+  mutate(observed = if_else(observed == "B", "Breeder", "Non-breeder"), 
+         observed_int = if_else(observed == "Breeder", 1, 0), 
+         pup_survived = pupseeneveragain > 0) %>% 
+  filter(age > 3, 
+         year > 1987) %>% 
+  mutate(animalID = factor(animalID),
+         year_fct = factor(year),
+         age10 = age / 10) %>% 
+  group_by(animalID) %>% 
+  mutate(longevity = ifelse(max(year) < 2020,
+                            max(age),
+                            NA),
+         longevity10 = longevity / 10,
+         recruit_age = ifelse(max(year) < 2020 && any(observed_int == 1),
+                              min(age[observed_int == 1]),
+                              NA)) %>% 
+  ungroup() %>% 
+  mutate(age2 = (age - 4) / 10) 
+
+bform <- bf(
+  # intercept and fixed effect of longevity
+  observed_int ~ b0 + bl * longevity +
+    # pre-change slope
+    b1 * (age2 - omega) * step(omega - age2) +
+    # post-change slope
+    b2 * (age2 - omega) * step(age2 - omega),
+  # intercept varies by animal and year
+  b0 ~ 1 + (1 | animalID + year),
+  # estimate longevity coef and change point for population
+  bl + b1 + b2 + omega ~ 1,
+  nl = TRUE
+)
+
+bprior <- prior(normal(0, 3), nlpar = "b0") +
+  prior(normal(0, 3), nlpar = "b1") +
+  prior(normal(0, 3), nlpar = "b2") +
+  prior(normal(0, 3), nlpar = "bl") + 
+  prior(uniform(0, 1.8), nlpar = "omega")
+
+fit <- brm(bform, data = sealdat, prior = bprior, chains = 4, iter = 10000)

### 2024-06-22 scratch/paternity.R

@@ -1,53 +1,0 @@
-# Is this the paper you do with Millie?
-
-seal_agent <- function(
-    # Parameterize mom's life history here
-    ) {
-  seal <- list(
-    age = 3,
-    sex = 0, # female
-    pups = 0,
-    grandpups = 0,
-    recruits = 0,
-    alive = TRUE
-  )
-  while (seal$alive) {
-    seal <- seal_agent_year(seal)
-  }
-  seal
-}
-
-seal_agent_year <- function(seal) {
-  survived <- rbinom(1, 1, some_probability(seal$age)) # from the age paper
-  bred <- rbinom(1, 1, some_probability(seal$age)) # from the age paper
-  if (bred == 1) {
-    pup = list(
-      # this is the strategy switching part
-      sex = rbinom(1, 1, some_probability(seal$age))
-    )
-    if (pup$sex == 0) { 
-      # if female
-      pup$recruited <- rbinom(1, 1, some_probability)
-      pup$lifetime_repro <- some_distribution() # parameterize using whole dataset?
-    } else {
-      # if male
-      pup$recruited <- one_in_nine_maybe() # ???
-      pup$lifetime_repro <- any_clue() # ???
-    }
-    seal$pups <- seal$pups + 1
-    seal$recruits <- seal$recruits + pup$recruited
-    seal$grandpups <- seal$grandpups + pup$lifetime_repro
-  }
-  if (survived) {
-    seal$age <- seal$age + 1
-  } else {
-    seal$alive <- FALSE
-  }
-  seal
-}
-
-##pup sex ratio change - taking on a riskier strategy as they get older. well established phenomenoon! 
-#seems to be a case of late-in-life change towards a riskier strategy, not maternal effect senescence
-#captial breeder? 
-#fitness benefits depend on the risk reward ratio for male vs female. we dont know this yet because we only track females
-#future work on paternity would allow us to quantitatively test optimal life history strategies in this model system

### 2024-06-22 scratch/phen_predict.R

@@ -1,20 +1,0 @@
-
-expand_grid(
-  age = 6:16) %>% 
-  mutate(ba = predict(phenology_models$model[[1]],
-                      newdata = .,
-                      re.form = NA),
-         bd = predict(phenology_models$model[[2]],
-                      newdata = .,
-                      re.form = NA), 
-         ma = predict(phenology_models$model[[3]],
-                      newdata = .,
-                      re.form = NA), 
-         md = predict(phenology_models$model[[4]],
-                      newdata = .,
-                      re.form = NA)) %>% 
-  mutate(short_trip = ma - bd, 
-         molt_dur = md - ma) %>%  
-  filter(age %in% c(6,16)) %>% 
-  view()
-

### 2024-06-22 scratch/sealaging.R

@@ -1,0 +1,79 @@
+library(posterior)
+library(rstan)
+library(tidyverse)
+
+sealdat <- read_csv(here::here("data/raw/128L pull 2023_12_05.csv"), 
+                    show_col_types = FALSE) %>% 
+  mutate(observed = if_else(observed == "B", "Breeder", "Non-breeder"), 
+         observed_int = if_else(observed == "Breeder", 1, 0), 
+         pup_survived = pupseeneveragain > 0) %>% 
+  filter(age > 3, 
+         year > 1987) %>% 
+  mutate(animalID = factor(animalID),
+         year_fct = factor(year),
+         age10 = age / 10) %>% 
+  group_by(animalID) %>% 
+  mutate(longevity = ifelse(max(year) < 2020,
+                            max(age),
+                            NA),
+         longevity10 = longevity / 10,
+         recruit_age = ifelse(max(year) < 2020 && any(observed_int == 1),
+                              min(age[observed_int == 1]),
+                              NA)) %>% 
+  ungroup()
+
+sealdat2 <- sealdat %>% 
+  select(animalID, year, age, observed_int) %>% 
+  mutate(age = (age - mean(age)) / sd(age),
+         across(c(animalID, year), \(x) as.integer(factor(x))))
+fit <- stan("scratch/sealaging.stan", "sealaging",
+            data = list(N = nrow(sealdat2),
+                        A = n_distinct(sealdat2$animalID),
+                        Y = n_distinct(sealdat2$year),
+                        b = sealdat2$observed_int,
+                        k = sealdat2$age,
+                        a = sealdat2$animalID,
+                        y = sealdat2$year),
+            iter = 4000, chains = 4, cores = 4)
+
+fit_draws <- subset_draws(
+  as_draws_df(fit), 
+  variable = c("tau", "beta_k", "alpha", "sigma_u", "sigma_v")
+)
+
+inv_logit <- \(x) exp(x) / (1 + exp(x))
+
+## Where is tau?
+foo <- fit_draws %>% 
+  mutate(tau = tau * sd(sealdat$age) + mean(sealdat$age))
+ggplot(foo, aes(tau)) +
+  geom_density(fill = "grey80") +
+  geom_vline(xintercept = mean(foo$tau), color = "firebrick") +
+  theme_classic()
+
[…]

### 2024-06-22 scratch/sealaging.R
selective disappearance model - not converging

@@ -23,8 +23,9 @@ sealdat <- read_csv(here::here("data/raw/128L pull 2023_12_05.csv"),
   ungroup()
 
 sealdat2 <- sealdat %>% 
-  select(animalID, year, age, observed_int) %>% 
-  mutate(age = (age - mean(age)) / sd(age),
+  select(animalID, year, age, observed_int, longevity) %>% 
+  drop_na(longevity) %>% 
+  mutate(across(c(age, longevity), \(x) (x - mean(x)) / sd(x)),
          across(c(animalID, year), \(x) as.integer(factor(x))))
 fit <- stan("scratch/sealaging.stan", "sealaging",
             data = list(N = nrow(sealdat2),
@@ -33,12 +34,13 @@ fit <- stan("scratch/sealaging.stan", "sealaging",
                         b = sealdat2$observed_int,
                         k = sealdat2$age,
                         a = sealdat2$animalID,
-                        y = sealdat2$year),
-            iter = 4000, chains = 4, cores = 4)
+                        y = sealdat2$year,
+                        l = sealdat2$longevity),
+            iter = 8000, chains = 4, cores = 4)
 
 fit_draws <- subset_draws(
   as_draws_df(fit), 
-  variable = c("tau", "beta_k", "alpha", "sigma_u", "sigma_v")
+  variable = c("alpha", "beta_k", "beta_l", "tau", "sigma_u", "sigma_v")
 )
 
 inv_logit <- \(x) exp(x) / (1 + exp(x))

### 2024-06-25 paper/paperv3.qmd
selective disappearance and partial detection

@@
     warning: false
+    toc: true
 editor: visual
@@
 
-We found that as maternal age increases, pup sex skews slightly towards males, but it is not significant (@fig-sex_ratio).
+We found that as maternal age increases, pup sex skews slightly towards males, but it is not significant (@fig-sex-ratio).
 
@@
 #| label: fig-mod-sens
-#| fig-cap: "Model sensitivity to onset of senescence.
+#| fig-cap: Model sensitivity to onset of senescence.
 Age at onset of senescence on the x-axis, p-value for the coefficient of the interaction (age category = post-senescence) x (response variable) on the y-axis.
@@
 Hypothesis H2a (maternal-effect senescence; offspring survival) was significantly supported for onset of senescence age 11+.
-At age 10, the p-vale for the coefficient of the interaction (age category = post-senescence) x (offspring survival) was 0.051."
+At age 10, the p-vale for the coefficient of the interaction (age category = post-senescence) x (offspring survival) was 0.051.
 
@@
   facet_wrap(~hypothesis)
-
-```
-
-## Haul-out duration by reproductive state
-
-```{r}
-#| label: fig-haul-dur
-
-haul_dur <- drop_na(sealdat, breeddur)
-haul_summ <- haul_dur %>% 
-  group_by(observed) %>% 
-  summarize(breeddur_10 = quantile(breeddur, 0.1),
-            breeddur_lbl = str_glue("10% < {breeddur_10} days"))
-ggplot(haul_dur, aes(breeddur)) +
-  geom_density(aes(fill = observed), alpha = 0.5) +
-  geom_density(aes(color = observed)) +
-  geom_vline(aes(color = observed, xintercept = breeddur_10), 
-             haul_summ) +
-  geom_text(aes(x = breeddur_10 + 0.5, label = breeddur_lbl), 
-            haul_summ, 
-            y = 0.09,
-            hjust = 0) +
-  scale_x_continuous("Haul-out duration (days)", 
-                     limits = c(0, 50)) +
-  scale_color_manual(values = c("darkgreen", "darkorchid")) +
-  scale_fill_manual(values = c("darkgreen", "darkorchid")) +
-  theme(legend.position = "none")
 
@@
 However, it is possible that selective appearance or disappearance of animals in the population can create the same pattern.
-For example, if animals the breed more frequently have lower survival, then older animals will have lower reproductive traits not because of senescence but rather because only infrequent breeders survive to older ages (i.e.
+For example, if animals the breed more frequently have lower survival, then older animals will reproduce less frequently not because of senescence but rather because infrequent breeders disproportionately survive to older ages (i.e.
 selective disappearance).
-We tested whether our significant results (reproductive senescence, maternal effect senescence in pup survival) were influenced by selective disappearance, following the methods @vandepol2006, by adding a term for longevity to our models.
-We did not test for selective appearance because delayed sexual maturity was very rare in this population.
+We tested whether our significant results (reproductive senescence, maternal effect senescence in pup survival) were influenced by selective disappearance, following the methods of @vandepol2006, by adding a term for longevity to our models.
+We did not test for selective *appearance*, which results from variability in age of recruitment, because delayed sexual maturity was very rare in this population.
 `r scales::label_percent(accuracy = 0.1)(recruit_by_6)` of seals recruited at ages 4, 5, or 6.
@@
 breed_sel_dis <- glmer(
-  observed_int ~ age10 * age_cat + longevity + (1 | animalID) + (1 | year_fct),
+  observed_int ~ age10 : age_cat + longevity + (1 | animalID) + (1 | year_fct),
   sealdat,
@@
   as_tibble() %>% 
-  mutate(age = x * 10,
+  mutate(age = x * 10 + age_senesce,
          age_cat = factor(group, levels = c("Young", "Old")),
@@
 surv_sel_dis <- glmer(
-  pup_surv_int ~ age10 * age_cat + longevity + (1 | animalID) + (1 | year_fct),
+  pup_surv_int ~ age10 : age_cat + longevity + (1 | animalID) + (1 | year_fct),
   surv_recr_data,
@@
 ```
+
+## Breeding state and detection probability
+
+Not all seals were detected in all years.
+A non-detection year during the seal's lifespan (i.e.
+an unobserved year between the first and last observed years) could represent an unobserved animal at Año Nuevo or temporary emigration.
+In either case, the animal may or may not have reproduced that year.
+By excluding these unobserved animals, our inferences for H1 (reproductive senescence) may be biased if detectability varies with age and/or by breeding state.
+Here, we assess those potential biases.
+
+```{r}
+#| label: missing-seals
+
+## Are we missing animals more or less with age?
+missing_seals <- sealdat %>% 
+  # Need to see a seal 3+ years for calculation to make sense
+  group_by(animalID) %>% 
+  mutate(n_years = n_distinct(year)) %>% 
+  ungroup() %>% 
+  filter(n_years >= 3,
+         age <= 20) %>% 
+  # Fill in non-resighted years
+  group_by(animalID) %>% 
+  arrange(year) %>% 
+  reframe(tibble(year2 = (min(year) + 1):(max(year) - 1),
+                 unobserved = as.integer(!year2 %in% year),
+                 age = year2 - yearborn[1],
+                 age10 = (age - age_senesce) / 10)) %>% 
+  ungroup() %>% 
+  rename(year = year2) %>% 
+  mutate(age_cat = factor(age >= age_senesce, 
+                          labels = c("Young", "Old")),
+         year_fct = factor(year)) %>% 
+  left_join(select(sealdat, animalID, year, observed_int),
+            by = c("animalID", "year"))
+
+missing_mod <- glmer(unobserved ~ age10 + (1 | year) + (1 | animalID), 
+                     missing_seals, 
+                     family = binomial)
+missing_pred <- ggpredict(missing_mod,
+                          terms = "age10 [all]") %>% 
+  as_tibble() %>% 
+  mutate(age = x * 10 + age_senesce,
+         unobserved = predicted)
+
+missing_age_coef <- sprintf(
+  "%0.3f \u00B1 %0.3f",
+  summary(missing_mod)$coef["age10", "Estimate"],
+  summary(missing_mod)$coef["age10", "Std. Error"] * 1.96
+)
+
+```
+
+First, we tested whether probability of detection was related to age.
+We found the proportion of unobserved animals increased with age.
+For this analysis, we excluded the first and last year of observation for each animal, because animals were necessarily observed in those years.
+For all years between, we determined whether animals were observed.
+Then we estimated the effect of age on probability of an animal being unobserved using a GLMM (binomial response, logit link) with random effects for year and animal ([@fig-unobserved]).
+The coefficient for age was `r missing_age_coef` (*p*\<0.0001).
+
+```{r}
+#| label: fig-unobserved
+#| fig-cap: The probability of an animal being unobserved increased with age.
+Points and error bars represent the proportion of unobserved animals within age classes plus/minus one standard error.
+The blue line and shaded area are the mean and 95% CI probability of an animal being unobserved estimated with a generalized linear mixed effect model.
+See text for model details.
+
+missing_seals %>% 
+  group_by(age) %>% 
+  summarize(unobserved_mean = mean(unobserved),
+            unobserved_se = sqrt(unobserved_mean * (1 - unobserved_mean) / n()),
+            unobserved_lwr = unobserved_mean - unobserved_se,
+            unobserved_upr = unobserved_mean + unobserved_se) %>% 
+  ggplot(aes(age)) +
+  geom_ribbon(aes(ymin = conf.low, ymax = conf.high),
+              missing_pred,
+              alpha = 0.2) +
+  geom_line(aes(y = unobserved), 
+            missing_pred, 
+            color = "cornflowerblue",
+            linewidth = 2) +
+  geom_pointrange(aes(y = unobserved_mean, 
+                      ymin = unobserved_lwr, ymax = unobserved_upr)) +
+  scale_x_continuous("Age", breaks = seq(5, 19, by = 4)) +
+  scale_y_continuous("Unobserved proportion", 
+                     labels = scales::label_percent(accuracy = 1))
+```
+
+```{r}
+#| label: haulout
+
+breed_haulout <- sprintf(
+  "%i - %i", 
+  quantile(sealdat$breeddur[sealdat$observed_int == 1], 0.25, na.rm = TRUE), 
+  quantile(sealdat$breeddur[sealdat$observed_int == 1], 0.75, na.rm = TRUE)
+)
+
+non_breed_haulout <- sprintf(
+  "%i - %i", 
+  quantile(sealdat$breeddur[sealdat$observed_int == 0], 0.25, na.rm = TRUE), 
+  quantile(sealdat$breeddur[sealdat$observed_int == 0], 0.75, na.rm = TRUE)
+)
+
+resight_by_observed <- resight %>% 
+  mutate(animalID = as.character(animalID)) %>% 
+  right_join(select(sealdat, animalID, year, observed_int), 
+             by = c("animalID", season = "year"),
+             relationship = "many-to-many") %>% 
+  count(animalID, season, observed_int) %>% 
+  group_by(observed_int) %>% 
+  summarize(n_q1 = quantile(n, 0.25), 
+            n_mean = mean(n), 
+            n_q3 = quantile(n, 0.75))
+
+breed_resight <- with(resight_by_observed[2, ],
+                      sprintf("%0.1f (%i - %i)", n_mean, n_q1, n_q3))
+nonbreed_resight <- with(resight_by_observed[1, ],
+                         sprintf("%0.1f (%i - %i)", n_mean, n_q1, n_q3))
+
+```
+
+The pattern of unobserved animals increasing with age means it is highly unlikely our inferences about reproductive senescence are due to partial detection.
+If there was no reproductive senescence in the population, then the apparent senescence we observed would have to result from older breeders being less detectable.
+However, the biology of northern elephant seals makes breeding a more detectable behavior than non-breeding.
+Animals must come to land for an extended period to give birth and wean their pups.
+Among the animals in this study, breeders hauled out for longer than non-breeders (1st - 3rd quartile: breeders `r breed_haulout` days; non-breeders `r non_breed_haulout` days).
+This extended period of time on the beach makes them more accessible to observers.
+Furthermore, northern elephant seals exhibit a high degree of site fidelity.
+Once they recruit to a breeding colony, temporary emigration is rare (cite me).
+The only plausible explanation for unobserved animals being disproportionately breeders would be crowding at the breeding colonies, which could obscure identifying marks.
+However, breeding animals were resighted `r breed_resight` times per year, compared to `r nonbreed_resight` for non-breeders (mean, 1st - 3rd quartile), which suggests that breeders are more readily observable than non-breeders.
+Therefore, it is biologically implausible that the unobserved seals were disproportionately breeders and our estimates for reproductive senescence are, if anything, conservative.

### 2024-06-27 paper/paperv3.qmd
bioage

--- 2024-06-25 paper/paperv3.qmd
@@
 
+## Chronological versus biological age
+
+```{r}
+#| label: bioage
+
+bioage_senesce <- -7
+sealdat2 <- sealdat %>% 
+  drop_na(longevity) %>% 
+  mutate(bioage = age - longevity,
+         bioage10 = (bioage - bioage_senesce) / 10,
+         bioage_cat = factor(bioage < bioage_senesce,
+                             labels = c("Old", "Young")))
+
+```
+
+The chronological age of an individual may not reflect their biological age because aging rates vary within populations [@martin2011].
+Years-to-death, a proxy for biological age, may therefore be a better predictor for senescence patterns than chronological age.
+We repeated our analysis for the reproductive senescence hypothesis using biological age, setting the breakpoint for senescence at biological age -7 years (i.e.
+7 years prior to final observation), based on visual inspection of the raw data.
+This required us to limit our sample to animals with known longevity (i.e., last observation in 2020 or earlier), which reduced our sample size to `r nrow(sealdat2)` observations of `r n_distinct(sealdat2$animalID)` distinct animals.
+Subsequently, we fit the model for breakpoints at all biological ages -10 to -3 years, to assess whether our results were sensitive to the choice of breakpoint.
+Finally, we re-fit the chronological reproductive senescence model using the known-longevity dataset and compared chronological versus biological age models using AIC.
+
+```{r}
+#| label: bioage-mod
+
+breed_bioage_mod <- glmer(
+  observed_int ~ bioage10 : bioage_cat + (1 | animalID) + (1 | year_fct),
+  sealdat2,
+  family = "binomial",
+  control = glmerControl(optimizer = "bobyqa")
+)
+
+summary(breed_bioage_mod)$coefficients
+
+bioage_est <- summary(breed_bioage_mod)$coef["bioage10:bioage_catOld", "Estimate"]
+bioage_se <- summary(breed_bioage_mod)$coef["bioage10:bioage_catOld", "Std. Error"]
+bioage_coef <- sprintf("%0.3f (%0.3f - %0.3f)", 
+                       bioage_est,
+                       bioage_est - 1.96 * bioage_se,
+                       bioage_est + 1.96 * bioage_se)
+
+bioage_pval <- sprintf("%0.3f", summary(breed_bioage_mod)$coef["bioage10:bioage_catOld", "Pr(>|z|)"])
+
+breed_age_mod2 <- glmer(
+  observed_int ~ age10 : age_cat + (1 | animalID) + (1 | year_fct),
+  drop_na(sealdat, longevity),
+  family = "binomial",
+  control = glmerControl(optimizer = "bobyqa")
+)
+
+p_by_tau <- map_dbl(-10:-3, \(tau) {
+  mod <- sealdat %>%
+    drop_na(longevity) %>%
+    mutate(bioage = age - longevity,
+           bioage10 = (bioage - tau) / 10,
+           bioage_cat = factor(bioage < tau,
+                               labels = c("Old", "Young"))) %>% 
+    glmer(
+      observed_int ~ bioage10 : bioage_cat + (1 | animalID) + (1 | year_fct),
+      .,
+      family = "binomial",
+      control = glmerControl(optimizer = "bobyqa")
+    )
+  summary(mod)$coef["bioage10:bioage_catOld", "Pr(>|z|)"]
+})
+
+chrono_bio_aic <- AIC(breed_bioage_mod, breed_age_mod2)
+daic <- chrono_bio_aic$AIC - min(chrono_bio_aic$AIC)
+chrono_bio_rellik <- exp(-0.5 * daic)
+chrono_bio_aicw <- chrono_bio_rellik / sum(chrono_bio_rellik)
+
+```
+
+```{r}
+#| label: fig-bio-age
+#| fig-cap: Breeding probability over biological age.
+Points and error bars are the mean ± SE of observed breeding proportions within age classes.
+The vertical dashed line indicates the threshold for senescence.
+Solid lines and ribbons are the mean and 95% CI of the population-level breeding probability, estimated by a GLMM with random effects for individual and year.
+The dashed lines represent the mean of yearly breeding probabilities, weighted by the number of seals observed each year, which should match the observed data more closely than the population-level probabilities.
+
+# Summarize the raw data
+bioage_summ <- sealdat2 %>% 
+  group_by(bioage) %>% 
+  summarize(n_obs = n(), 
+            perc_breed = mean(observed_int), 
+            se = sqrt(perc_breed * (1 - perc_breed) / n_obs))
+
+# Create POPULATION-LEVEL (i.e. without rand effects) predictions
+bioage_pred_pop <- ggpredict(breed_bioage_mod, 
+                             terms = c("bioage10 [all]", "bioage_cat")) %>% 
+  as_tibble() %>% 
+  mutate(bioage = x * 10 + bioage_senesce,
+         bioage_cat = factor(group, levels = c("Young", "Old"))) %>% 
+  filter((bioage_cat == "Young" & bioage < bioage_senesce) |
+           (bioage_cat == "Old" & bioage >= bioage_senesce))
+
+# Create WEIGHTED-AVERAGE predictions using years' random effects
+bioage_pred_wgt <- ggpredict(
+  breed_bioage_mod,
+  terms = c("bioage10 [all]", "bioage_cat", "year_fct [all]"),
+  type = "random",
+  interval = "confidence"
+) %>% 
+  as_tibble() %>% 
+  mutate(bioage = x * 10 + bioage_senesce,
+         bioage_cat = factor(group, levels = c("Young", "Old")),
+         year_fct = facet) %>% 
+  filter((bioage_cat == "Young" & bioage < bioage_senesce) |
+           (bioage_cat == "Old" & bioage >= bioage_senesce)) %>% 
+  left_join(seal_years, by = "year_fct") %>% 
+  group_by(bioage, bioage_cat) %>% 
+  summarize(predicted = weighted.mean(predicted, n),
+            .groups = "drop")
+
+ggplot(bioage_pred_pop, aes(bioage, predicted)) +
+  geom_ribbon(aes(fill = bioage_cat, ymin = conf.low, ymax = conf.high), 
+              alpha = 0.2) +
+  geom_line(aes(color = bioage_cat), linewidth = 1.2) +
+  geom_line(aes(color = bioage_cat),
+            bioage_pred_wgt, 
+            linewidth = 1.2,
+            linetype = "dashed") +
+  geom_pointrange(aes(y = perc_breed, 
+                      ymin = perc_breed - se, 
+                      ymax = perc_breed + se),
+                  bioage_summ) +
+  geom_vline(xintercept = bioage_senesce - 0.5, linetype = "dashed") +
+  scale_y_continuous("Breeding", labels = scales::percent) +
+  scale_color_manual(values = c("#7fbc41", "#de77ae")) +
+  scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
+  labs(x = "Biological age (years-to-death)") +
+  theme(legend.position = "none")
+
+```
+
+The biological age model did not support the reproductive senescence hypothesis.
+The coefficient for biological age among older animals (biological age \> -7 years) was `r bioage_coef` (95% CI) which was not significantly different than 0 (*p* = `r bioage_pval`).
+Choice of breakpoint for senescence did not affect the significance of this coefficient.
+For breakpoints -10 to -3 years, *p* reached a minimum at biological age -7 years.
+
+The chronological age model outperformed the biological age model according to AIC.
+The difference in AIC between the two models was `r sprintf("%0.1f", -diff(chrono_bio_aic$AIC))` and the chronological age model had an AIC weight of `r sprintf("%0.3f", chrono_bio_aicw[2])`.
+
+Based on these results, we find there is evidence to support the reproductive senescence hypothesis based on chronological age.
+The biological age proxy, years-to-death, should be a better predictor of senescence than chronological age if most mortality is age-related.
+However, adult elephant seal mortality is largely influenced by extrinsic factors, primarily food availability [@colegrove2005; @holser2021].
+Poor foraging conditions during El Niño years, for example, may increase mortality rates among animals whose biological age is not yet advanced.
+As result, the years-to-death proxy for biological age likely conflates many biologically young and old animals that died due to extrinsic factors.
+More accurate measures of biological age, such as telomere length [@jylhävä2017], may be necessary to investigate senescence in elephant seals and other species with high extrinsic sources of mortality from a non-chronological perspective.
+
 ## Breeding state and detection probability

### 2024-06-27 paper/paperv3.qmd
Changing some line types and colors

--- 2024-06-25 paper/paperv3.qmd
@@
 The solid lines and shaded areas represent the mean and 95% confidence interval of the population-level model predictions.
-The observed data are unequally distributed across years, so the dashed lines indicate the predicted breeding rates averaged over years (i.e., including the random effect of year), weighted by the number of seals observed in each year."
+The observed data are unequally distributed across years, so the dotted lines indicate the predicted breeding rates averaged over years (i.e., including the random effect of year), weighted by the number of seals observed in each year."
 
@@
             linewidth = 1.2,
-            linetype = "dashed") +
+            linetype = "dotted") +
   geom_pointrange(aes(y = perc_breed, 
@@
               missing_pred,
-              alpha = 0.2) +
+              alpha = 0.2, 
+              fill = "#de77ae") +
   geom_line(aes(y = unobserved), 
             missing_pred, 
-            color = "cornflowerblue",
+            color = "#de77ae",
             linewidth = 2) +

### 2024-07-08 paper/paperv3.qmd
Selective appearance + disappearance

--- 2024-06-27 paper/paperv3.qmd
@@
   group_by(animalID) %>% 
-  mutate(longevity = ifelse(max(year) < 2020,
+  mutate(firstrepro = min(age),
+         lastrepro = ifelse(max(year) < 2020,
                             max(age),
                             NA),
-         longevity10 = longevity / 10,
+         lastrepro10 = lastrepro / 10,
          recruit_age = ifelse(max(year) < 2020 && any(observed_int == 1),
@@
   group_by(animalID) %>% 
-  mutate(longevity = max(age),
+  mutate(lastrepro = max(age),
          yearborn = na.omit(yearborn)[1]) %>% 
   ungroup() %>% 
-  arrange(yearborn, longevity, animalID) %>% 
+  arrange(yearborn, lastrepro, animalID) %>% 
   mutate(animalID = factor(animalID, levels = unique(animalID))) %>% 
@@
 The solid lines and shaded areas represent the mean and 95% confidence interval of the population-level model predictions.
-The observed data are unequally distributed across years, so the dotted lines indicate the predicted breeding rates averaged over years (i.e., including the random effect of year), weighted by the number of seals observed in each year."
+The observed data are unequally distributed across years, so the dashed lines indicate the predicted breeding rates averaged over years (i.e., including the random effect of year), weighted by the number of seals observed in each year."
 
@@
             linewidth = 1.2,
-            linetype = "dotted") +
+            linetype = "dashed") +
   geom_pointrange(aes(y = perc_breed, 
@@
 selective disappearance).
-We tested whether our significant results (reproductive senescence, maternal effect senescence in pup survival) were influenced by selective disappearance, following the methods of @vandepol2006, by adding a term for longevity to our models.
-We did not test for selective *appearance*, which results from variability in age of recruitment, because delayed sexual maturity was very rare in this population.
-`r scales::label_percent(accuracy = 0.1)(recruit_by_6)` of seals recruited at ages 4, 5, or 6.
+We tested whether our significant results (reproductive senescence, maternal effect senescence in pup survival) were influenced by selective appearance and disappearance, following the methods of @vandepol2006, by adding terms for age at first and last reproduction, respectively.
+We used AIC to compare the base model to models with selective appearance and disappearance.
 
@@
 ```{r}
-#| label: sel-dis-h1
-
-breed_sel_dis <- glmer(
-  observed_int ~ age10 : age_cat + longevity + (1 | animalID) + (1 | year_fct),
-  sealdat,
+#| label: selective-h1
+
+h1_selective_models <- map(
+  list(
+    # Base model
+    observed_int ~ age10 : age_cat + (1 | animalID) + (1 | year_fct),
+    # Selective appearance
+    observed_int ~ age10 : age_cat + firstrepro + 
+      (1 | animalID) + (1 | year_fct),
+    # Selective disappearance
+    observed_int ~ age10 : age_cat + lastrepro + 
+      (1 | animalID) + (1 | year_fct),
+    # Selective appearance and disappearance
+    observed_int ~ age10 : age_cat + firstrepro + lastrepro +
+      (1 | animalID) + (1 | year_fct)
+  ),
+  \(form) {
+    glmer(form, 
+          drop_na(sealdat, lastrepro), 
           family = "binomial", 
-  control = glmerControl(optimizer = "bobyqa")
-)
-
-breed_long_est <- summary(breed_sel_dis)$coefficients["longevity", "Estimate"]
-breed_long_se <- summary(breed_sel_dis)$coefficients["longevity", "Std. Error"]
-breed_long_coef <- sprintf(
-  "%0.5f (95%% CI: %0.5f - %0.5f)", 
-  breed_long_est,
-  breed_long_est - 1.96 * breed_long_se,
-  breed_long_est + 1.96 * breed_long_se
-)
-
-breed_long_p <- sprintf(
-  "%0.3f", 
-  summary(breed_sel_dis)$coefficients["longevity", "Pr(>|z|)"]
-)
-
-summary(breed_sel_dis)$coefficients
-
-```
-
-Selective disappearance did not explain the observed pattern of declining breeding probabilities with age.
-The coefficient for longevity was estimated at `r breed_long_coef`, which was not significantly different than 0 ($p$ = `r breed_long_p`).
-The predicted breeding probabilities for seals with longevity of 5, 10, and 17 years are shown in @fig-sel-dis-h1.
-
-```{r}
-#| label: fig-sel-dis-h1
-#| fig-cap: Longevity did not explain changes in breeding probabilities with age.
-Predicted breeding probabilities for seals with longevity of 5, 10, and 17 years shown with solid, dotted, and dashed lines, respectively.
-95% confidence intervals shown in shaded regions.
-
-# Create POPULATION-LEVEL (i.e. without rand effects) predictions
-breed_pred_pop <- ggpredict(
-  breed_sel_dis, 
-  terms = c("age10 [all]", "age_cat", "longevity [5,10,17]")
+          control = glmerControl(optimizer = "bobyqa"))
+  }
+)
+
+do.call(AIC, h1_selective_models) %>% 
+  as_tibble() %>% 
+  mutate(model = c("Base", 
+                   "Selective appearance", 
+                   "Selective disappearance", 
+                   "Selective appearance and disappearance"),
+         dAIC = AIC - min(AIC),
+         rellik = exp(-0.5 * dAIC),
+         weight = rellik / sum(rellik)) %>% 
+  relocate(model)
+```
+
+```{r}
+#| label: fig-selective-h1
+#| fig-cap: Predicted breeding probabilities of (**A**) selective appearance and (**B**) selective disappearance models.
+The semi-transparent solid lines are predictions of the base model.
+The dashed and dotted lines represent predictions for the 10th and 90th percentiles of observed age at first (**A**) and last (**B**) reproduction.
+In **A**, these correspond to first reproduction at 4 and 6 years old, respectively.
+In **B**, these correspond to last reproduction at 5 and 17 years old, respectively.
+Confidence intervals not shown for clarity.
+
+# Compare model predictions
+base_pred <- ggpredict(
+  h1_selective_models[[1]],
+  terms = c("age10 [all]", "age_cat")
 ) %>% 
@@
   mutate(age = x * 10 + age_senesce,
-         age_cat = factor(group, levels = c("Young", "Old")),
-         longevity = facet,
-         longevity_fct = factor(longevity)) %>% 
+         age_cat = factor(group, levels = c("Young", "Old"))) %>% 
   filter((age_cat == "Young" & age < age_senesce) |
@@
 
-ggplot(breed_pred_pop, aes(age, predicted)) +
-  geom_ribbon(aes(fill = age_cat, 
-                  group = interaction(age_cat, longevity_fct),
-                  ymin = conf.low,
-                  ymax = conf.high),
-              alpha = 0.2) +
+# selective appearance for firstrepro=4,6 (10% and 90% of observed)
+selapp_pred <- ggpredict(
+  h1_selective_models[[2]],
+  terms = c("age10 [all]", "age_cat", "firstrepro [4,6]")
+) %>% 
+  as_tibble() %>% 
+  mutate(age = x * 10 + age_senesce,
+         age_cat = factor(group, levels = c("Young", "Old")),
+         firstrepro = facet) %>% 
+  filter((age_cat == "Young" & age < age_senesce) |
+           (age_cat == "Old" & age >= age_senesce),
+         age <= 20)
+
+# selective disappearance for lastrepro=5,17 (10% and 90% of observed) 
+seldis_pred <- ggpredict(
+  h1_selective_models[[4]],
+  terms = c("age10 [all]", "age_cat", "lastrepro [5,17]")
+) %>% 
+  as_tibble() %>% 
+  mutate(age = x * 10 + age_senesce,
+         age_cat = factor(group, levels = c("Young", "Old")),
+         lastrepro = facet) %>% 
+  filter((age_cat == "Young" & age < age_senesce) |
+           (age_cat == "Old" & age >= age_senesce),
+         age <= 20)
+
+# Selective appearance predictions
+fig_selapp <- ggplot(base_pred, aes(age, predicted)) +
+  # Base predictions
+  geom_line(aes(color = age_cat), alpha = 0.5) +
+  # Selective appearance ages 4 and 6
   geom_line(aes(color = age_cat, 
-                linetype = longevity_fct)) +
+                linetype = firstrepro),
+            selapp_pred) +
   geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
@@
   scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
+  scale_linetype_manual(values = c("dashed", "dotted")) +
   labs(x = "Age (years)") +
@@
 
-```
+fig_seldis <- ggplot(base_pred, aes(age, predicted)) +
+  # Base predictions
+  geom_line(aes(color = age_cat), alpha = 0.5) +
+  # Selective disappearance ages 5 and 17
+  geom_line(aes(color = age_cat, 
+                linetype = lastrepro),
+            seldis_pred) +
+  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
+  scale_y_continuous("Breeding", 
+                     labels = scales::percent) +
+  scale_color_manual(values = c("#7fbc41", "#de77ae")) +
+  scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
+  scale_linetype_manual(values = c("dashed", "dotted")) +
+  labs(x = "Age (years)") +
+  theme(legend.position = "none")
+
+(fig_selapp | fig_seldis) +
+  plot_annotation(tag_levels = "A")
+```
+
+Selective appearance and disappearance did not explain the observed pattern of declining breeding probabilities with age.
+The AIC weights for the base, selective appearance, selective disappearance, and combined models were 0.526, 0.197, 0.201, and 0.076, respectively.
+These weights indicate the most likely explanation for the observed pattern is within-individual change, rather than selective appearance or disappearance.
+In the selective appearance and disappearance models, the coefficients for age at first reproduction (95% CI: -0.091 - 0.112) and last reproduction (95% CI: -0.037 - 0.051) were not significantly different than 0, respectively.
+Furthermore, predictions of the selective appearance and disappearance models did not meaningfully differ from the base model @fig-selective-h1.
 
@@
 ```{r}
-#| label: sel-dis-h2
-
-# Fit pup *survival* model
-surv_sel_dis <- glmer(
-  pup_surv_int ~ age10 : age_cat + longevity + (1 | animalID) + (1 | year_fct),
-  surv_recr_data,
+#| label: selective-h2
+
+h2_selective_models <- map(
+  list(
+    # Base model
+    pup_surv_int ~ age10 : age_cat + (1 | animalID) + (1 | year_fct),
+    # Selective appearance
+    pup_surv_int ~ age10 : age_cat + firstrepro + 
+      (1 | animalID) + (1 | year_fct),
+    # Selective disappearance
+    pup_surv_int ~ age10 : age_cat + lastrepro + 
+      (1 | animalID) + (1 | year_fct),
+    # Selective appearance and disappearance
+    pup_surv_int ~ age10 : age_cat + firstrepro + lastrepro +
+      (1 | animalID) + (1 | year_fct)
+  ),
+  \(form) {
+    glmer(form, 
+          drop_na(surv_recr_data, lastrepro), 
           family = "binomial", 
-  control = glmerControl(optimizer = "bobyqa")
-)
-
-surv_long_est <- summary(surv_sel_dis)$coefficients["longevity", "Estimate"]
-surv_long_se <- summary(surv_sel_dis)$coefficients["longevity", "Std. Error"]
-surv_long_coef <- sprintf(
-  "%0.5f (95%% CI: %0.5f - %0.5f)", 
-  surv_long_est,
-  surv_long_est - 1.96 * surv_long_se,
-  surv_long_est + 1.96 * surv_long_se
-)
-
-surv_long_p <- sprintf(
-  "%0.3f", 
-  summary(surv_sel_dis)$coefficients["longevity", "Pr(>|z|)"]
-)
-
-summary(surv_sel_dis)$coefficients
-
-```
-
-Selective disappearance did not explain the observed pattern of declining offspring survival with maternal age.
-The coefficient for longevity was estimated at `r surv_long_coef`, which was not significantly different than 0 ($p$ = `r surv_long_p`).
-The predicted breeding probabilities for seals with longevity of 5, 10, and 17 years are shown in @fig-sel-dis-h2.
-
-```{r}
-#| label: fig-sel-dis-h2
-#| fig-cap: Longevity did not explain changes in offspring survival or recruitment with maternal age.
-Predicted offspring survival and recruitment probabilities for seals with longevity of 5, 10, and 17 years shown with solid, dotted, and dashed lines, respectively.
-95% confidence intervals shown in shaded regions.
-
-surv_pred_long <- ggpredict(
-  surv_sel_dis, 
-  terms = c("age10 [all]", "age_cat", "longevity [5,10,17]"),
-  type = "fixed",
-  interval = "confidence"
+          control = glmerControl(optimizer = "bobyqa"))
+  }
+)
+
+do.call(AIC, h2_selective_models) %>% 
+  as_tibble() %>% 
+  mutate(model = c("Base", 
+                   "Selective appearance", 
+                   "Selective disappearance", 
+                   "Selective appearance and disappearance"),
+         dAIC = AIC - min(AIC),
+         rellik = exp(-0.5 * dAIC),
+         weight = rellik / sum(rellik)) %>% 
+  relocate(model)
+
+```
+
+```{r}
+#| label: fig-selective-h2
+#| fig-cap: Predicted breeding probabilities of (**A**) selective appearance and (**B**) selective disappearance models.
+The semi-transparent solid lines are predictions of the base model.
+The dashed and dotted lines represent predictions for the 10th and 90th percentiles of observed age at first (**A**) and last (**B**) reproduction.
+In **A**, these correspond to first reproduction at 4 and 6 years old, respectively.
+In **B**, these correspond to last reproduction at 5 and 17 years old, respectively.
+Confidence intervals not shown for clarity.
+
+# Compare model predictions
+base_pred <- ggpredict(
+  h2_selective_models[[1]],
+  terms = c("age10 [all]", "age_cat")
 ) %>% 
@@
   mutate(age = x * 10 + age_senesce,
-         age_cat = group,
-         longevity = facet,
-         longevity_fct = factor(longevity)) %>% 
+         age_cat = factor(group, levels = c("Young", "Old"))) %>% 
   filter((age_cat == "Young" & age < age_senesce) |
@@
 
-ggplot(surv_pred_long, aes(age, predicted)) +
-  geom_ribbon(aes(fill = age_cat, 
-                  group = interaction(age_cat, longevity_fct),
-                  ymin = conf.low, 
-                  ymax = conf.high), 
-              alpha = 0.2) +
-  geom_line(aes(color = age_cat, linetype = longevity_fct)) +
+# selective appearance for firstrepro=4,6 (10% and 90% of observed)
+selapp_pred <- ggpredict(
+  h2_selective_models[[2]],
+  terms = c("age10 [all]", "age_cat", "firstrepro [4,6]")
+) %>% 
+  as_tibble() %>% 
+  mutate(age = x * 10 + age_senesce,
+         age_cat = factor(group, levels = c("Young", "Old")),
+         firstrepro = facet) %>% 
+  filter((age_cat == "Young" & age < age_senesce) |
+           (age_cat == "Old" & age >= age_senesce),
+         age <= 20)
+
+# selective disappearance for lastrepro=5,17 (10% and 90% of observed) 
+seldis_pred <- ggpredict(
+  h2_selective_models[[4]],
+  terms = c("age10 [all]", "age_cat", "lastrepro [5,17]")
+) %>% 
+  as_tibble() %>% 
+  mutate(age = x * 10 + age_senesce,
+         age_cat = factor(group, levels = c("Young", "Old")),
+         lastrepro = facet) %>% 
+  filter((age_cat == "Young" & age < age_senesce) |
+           (age_cat == "Old" & age >= age_senesce),
+         age <= 20)
+
+# Selective appearance predictions
+fig_selapp <- ggplot(base_pred, aes(age, predicted)) +
+  # Base predictions
+  geom_line(aes(color = age_cat), alpha = 0.5) +
+  # Selective appearance ages 4 and 6
+  geom_line(aes(color = age_cat, 
+                linetype = firstrepro),
+            selapp_pred) +
   geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
   scale_y_continuous("Offspring survival", 
-                     labels = scales::percent,
-                     limits = c(0, 0.62)) +
+                     labels = scales::percent) +
   scale_color_manual(values = c("#7fbc41", "#de77ae")) +
   scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
-  labs(x = "Age (years)") +
+  scale_linetype_manual(values = c("dashed", "dotted")) +
+  labs(x = "Maternal age (years)") +
   theme(legend.position = "none")
 
-```
+fig_seldis <- ggplot(base_pred, aes(age, predicted)) +
+  # Base predictions
+  geom_line(aes(color = age_cat), alpha = 0.5) +
+  # Selective disappearance ages 5 and 17
+  geom_line(aes(color = age_cat, 
+                linetype = lastrepro),
+            seldis_pred) +
+  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
+  scale_y_continuous("Offspring survival", 
+                     labels = scales::percent) +
+  scale_color_manual(values = c("#7fbc41", "#de77ae")) +
+  scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
+  scale_linetype_manual(values = c("dashed", "dotted")) +
+  labs(x = "Maternal age (years)") +
+  theme(legend.position = "none")
+
+fig_selapp | fig_seldis +
+  plot_annotation(tag_levels = "A")
+```
+
+Maternal effect senescence was the most likely explanation for declining offspring survival with maternal age, although the evidence for selective appearance and disappearance was stronger for offspring survival than for maternal reproduction.
+The AIC weights for the base, selective appearance, selective disappearance, and combined models were 0.421, 0.254, 0.206, and 0.119, respectively, which indicates selective appearance and disappearance cannot be disregarded.
+However, as with maternal reproduction, the coefficients for ages at first (95% CI: -0.143 - 0.047) and last (95% CI: -0.263 - 1.09) reproduction were not significantly greater than 0 and the selective appearance/disappearance models did not generate qualitatively different predictions than the senescence-only model \[@fig-selective-h2\].
 
@@
 sealdat2 <- sealdat %>% 
-  drop_na(longevity) %>% 
-  mutate(bioage = age - longevity,
+  drop_na(lastrepro) %>% 
+  mutate(bioage = age - lastrepro,
          bioage10 = (bioage - bioage_senesce) / 10,
@@
   observed_int ~ age10 : age_cat + (1 | animalID) + (1 | year_fct),
-  drop_na(sealdat, longevity),
+  drop_na(sealdat, lastrepro),
   family = "binomial",
@@
   mod <- sealdat %>%
-    drop_na(longevity) %>%
-    mutate(bioage = age - longevity,
+    drop_na(lastrepro) %>%
+    mutate(bioage = age - lastrepro,
            bioage10 = (bioage - tau) / 10,
@@
               missing_pred,
-              alpha = 0.2, 
-              fill = "#de77ae") +
+              alpha = 0.2) +
   geom_line(aes(y = unobserved), 
             missing_pred, 
-            color = "#de77ae",
+            color = "cornflowerblue",
             linewidth = 2) +

### 2024-07-09 paper/paperv3.qmd
incorporate marm's edits

@@
   group_by(animalID) %>% 
+  # FIXME
   mutate(firstrepro = min(age),
@@
           prerecruit = year < first_obs) %>% 
-  filter(year > 1987) %>% 
   left_join(sealdat %>% 
@@
 seal_years <- count(sealdat, year_fct)
-breed_pred_wgt <- ggpredict(
+breed_pred_years <- ggpredict(
   breed_age_mod,
@@
            (age_cat == "Old" & age >= age_senesce),
-         age <= 20) %>% 
+         age <= 20)
+breed_pred_wgt <- breed_pred_years %>% 
   left_join(seal_years, by = "year_fct") %>% 
@@
 ggplot(breed_pred_pop, aes(age, predicted)) +
+  geom_line(aes(group = interaction(age_cat, year_fct)),
+            breed_pred_years,
+            alpha = 0.1) +
   geom_ribbon(aes(fill = age_cat, ymin = conf.low, ymax = conf.high), 
-              alpha = 0.2) +
+              alpha = 0.5) +
   geom_line(aes(color = age_cat), linewidth = 1.2) +
@@
   theme(legend.position = "none")
-
-ggsave("outputs/breeding_success.svg", height = 6, width = 10, units = "in")
 ```
@@
 
+# Predictions for each year (random effect)
+surv_pred_years <- ggpredict(
+  pup_surv_mod,
+  terms = c("age10 [all]", "age_cat", "year_fct [all]"),
+  type = "random",
+  interval = "confidence"
+) %>% 
+  as_tibble() %>% 
+  mutate(age = x * 10 + age_senesce,
+         age_cat = group,
+         year_fct = facet) %>% 
+  filter((age_cat == "Young" & age < age_senesce) |
+           (age_cat == "Old" & age >= age_senesce),
+         age <= 20)
+
 fig_surv <- ggplot(surv_pred_pop, aes(age, predicted)) +
+  geom_line(aes(group = interaction(age_cat, year_fct)),
+            surv_pred_years,
+            alpha = 0.1) +
   geom_ribbon(aes(fill = age_cat, ymin = conf.low, ymax = conf.high), 
@@
 
+# Predictions for each year (random effect)
+recr_pred_years <- ggpredict(
+  pup_recr_mod,
+  terms = c("age10 [all]", "age_cat", "year_fct [all]"),
+  type = "random",
+  interval = "confidence"
+) %>% 
+  as_tibble() %>% 
+  mutate(age = x * 10 + age_senesce,
+         age_cat = group,
+         year_fct = facet) %>% 
+  filter((age_cat == "Young" & age < age_senesce) |
+           (age_cat == "Old" & age >= age_senesce),
+         age <= 20)
+
 fig_recr <- ggplot(recr_pred_pop, aes(age, predicted)) +
+  geom_line(aes(group = interaction(age_cat, year_fct)),
+            recr_pred_years,
+            alpha = 0.1) +
   geom_ribbon(aes(fill = age_cat, ymin = conf.low, ymax = conf.high), 
@@
 
+# Predictions for each year (random effect)
+sexrat_pred_years <- ggpredict(
+  sexrat_age_mod,
+  terms = c("age10 [all]", "age_cat", "year_fct [all]"),
+  type = "random",
+  interval = "confidence"
+) %>% 
+  as_tibble() %>% 
+  mutate(age = x * 10 + age_senesce,
+         age_cat = group,
+         year_fct = facet) %>% 
+  filter((age_cat == "Young" & age < age_senesce) |
+           (age_cat == "Old" & age >= age_senesce),
+         age <= 20)
+
 ggplot(sexrat_pred, aes(age, predicted)) +
+  geom_line(aes(group = interaction(age_cat, year_fct)),
+            sexrat_pred_years,
+            alpha = 0.1) +
   geom_ribbon(aes(fill = age_cat, ymin = conf.low, ymax = conf.high), 
@@
 
+## Detailed model outputs
+
+### H1: Reproductive senescence
+
+```{r}
+#| label: detailed-h1
+
+summary(breed_age_mod)
+
+```
+
+### H2: Maternal effect senescence
+
+```{r}
+#| label: detailed-h2
+
+summary(pup_surv_mod)
+summary(pup_recr_mod)
+
+```
+
+### H3: Offspring sex ratio
+
+```{r}
+#| label: detailed-h3
+
+summary(sexrat_age_mod)
+
+```
+
 ## Models' sensitivity to age of senescence
@@
   breed_age_mod <- glmer(
-    observed_int ~ age10 * age_cat + (1 | animalID) + (1 | year_fct),
+    observed_int ~ age10 : age_cat + (1 | animalID) + (1 | year_fct),
     sealdat,
@@
   pup_surv_mod <- glmer(
-    pup_surv_int ~ age10 * age_cat + (1 | animalID) + (1 | year_fct),
+    pup_surv_int ~ age10 : age_cat + (1 | animalID) + (1 | year_fct),
     surv_recr_data,
@@
   pup_recr_mod <- glmer(
-    pup_recr_int ~ age10 * age_cat + (1 | year_fct),
+    pup_recr_int ~ age10 : age_cat + (1 | year_fct),
     filter(surv_recr_data, pupsex == "F"),
@@
   sexrat_age_mod <- glmer(
-    is_male ~ age10 * age_cat + (1 | animalID) + (1 | year_fct),
+    is_male ~ age10 : age_cat + (1 | animalID) + (1 | year_fct),
     pup_sex_data,
@@
 Hypothesis H2a (maternal-effect senescence; offspring survival) was significantly supported for onset of senescence age 11+.
-At age 10, the p-vale for the coefficient of the interaction (age category = post-senescence) x (offspring survival) was 0.051.
+At age 10, the p-vale for the coefficient of the interaction (age category = post-senescence) x (offspring survival) was 0.058.
 
@@
   geom_hline(yintercept = 0.05, linetype = "dashed") +
+  geom_text(aes(x, y, label = tag),
+            tibble(hypothesis = colnames(p_tbl)[-1],
+                   tag = LETTERS[1:4],
+                   x = 8, y = 0.75),
+            vjust = -0.2,
+            fontface = "bold") +
   scale_y_log10(labels = scales::number) +
   labs(x = "Senescence onset (years)") +
+  expand_limits(y = 1.3) +
   facet_wrap(~hypothesis)
@@
 The AIC weights for the base, selective appearance, selective disappearance, and combined models were 0.421, 0.254, 0.206, and 0.119, respectively, which indicates selective appearance and disappearance cannot be disregarded.
-However, as with maternal reproduction, the coefficients for ages at first (95% CI: -0.143 - 0.047) and last (95% CI: -0.263 - 1.09) reproduction were not significantly greater than 0 and the selective appearance/disappearance models did not generate qualitatively different predictions than the senescence-only model \[@fig-selective-h2\].
+However, as with maternal reproduction, the coefficients for ages at first (95% CI: -0.143 - 0.047) and last (95% CI: -0.263 - 1.09) reproduction were not significantly greater than 0 and the selective appearance/disappearance models did not generate qualitatively different predictions than the senescence-only model [@fig-selective-h2].
 

### 2024-07-09 paper/paperv3.qmd
selective app/disapp, dashed line in h2a

@@
   group_by(animalID) %>% 
-  # FIXME
-  mutate(firstrepro = min(age),
-         lastrepro = ifelse(max(year) < 2020,
-                            max(age),
+  mutate(firstrepro = ifelse(max(year) < 2020 && any(observed_int == 1),
+                             min(age[observed_int == 1]),
                              NA),
-         lastrepro10 = lastrepro / 10,
-         recruit_age = ifelse(max(year) < 2020 && any(observed_int == 1),
-                              min(age[observed_int == 1]),
+         lastrepro = ifelse(max(year) < 2020 && any(observed_int == 1),
+                            max(age[observed_int == 1]),
                             NA)) %>% 
@@
               alpha = 0.2) +
-  geom_line(aes(color = age_cat), linewidth = 1.2) +
+  geom_line(aes(color = age_cat, linetype = age_cat), linewidth = 1.2) +
   geom_pointrange(aes(y = perc_surv, 
@@
   scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
+  scale_linetype_manual(values = c("dashed", "solid")) +
   labs(x = "Age (years)") +
@@
 
-```{r}
-#| label: selective_values
-
-recruit_by_6 <- mean(sealdat$recruit_age <= 6, na.rm = TRUE)
-
-```
-
 Age-dependent life history shifts can result from within- or between-animal changes [@vandepol2006].
@@
 
-do.call(AIC, h1_selective_models) %>% 
+h1_aic <- do.call(AIC, h1_selective_models) %>% 
   as_tibble() %>% 
@@
   relocate(model)
+
+# Selective appearance model, first repro 95% ci
+h1_selapp_coef <- summary(h1_selective_models[[2]])$coef["firstrepro", ]
+make_95 <- function(ci95) {
+  sprintf("95%% CI: %0.3f - %0.3f", ci95[1], ci95[2])
+}
+h1_firstrepro_95ci <- (h1_selapp_coef["Estimate"] + c(-1.96, 1.96) * h1_selapp_coef["Std. Error"]) %>% 
+  make_95()
+
+h1_aic
 ```
@@
 The dashed and dotted lines represent predictions for the 10th and 90th percentiles of observed age at first (**A**) and last (**B**) reproduction.
-In **A**, these correspond to first reproduction at 4 and 6 years old, respectively.
+In **A**, these correspond to first reproduction at 4 and 7 years old, respectively.
 In **B**, these correspond to last reproduction at 5 and 17 years old, respectively.
 Confidence intervals not shown for clarity.
+#| fig-width: 9
 
@@
   h1_selective_models[[2]],
-  terms = c("age10 [all]", "age_cat", "firstrepro [4,6]")
+  terms = c("age10 [all]", "age_cat", "firstrepro [4,7]")
 ) %>% 
@@
 
-Selective appearance and disappearance did not explain the observed pattern of declining breeding probabilities with age.
-The AIC weights for the base, selective appearance, selective disappearance, and combined models were 0.526, 0.197, 0.201, and 0.076, respectively.
-These weights indicate the most likely explanation for the observed pattern is within-individual change, rather than selective appearance or disappearance.
-In the selective appearance and disappearance models, the coefficients for age at first reproduction (95% CI: -0.091 - 0.112) and last reproduction (95% CI: -0.037 - 0.051) were not significantly different than 0, respectively.
-Furthermore, predictions of the selective appearance and disappearance models did not meaningfully differ from the base model @fig-selective-h1.
+There was no evidence for selective appearance or disappearance influencing the relationship between maternal age and probability of reproduction.
+The selective appearance model (i.e., included a term for age-of-first-reproduction) had a similar AIC weight as the base model.
+However, in the selective appearance model, the coefficient for age at first reproduction was not significantly different than 0 (`r h1_firstrepro_95ci`).
+As a result, the predicted breeding probabilities for animals that first recruited at ages 4 and 7 (the 10th and 90th observed percentiles in the population) were not qualitatively different than the predictions of the base model [@fig-selective-h1].
+Therefore the observed pattern is consistent with the senescence hypothesis (i.e., within-animal change with age).
 
@@
 
-do.call(AIC, h2_selective_models) %>% 
+h2_aic <- do.call(AIC, h2_selective_models) %>% 
   as_tibble() %>% 
@@
 
+# Selective appearance model, first repro 95% ci
+h2_selapp_coef <- summary(h2_selective_models[[2]])$coef["firstrepro", ]
+h2_firstrepro_95ci <- (h2_selapp_coef["Estimate"] + c(-1.96, 1.96) * h2_selapp_coef["Std. Error"]) %>% 
+  make_95()
+
+h2_aic
+
 ```
@@
   h2_selective_models[[2]],
-  terms = c("age10 [all]", "age_cat", "firstrepro [4,6]")
+  terms = c("age10 [all]", "age_cat", "firstrepro [4,7]")
 ) %>% 
@@
 
-Maternal effect senescence was the most likely explanation for declining offspring survival with maternal age, although the evidence for selective appearance and disappearance was stronger for offspring survival than for maternal reproduction.
-The AIC weights for the base, selective appearance, selective disappearance, and combined models were 0.421, 0.254, 0.206, and 0.119, respectively, which indicates selective appearance and disappearance cannot be disregarded.
-However, as with maternal reproduction, the coefficients for ages at first (95% CI: -0.143 - 0.047) and last (95% CI: -0.263 - 1.09) reproduction were not significantly greater than 0 and the selective appearance/disappearance models did not generate qualitatively different predictions than the senescence-only model [@fig-selective-h2].
+There was no evidence for selective appearance or disappearance influencing the relationship between maternal age and offspring survival.
+The selective appearance model (i.e., included a term for age-of-first-reproduction) had a slightly greater AIC weight than the base model.
+However, in the selective appearance model, the coefficient for age at first reproduction was not significantly different than 0 (`r h2_firstrepro_95ci`).
+As a result, the predicted breeding probabilities for animals that first recruited at ages 4 and 7 (the 10th and 90th observed percentiles in the population) were not qualitatively different than the predictions of the base model [@fig-selective-h2].
+Therefore the observed pattern is consistent with the senescence hypothesis (i.e., within-animal change with age).
 

### 2024-07-09 paper/paperv3.qmd
updating the biological age figure

@@
               alpha = 0.2) +
-  geom_line(aes(color = bioage_cat), linewidth = 1.2) +
+  geom_line(aes(color = bioage_cat), linewidth = 1.2, linetype = "dashed") +
   geom_line(aes(color = bioage_cat),
@@
   geom_vline(xintercept = bioage_senesce - 0.5, linetype = "dashed") +
-  scale_y_continuous("Breeding", labels = scales::percent) +
+  scale_y_continuous("Breeding", labels = scales::percent, limits = c(.7, 1)) +
   scale_color_manual(values = c("#7fbc41", "#de77ae")) +
   scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
+  scale_x_continuous(limits = c(-17, 0)) + 
   labs(x = "Biological age (years-to-death)") +

### 2024-07-10 paper/paperv3.qmd
weighted model CIs

--- 2024-07-09 paper/paperv3.qmd
@@
 
-# Create WEIGHTED-AVERAGE predictions using years' random effects
-seal_years <- count(sealdat, year_fct)
+# Create ANNUAL (i.e. rand effect of year only) predictions
 breed_pred_years <- ggpredict(
   breed_age_mod, 
-  terms = c("age10 [all]", "age_cat", "year_fct [all]"),
-  type = "random",
-  interval = "confidence"
-) %>% 
-  as_tibble() %>% 
-  mutate(age = x * 10 + age_senesce,
-         age_cat = group,
-         year_fct = facet) %>% 
+  terms = c("age10 [all]", "age_cat", "year_fct"),
+  type = "random"
+) %>% 
+  as_tibble() %>% 
+  mutate(age = x * 10 + age_senesce,
+         age_cat = factor(group, levels = c("Young", "Old")),
+         year_fct = factor(facet)) %>% 
   filter((age_cat == "Young" & age < age_senesce) |
@@
          age <= 20)
-breed_pred_wgt <- breed_pred_years %>% 
+
+# Create WEIGHTED-AVERAGE predictions using years' random effects
+# ggpredict can't do that for confidence intervals
+# create prediction grid
+breed_grid <- expand_grid(
+  age = 4:20,
+  year_fct = factor(1988:2023)
+) %>% 
+  mutate(age10 = (age - age_senesce) / 10,
+         age_cat = factor(ifelse(age < age_senesce, "Young", "Old"),
+                          levels = c("Young", "Old")))
+
+# fixed + random effect predictions
+year_ranef <- ranef(breed_age_mod)$year_fct %>% 
+  as_tibble(rownames = "year_fct") %>% 
+  rename(year_ranef = `(Intercept)`)
+
+breed_pred <- breed_grid %>% 
+  mutate(breed_fixef = predict(breed_age_mod, 
+                               newdata = breed_grid,
+                               re.form = NA)) %>% 
+  left_join(year_ranef, by = "year_fct") %>% 
+  mutate(mean_breed = breed_fixef + year_ranef)
+
+# add confidence intervals
+mm <- model.matrix(~ age_cat:age10, breed_grid)
+pvar1 <- diag(mm %*% tcrossprod(vcov(breed_age_mod), mm))
+breed_pred$ci_low <- breed_pred$mean_breed - 1.96 * sqrt(pvar1)
+breed_pred$ci_high <- breed_pred$mean_breed + 1.96 * sqrt(pvar1)
+
+# Apply weighted average
+seal_years <- count(sealdat, year_fct)
+breed_pred_wgt <- breed_pred %>% 
   left_join(seal_years, by = "year_fct") %>% 
   group_by(age, age_cat) %>% 
-  summarize(predicted = weighted.mean(predicted, n),
-            .groups = "drop")
-
-ggplot(breed_pred_pop, aes(age, predicted)) +
-  geom_line(aes(group = interaction(age_cat, year_fct)),
+  summarize(across(c(mean_breed, ci_low, ci_high), \(x) weighted.mean(x, n)),
+            .groups = "drop") %>% 
+  # invert logit
+  mutate(across(c(mean_breed, ci_low, ci_high), 
+                family(breed_age_mod)$linkinv))
+
+ggplot(breed_pred_wgt, aes(age)) +
+  geom_line(aes(y = predicted, group = interaction(age_cat, year_fct)),
             breed_pred_years,
             alpha = 0.1) +
-  geom_ribbon(aes(fill = age_cat, ymin = conf.low, ymax = conf.high), 
+  geom_ribbon(aes(fill = age_cat, ymin = ci_low, ymax = ci_high), 
               alpha = 0.5) +
-  geom_line(aes(color = age_cat), linewidth = 1.2) +
-  geom_line(aes(color = age_cat),
-            breed_pred_wgt, 
-            linewidth = 1.2,
-            linetype = "dashed") +
+  geom_line(aes(y = mean_breed, color = age_cat),
+            linewidth = 1.2) +
   geom_pointrange(aes(y = perc_breed,

### 2024-07-11 paper/paperv3.qmd
Updating plots with sample sizes and making other edits

--- 2024-07-09 paper/paperv3.qmd
@@
     echo: false
-    fig-width: 5
+    fig-width: 7
     fig-height: 5
@@
 
+#Create sample size data frame
+n_breed <- sealdat %>% 
+  filter(age <= 20) %>% 
+  count(age)
+
 # Create POPULATION-LEVEL (i.e. without rand effects) predictions
@@
            (age_cat == "Old" & age >= age_senesce),
-         age <= 20)
+         age <= 20) %>% 
+  left_join(n_breed, by = "age")
 
@@
   geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
-  scale_y_continuous("Breeding", labels = scales::percent) +
+  scale_y_continuous("Breeding probability", labels = scales::percent) +
   scale_color_manual(values = c("#7fbc41", "#de77ae")) +
@@
   labs(x = "Age (years)") +
-  theme(legend.position = "none")
+  theme(legend.position = "none") + 
+  geom_text(aes(label = n, y = .55))
+
+ggsave("outputs/repro_senesce.svg", width = 11, height = 6, units = "in")
 ```
@@
 #| label: fig-repro-age
-#| fig-width: 9
-
-n_recruit <- surv_recr_data %>% 
-  filter(pupsex == "F")
+#| fig-width: 8
+#| fig-height: 8
 
@@
 
+#Create sample size data frame
+n_surv <- surv_recr_data %>% 
+  group_by(age) %>% 
+  count(age)
+
 # Create pup survival POPULATION-LEVEL (i.e. without rand effects) predictions
@@
            (age_cat == "Old" & age >= age_senesce),
-         age <= 20)
+         age <= 20) %>% 
+  left_join(n_surv, by = "age")
 
@@
                      labels = scales::percent,
-                     limits = c(0, 0.62)) +
+                     limits = c(0, 0.68)) +
   scale_color_manual(values = c("#7fbc41", "#de77ae")) +
@@
   labs(x = "Age (years)") +
-  theme(legend.position = "none")
+  theme(legend.position = "none") + 
+  geom_text(aes(label = n, y = .65))
+
+#Create sample size data frame
+n_recruit <- surv_recr_data %>% 
+  filter(pupsex == "F") %>% 
+  count(age)
 
@@
            (age_cat == "Old" & age >= age_senesce),
-         age <= 20)
+         age <= 20) %>% 
+  left_join(n_recruit, by = "age")
 
@@
   labs(x = "Age (years)") +
-  theme(legend.position = "none")
-
-(fig_surv | fig_recr) + plot_annotation(tag_levels = "A")
+  theme(legend.position = "none") + 
+  geom_text(aes(label = n, y = .65))
+
+(fig_surv | fig_recr) + plot_annotation(tag_levels = "A") + plot_layout(ncol = 1, nrow = 2)
 
@@
 
+#Make sample size data frame
+n_sexrat <- pup_sex_data %>% 
+  count(age)
+
 # Create POPULATION-LEVEL (i.e. without rand effects) predictions
@@
   filter((age_cat == "Young" & age < age_senesce) |
-           (age_cat == "Old" & age >= age_senesce))
+           (age_cat == "Old" & age >= age_senesce)) %>% 
+  left_join(n_sexrat, by = "age")
 
@@
   labs(x = "Age (years)") +
-  theme(legend.position = "none")
+  theme(legend.position = "none") +
+  geom_text(aes(label = n, y = .95))
+
 ```
@@
 B: Raw data for breeding and molting durations (mean ± se) and predictions (mean and 95% CI) from negative binomial weighted by number of seals obersved per year."
-#| fig-width: 9
-
-phenology_n <- unique(seal_phenology$animalID)
+#| fig-width: 4
+#| fig-height: 9
+
+phenology_unique_animals <- unique(seal_phenology$animalID)
+n_phenology <- seal_phenology %>% 
+  count(year)
+
+#Make sample size data frame
+n_phen <- seal_phenology %>% 
+  count(age)
 
@@
                                           "Molt haulout", 
-                                          "foraging2")))
+                                          "foraging2"))) %>% 
+    left_join(n_phen, by = "age")
 
@@
         legend.key.size = unit(3, "mm")) 
-
-n_phenology <- seal_phenology %>% 
-  count(year)
+  # geom_text(aes(label = n, color = "white", y = 300))
 
@@
 
-duration_plot <- ggpredict(
+phen_pred_years <- ggpredict(
+  pheno_model,
+  terms = c("age10 [all]", "phase", "year [all]"),
+  type = "random",
+  interval = "confidence"
+) %>% 
+  as_tibble() %>% 
+  mutate(age = x * 10,
+         phase = group)
+
+breed_pheno_raw <- pheno_raw[pheno_raw$phase == "breeddur",]
+molt_pheno_raw <- pheno_raw[pheno_raw$phase == "moltdur",]
+
+breed_duration_plot <- ggpredict(
+  pheno_model, 
+  terms = c("age10 [all]", "phase", "year [all]")
+) %>% 
+  as_tibble() %>% 
+  transmute(age = x * 10,
+            dur_days = predicted,
+            dur_days_lwr = conf.low,
+            dur_days_upr = conf.high,
+            phase = group,
+            year = as.numeric(as.character(facet))) %>% 
+  left_join(n_phenology, by = "year") %>% 
+  group_by(age) %>% 
+  filter(phase == "breeddur") %>% 
+  summarize(across(starts_with("dur_days"), \(x) weighted.mean(x, n)),
+            .groups = "drop") %>% 
+  ggplot(aes(age)) +
+  geom_line(aes(age, predicted, group = interaction(phase, facet)), 
+            phen_pred_years, 
+            alpha = 0.1) + 
+  geom_ribbon(aes(ymin = dur_days_lwr, ymax = dur_days_upr),
+              alpha = 0.25, 
+              fill = "#D05794") +
+  geom_pointrange(aes(y = dur_days, 
+                      ymin = dur_days_lwr, 
+                      ymax = dur_days_upr),
+                  breed_pheno_raw,
+                  position = position_dodge(width = 0.5), 
+                  color = "#D05794") +
+  geom_line(aes(y = dur_days), 
+            linetype = "dashed", 
+            color = "#D05794") +
+  labs(x = "Maternal age (years)",
+       y = "Breeding haul out (days)") +
+  theme(legend.position = "none",
+        axis.text.y = element_text(size = 8))
+
+molt_duration_plot <- ggpredict(
   pheno_model,
@@
   group_by(age, phase) %>%
+  filter(phase == "moltdur") %>%
   summarize(across(starts_with("dur_days"), \(x) weighted.mean(x, n)),
@@
   ggplot(aes(age)) +
-  geom_ribbon(aes(ymin = dur_days_lwr, ymax = dur_days_upr, fill = phase),
-              alpha = 0.25) +
+  geom_line(aes(age, predicted, group = interaction(phase, facet)), 
+            phen_pred_years, 
+            alpha = 0.1) + 
+  geom_ribbon(aes(ymin = dur_days_lwr, ymax = dur_days_upr),
+              alpha = 0.25, 
+              fill = "#E09A2D") +
   geom_pointrange(aes(y = dur_days, 
                       ymin = dur_days_lwr, 
-                      ymax = dur_days_upr,
-                      color = phase),
-                  pheno_raw,
-                  position = position_dodge(width = 0.5)) +
-  geom_line(aes(y = dur_days, color = phase, linetype = phase)) +
-  annotate("text", 12, 14, label = "Molting", color = "#E09A2D", hjust = 0) +
-  annotate("text", 12, 32, label = "Breeding", color = "#D05794", hjust = 0) +
-  scale_fill_manual(values = c("#D05794", "#E09A2D")) +
-  scale_color_manual(values = c("#D05794", "#E09A2D")) +
-  scale_linetype_manual(values = c("dashed", "solid")) +
+                      ymax = dur_days_upr),
+                  molt_pheno_raw,
+                  position = position_dodge(width = 0.5), 
+                  color = "#E09A2D") +
+  geom_line(aes(y = dur_days), 
+            color = "#E09A2D") +
   labs(x = "Maternal age (years)",
-       y = "Phase duration (days)") +
-  theme(legend.position = "none")
-
-(annual_plot | duration_plot) + plot_annotation(tag_levels = "A") + plot_layout(widths = c(1.25, 2))
-
-```
-
-Sample size n = `r nrow(seal_phenology)` observations of `r n_distinct(phenology_n)` individuals.
+       y = "Molting haul out (days)") +
+  theme(legend.position = "none", 
+        axis.text.y = element_text(size = 8))
+
+(annual_plot | breed_duration_plot | molt_duration_plot) + plot_annotation(tag_levels = "A") + plot_layout(ncol = 1, nrow = 3)
+```
+
+Sample size n = `r nrow(seal_phenology)` observations of `r n_distinct(phenology_unique_animals)` individuals.
 
@@
   geom_vline(xintercept = bioage_senesce - 0.5, linetype = "dashed") +
-  scale_y_continuous("Breeding", labels = scales::percent) +
+  scale_y_continuous("Breeding probability", labels = scales::percent, limits = c(.7, 1)) +
   scale_color_manual(values = c("#7fbc41", "#de77ae")) +
   scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
+  scale_x_continuous(limits = c(-17, 0)) + 
   labs(x = "Biological age (years-to-death)") +

### 2024-07-11 paper/paperv3.qmd
Display weighted model CIs according to Marm's suggestions

--- 2024-07-10 paper/paperv3.qmd
@@
 Based on the literature, we chose age `r age_senesce` to be the end of prime age and indicate the onset of senescence.
-If breeding success declined with age, we would expect to see a negative difference between the slopes of the two models (pre and post senescence) (@fig-breeding-mod).
+If breeding success declined with age, we would expect to see a negative slope post-senescence (@fig-breeding-mod).
 
@@
 
-# fixed + random effect predictions
-year_ranef <- ranef(breed_age_mod)$year_fct %>% 
-  as_tibble(rownames = "year_fct") %>% 
-  rename(year_ranef = `(Intercept)`)
-
-breed_pred <- breed_grid %>% 
-  mutate(breed_fixef = predict(breed_age_mod, 
+seal_years <- count(sealdat, year_fct)
+breed_pred_wgt <- breed_grid %>% 
+  mutate(predicted = predict(breed_age_mod, 
                              newdata = breed_grid,
-                               re.form = NA)) %>% 
-  left_join(year_ranef, by = "year_fct") %>% 
-  mutate(mean_breed = breed_fixef + year_ranef)
-
-# add confidence intervals
-mm <- model.matrix(~ age_cat:age10, breed_grid)
-pvar1 <- diag(mm %*% tcrossprod(vcov(breed_age_mod), mm))
-breed_pred$ci_low <- breed_pred$mean_breed - 1.96 * sqrt(pvar1)
-breed_pred$ci_high <- breed_pred$mean_breed + 1.96 * sqrt(pvar1)
-
-# Apply weighted average
-seal_years <- count(sealdat, year_fct)
-breed_pred_wgt <- breed_pred %>% 
+                             re.form = ~ (1 | year_fct)),
+         predicted_se = suppressWarnings(
+           # warning: "se.fit computation uses an approximation to estimate the sampling distribution of the parameters"
+           predict(breed_age_mod, 
+                   newdata = breed_grid,
+                   re.form = NA,
+                   se.fit = TRUE)
+         )$se.fit,
+         predicted_pop = predict(breed_age_mod, 
+                                 newdata = breed_grid,
+                                 re.form = NA),
+         conf_lo = predicted - 1.96 * predicted_se,
+         conf_hi = predicted + 1.96 * predicted_se) %>% 
   left_join(seal_years, by = "year_fct") %>% 
   group_by(age, age_cat) %>% 
-  summarize(across(c(mean_breed, ci_low, ci_high), \(x) weighted.mean(x, n)),
+  summarize(across(c(predicted, predicted_pop, conf_lo, conf_hi), 
+                   \(x) weighted.mean(x, n)),
             .groups = "drop") %>% 
   # invert logit
-  mutate(across(c(mean_breed, ci_low, ci_high), 
+  mutate(across(c(predicted, predicted_pop, conf_lo, conf_hi), 
                 family(breed_age_mod)$linkinv))
 
-ggplot(breed_pred_wgt, aes(age)) +
-  geom_line(aes(y = predicted, group = interaction(age_cat, year_fct)),
+ggplot(breed_pred_wgt, aes(age, predicted)) +
+  # Predictions for individual years
+  geom_line(aes(group = interaction(age_cat, year_fct)),
             breed_pred_years,
             alpha = 0.1) +
-  geom_ribbon(aes(fill = age_cat, ymin = ci_low, ymax = ci_high), 
+  # Weighted model (CI and mean response)
+  geom_ribbon(aes(fill = age_cat, ymin = conf_lo, ymax = conf_hi), 
               alpha = 0.5) +
-  geom_line(aes(y = mean_breed, color = age_cat),
+  geom_line(aes(color = age_cat),
             linewidth = 1.2) +
+  # Population-level mean response
+  geom_line(aes(y = predicted_pop, group = age_cat),
+            linetype = "dotted",
+            linewidth = 1.2,
+            alpha = 0.5) +
   geom_pointrange(aes(y = perc_breed,
-                      ymin = perc_breed - se,
-                      ymax = perc_breed + se),
+                      ymin = perc_breed - 1.96 * se,
+                      ymax = perc_breed + 1.96 * se),
                   breed_nonbreed) +
@@
             perc_recr = sum(pup_recr_int) / n_fpups, 
-            se_recr = (perc_recr * (1 - perc_recr) / n_fpups)^0.5) %>% 
-  filter(age <= 20)
-
-# Create pup survival POPULATION-LEVEL (i.e. without rand effects) predictions
-surv_pred_pop <- ggpredict(
-  pup_surv_mod, 
-  terms = c("age10 [all]", "age_cat"),
-  type = "fixed",
-  interval = "confidence"
-) %>% 
-  as_tibble() %>% 
-  mutate(age = x * 10 + age_senesce,
-         age_cat = group) %>% 
-  filter((age_cat == "Young" & age < age_senesce) |
-           (age_cat == "Old" & age >= age_senesce),
-         age <= 20)
+            se_recr = (perc_recr * (1 - perc_recr) / n_fpups)^0.5)
+
+# Create pup survival predictions (population-level and weighted by year)
+surv_grid <- expand_grid(
+  age = 4:19,
+  year_fct = factor(1988:2016)
+) %>% 
+  mutate(age10 = (age - age_senesce) / 10,
+         age_cat = factor(ifelse(age < age_senesce, "Young", "Old"),
+                          levels = c("Young", "Old")))
+
+surv_years <- count(surv_recr_data, year_fct)
+surv_pred_wgt <- surv_grid %>% 
+  mutate(predicted = predict(pup_surv_mod, 
+                             newdata = surv_grid,
+                             re.form = ~ (1 | year_fct)),
+         predicted_se = suppressWarnings(
+           # warning: "se.fit computation uses an approximation to estimate the sampling distribution of the parameters"
+           predict(pup_surv_mod, 
+                   newdata = surv_grid,
+                   re.form = NA,
+                   se.fit = TRUE),
+         )$se.fit,
+         predicted_pop = predict(pup_surv_mod, 
+                                 newdata = surv_grid,
+                                 re.form = NA),
+         conf_lo = predicted - 1.96 * predicted_se,
+         conf_hi = predicted + 1.96 * predicted_se) %>% 
+  left_join(surv_years, by = "year_fct") %>% 
+  group_by(age, age_cat) %>% 
+  summarize(across(c(predicted, predicted_pop, conf_lo, conf_hi), 
+                   \(x) weighted.mean(x, n)),
+            .groups = "drop") %>% 
+  # invert logit
+  mutate(across(c(predicted, predicted_pop, conf_lo, conf_hi), 
+                family(pup_surv_mod)$linkinv))
 
@@
 
-fig_surv <- ggplot(surv_pred_pop, aes(age, predicted)) +
+fig_surv <- ggplot(surv_pred_wgt, aes(age, predicted)) +
+  # Individual years
   geom_line(aes(group = interaction(age_cat, year_fct)),
@@
             alpha = 0.1) +
-  geom_ribbon(aes(fill = age_cat, ymin = conf.low, ymax = conf.high), 
+  # Weighted model (confidence interval and mean response)
+  geom_ribbon(aes(fill = age_cat, ymin = conf_lo, ymax = conf_hi), 
               alpha = 0.2) +
   geom_line(aes(color = age_cat, linetype = age_cat), linewidth = 1.2) +
+  # Population-level mean response
+  geom_line(aes(y = predicted_pop, group = age_cat),
+            linetype = "dotted", linewidth = 1.2, alpha = 0.5) +
+  # Raw data
   geom_pointrange(aes(y = perc_surv, 
@@
                   surv_recr_summ) +
+  # Senescence threshold
   geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
@@
 
-# Create pup recruit POPULATION-LEVEL (i.e. without rand effects) predictions
-recr_pred_pop <- ggpredict(
-  pup_recr_mod, 
-  terms = c("age10 [all]", "age_cat"),
-  type = "fixed",
-  interval = "confidence"
-) %>% 
-  as_tibble() %>% 
-  mutate(age = x * 10 + age_senesce,
-         age_cat = group) %>% 
-  filter((age_cat == "Young" & age < age_senesce) |
-           (age_cat == "Old" & age >= age_senesce),
-         age <= 20)
+# Create pup recruitment predictions (population-level and weighted by year)
+recr_grid <- expand_grid(
+  age = 4:19,
+  year_fct = factor(1988:2016)
+) %>% 
+  mutate(age10 = (age - age_senesce) / 10,
+         age_cat = factor(ifelse(age < age_senesce, "Young", "Old"),
+                          levels = c("Young", "Old")))
+
+recr_years <- surv_recr_data %>% 
+  filter(pupsex == "F") %>% 
+  count(year_fct)
+recr_pred <- recr_grid %>% 
+  mutate(predicted = predict(pup_recr_mod, 
+                             newdata = recr_grid,
+                             re.form = ~ (1 | year_fct)),
+         predicted_se = suppressWarnings(
+           # warning: "se.fit computation uses an approximation to estimate the sampling distribution of the parameters"
+           predict(pup_recr_mod, 
+                   newdata = recr_grid,
+                   re.form = NA,
+                   se.fit = TRUE),
+         )$se.fit,
+         predicted_pop = predict(pup_recr_mod, 
+                                 newdata = recr_grid,
+                                 re.form = NA),
+         conf_lo = predicted - 1.96 * predicted_se,
+         conf_hi = predicted + 1.96 * predicted_se) %>% 
+  left_join(recr_years, by = "year_fct") %>% 
+  group_by(age, age_cat) %>% 
+  summarize(across(c(predicted, predicted_pop, conf_lo, conf_hi), 
+                   \(x) weighted.mean(x, n)),
+            .groups = "drop") %>% 
+  # invert logit
+  mutate(across(c(predicted, predicted_pop, conf_lo, conf_hi), 
+                family(pup_recr_mod)$linkinv))
 
@@
 
-fig_recr <- ggplot(recr_pred_pop, aes(age, predicted)) +
+fig_recr <- ggplot(recr_pred, aes(age, predicted)) +
+  # Individual years
   geom_line(aes(group = interaction(age_cat, year_fct)),
@@
             alpha = 0.1) +
-  geom_ribbon(aes(fill = age_cat, ymin = conf.low, ymax = conf.high), 
+  # Weighted model (confidence interval and mean response)
+  geom_ribbon(aes(fill = age_cat, ymin = conf_lo, ymax = conf_hi), 
               alpha = 0.2) +
-  geom_line(aes(color = age_cat), linewidth = 1.2, linetype = "dashed") +
+  geom_line(aes(color = age_cat, linetype = age_cat), linewidth = 1.2) +
+  # Population-level mean response
+  geom_line(aes(y = predicted_pop, group = age_cat),
+            linetype = "dotted", linewidth = 1.2, alpha = 0.5) +
+  # Raw data
   geom_pointrange(aes(y = perc_recr, 
@@
 
-# Create POPULATION-LEVEL (i.e. without rand effects) predictions
-sexrat_pred <- ggpredict(sexrat_age_mod, 
-                         terms = c("age10 [all]", "age_cat")) %>% 
-  as_tibble() %>% 
-  mutate(age = x * 10 + age_senesce,
-         age_cat = factor(group, levels = c("Young", "Old"))) %>% 
-  filter((age_cat == "Young" & age < age_senesce) |
-           (age_cat == "Old" & age >= age_senesce))
+# Create offspring sex ratio predictions (population-level and weighted by year)
+sexrat_grid <- expand_grid(
+  age = 4:18,
+  year_fct = factor(1988:2023)
+) %>% 
+  mutate(age10 = (age - age_senesce) / 10,
+         age_cat = factor(ifelse(age < age_senesce, "Young", "Old"),
+                          levels = c("Young", "Old")))
+
+sexrat_years <- count(pup_sex_data, year_fct)
+sexrat_pred <- sexrat_grid %>% 
+  mutate(predicted = predict(sexrat_age_mod, 
+                             newdata = sexrat_grid,
+                             re.form = ~ (1 | year_fct)),
+         predicted_se = suppressWarnings(
+           # warning: "se.fit computation uses an approximation to estimate the sampling distribution of the parameters"
+           predict(sexrat_age_mod, 
+                   newdata = sexrat_grid,
+                   re.form = NA,
+                   se.fit = TRUE),
+         )$se.fit,
+         predicted_pop = predict(sexrat_age_mod, 
+                                 newdata = sexrat_grid,
+                                 re.form = NA),
+         conf_lo = predicted - 1.96 * predicted_se,
+         conf_hi = predicted + 1.96 * predicted_se) %>% 
+  left_join(sexrat_years, by = "year_fct") %>% 
+  group_by(age, age_cat) %>% 
+  summarize(across(c(predicted, predicted_pop, conf_lo, conf_hi), 
+                   \(x) weighted.mean(x, n)),
+            .groups = "drop") %>% 
+  # invert logit
+  mutate(across(c(predicted, predicted_pop, conf_lo, conf_hi), 
+                family(sexrat_age_mod)$linkinv))
 
 # Predictions for each year (random effect)
-sexrat_pred_years <- ggpredict(
+sexrat_years <- ggpredict(
   sexrat_age_mod,
@@
 ggplot(sexrat_pred, aes(age, predicted)) +
+  # Individual years
   geom_line(aes(group = interaction(age_cat, year_fct)),
-            sexrat_pred_years,
+            sexrat_years,
             alpha = 0.1) +
-  geom_ribbon(aes(fill = age_cat, ymin = conf.low, ymax = conf.high), 
+  # Weighted model (confidence interval and mean response)
+  geom_ribbon(aes(fill = age_cat, ymin = conf_lo, ymax = conf_hi), 
               alpha = 0.2) +
-  geom_line(aes(color = age_cat), linewidth = 1.2, linetype = "dashed") +
+  geom_line(aes(color = age_cat, group = age_cat),
+            linetype = "dashed", linewidth = 1.2) +
+  # Population-level mean response
+  geom_line(aes(y = predicted_pop, group = age_cat),
+            linetype = "dotted", linewidth = 1.2, alpha = 0.5) +
+  # Raw data
   geom_pointrange(aes(y = perc_male, 
@@
   geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
+  scale_y_continuous("Male offspring", labels = scales::percent) +
   scale_color_manual(values = c("#7fbc41", "#de77ae")) +
   scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
-  scale_y_continuous("Male offspring", labels = scales::percent) +
   labs(x = "Age (years)") +
   theme(legend.position = "none")
+
 ```

### 2024-07-11 paper/paperv3.qmd

--- 2024-07-11 paper/paperv3.qmd
@@
 
-# Create POPULATION-LEVEL (i.e. without rand effects) predictions
-breed_pred_pop <- ggpredict(breed_age_mod, 
-                            terms = c("age10 [all]", "age_cat")) %>% 
-  as_tibble() %>% 
-  mutate(age = x * 10 + age_senesce,
-         age_cat = factor(group, levels = c("Young", "Old"))) %>% 
-  filter((age_cat == "Young" & age < age_senesce) |
-           (age_cat == "Old" & age >= age_senesce),
-         age <= 20) %>% 
-  left_join(n_breed, by = "age")
-
-# Create WEIGHTED-AVERAGE predictions using years' random effects
-seal_years <- count(sealdat, year_fct)
+# # Create POPULATION-LEVEL (i.e. without rand effects) predictions
+# breed_pred_pop <- ggpredict(breed_age_mod, 
+#                             terms = c("age10 [all]", "age_cat")) %>% 
+#   as_tibble() %>% 
+#   mutate(age = x * 10 + age_senesce,
+#          age_cat = factor(group, levels = c("Young", "Old"))) %>% 
+#   filter((age_cat == "Young" & age < age_senesce) |
+#            (age_cat == "Old" & age >= age_senesce),
+#          age <= 20) %>% 
+#   left_join(n_breed, by = "age")
+
+# Create ANNUAL (i.e. rand effect of year only) predictions
 breed_pred_years <- ggpredict(
   breed_age_mod, 
-  terms = c("age10 [all]", "age_cat", "year_fct [all]"),
-  type = "random",
-  interval = "confidence"
+  terms = c("age10 [all]", "age_cat", "year_fct"),
+  type = "random"
 ) %>% 
@@
   mutate(age = x * 10 + age_senesce,
-         age_cat = group,
-         year_fct = facet) %>% 
+         age_cat = factor(group, levels = c("Young", "Old")),
+         year_fct = factor(facet)) %>% 
   filter((age_cat == "Young" & age < age_senesce) |
@@
          age <= 20)
-breed_pred_wgt <- breed_pred_years %>% 
+
+# Create WEIGHTED-AVERAGE predictions using years' random effects
+# ggpredict can't do that for confidence intervals
+# create prediction grid
+breed_grid <- expand_grid(
+  age = 4:20,
+  year_fct = factor(1988:2023)
+) %>% 
+  mutate(age10 = (age - age_senesce) / 10,
+         age_cat = factor(ifelse(age < age_senesce, "Young", "Old"),
+                          levels = c("Young", "Old")))
+
+seal_years <- count(sealdat, year_fct)
+breed_pred_wgt <- breed_grid %>% 
+  mutate(predicted = predict(breed_age_mod, 
+                             newdata = breed_grid,
+                             re.form = ~ (1 | year_fct)),
+         predicted_se = suppressWarnings(
+           # warning: "se.fit computation uses an approximation to estimate the sampling distribution of the parameters"
+           predict(breed_age_mod, 
+                   newdata = breed_grid,
+                   re.form = NA,
+                   se.fit = TRUE)
+         )$se.fit,
+         predicted_pop = predict(breed_age_mod, 
+                                 newdata = breed_grid,
+                                 re.form = NA),
+         conf_lo = predicted - 1.96 * predicted_se,
+         conf_hi = predicted + 1.96 * predicted_se) %>% 
   left_join(seal_years, by = "year_fct") %>% 
   group_by(age, age_cat) %>% 
-  summarize(predicted = weighted.mean(predicted, n),
-            .groups = "drop")
-
-ggplot(breed_pred_pop, aes(age, predicted)) +
+  summarize(across(c(predicted, predicted_pop, conf_lo, conf_hi), 
+                   \(x) weighted.mean(x, n)),
+            .groups = "drop") %>% 
+  # invert logit
+  mutate(across(c(predicted, predicted_pop, conf_lo, conf_hi), 
+                family(breed_age_mod)$linkinv)) %>% 
+    left_join(n_breed, by = "age")
+
+ggplot(breed_pred_wgt, aes(age, predicted)) +
+  # Predictions for individual years
   geom_line(aes(group = interaction(age_cat, year_fct)),
@@
             alpha = 0.1) +
-  geom_ribbon(aes(fill = age_cat, ymin = conf.low, ymax = conf.high), 
+  # Weighted model (CI and mean response)
+  geom_ribbon(aes(fill = age_cat, ymin = conf_lo, ymax = conf_hi), 
               alpha = 0.5) +
-  geom_line(aes(color = age_cat), linewidth = 1.2) +
   geom_line(aes(color = age_cat),
-            breed_pred_wgt, 
+            linewidth = 1.2) +
+  # Population-level mean response
+  geom_line(aes(y = predicted_pop, group = age_cat),
+            linetype = "dotted",
             linewidth = 1.2,
-            linetype = "dashed") +
+            alpha = 0.5) +
   geom_pointrange(aes(y = perc_breed,
-                      ymin = perc_breed - se,
-                      ymax = perc_breed + se),
+                      ymin = perc_breed - 1.96 * se,
+                      ymax = perc_breed + 1.96 * se),
                   breed_nonbreed) +
@@
   theme(legend.position = "none") + 
-  geom_text(aes(label = n, y = .55))
+  geom_text(aes(label = n, y = 1.1))
 
@@
   group_by(age) %>% 
+  filter(age <= 19) %>% 
   summarize(n_pups = n(), 
@@
             perc_recr = sum(pup_recr_int) / n_fpups, 
-            se_recr = (perc_recr * (1 - perc_recr) / n_fpups)^0.5) %>% 
-  filter(age <= 20)
+            se_recr = (perc_recr * (1 - perc_recr) / n_fpups)^0.5)
 
@@
   group_by(age) %>% 
+  filter(age <= 19) %>% 
   count(age)
 
-# Create pup survival POPULATION-LEVEL (i.e. without rand effects) predictions
-surv_pred_pop <- ggpredict(
-  pup_surv_mod, 
-  terms = c("age10 [all]", "age_cat"),
-  type = "fixed",
-  interval = "confidence"
-) %>% 
-  as_tibble() %>% 
-  mutate(age = x * 10 + age_senesce,
-         age_cat = group) %>% 
-  filter((age_cat == "Young" & age < age_senesce) |
-           (age_cat == "Old" & age >= age_senesce),
-         age <= 20) %>% 
+# Create pup survival predictions (population-level and weighted by year)
+surv_grid <- expand_grid(
+  age = 4:19,
+  year_fct = factor(1988:2016)
+) %>% 
+  mutate(age10 = (age - age_senesce) / 10,
+         age_cat = factor(ifelse(age < age_senesce, "Young", "Old"),
+                          levels = c("Young", "Old")))
+
+surv_years <- count(surv_recr_data, year_fct)
+surv_pred_wgt <- surv_grid %>% 
+  mutate(predicted = predict(pup_surv_mod, 
+                             newdata = surv_grid,
+                             re.form = ~ (1 | year_fct)),
+         predicted_se = suppressWarnings(
+           # warning: "se.fit computation uses an approximation to estimate the sampling distribution of the parameters"
+           predict(pup_surv_mod, 
+                   newdata = surv_grid,
+                   re.form = NA,
+                   se.fit = TRUE),
+         )$se.fit,
+         predicted_pop = predict(pup_surv_mod, 
+                                 newdata = surv_grid,
+                                 re.form = NA),
+         conf_lo = predicted - 1.96 * predicted_se,
+         conf_hi = predicted + 1.96 * predicted_se) %>% 
+  left_join(surv_years, by = "year_fct") %>% 
+  group_by(age, age_cat) %>% 
+  summarize(across(c(predicted, predicted_pop, conf_lo, conf_hi), 
+                   \(x) weighted.mean(x, n)),
+            .groups = "drop") %>% 
+  # invert logit
+  mutate(across(c(predicted, predicted_pop, conf_lo, conf_hi), 
+                family(pup_surv_mod)$linkinv)) %>% 
       left_join(n_surv, by = "age")
@@
 
-# Create pup recruit POPULATION-LEVEL (i.e. without rand effects) predictions
-recr_pred_pop <- ggpredict(
-  pup_recr_mod, 
-  terms = c("age10 [all]", "age_cat"),
-  type = "fixed",
-  interval = "confidence"
-) %>% 
-  as_tibble() %>% 
-  mutate(age = x * 10 + age_senesce,
-         age_cat = group) %>% 
-  filter((age_cat == "Young" & age < age_senesce) |
-           (age_cat == "Old" & age >= age_senesce),
-         age <= 20) %>% 
+# Create pup recruitment predictions (population-level and weighted by year)
+recr_grid <- expand_grid(
+  age = 4:19,
+  year_fct = factor(1988:2016)
+) %>% 
+  mutate(age10 = (age - age_senesce) / 10,
+         age_cat = factor(ifelse(age < age_senesce, "Young", "Old"),
+                          levels = c("Young", "Old")))
+
+recr_years <- surv_recr_data %>% 
+  filter(pupsex == "F") %>% 
+  count(year_fct)
+recr_pred <- recr_grid %>% 
+  mutate(predicted = predict(pup_recr_mod, 
+                             newdata = recr_grid,
+                             re.form = ~ (1 | year_fct)),
+         predicted_se = suppressWarnings(
+           # warning: "se.fit computation uses an approximation to estimate the sampling distribution of the parameters"
+           predict(pup_recr_mod, 
+                   newdata = recr_grid,
+                   re.form = NA,
+                   se.fit = TRUE),
+         )$se.fit,
+         predicted_pop = predict(pup_recr_mod, 
+                                 newdata = recr_grid,
+                                 re.form = NA),
+         conf_lo = predicted - 1.96 * predicted_se,
+         conf_hi = predicted + 1.96 * predicted_se) %>% 
+  left_join(recr_years, by = "year_fct") %>% 
+  group_by(age, age_cat) %>% 
+  summarize(across(c(predicted, predicted_pop, conf_lo, conf_hi), 
+                   \(x) weighted.mean(x, n)),
+            .groups = "drop") %>% 
+  # invert logit
+  mutate(across(c(predicted, predicted_pop, conf_lo, conf_hi), 
+                family(pup_recr_mod)$linkinv)) %>% 
         left_join(n_recruit, by = "age")
@@
             alpha = 0.1) +
-  geom_ribbon(aes(fill = age_cat, ymin = conf.low, ymax = conf.high), 
+  # Weighted model (confidence interval and mean response)
+  geom_ribbon(aes(fill = age_cat, ymin = conf_lo, ymax = conf_hi), 
               alpha = 0.2) +
-  geom_line(aes(color = age_cat), linewidth = 1.2, linetype = "dashed") +
+  geom_line(aes(color = age_cat), linetype = "dashed", linewidth = 1.2) +
+  # Population-level mean response
+  geom_line(aes(y = predicted_pop, group = age_cat),
+            linetype = "dotted", linewidth = 1.2, alpha = 0.5) +
+  # Raw data
   geom_pointrange(aes(y = perc_recr, 
@@
 
-# Create POPULATION-LEVEL (i.e. without rand effects) predictions
-sexrat_pred <- ggpredict(sexrat_age_mod, 
-                         terms = c("age10 [all]", "age_cat")) %>% 
-  as_tibble() %>% 
-  mutate(age = x * 10 + age_senesce,
-         age_cat = factor(group, levels = c("Young", "Old"))) %>% 
-  filter((age_cat == "Young" & age < age_senesce) |
-           (age_cat == "Old" & age >= age_senesce)) %>% 
+# Create offspring sex ratio predictions (population-level and weighted by year)
+sexrat_grid <- expand_grid(
+  age = 4:18,
+  year_fct = factor(1988:2023)
+) %>% 
+  mutate(age10 = (age - age_senesce) / 10,
+         age_cat = factor(ifelse(age < age_senesce, "Young", "Old"),
+                          levels = c("Young", "Old")))
+
+sexrat_years <- count(pup_sex_data, year_fct)
+sexrat_pred <- sexrat_grid %>% 
+  mutate(predicted = predict(sexrat_age_mod, 
+                             newdata = sexrat_grid,
+                             re.form = ~ (1 | year_fct)),
+         predicted_se = suppressWarnings(
+           # warning: "se.fit computation uses an approximation to estimate the sampling distribution of the parameters"
+           predict(sexrat_age_mod, 
+                   newdata = sexrat_grid,
+                   re.form = NA,
+                   se.fit = TRUE),
+         )$se.fit,
+         predicted_pop = predict(sexrat_age_mod, 
+                                 newdata = sexrat_grid,
+                                 re.form = NA),
+         conf_lo = predicted - 1.96 * predicted_se,
+         conf_hi = predicted + 1.96 * predicted_se) %>% 
+  left_join(sexrat_years, by = "year_fct") %>% 
+  group_by(age, age_cat) %>% 
+  summarize(across(c(predicted, predicted_pop, conf_lo, conf_hi), 
+                   \(x) weighted.mean(x, n)),
+            .groups = "drop") %>% 
+  # invert logit
+  mutate(across(c(predicted, predicted_pop, conf_lo, conf_hi), 
+                family(sexrat_age_mod)$linkinv)) %>% 
   left_join(n_sexrat, by = "age")
@@
 # Predictions for each year (random effect)
-sexrat_pred_years <- ggpredict(
+sexrat_years <- ggpredict(
   sexrat_age_mod,
@@
   geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
+  scale_y_continuous("Male offspring", labels = scales::percent) +
   scale_color_manual(values = c("#7fbc41", "#de77ae")) +
   scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
-  scale_y_continuous("Male offspring", labels = scales::percent) +
   labs(x = "Age (years)") +

### 2024-07-12 paper/paperv3.qmd
more edits from marm

@@
 
-ggsave("outputs/repro_senesce.svg", width = 11, height = 6, units = "in")
+ggsave("outputs/repro_senesce.svg", width = 13, height = 6, units = "in")
 ```
@@
 seal_phenology <- sealdat %>% 
+  filter(year >= 2011) %>% 
   mutate(BA = firstobsbreeddoy, 
@@
 
-h1_aic <- do.call(AIC, h1_selective_models) %>% 
-  as_tibble() %>% 
-  mutate(model = c("Base", 
-                   "Selective appearance", 
-                   "Selective disappearance", 
-                   "Selective appearance and disappearance"),
-         dAIC = AIC - min(AIC),
-         rellik = exp(-0.5 * dAIC),
-         weight = rellik / sum(rellik)) %>% 
-  relocate(model)
-
-# Selective appearance model, first repro 95% ci
-h1_selapp_coef <- summary(h1_selective_models[[2]])$coef["firstrepro", ]
-make_95 <- function(ci95) {
-  sprintf("95%% CI: %0.3f - %0.3f", ci95[1], ci95[2])
-}
-h1_firstrepro_95ci <- (h1_selapp_coef["Estimate"] + c(-1.96, 1.96) * h1_selapp_coef["Std. Error"]) %>% 
-  make_95()
-
-h1_aic
-```
-
-```{r}
-#| label: fig-selective-h1
-#| fig-cap: Predicted breeding probabilities of (**A**) selective appearance and (**B**) selective disappearance models.
-The semi-transparent solid lines are predictions of the base model.
-The dashed and dotted lines represent predictions for the 10th and 90th percentiles of observed age at first (**A**) and last (**B**) reproduction.
-In **A**, these correspond to first reproduction at 4 and 7 years old, respectively.
-In **B**, these correspond to last reproduction at 5 and 17 years old, respectively.
-Confidence intervals not shown for clarity.
-#| fig-width: 9
-
-# Compare model predictions
-base_pred <- ggpredict(
-  h1_selective_models[[1]],
-  terms = c("age10 [all]", "age_cat")
-) %>% 
-  as_tibble() %>% 
-  mutate(age = x * 10 + age_senesce,
-         age_cat = factor(group, levels = c("Young", "Old"))) %>% 
-  filter((age_cat == "Young" & age < age_senesce) |
-           (age_cat == "Old" & age >= age_senesce),
-         age <= 20)
-
-# selective appearance for firstrepro=4,6 (10% and 90% of observed)
-selapp_pred <- ggpredict(
-  h1_selective_models[[2]],
-  terms = c("age10 [all]", "age_cat", "firstrepro [4,7]")
-) %>% 
-  as_tibble() %>% 
-  mutate(age = x * 10 + age_senesce,
-         age_cat = factor(group, levels = c("Young", "Old")),
-         firstrepro = facet) %>% 
-  filter((age_cat == "Young" & age < age_senesce) |
-           (age_cat == "Old" & age >= age_senesce),
-         age <= 20)
-
-# selective disappearance for lastrepro=5,17 (10% and 90% of observed) 
-seldis_pred <- ggpredict(
-  h1_selective_models[[4]],
-  terms = c("age10 [all]", "age_cat", "lastrepro [5,17]")
-) %>% 
-  as_tibble() %>% 
-  mutate(age = x * 10 + age_senesce,
-         age_cat = factor(group, levels = c("Young", "Old")),
-         lastrepro = facet) %>% 
-  filter((age_cat == "Young" & age < age_senesce) |
-           (age_cat == "Old" & age >= age_senesce),
-         age <= 20)
-
-# Selective appearance predictions
-fig_selapp <- ggplot(base_pred, aes(age, predicted)) +
-  # Base predictions
-  geom_line(aes(color = age_cat), alpha = 0.5) +
-  # Selective appearance ages 4 and 6
-  geom_line(aes(color = age_cat, 
-                linetype = firstrepro),
-            selapp_pred) +
-  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
-  scale_y_continuous("Breeding", 
-                     labels = scales::percent) +
-  scale_color_manual(values = c("#7fbc41", "#de77ae")) +
-  scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
-  scale_linetype_manual(values = c("dashed", "dotted")) +
-  labs(x = "Age (years)") +
-  theme(legend.position = "none")
-
-fig_seldis <- ggplot(base_pred, aes(age, predicted)) +
-  # Base predictions
-  geom_line(aes(color = age_cat), alpha = 0.5) +
-  # Selective disappearance ages 5 and 17
-  geom_line(aes(color = age_cat, 
-                linetype = lastrepro),
-            seldis_pred) +
-  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
-  scale_y_continuous("Breeding", 
-                     labels = scales::percent) +
-  scale_color_manual(values = c("#7fbc41", "#de77ae")) +
-  scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
-  scale_linetype_manual(values = c("dashed", "dotted")) +
-  labs(x = "Age (years)") +
-  theme(legend.position = "none")
-
-(fig_selapp | fig_seldis) +
-  plot_annotation(tag_levels = "A")
-```
-
-There was no evidence for selective appearance or disappearance influencing the relationship between maternal age and probability of reproduction.
-The selective appearance model (i.e., included a term for age-of-first-reproduction) had a similar AIC weight as the base model.
-However, in the selective appearance model, the coefficient for age at first reproduction was not significantly different than 0 (`r h1_firstrepro_95ci`).
-As a result, the predicted breeding probabilities for animals that first recruited at ages 4 and 7 (the 10th and 90th observed percentiles in the population) were not qualitatively different than the predictions of the base model [@fig-selective-h1].
-Therefore the observed pattern is consistent with the senescence hypothesis (i.e., within-animal change with age).
-
-### H2: Maternal effect senescence
-
-```{r}
-#| label: selective-h2
-
 h2_selective_models <- map(
@@
 )
+
+get_old_coef <- function(model) {
+  old_coef <- summary(model)$coefficients["age10:age_catOld", "Estimate"]
+  old_se <- summary(model)$coefficients["age10:age_catOld", "Std. Error"]
+  ci <- old_coef + c(-1.96, 1.96) * old_se
+  tibble(mean = old_coef, 
+         lwr = ci[1], 
+         upr = ci[2])
+}
+
+h1_selective_coef <- map(h1_selective_models, get_old_coef) %>% 
+  list_rbind() %>% 
+  mutate(model = c("Base", "Appearance", "Disappearance", "Combined"),
+         hypothesis = "H1")
+
+h2_selective_coef <- map(h2_selective_models, get_old_coef) %>% 
+  list_rbind() %>% 
+  mutate(model = c("Base", "Appearance", "Disappearance", "Combined"), 
+         hypothesis = "H2")
+
+selective_coef <- rbind(h1_selective_coef, h2_selective_coef) %>% 
+  mutate(model = factor(model, levels = c("Base", "Appearance", "Disappearance", "Combined")))
+
+ggplot(selective_coef, aes(model, mean, color = hypothesis)) + 
+  geom_pointrange(aes(ymin = lwr, ymax = upr), position = position_dodge(width = 0.2)) +
+  geom_hline(yintercept = 0, linetype = "dashed") +
+  scale_color_manual(values = c("firebrick", "cornflowerblue")) +
+  expand_limits(y = 0) + 
+  labs(y = "Coefficient", x = "Model") + 
+  theme(legend.position = "none")
+
+```
+
+```{r}
+#| label: fig-selective-h1
+#| fig-cap: Predicted breeding probabilities of (**A**) selective appearance and (**B**) selective disappearance models.
+The semi-transparent solid lines are predictions of the base model.
+The dashed and dotted lines represent predictions for the 10th and 90th percentiles of observed age at first (**A**) and last (**B**) reproduction.
+In **A**, these correspond to first reproduction at 4 and 7 years old, respectively.
+In **B**, these correspond to last reproduction at 5 and 17 years old, respectively.
+Confidence intervals not shown for clarity.
+#| fig-width: 9
+
+# Compare model predictions
+base_pred <- ggpredict(
+  h1_selective_models[[1]],
+  terms = c("age10 [all]", "age_cat")
+) %>% 
+  as_tibble() %>% 
+  mutate(age = x * 10 + age_senesce,
+         age_cat = factor(group, levels = c("Young", "Old"))) %>% 
+  filter((age_cat == "Young" & age < age_senesce) |
+           (age_cat == "Old" & age >= age_senesce),
+         age <= 20)
+
+# selective appearance for firstrepro=4,6 (10% and 90% of observed)
+selapp_pred <- ggpredict(
+  h1_selective_models[[2]],
+  terms = c("age10 [all]", "age_cat", "firstrepro [4,7]")
+) %>% 
+  as_tibble() %>% 
+  mutate(age = x * 10 + age_senesce,
+         age_cat = factor(group, levels = c("Young", "Old")),
+         firstrepro = facet) %>% 
+  filter((age_cat == "Young" & age < age_senesce) |
+           (age_cat == "Old" & age >= age_senesce),
+         age <= 20)
+
+# selective disappearance for lastrepro=5,17 (10% and 90% of observed) 
+seldis_pred <- ggpredict(
+  h1_selective_models[[4]],
+  terms = c("age10 [all]", "age_cat", "lastrepro [5,17]")
+) %>% 
+  as_tibble() %>% 
+  mutate(age = x * 10 + age_senesce,
+         age_cat = factor(group, levels = c("Young", "Old")),
+         lastrepro = facet) %>% 
+  filter((age_cat == "Young" & age < age_senesce) |
+           (age_cat == "Old" & age >= age_senesce),
+         age <= 20)
+
+# Selective appearance predictions
+fig_selapp <- ggplot(base_pred, aes(age, predicted)) +
+  # Base predictions
+  geom_line(aes(color = age_cat), alpha = 0.5) +
+  # Selective appearance ages 4 and 6
+  geom_line(aes(color = age_cat, 
+                linetype = firstrepro),
+            selapp_pred) +
+  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
+  scale_y_continuous("Breeding", 
+                     labels = scales::percent) +
+  scale_color_manual(values = c("#7fbc41", "#de77ae")) +
+  scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
+  scale_linetype_manual(values = c("dashed", "dotted")) +
+  labs(x = "Age (years)") +
+  theme(legend.position = "none")
+
+fig_seldis <- ggplot(base_pred, aes(age, predicted)) +
+  # Base predictions
+  geom_line(aes(color = age_cat), alpha = 0.5) +
+  # Selective disappearance ages 5 and 17
+  geom_line(aes(color = age_cat, 
+                linetype = lastrepro),
+            seldis_pred) +
+  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
+  scale_y_continuous("Breeding", 
+                     labels = scales::percent) +
+  scale_color_manual(values = c("#7fbc41", "#de77ae")) +
+  scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
+  scale_linetype_manual(values = c("dashed", "dotted")) +
+  labs(x = "Age (years)") +
+  theme(legend.position = "none")
+
+(fig_selapp | fig_seldis) +
+  plot_annotation(tag_levels = "A")
+```
+
+There was no evidence for selective appearance or disappearance influencing the relationship between maternal age and probability of reproduction.
+The selective appearance model (i.e., included a term for age-of-first-reproduction) had a similar AIC weight as the base model.
+However, in the selective appearance model, the coefficient for age at first reproduction was not significantly different than 0 (`r h1_firstrepro_95ci`).
+As a result, the predicted breeding probabilities for animals that first recruited at ages 4 and 7 (the 10th and 90th observed percentiles in the population) were not qualitatively different than the predictions of the base model [@fig-selective-h1].
+Therefore the observed pattern is consistent with the senescence hypothesis (i.e., within-animal change with age).
+
+### H2: Maternal effect senescence
+
+```{r}
+#| label: selective-h2
 
@@
 
-missing_mod <- glmer(unobserved ~ age10 + (1 | year) + (1 | animalID), 
+missing_mod <- glmer(unobserved ~ age10 + (1 | year_fct) + (1 | animalID), 
                      missing_seals, 
                      family = binomial)
-missing_pred <- ggpredict(missing_mod,
-                          terms = "age10 [all]") %>% 
-  as_tibble() %>% 
-  mutate(age = x * 10 + age_senesce,
-         unobserved = predicted)
-
-missing_age_coef <- sprintf(
-  "%0.3f \u00B1 %0.3f",
-  summary(missing_mod)$coef["age10", "Estimate"],
-  summary(missing_mod)$coef["age10", "Std. Error"] * 1.96
-)
+
+summary(missing_mod)
 
@@
 See text for model details.
+
+missing_years <- count(missing_seals, year_fct)
+missing_pred <- expand_grid(age = 5:19, 
+                            year_fct = factor(1989:2022)) %>% 
+  mutate(age10 = (age - age_senesce) / 10) %>% 
+  mutate(predicted = predict(missing_mod, 
+                             newdata = missing_grid,
+                             re.form = ~ (1 | year_fct)),
+         se = predict(missing_mod, 
+                      newdata = missing_grid,
+                      se.fit = TRUE,
+                      re.form = NA)$se.fit,
+         conf.low = predicted - 1.96 * se,
+         conf.high = predicted + 1.96 * se) %>% 
+  left_join(missing_years, by = "year_fct") %>% 
+  group_by(age) %>% 
+  summarize(across(c(predicted, conf.low, conf.high), mean)) %>% 
+  mutate(across(c(predicted, conf.low, conf.high), family(missing_mod)$linkinv))
+
+missing_pred_years <- expand_grid(age = 5:19, 
+                            year_fct = factor(1989:2022)) %>% 
+  mutate(age10 = (age - age_senesce) / 10) %>% 
+  mutate(predicted = predict(missing_mod, 
+                             newdata = missing_grid,
+                             type = "response",
+                             re.form = ~ (1 | year_fct)))
+
+missing_pred
 
@@
   ggplot(aes(age)) +
+  geom_line(aes(y = predicted, group = year_fct), 
+            missing_pred_years, 
+            alpha = 0.2) +
   geom_ribbon(aes(ymin = conf.low, ymax = conf.high),
@@
               fill = "#de77ae") +
-  geom_line(aes(y = unobserved), 
+  geom_line(aes(y = predicted),
             missing_pred,
@@
   scale_y_continuous("Unobserved proportion", 
-                     labels = scales::label_percent(accuracy = 1))
+                     labels = scales::label_percent(accuracy = 1),
+                     limits = c(0, NA))
 ```

### 2024-07-15 paper/paperv3.qmd
more small changes

@@
   count(age)
-
-# # Create POPULATION-LEVEL (i.e. without rand effects) predictions
-# breed_pred_pop <- ggpredict(breed_age_mod, 
-#                             terms = c("age10 [all]", "age_cat")) %>% 
-#   as_tibble() %>% 
-#   mutate(age = x * 10 + age_senesce,
-#          age_cat = factor(group, levels = c("Young", "Old"))) %>% 
-#   filter((age_cat == "Young" & age < age_senesce) |
-#            (age_cat == "Old" & age >= age_senesce),
-#          age <= 20) %>% 
-#   left_join(n_breed, by = "age")
 
@@
 
-(fig_selapp | fig_seldis) +
-  plot_annotation(tag_levels = "A")
 ```
@@
 The selective appearance model (i.e., included a term for age-of-first-reproduction) had a similar AIC weight as the base model.
-However, in the selective appearance model, the coefficient for age at first reproduction was not significantly different than 0 (`r h1_firstrepro_95ci`).
+However, in the selective appearance model, the coefficient for age at first reproduction was not significantly different than 0.
 As a result, the predicted breeding probabilities for animals that first recruited at ages 4 and 7 (the 10th and 90th observed percentiles in the population) were not qualitatively different than the predictions of the base model [@fig-selective-h1].
@@
 ### H2: Maternal effect senescence
-
-```{r}
-#| label: selective-h2
-
-h2_aic <- do.call(AIC, h2_selective_models) %>% 
-  as_tibble() %>% 
-  mutate(model = c("Base", 
-                   "Selective appearance", 
-                   "Selective disappearance", 
-                   "Selective appearance and disappearance"),
-         dAIC = AIC - min(AIC),
-         rellik = exp(-0.5 * dAIC),
-         weight = rellik / sum(rellik)) %>% 
-  relocate(model)
-
-# Selective appearance model, first repro 95% ci
-h2_selapp_coef <- summary(h2_selective_models[[2]])$coef["firstrepro", ]
-h2_firstrepro_95ci <- (h2_selapp_coef["Estimate"] + c(-1.96, 1.96) * h2_selapp_coef["Std. Error"]) %>% 
-  make_95()
-
-h2_aic
-
-```
 
@@
 
-fig_selapp | fig_seldis +
-  plot_annotation(tag_levels = "A")
 ```
@@
 The selective appearance model (i.e., included a term for age-of-first-reproduction) had a slightly greater AIC weight than the base model.
-However, in the selective appearance model, the coefficient for age at first reproduction was not significantly different than 0 (`r h2_firstrepro_95ci`).
+However, in the selective appearance model, the coefficient for age at first reproduction was not significantly different than 0.
 As a result, the predicted breeding probabilities for animals that first recruited at ages 4 and 7 (the 10th and 90th observed percentiles in the population) were not qualitatively different than the predictions of the base model [@fig-selective-h2].
@@
 Then we estimated the effect of age on probability of an animal being unobserved using a GLMM (binomial response, logit link) with random effects for year and animal ([@fig-unobserved]).
-The coefficient for age was `r missing_age_coef` (*p*\<0.0001).
+The coefficient for age was r missing_age_coef (*p*\<0.0001).
 
@@
   mutate(predicted = predict(missing_mod, 
-                             newdata = missing_grid,
+                             newdata = .,
                              re.form = ~ (1 | year_fct)),
          se = predict(missing_mod, 
-                      newdata = missing_grid,
+                      newdata = .,
                       se.fit = TRUE,
@@
   mutate(predicted = predict(missing_mod, 
-                             newdata = missing_grid,
+                             newdata = .,
                              type = "response",
@@
 
-missing_pred
+missing_pred_pop <- tibble(age = 5:19,
+                           age10 = (age - age_senesce) / 10) %>% 
+  mutate(predicted = predict(missing_mod, 
+                             newdata = .,
+                             type = "response",
+                             re.form = NA))
 
@@
             linewidth = 2) +
+  geom_line(aes(y = predicted),
+            missing_pred_pop,
+            linewidth = 1.5,
+            linetype = "dotted",
+            alpha = 0.5) +
   geom_pointrange(aes(y = unobserved_mean, 

### 2024-07-15 paper/paperv3.qmd
H1, H2 binom test for CIs

--- 2024-07-12 paper/paperv3.qmd
@@
   summarize(n_observed = n(),
-            perc_breed = sum(observed_int) / n_observed,
-            se = (perc_breed * (1 - perc_breed) / n_observed)^0.5) %>%
+            n_breed = sum(observed_int),
+            perc_breed = n_breed / n_observed,
+            breed_lwr = binom.test(n_breed, n_observed)$conf.int[1],
+            breed_upr = binom.test(n_breed, n_observed)$conf.int[2]) %>%
   filter(age <= 20)
@@
             alpha = 0.5) +
+  # Summarized raw data
   geom_pointrange(aes(y = perc_breed,
-                      ymin = perc_breed - 1.96 * se,
-                      ymax = perc_breed + 1.96 * se),
+                      ymin = breed_lwr,
+                      ymax = breed_upr),
                   breed_nonbreed) +
   geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
-  scale_y_continuous("Breeding probability", labels = scales::percent) +
+  scale_y_continuous("Breeding probability", 
+                     labels = scales::percent,
+                     limits = c(NA, 1.02),
+                     breaks = seq(0.4, 1.0, by = 0.2)) +
   scale_color_manual(values = c("#7fbc41", "#de77ae")) +
@@
   theme(legend.position = "none") + 
-  geom_text(aes(label = n, y = 1.1))
+  geom_text(aes(label = n), y = 1, vjust = -0.5)
 
 ggsave("outputs/repro_senesce.svg", width = 13, height = 6, units = "in")
+```
+
+```{r}
+#| label: h1-pred-deets
+
+tibble(age = c(11, 19),
+       age10 = (age - age_senesce) / 10,
+       age_cat = "Old") %>% 
+  # logit-scale predictions
+  mutate(predicted = predict(breed_age_mod, 
+                             newdata = .,
+                             re.form = NA),
+         se = suppressWarnings(predict(breed_age_mod, 
+                                       newdata = .,
+                                       se.fit = TRUE,
+                                       re.form = NA)$se.fit),
+         pred_lwr = predicted - 1.96 * se,
+         pred_upr = predicted + 1.96 * se,
+         # Invert the logit
+         across(c(predicted, pred_upr, pred_lwr), 
+                family(breed_age_mod)$linkinv))
 ```
@@
   filter(pupyearborn <= 2016) %>% 
+  filter(age < 20) %>% 
   mutate(pup_surv_int = as.integer(pup_survived),
@@
 # Summarize the raw data
-surv_recr_summ <- surv_recr_data %>% 
+surv_summ <- surv_recr_data %>% 
   group_by(age) %>% 
@@
   summarize(n_pups = n(), 
-            n_fpups = sum(pupsex == "F"),
-            perc_surv = sum(pup_surv_int) / n_pups, 
-            se_surv = (perc_surv * (1 - perc_surv) / n_pups)^0.5,
-            perc_recr = sum(pup_recr_int) / n_fpups, 
-            se_recr = (perc_recr * (1 - perc_recr) / n_fpups)^0.5)
+            n_surv = sum(pup_surv_int),
+            perc_surv = n_surv / n_pups, 
+            surv_lwr = binom.test(n_surv, n_pups)$conf.int[1],
+            surv_upr = binom.test(n_surv, n_pups)$conf.int[2])
+recr_summ <- surv_recr_data %>% 
+  filter(pupsex == "F") %>% 
+  group_by(age) %>% 
+  summarize(n_fpups = n(), 
+            n_recr = sum(pup_recr_int),
+            perc_recr = n_recr / n_fpups, 
+            recr_lwr = binom.test(n_recr, n_fpups)$conf.int[1],
+            recr_upr = binom.test(n_recr, n_fpups)$conf.int[2])
 
@@
 n_surv <- surv_recr_data %>% 
-  group_by(age) %>% 
   filter(age <= 19) %>% 
@@
   geom_pointrange(aes(y = perc_surv, 
-                      ymin = perc_surv - se_surv, 
-                      ymax = perc_surv + se_surv),
-                  surv_recr_summ) +
+                      ymin = surv_lwr, 
+                      ymax = surv_upr),
+                  surv_summ) +
   # Senescence threshold
@@
                      labels = scales::percent,
-                     limits = c(0, 0.68)) +
+                     limits = c(0, 1)) +
   scale_color_manual(values = c("#7fbc41", "#de77ae")) +
@@
   theme(legend.position = "none") +
-  geom_text(aes(label = n, y = .65))
+  geom_text(aes(label = n), y = 1, vjust = 1)
 
@@
   geom_pointrange(aes(y = perc_recr, 
-                      ymin = perc_recr - se_recr, 
-                      ymax = perc_recr + se_recr),
-                  surv_recr_summ) +
+                      ymin = recr_lwr, 
+                      ymax = recr_upr),
+                  recr_summ) +
   geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
@@
   theme(legend.position = "none") + 
-  geom_text(aes(label = n, y = .65))
+  geom_text(aes(label = n), y = 1, vjust = 1)
 
@@
   summarize(n_observed = n(), 
-            perc_male = mean(is_male), 
+            n_male = sum(is_male),
+            perc_male = n_male / n_observed, 
+            male_lwr = binom.test()
             se = (perc_male * (1 - perc_male) / n_observed)^0.5) %>% 
@@
 Therefore, it is biologically implausible that the unobserved seals were disproportionately breeders and our estimates for reproductive senescence are, if anything, conservative.
+
+## Threshold vs

### 2024-07-15 paper/paperv3.qmd
H3 binom conf int

--- 2024-07-15 paper/paperv3.qmd
@@
             perc_male = n_male / n_observed, 
-            male_lwr = binom.test()
-            se = (perc_male * (1 - perc_male) / n_observed)^0.5) %>% 
+            male_lwr = binom.test(n_male, n_observed)$conf.int[1], 
+            male_upr = binom.test(n_male, n_observed)$conf.int[2]) %>% 
   filter(age <= 20)
@@
   geom_pointrange(aes(y = perc_male, 
-                      ymin = perc_male - se, 
-                      ymax = perc_male + se),
+                      ymin = male_lwr, 
+                      ymax = male_upr),
                   sexrat_summ) +
   geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
-  scale_y_continuous("Male offspring", labels = scales::percent) +
+  scale_y_continuous("Male offspring", 
+                     labels = scales::percent,
+                     limits = c(NA, 0.97)) +
   scale_color_manual(values = c("#7fbc41", "#de77ae")) +
@@
   theme(legend.position = "none") +
-  geom_text(aes(label = n, y = .95))
+  geom_text(aes(label = n), y = .97, vjust = 0)
 

### 2024-07-15 paper/paperv3.qmd
adding 11/19 comparisons for h2a

--- 2024-07-15 paper/paperv3.qmd
@@
 
+#Overall decline 11-19
 tibble(age = c(11, 19),
@@
                 family(breed_age_mod)$linkinv))
+
+#Difference between years 
+best_worst_yr <- ranef(breed_age_mod)$year_fct %>% 
+  as_tibble(rownames = "year_fct") %>% 
+  arrange(`(Intercept)`) %>% 
+  slice(1, nrow(.))
+
+tibble(age = c(11, 19),
+       age10 = (age - age_senesce) / 10,
+       age_cat = "Old") %>% 
+  cross_join(best_worst_yr) %>% 
+  mutate(predicted = predict(breed_age_mod, 
+                             newdata = .,
+                             type = "response", 
+                             re.form = ~ (1 | year_fct)))
 ```
@@
 Sample size for offspring recruitment was n = `r nrow(n_recruit)` observations of `r n_distinct(n_recruit$animalID)` female individuals.
+
+```{r}
+#| label: h2-pred-deets
+
+#Overall decline 11-19
+tibble(age = c(11, 19),
+       age10 = (age - age_senesce) / 10,
+       age_cat = "Old") %>% 
+  # logit-scale predictions
+  mutate(predicted = predict(pup_surv_mod, 
+                             newdata = .,
+                             re.form = NA),
+         se = suppressWarnings(predict(pup_surv_mod, 
+                                       newdata = .,
+                                       se.fit = TRUE,
+                                       re.form = NA)$se.fit),
+         pred_lwr = predicted - 1.96 * se,
+         pred_upr = predicted + 1.96 * se,
+         # Invert the logit
+         across(c(predicted, pred_upr, pred_lwr), 
+                family(pup_surv_mod)$linkinv))
+
+#Difference between years 
+best_worst_yr_surv <- ranef(pup_surv_mod)$year_fct %>% 
+  as_tibble(rownames = "year_fct") %>% 
+  arrange(`(Intercept)`) %>% 
+  slice(1, nrow(.))
+
+tibble(age = c(11, 19),
+       age10 = (age - age_senesce) / 10,
+       age_cat = "Old") %>% 
+  cross_join(best_worst_yr_surv) %>% 
+  mutate(predicted = predict(pup_surv_mod, 
+                             newdata = .,
+                             type = "response", 
+                             re.form = ~ (1 | year_fct)))
+```
 

### 2024-07-15 paper/paperv3.qmd
adding histogram for sample size of phenology

@@
 B: Raw data for breeding and molting durations (mean ± se) and predictions (mean and 95% CI) from negative binomial weighted by number of seals obersved per year."
-#| fig-width: 4
+#| fig-width: 9
 #| fig-height: 9
@@
   count(age)
+
+phen_dist <- seal_phenology %>% 
+  # group_by(age) %>% 
+  count(age) %>% 
+  ggplot(aes(x = age, y = n)) + 
+  geom_col(fill = "grey50") + 
+  labs(x = "Age (Years)", y = "Number of individuals") + 
+  theme()
 
@@
 
-(annual_plot | breed_duration_plot | molt_duration_plot) + plot_annotation(tag_levels = "A") + plot_layout(ncol = 1, nrow = 3)
+(annual_plot | phen_dist | breed_duration_plot | molt_duration_plot) + plot_annotation(tag_levels = "A") + plot_layout(ncol = 2, nrow = 2)
 ```

### 2024-07-15 paper/paperv3.qmd
quadratic vs threshold

--- 2024-07-15 paper/paperv3.qmd
@@
 
-## Threshold vs
+## Breakpoint vs quadratic
+
+### H1: Reproductive senescence
+
+```{r}
+#| label: quad-h1
+
+breed_age_quad <- glmer(
+  observed_int ~ poly(age10, 2, raw = TRUE) + (1 | animalID) + (1 | year_fct),
+  sealdat,
+  family = "binomial",
+  control = glmerControl(optimizer = "bobyqa")
+)
+
+h1_age10_peak <- -summary(breed_age_quad)$coefficients[2, 1] / 
+  (2 * summary(breed_age_quad)$coefficients[3, 1])
+h1_age_peak <- h1_age10_peak * 10 + age_senesce
+
+summary(breed_age_quad)$coefficients
+
+sprintf("Age peak (H1) = %0.1f", h1_age_peak)
+
+```
+
+### H2: Maternal effect senescence
+
+```{r}
+#| label: quad-h2
+
+# H2a
+surv_age_quad <- glmer(
+  pup_surv_int ~ poly(age10, 2, raw = TRUE) + (1 | animalID) + (1 | year_fct),
+  surv_recr_data,
+  family = "binomial",
+  control = glmerControl(optimizer = "bobyqa")
+)
+
+h2a_age10_peak <- -summary(surv_age_quad)$coefficients[2, 1] / 
+  (2 * summary(surv_age_quad)$coefficients[3, 1])
+h2a_age_peak <- h2a_age10_peak * 10 + age_senesce
+
+summary(surv_age_quad)$coefficients
+
+sprintf("Age peak (H2a) = %0.1f", h2a_age_peak)
+
+# H2b
+
+recr_age_quad <- glmer(
+  pup_recr_int ~ poly(age10, 2, raw = TRUE) + (1 | year_fct),
+  filter(surv_recr_data, pupsex == "F"),
+  family = "binomial",
+  control = glmerControl(optimizer = "bobyqa")
+)
+
+h2b_age10_peak <- -summary(recr_age_quad)$coefficients[2, 1] / 
+  (2 * summary(recr_age_quad)$coefficients[3, 1])
+h2b_age_peak <- h2b_age10_peak * 10 + age_senesce
+
+summary(recr_age_quad)$coefficients
+
+sprintf("Age peak (H2b) = %0.1f", h2b_age_peak)
+
+```

### 2024-07-15 paper/paperv3.qmd
selective (dis)appearance, bio vs chrono age

@@
                             max(age[observed_int == 1]),
-                            NA)) %>% 
+                            NA),
+         lastobs = ifelse(max(year) < 2020, max(age), NA)) %>% 
   ungroup()
@@
 
-# # Create POPULATION-LEVEL (i.e. without rand effects) predictions
-# breed_pred_pop <- ggpredict(breed_age_mod, 
-#                             terms = c("age10 [all]", "age_cat")) %>% 
-#   as_tibble() %>% 
-#   mutate(age = x * 10 + age_senesce,
-#          age_cat = factor(group, levels = c("Young", "Old"))) %>% 
-#   filter((age_cat == "Young" & age < age_senesce) |
-#            (age_cat == "Old" & age >= age_senesce),
-#          age <= 20) %>% 
-#   left_join(n_breed, by = "age")
-
 # Create ANNUAL (i.e. rand effect of year only) predictions
@@
 We tested whether our significant results (reproductive senescence, maternal effect senescence in pup survival) were influenced by selective appearance and disappearance, following the methods of @vandepol2006, by adding terms for age at first and last reproduction, respectively.
-We used AIC to compare the base model to models with selective appearance and disappearance.
-
-### H1: Reproductive senescence
-
-```{r}
-#| label: selective-h1
+If our results are due to selective appearance/disappearance instead of senescence, then the 95% CI for the coefficient of age in the old age class in these models will overlap 0.
+
+```{r}
+#| label: fig-selective
+#| fig-cap: The coefficient for age in the old age class in models accounting for selective appearance and disappearance.
+Red and blue indicate  reproductive senescence (H1) and maternal effect senescence (H2a), respectively.
+Compared to the base model, the appearance model includes a term for age at first reproduction, the disappearance model includes a term for age at last reproduction, and the combined model includes both.
 
@@
 
-```{r}
-#| label: fig-selective-h1
-#| fig-cap: Predicted breeding probabilities of (**A**) selective appearance and (**B**) selective disappearance models.
-The semi-transparent solid lines are predictions of the base model.
-The dashed and dotted lines represent predictions for the 10th and 90th percentiles of observed age at first (**A**) and last (**B**) reproduction.
-In **A**, these correspond to first reproduction at 4 and 7 years old, respectively.
-In **B**, these correspond to last reproduction at 5 and 17 years old, respectively.
-Confidence intervals not shown for clarity.
-#| fig-width: 9
-
-# Compare model predictions
-base_pred <- ggpredict(
-  h1_selective_models[[1]],
-  terms = c("age10 [all]", "age_cat")
-) %>% 
-  as_tibble() %>% 
-  mutate(age = x * 10 + age_senesce,
-         age_cat = factor(group, levels = c("Young", "Old"))) %>% 
-  filter((age_cat == "Young" & age < age_senesce) |
-           (age_cat == "Old" & age >= age_senesce),
-         age <= 20)
-
-# selective appearance for firstrepro=4,6 (10% and 90% of observed)
-selapp_pred <- ggpredict(
-  h1_selective_models[[2]],
-  terms = c("age10 [all]", "age_cat", "firstrepro [4,7]")
-) %>% 
-  as_tibble() %>% 
-  mutate(age = x * 10 + age_senesce,
-         age_cat = factor(group, levels = c("Young", "Old")),
-         firstrepro = facet) %>% 
-  filter((age_cat == "Young" & age < age_senesce) |
-           (age_cat == "Old" & age >= age_senesce),
-         age <= 20)
-
-# selective disappearance for lastrepro=5,17 (10% and 90% of observed) 
-seldis_pred <- ggpredict(
-  h1_selective_models[[4]],
-  terms = c("age10 [all]", "age_cat", "lastrepro [5,17]")
-) %>% 
-  as_tibble() %>% 
-  mutate(age = x * 10 + age_senesce,
-         age_cat = factor(group, levels = c("Young", "Old")),
-         lastrepro = facet) %>% 
-  filter((age_cat == "Young" & age < age_senesce) |
-           (age_cat == "Old" & age >= age_senesce),
-         age <= 20)
-
-# Selective appearance predictions
-fig_selapp <- ggplot(base_pred, aes(age, predicted)) +
-  # Base predictions
-  geom_line(aes(color = age_cat), alpha = 0.5) +
-  # Selective appearance ages 4 and 6
-  geom_line(aes(color = age_cat, 
-                linetype = firstrepro),
-            selapp_pred) +
-  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
-  scale_y_continuous("Breeding", 
-                     labels = scales::percent) +
-  scale_color_manual(values = c("#7fbc41", "#de77ae")) +
-  scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
-  scale_linetype_manual(values = c("dashed", "dotted")) +
-  labs(x = "Age (years)") +
-  theme(legend.position = "none")
-
-fig_seldis <- ggplot(base_pred, aes(age, predicted)) +
-  # Base predictions
-  geom_line(aes(color = age_cat), alpha = 0.5) +
-  # Selective disappearance ages 5 and 17
-  geom_line(aes(color = age_cat, 
-                linetype = lastrepro),
-            seldis_pred) +
-  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
-  scale_y_continuous("Breeding", 
-                     labels = scales::percent) +
-  scale_color_manual(values = c("#7fbc41", "#de77ae")) +
-  scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
-  scale_linetype_manual(values = c("dashed", "dotted")) +
-  labs(x = "Age (years)") +
-  theme(legend.position = "none")
-
-(fig_selapp | fig_seldis) +
-  plot_annotation(tag_levels = "A")
-```
-
-There was no evidence for selective appearance or disappearance influencing the relationship between maternal age and probability of reproduction.
-The selective appearance model (i.e., included a term for age-of-first-reproduction) had a similar AIC weight as the base model.
-However, in the selective appearance model, the coefficient for age at first reproduction was not significantly different than 0 (`r h1_firstrepro_95ci`).
-As a result, the predicted breeding probabilities for animals that first recruited at ages 4 and 7 (the 10th and 90th observed percentiles in the population) were not qualitatively different than the predictions of the base model [@fig-selective-h1].
-Therefore the observed pattern is consistent with the senescence hypothesis (i.e., within-animal change with age).
-
-### H2: Maternal effect senescence
-
-```{r}
-#| label: selective-h2
-
-h2_aic <- do.call(AIC, h2_selective_models) %>% 
-  as_tibble() %>% 
-  mutate(model = c("Base", 
-                   "Selective appearance", 
-                   "Selective disappearance", 
-                   "Selective appearance and disappearance"),
-         dAIC = AIC - min(AIC),
-         rellik = exp(-0.5 * dAIC),
-         weight = rellik / sum(rellik)) %>% 
-  relocate(model)
-
-# Selective appearance model, first repro 95% ci
-h2_selapp_coef <- summary(h2_selective_models[[2]])$coef["firstrepro", ]
-h2_firstrepro_95ci <- (h2_selapp_coef["Estimate"] + c(-1.96, 1.96) * h2_selapp_coef["Std. Error"]) %>% 
-  make_95()
-
-h2_aic
-
-```
-
-```{r}
-#| label: fig-selective-h2
-#| fig-cap: Predicted breeding probabilities of (**A**) selective appearance and (**B**) selective disappearance models.
-The semi-transparent solid lines are predictions of the base model.
-The dashed and dotted lines represent predictions for the 10th and 90th percentiles of observed age at first (**A**) and last (**B**) reproduction.
-In **A**, these correspond to first reproduction at 4 and 7 years old, respectively.
-In **B**, these correspond to last reproduction at 5 and 17 years old, respectively.
-Confidence intervals not shown for clarity.
-#| fig-width: 9
-
-# Compare model predictions
-base_pred <- ggpredict(
-  h2_selective_models[[1]],
-  terms = c("age10 [all]", "age_cat")
-) %>% 
-  as_tibble() %>% 
-  mutate(age = x * 10 + age_senesce,
-         age_cat = factor(group, levels = c("Young", "Old"))) %>% 
-  filter((age_cat == "Young" & age < age_senesce) |
-           (age_cat == "Old" & age >= age_senesce),
-         age <= 20)
-
-# selective appearance for firstrepro=4,6 (10% and 90% of observed)
-selapp_pred <- ggpredict(
-  h2_selective_models[[2]],
-  terms = c("age10 [all]", "age_cat", "firstrepro [4,7]")
-) %>% 
-  as_tibble() %>% 
-  mutate(age = x * 10 + age_senesce,
-         age_cat = factor(group, levels = c("Young", "Old")),
-         firstrepro = facet) %>% 
-  filter((age_cat == "Young" & age < age_senesce) |
-           (age_cat == "Old" & age >= age_senesce),
-         age <= 20)
-
-# selective disappearance for lastrepro=5,17 (10% and 90% of observed) 
-seldis_pred <- ggpredict(
-  h2_selective_models[[4]],
-  terms = c("age10 [all]", "age_cat", "lastrepro [5,17]")
-) %>% 
-  as_tibble() %>% 
-  mutate(age = x * 10 + age_senesce,
-         age_cat = factor(group, levels = c("Young", "Old")),
-         lastrepro = facet) %>% 
-  filter((age_cat == "Young" & age < age_senesce) |
-           (age_cat == "Old" & age >= age_senesce),
-         age <= 20)
-
-# Selective appearance predictions
-fig_selapp <- ggplot(base_pred, aes(age, predicted)) +
-  # Base predictions
-  geom_line(aes(color = age_cat), alpha = 0.5) +
-  # Selective appearance ages 4 and 6
-  geom_line(aes(color = age_cat, 
-                linetype = firstrepro),
-            selapp_pred) +
-  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
-  scale_y_continuous("Offspring survival", 
-                     labels = scales::percent) +
-  scale_color_manual(values = c("#7fbc41", "#de77ae")) +
-  scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
-  scale_linetype_manual(values = c("dashed", "dotted")) +
-  labs(x = "Maternal age (years)") +
-  theme(legend.position = "none")
-
-fig_seldis <- ggplot(base_pred, aes(age, predicted)) +
-  # Base predictions
-  geom_line(aes(color = age_cat), alpha = 0.5) +
-  # Selective disappearance ages 5 and 17
-  geom_line(aes(color = age_cat, 
-                linetype = lastrepro),
-            seldis_pred) +
-  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
-  scale_y_continuous("Offspring survival", 
-                     labels = scales::percent) +
-  scale_color_manual(values = c("#7fbc41", "#de77ae")) +
-  scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
-  scale_linetype_manual(values = c("dashed", "dotted")) +
-  labs(x = "Maternal age (years)") +
-  theme(legend.position = "none")
-
-fig_selapp | fig_seldis +
-  plot_annotation(tag_levels = "A")
-```
-
-There was no evidence for selective appearance or disappearance influencing the relationship between maternal age and offspring survival.
-The selective appearance model (i.e., included a term for age-of-first-reproduction) had a slightly greater AIC weight than the base model.
-However, in the selective appearance model, the coefficient for age at first reproduction was not significantly different than 0 (`r h2_firstrepro_95ci`).
-As a result, the predicted breeding probabilities for animals that first recruited at ages 4 and 7 (the 10th and 90th observed percentiles in the population) were not qualitatively different than the predictions of the base model [@fig-selective-h2].
-Therefore the observed pattern is consistent with the senescence hypothesis (i.e., within-animal change with age).
+There was no evidence for selective appearance or disappearance influencing the relationship between maternal age and probability of reproduction or offspring survival.
+The 95% CI of the coefficient for age in the old age class did not overlap 0 for either hypothesis in any of the selective (dis)appearance models [@fig-selective].
+Therefore the observed patterns are consistent with the senescence hypothesis (i.e., within-animal change with age).
 
@@
   drop_na(lastrepro) %>% 
-  mutate(bioage = age - lastrepro,
+  mutate(bioage = age - lastobs,
          bioage10 = (bioage - bioage_senesce) / 10,
@@
 This required us to limit our sample to animals with known longevity (i.e., last observation in 2020 or earlier), which reduced our sample size to `r nrow(sealdat2)` observations of `r n_distinct(sealdat2$animalID)` distinct animals.
-Subsequently, we fit the model for breakpoints at all biological ages -10 to -3 years, to assess whether our results were sensitive to the choice of breakpoint.
+Subsequently, we fit the model for breakpoints at all biological ages -12 to -3 years, to assess whether our results were sensitive to the choice of breakpoint.
 Finally, we re-fit the chronological reproductive senescence model using the known-longevity dataset and compared chronological versus biological age models using AIC.
@@
 
+# Biological age model
 breed_bioage_mod <- glmer(
@@
 
-bioage_pval <- sprintf("%0.3f", summary(breed_bioage_mod)$coef["bioage10:bioage_catOld", "Pr(>|z|)"])
-
-breed_age_mod2 <- glmer(
-  observed_int ~ age10 : age_cat + (1 | animalID) + (1 | year_fct),
-  drop_na(sealdat, lastrepro),
-  family = "binomial",
-  control = glmerControl(optimizer = "bobyqa")
-)
-
-p_by_tau <- map_dbl(-10:-3, \(tau) {
+old_z <- summary(breed_bioage_mod)$coefficients["bioage10:bioage_catOld", "z value"]
+bioage_pval <- sprintf("%0.3f", pnorm(-old_z))
+
+# Breakpoint comparison
+p_by_tau <- map_dbl(-12:-3, \(tau) {
   mod <- sealdat %>%
     drop_na(lastrepro) %>%
-    mutate(bioage = age - lastrepro,
+    mutate(bioage = age - lastobs,
            bioage10 = (bioage - tau) / 10,
@@
     )
-  summary(mod)$coef["bioage10:bioage_catOld", "Pr(>|z|)"]
+  old_z <- summary(mod)$coef["bioage10:bioage_catOld", "z value"] 
+  pnorm(old_z)
 })
+
+tibble(`Biological age threshold` = -12:-3,
+       `p value` = p_by_tau)
+
+# Model comparison
+breed_age_mod2 <- glmer(
+  observed_int ~ age10 : age_cat + (1 | animalID) + (1 | year_fct),
+  drop_na(sealdat2, lastrepro),
+  family = "binomial",
+  control = glmerControl(optimizer = "bobyqa")
+)
 
@@
   summarize(n_obs = n(), 
-            perc_breed = mean(observed_int), 
-            se = sqrt(perc_breed * (1 - perc_breed) / n_obs))
-
-# Create POPULATION-LEVEL (i.e. without rand effects) predictions
-bioage_pred_pop <- ggpredict(breed_bioage_mod, 
-                             terms = c("bioage10 [all]", "bioage_cat")) %>% 
-  as_tibble() %>% 
-  mutate(bioage = x * 10 + bioage_senesce,
-         bioage_cat = factor(group, levels = c("Young", "Old"))) %>% 
-  filter((bioage_cat == "Young" & bioage < bioage_senesce) |
-           (bioage_cat == "Old" & bioage >= bioage_senesce))
-
-# Create WEIGHTED-AVERAGE predictions using years' random effects
-bioage_pred_wgt <- ggpredict(
+            n_breed = sum(observed_int),
+            perc_breed = n_breed / n_obs, 
+            breed_lwr = binom.test(n_breed, n_obs)$conf.int[1], 
+            breed_upr = binom.test(n_breed, n_obs)$conf.int[2])
+
+# Create ANNUAL (i.e. rand effect of year only) predictions
+bioage_pred_years <- ggpredict(
   breed_bioage_mod, 
-  terms = c("bioage10 [all]", "bioage_cat", "year_fct [all]"),
-  type = "random",
-  interval = "confidence"
+  terms = c("bioage10 [all]", "bioage_cat", "year_fct"),
+  type = "random"
 ) %>% 
@@
          bioage_cat = factor(group, levels = c("Young", "Old")),
-         year_fct = facet) %>% 
+         year_fct = factor(facet)) %>% 
   filter((bioage_cat == "Young" & bioage < bioage_senesce) |
-           (bioage_cat == "Old" & bioage >= bioage_senesce)) %>% 
-  left_join(seal_years, by = "year_fct") %>% 
+           (bioage_cat == "Old" & bioage >= bioage_senesce),
+         bioage <= 20)
+
+# Create WEIGHTED-AVERAGE predictions using years' random effects
+# ggpredict can't do that for confidence intervals
+# create prediction grid
+bioage_grid <- expand_grid(
+  bioage = -17:0,
+  year_fct = factor(1988:2019)
+) %>% 
+  mutate(bioage10 = (bioage - bioage_senesce) / 10,
+         bioage_cat = factor(ifelse(bioage < bioage_senesce, "Young", "Old"),
+                          levels = c("Young", "Old")))
+
+bioage_years <- count(sealdat2, year_fct)
+bioage_pred_wgt <- bioage_grid %>% 
+  mutate(predicted = predict(breed_bioage_mod, 
+                             newdata = bioage_grid,
+                             re.form = ~ (1 | year_fct)),
+         predicted_se = suppressWarnings(
+           # warning: "se.fit computation uses an approximation to estimate the sampling distribution of the parameters"
+           predict(breed_bioage_mod, 
+                   newdata = bioage_grid,
+                   re.form = NA,
+                   se.fit = TRUE)
+         )$se.fit,
+         predicted_pop = predict(breed_bioage_mod, 
+                                 newdata = bioage_grid,
+                                 re.form = NA),
+         conf_lo = predicted - 1.96 * predicted_se,
+         conf_hi = predicted + 1.96 * predicted_se) %>% 
+  left_join(bioage_years, by = "year_fct") %>% 
   group_by(bioage, bioage_cat) %>% 
-  summarize(predicted = weighted.mean(predicted, n),
-            .groups = "drop")
-
-ggplot(bioage_pred_pop, aes(bioage, predicted)) +
-  geom_ribbon(aes(fill = bioage_cat, ymin = conf.low, ymax = conf.high), 
-              alpha = 0.2) +
-  geom_line(aes(color = bioage_cat), linewidth = 1.2, linetype = "dashed") +
+  summarize(across(c(predicted, predicted_pop, conf_lo, conf_hi), 
+                   \(x) weighted.mean(x, n)),
+            .groups = "drop") %>% 
+  # invert logit
+  mutate(across(c(predicted, predicted_pop, conf_lo, conf_hi), 
+                family(breed_bioage_mod)$linkinv)) %>% 
+  left_join(count(sealdat2, bioage), by = "bioage")
+
+ggplot(bioage_pred_wgt, aes(bioage, predicted)) +
+  # Predictions for individual years
+  geom_line(aes(group = interaction(bioage_cat, year_fct)),
+            bioage_pred_years,
+            alpha = 0.1) +
+  # Weighted model (CI and mean response)
+  geom_ribbon(aes(fill = bioage_cat, ymin = conf_lo, ymax = conf_hi), 
+              alpha = 0.5) +
   geom_line(aes(color = bioage_cat),
-            bioage_pred_wgt, 
+            linewidth = 1.2) +
+  # Population-level mean response
+  geom_line(aes(y = predicted_pop, group = bioage_cat),
+            linetype = "dotted",
             linewidth = 1.2,
-            linetype = "dotted") +
+            alpha = 0.5) +
+  # Summarized raw data
   geom_pointrange(aes(y = perc_breed,
-                      ymin = perc_breed - se, 
-                      ymax = perc_breed + se),
+                      ymin = breed_lwr,
+                      ymax = breed_upr),
                   bioage_summ) +
   geom_vline(xintercept = bioage_senesce - 0.5, linetype = "dashed") +
-  scale_y_continuous("Breeding probability", labels = scales::percent, limits = c(.7, 1)) +
+  scale_y_continuous("Breeding probability", 
+                     labels = scales::percent,
+                     limits = c(NA, 1.02),
+                     breaks = seq(0.4, 1.0, by = 0.2)) +
   scale_color_manual(values = c("#7fbc41", "#de77ae")) +
   scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
-  scale_x_continuous(limits = c(-17, 0)) + 
-  labs(x = "Biological age (years-to-death)") +
-  theme(legend.position = "none")
+  labs(x = "Biological age (years)") +
+  theme(legend.position = "none") + 
+  geom_text(aes(label = n), y = 1, vjust = -0.5)
 
@@
 Choice of breakpoint for senescence did not affect the significance of this coefficient.
-For breakpoints -10 to -3 years, *p* reached a minimum at biological age -7 years.
+For breakpoints -12 to -3 years, *p* reached a minimum at biological age -7 years (see table above).
 
@@
 More accurate measures of biological age, such as telomere length [@jylhävä2017], may be necessary to investigate senescence in elephant seals and other species with high extrinsic sources of mortality from a non-chronological perspective.
+
+::: callout-note
+Max's commentary: almost all seals bred in the last year they were observed.
+As a result, 96.6% (95% CI: 95.2% - 97.7%) of seals in the bioage 0 group bred.
+That proportion was greater than all bioages except -17 and -16, which had much smaller sample sizes (0: 903, -17: 1, -16: 4).
+With such a high probability in bioage 0, it's just about impossible to get a negative slope in the old age category, because that one data point is pulling the right side of the line up.
+FWIW, 53.7% of the bioage 0 animals have a chronological age 4-7, so that part of the "old" biological age category is mostly chronologically young animals.
+:::
 
@@
 
+missing_age_se <- summary(missing_mod)$coef["age10", "Std. Error"]
+missing_age_est <- summary(missing_mod)$coef["age10", "Estimate"]
+missing_age_coef <- sprintf("%0.2f (95%% CI: %0.2f - %0.2f)",
+                            missing_age_est,
+                            missing_age_est - 1.96 * missing_age_se,
+                            missing_age_est + 1.96 * missing_age_se)
+
 ```
@@
 missing_years <- count(missing_seals, year_fct)
-missing_pred <- expand_grid(age = 5:19, 
+missing_grid <- expand_grid(age = 5:19, 
                             year_fct = factor(1989:2022)) %>% 
-  mutate(age10 = (age - age_senesce) / 10) %>% 
+  mutate(age10 = (age - age_senesce) / 10)
+
+missing_pred <- missing_grid %>% 
   mutate(predicted = predict(missing_mod, 
-                             newdata = missing_grid,
+                             newdata = .,
                              re.form = ~ (1 | year_fct)),
-         se = predict(missing_mod, 
-                      newdata = missing_grid,
+         se = suppressWarnings(predict(missing_mod, 
+                                       newdata = .,
                                        se.fit = TRUE,
-                      re.form = NA)$se.fit,
+                                       re.form = NA)$se.fit),
          conf.low = predicted - 1.96 * se,
@@
 
-missing_pred_years <- expand_grid(age = 5:19, 
-                            year_fct = factor(1989:2022)) %>% 
-  mutate(age10 = (age - age_senesce) / 10) %>% 
+missing_pred_years <- missing_grid %>% 
   mutate(predicted = predict(missing_mod, 
-                             newdata = missing_grid,
+                             newdata = .,
                              type = "response",
@@
 
-missing_pred
-
 missing_seals %>% 
   group_by(age) %>% 
-  summarize(unobserved_mean = mean(unobserved),
-            unobserved_se = sqrt(unobserved_mean * (1 - unobserved_mean) / n()),
-            unobserved_lwr = unobserved_mean - unobserved_se,
-            unobserved_upr = unobserved_mean + unobserved_se) %>% 
+  summarize(n = n(),
+            n_unobserved = sum(unobserved),
+            unobserved_mean = n_unobserved / n,
+            unobserved_lwr = binom.test(n_unobserved, n)$conf.int[1],
+            unobserved_upr = binom.test(n_unobserved, n)$conf.int[2]) %>% 
   ggplot(aes(age)) +

### 2024-07-15 paper/paperv3.qmd

@@
 
+```{r}
+#| label: h2-pred-deets
+
+#Overall decline 11-19
+tibble(age = c(11, 19),
+       age10 = (age - age_senesce) / 10,
+       age_cat = "Old") %>% 
+  # logit-scale predictions
+  mutate(predicted = predict(pup_surv_mod, 
+                             newdata = .,
+                             re.form = NA),
+         se = suppressWarnings(predict(pup_surv_mod, 
+                                       newdata = .,
+                                       se.fit = TRUE,
+                                       re.form = NA)$se.fit),
+         pred_lwr = predicted - 1.96 * se,
+         pred_upr = predicted + 1.96 * se,
+         # Invert the logit
+         across(c(predicted, pred_upr, pred_lwr), 
+                family(pup_surv_mod)$linkinv))
+
+#Difference between years 
+best_worst_yr_surv <- ranef(pup_surv_mod)$year_fct %>% 
+  as_tibble(rownames = "year_fct") %>% 
+  arrange(`(Intercept)`) %>% 
+  slice(1, nrow(.))
+
+tibble(age = c(11, 19),
+       age10 = (age - age_senesce) / 10,
+       age_cat = "Old") %>% 
+  cross_join(best_worst_yr_surv) %>% 
+  mutate(predicted = predict(pup_surv_mod, 
+                             newdata = .,
+                             type = "response", 
+                             re.form = ~ (1 | year_fct)))
+```
+
 ### H3: Changes in pup sex ratio with increased maternal age
@@
 
+missing_pred_pop <- tibble(age = 5:19,
+                           age10 = (age - age_senesce) / 10) %>% 
+  mutate(predicted = predict(missing_mod, 
+                             newdata = .,
+                             type = "response",
+                             re.form = NA))
 missing_seals %>% 

### 2024-07-15 paper/paperv3.qmd
Fix observation freq

--- 2024-07-15 paper/paperv3.qmd
@@
 # Are hypotheses supported by the model?
-extract_p <- \(h) map_dbl(mod_sens, \(m) m[[h]]["age10:age_catOld", "Pr(>|z|)"])
+extract_p <- \(h) map_dbl(mod_sens, \(m) pnorm(m[[h]]["age10:age_catOld", "z value"]))
 p_tbl <- tibble(age_cutoff = cutoff, 
@@
                    tag = LETTERS[1:4],
-                   x = 8, y = 0.75),
-            vjust = -0.2,
+                   x = 8, y = 3),
+            vjust = 1,
             fontface = "bold") +
@@
   labs(x = "Senescence onset (years)") +
-  expand_limits(y = 1.3) +
+  expand_limits(y = 3) +
   facet_wrap(~hypothesis)
@@
 sealdat2 <- sealdat %>% 
-  drop_na(lastrepro) %>% 
+  drop_na(lastobs) %>% 
   mutate(bioage = age - lastobs,
@@
   mod <- sealdat %>%
-    drop_na(lastrepro) %>%
+    drop_na(lastobs) %>%
     mutate(bioage = age - lastobs,
@@
   observed_int ~ age10 : age_cat + (1 | animalID) + (1 | year_fct),
-  drop_na(sealdat2, lastrepro),
+  drop_na(sealdat2, lastobs),
   family = "binomial",
@@
 
-::: callout-note
-Max's commentary: almost all seals bred in the last year they were observed.
-As a result, 96.6% (95% CI: 95.2% - 97.7%) of seals in the bioage 0 group bred.
-That proportion was greater than all bioages except -17 and -16, which had much smaller sample sizes (0: 903, -17: 1, -16: 4).
-With such a high probability in bioage 0, it's just about impossible to get a negative slope in the old age category, because that one data point is pulling the right side of the line up.
-FWIW, 53.7% of the bioage 0 animals have a chronological age 4-7, so that part of the "old" biological age category is mostly chronologically young animals.
-:::
-
-## Breeding state and detection probability
+## Observation frequency
 
@@
 
+missing_mod_old <- glmer(
+  unobserved ~ age10 + (1 | year_fct),
+  filter(missing_seals, age >= 11),
+  family = binomial
+)
+
+summary(missing_mod_old)
+
+missing_old_se <- summary(missing_mod_old)$coef["age10", "Std. Error"]
+missing_old_est <- summary(missing_mod_old)$coef["age10", "Estimate"]
+missing_old_coef <- sprintf("%0.2f (95%% CI: %0.2f - %0.2f)",
+                            missing_old_est,
+                            missing_old_est - 1.96 * missing_old_se,
+                            missing_old_est + 1.96 * missing_old_se)
+
 ```
@@
 The coefficient for age was `r missing_age_coef` (*p*\<0.0001).
+The old-animals-only model estimated the age coefficient to be `r missing_old_coef`.
 
@@
 
-missing_years <- count(missing_seals, year_fct)
 missing_grid <- expand_grid(age = 5:19, 
@@
          conf.high = predicted + 1.96 * se) %>% 
-  left_join(missing_years, by = "year_fct") %>% 
+  left_join(count(missing_seals, year_fct), by = "year_fct") %>% 
   group_by(age) %>% 
@@
                              re.form = ~ (1 | year_fct)))
+
+missing_pred_old <- expand_grid(age = 11:19, 
+                                year_fct = factor(1989:2022)) %>% 
+  mutate(age10 = (age - age_senesce) / 10) %>% 
+  mutate(predicted = predict(missing_mod_old, 
+                             newdata = .,
+                             re.form = ~ (1 | year_fct)),
+         se = suppressWarnings(predict(missing_mod, 
+                                       newdata = .,
+                                       se.fit = TRUE,
+                                       re.form = NA)$se.fit),
+         conf.low = predicted - 1.96 * se,
+         conf.high = predicted + 1.96 * se) %>% 
+  left_join(count(filter(missing_seals, age >= 11), year_fct), 
+            by = "year_fct") %>% 
+  group_by(age) %>% 
+  summarize(across(c(predicted, conf.low, conf.high), mean)) %>% 
+  mutate(across(c(predicted, conf.low, conf.high), family(missing_mod)$linkinv))
 
@@
             alpha = 0.2) +
+  # Overall model
   geom_ribbon(aes(ymin = conf.low, ymax = conf.high),
               missing_pred,
-              alpha = 0.2,
-              fill = "#de77ae") +
+              alpha = 0.3,
+              fill = "firebrick") +
   geom_line(aes(y = predicted),
             missing_pred,
-            color = "#de77ae",
+            color = "firebrick",
             linewidth = 2) +
+  # Old only
+  geom_ribbon(aes(ymin = conf.low, ymax = conf.high),
+              missing_pred_old,
+              alpha = 0.3,
+              fill = "cornflowerblue") +
+  geom_line(aes(y = predicted),
+            missing_pred_old,
+            color = "cornflowerblue",
+            linewidth = 2) +
+  # Raw data
   geom_pointrange(aes(y = unobserved_mean, 

### 2024-07-15 paper/paperv3.qmd

@@
 
+missing_pred_pop <- tibble(age = 5:19,
+                           age10 = (age - age_senesce) / 10) %>% 
+  mutate(predicted = predict(missing_mod, 
+                             newdata = .,
+                             type = "response",
+                             re.form = NA))
+
 missing_seals %>% 
@@
             missing_pred_old,
+            linetype = "dashed",
             color = "cornflowerblue",

### 2024-07-15 paper/temp.qmd

--- 2024-07-15 paper/paperv3.qmd
@@
                             max(age[observed_int == 1]),
-                            NA),
-         lastobs = ifelse(max(year) < 2020, max(age), NA)) %>% 
+                            NA)) %>% 
   ungroup()
@@
             perc_male = n_male / n_observed, 
-            male_lwr = binom.test(n_male, n_observed)$conf.int[1], 
-            male_upr = binom.test(n_male, n_observed)$conf.int[2]) %>% 
+            male_lwr = binom.test()
+            se = (perc_male * (1 - perc_male) / n_observed)^0.5) %>% 
   filter(age <= 20)
@@
   geom_pointrange(aes(y = perc_male, 
-                      ymin = male_lwr, 
-                      ymax = male_upr),
+                      ymin = perc_male - se, 
+                      ymax = perc_male + se),
                   sexrat_summ) +
   geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
-  scale_y_continuous("Male offspring", 
-                     labels = scales::percent,
-                     limits = c(NA, 0.97)) +
+  scale_y_continuous("Male offspring", labels = scales::percent) +
   scale_color_manual(values = c("#7fbc41", "#de77ae")) +
@@
   theme(legend.position = "none") +
-  geom_text(aes(label = n), y = .97, vjust = 0)
+  geom_text(aes(label = n, y = .95))
 
@@
   scale_y_continuous("Day of year", expand = expansion(mult = 0)) +
+  geom_text(aes(label = "Breeding haul out"), 
+            x = 8,
+            y = 30,
+            color = "white") +
+    geom_text(aes(label = "Molting haul out"), 
+            x = 8,
+            y = 135,
+            color = "white") + 
+      geom_text(aes(label = "Foraging"), 
+            x = 8,
+            y = 250,
+            color = "white") + 
   theme(panel.spacing = unit(1, "lines"), 
-        legend.position = "bottom",
-        legend.title = element_blank(),
-        legend.text = element_text(size = 8),
-        legend.key.size = unit(3, "mm")) 
-  # geom_text(aes(label = n, color = "white", y = 300))
+        legend.position = "none")
 
@@
          phase = group)
+
+breed_phen_years <- phen_pred_years %>% filter(phase == "breeddur")
+molt_phen_years <- phen_pred_years %>% filter(phase == "moltdur")
 
@@
   ggplot(aes(age)) +
-  geom_line(aes(age, predicted, group = interaction(phase, facet)), 
-            phen_pred_years, 
+  geom_line(aes(age, predicted, group = facet), 
+            breed_phen_years, 
             alpha = 0.1) + 
@@
   ggplot(aes(age)) +
-  geom_line(aes(age, predicted, group = interaction(phase, facet)), 
-            phen_pred_years, 
+  geom_line(aes(age, predicted, group = facet), 
+            molt_phen_years, 
             alpha = 0.1) + 
@@
 We tested whether our significant results (reproductive senescence, maternal effect senescence in pup survival) were influenced by selective appearance and disappearance, following the methods of @vandepol2006, by adding terms for age at first and last reproduction, respectively.
-If our results are due to selective appearance/disappearance instead of senescence, then the 95% CI for the coefficient of age in the old age class in these models will overlap 0.
-
-```{r}
-#| label: fig-selective
-#| fig-cap: The coefficient for age in the old age class in models accounting for selective appearance and disappearance.
-Red and blue indicate  reproductive senescence (H1) and maternal effect senescence (H2a), respectively.
-Compared to the base model, the appearance model includes a term for age at first reproduction, the disappearance model includes a term for age at last reproduction, and the combined model includes both.
+We used AIC to compare the base model to models with selective appearance and disappearance.
+
+### H1: Reproductive senescence
+
+```{r}
+#| label: selective-h1
 
@@
 
-There was no evidence for selective appearance or disappearance influencing the relationship between maternal age and probability of reproduction or offspring survival.
-The 95% CI of the coefficient for age in the old age class did not overlap 0 for either hypothesis in any of the selective (dis)appearance models [@fig-selective].
-Therefore the observed patterns are consistent with the senescence hypothesis (i.e., within-animal change with age).
+```{r}
+#| label: fig-selective-h1
+#| fig-cap: Predicted breeding probabilities of (**A**) selective appearance and (**B**) selective disappearance models.
+The semi-transparent solid lines are predictions of the base model.
+The dashed and dotted lines represent predictions for the 10th and 90th percentiles of observed age at first (**A**) and last (**B**) reproduction.
+In **A**, these correspond to first reproduction at 4 and 7 years old, respectively.
+In **B**, these correspond to last reproduction at 5 and 17 years old, respectively.
+Confidence intervals not shown for clarity.
+#| fig-width: 9
+
+# Compare model predictions
+base_pred <- ggpredict(
+  h1_selective_models[[1]],
+  terms = c("age10 [all]", "age_cat")
+) %>% 
+  as_tibble() %>% 
+  mutate(age = x * 10 + age_senesce,
+         age_cat = factor(group, levels = c("Young", "Old"))) %>% 
+  filter((age_cat == "Young" & age < age_senesce) |
+           (age_cat == "Old" & age >= age_senesce),
+         age <= 20)
+
+# selective appearance for firstrepro=4,6 (10% and 90% of observed)
+selapp_pred <- ggpredict(
+  h1_selective_models[[2]],
+  terms = c("age10 [all]", "age_cat", "firstrepro [4,7]")
+) %>% 
+  as_tibble() %>% 
+  mutate(age = x * 10 + age_senesce,
+         age_cat = factor(group, levels = c("Young", "Old")),
+         firstrepro = facet) %>% 
+  filter((age_cat == "Young" & age < age_senesce) |
+           (age_cat == "Old" & age >= age_senesce),
+         age <= 20)
+
+# selective disappearance for lastrepro=5,17 (10% and 90% of observed) 
+seldis_pred <- ggpredict(
+  h1_selective_models[[4]],
+  terms = c("age10 [all]", "age_cat", "lastrepro [5,17]")
+) %>% 
+  as_tibble() %>% 
+  mutate(age = x * 10 + age_senesce,
+         age_cat = factor(group, levels = c("Young", "Old")),
+         lastrepro = facet) %>% 
+  filter((age_cat == "Young" & age < age_senesce) |
+           (age_cat == "Old" & age >= age_senesce),
+         age <= 20)
+
+# Selective appearance predictions
+fig_selapp <- ggplot(base_pred, aes(age, predicted)) +
+  # Base predictions
+  geom_line(aes(color = age_cat), alpha = 0.5) +
+  # Selective appearance ages 4 and 6
+  geom_line(aes(color = age_cat, 
+                linetype = firstrepro),
+            selapp_pred) +
+  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
+  scale_y_continuous("Breeding", 
+                     labels = scales::percent) +
+  scale_color_manual(values = c("#7fbc41", "#de77ae")) +
+  scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
+  scale_linetype_manual(values = c("dashed", "dotted")) +
+  labs(x = "Age (years)") +
+  theme(legend.position = "none")
+
+fig_seldis <- ggplot(base_pred, aes(age, predicted)) +
+  # Base predictions
+  geom_line(aes(color = age_cat), alpha = 0.5) +
+  # Selective disappearance ages 5 and 17
+  geom_line(aes(color = age_cat, 
+                linetype = lastrepro),
+            seldis_pred) +
+  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
+  scale_y_continuous("Breeding", 
+                     labels = scales::percent) +
+  scale_color_manual(values = c("#7fbc41", "#de77ae")) +
+  scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
+  scale_linetype_manual(values = c("dashed", "dotted")) +
+  labs(x = "Age (years)") +
+  theme(legend.position = "none")
+
+```
+
+There was no evidence for selective appearance or disappearance influencing the relationship between maternal age and probability of reproduction.
+The selective appearance model (i.e., included a term for age-of-first-reproduction) had a similar AIC weight as the base model.
+However, in the selective appearance model, the coefficient for age at first reproduction was not significantly different than 0.
+As a result, the predicted breeding probabilities for animals that first recruited at ages 4 and 7 (the 10th and 90th observed percentiles in the population) were not qualitatively different than the predictions of the base model [@fig-selective-h1].
+Therefore the observed pattern is consistent with the senescence hypothesis (i.e., within-animal change with age).
+
+### H2: Maternal effect senescence
+
+```{r}
+#| label: fig-selective-h2
+#| fig-cap: Predicted breeding probabilities of (**A**) selective appearance and (**B**) selective disappearance models.
+The semi-transparent solid lines are predictions of the base model.
+The dashed and dotted lines represent predictions for the 10th and 90th percentiles of observed age at first (**A**) and last (**B**) reproduction.
+In **A**, these correspond to first reproduction at 4 and 7 years old, respectively.
+In **B**, these correspond to last reproduction at 5 and 17 years old, respectively.
+Confidence intervals not shown for clarity.
+#| fig-width: 9
+
+# Compare model predictions
+base_pred <- ggpredict(
+  h2_selective_models[[1]],
+  terms = c("age10 [all]", "age_cat")
+) %>% 
+  as_tibble() %>% 
+  mutate(age = x * 10 + age_senesce,
+         age_cat = factor(group, levels = c("Young", "Old"))) %>% 
+  filter((age_cat == "Young" & age < age_senesce) |
+           (age_cat == "Old" & age >= age_senesce),
+         age <= 20)
+
+# selective appearance for firstrepro=4,6 (10% and 90% of observed)
+selapp_pred <- ggpredict(
+  h2_selective_models[[2]],
+  terms = c("age10 [all]", "age_cat", "firstrepro [4,7]")
+) %>% 
+  as_tibble() %>% 
+  mutate(age = x * 10 + age_senesce,
+         age_cat = factor(group, levels = c("Young", "Old")),
+         firstrepro = facet) %>% 
+  filter((age_cat == "Young" & age < age_senesce) |
+           (age_cat == "Old" & age >= age_senesce),
+         age <= 20)
+
+# selective disappearance for lastrepro=5,17 (10% and 90% of observed) 
+seldis_pred <- ggpredict(
+  h2_selective_models[[4]],
+  terms = c("age10 [all]", "age_cat", "lastrepro [5,17]")
+) %>% 
+  as_tibble() %>% 
+  mutate(age = x * 10 + age_senesce,
+         age_cat = factor(group, levels = c("Young", "Old")),
+         lastrepro = facet) %>% 
+  filter((age_cat == "Young" & age < age_senesce) |
+           (age_cat == "Old" & age >= age_senesce),
+         age <= 20)
+
+# Selective appearance predictions
+fig_selapp <- ggplot(base_pred, aes(age, predicted)) +
+  # Base predictions
+  geom_line(aes(color = age_cat), alpha = 0.5) +
+  # Selective appearance ages 4 and 6
+  geom_line(aes(color = age_cat, 
+                linetype = firstrepro),
+            selapp_pred) +
+  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
+  scale_y_continuous("Offspring survival", 
+                     labels = scales::percent) +
+  scale_color_manual(values = c("#7fbc41", "#de77ae")) +
+  scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
+  scale_linetype_manual(values = c("dashed", "dotted")) +
+  labs(x = "Maternal age (years)") +
+  theme(legend.position = "none")
+
+fig_seldis <- ggplot(base_pred, aes(age, predicted)) +
+  # Base predictions
+  geom_line(aes(color = age_cat), alpha = 0.5) +
+  # Selective disappearance ages 5 and 17
+  geom_line(aes(color = age_cat, 
+                linetype = lastrepro),
+            seldis_pred) +
+  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
+  scale_y_continuous("Offspring survival", 
+                     labels = scales::percent) +
+  scale_color_manual(values = c("#7fbc41", "#de77ae")) +
+  scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
+  scale_linetype_manual(values = c("dashed", "dotted")) +
+  labs(x = "Maternal age (years)") +
+  theme(legend.position = "none")
+
+```
+
+There was no evidence for selective appearance or disappearance influencing the relationship between maternal age and offspring survival.
+The selective appearance model (i.e., included a term for age-of-first-reproduction) had a slightly greater AIC weight than the base model.
+However, in the selective appearance model, the coefficient for age at first reproduction was not significantly different than 0.
+As a result, the predicted breeding probabilities for animals that first recruited at ages 4 and 7 (the 10th and 90th observed percentiles in the population) were not qualitatively different than the predictions of the base model [@fig-selective-h2].
+Therefore the observed pattern is consistent with the senescence hypothesis (i.e., within-animal change with age).
 
@@
   drop_na(lastrepro) %>% 
-  mutate(bioage = age - lastobs,
+  mutate(bioage = age - lastrepro,
          bioage10 = (bioage - bioage_senesce) / 10,
@@
 This required us to limit our sample to animals with known longevity (i.e., last observation in 2020 or earlier), which reduced our sample size to `r nrow(sealdat2)` observations of `r n_distinct(sealdat2$animalID)` distinct animals.
-Subsequently, we fit the model for breakpoints at all biological ages -12 to -3 years, to assess whether our results were sensitive to the choice of breakpoint.
+Subsequently, we fit the model for breakpoints at all biological ages -10 to -3 years, to assess whether our results were sensitive to the choice of breakpoint.
 Finally, we re-fit the chronological reproductive senescence model using the known-longevity dataset and compared chronological versus biological age models using AIC.
@@
 
-# Biological age model
 breed_bioage_mod <- glmer(
@@
 
-old_z <- summary(breed_bioage_mod)$coefficients["bioage10:bioage_catOld", "z value"]
-bioage_pval <- sprintf("%0.3f", pnorm(-old_z))
-
-# Breakpoint comparison
-p_by_tau <- map_dbl(-12:-3, \(tau) {
+bioage_pval <- sprintf("%0.3f", summary(breed_bioage_mod)$coef["bioage10:bioage_catOld", "Pr(>|z|)"])
+
+breed_age_mod2 <- glmer(
+  observed_int ~ age10 : age_cat + (1 | animalID) + (1 | year_fct),
+  drop_na(sealdat, lastrepro),
+  family = "binomial",
+  control = glmerControl(optimizer = "bobyqa")
+)
+
+p_by_tau <- map_dbl(-10:-3, \(tau) {
   mod <- sealdat %>%
     drop_na(lastrepro) %>%
-    mutate(bioage = age - lastobs,
+    mutate(bioage = age - lastrepro,
            bioage10 = (bioage - tau) / 10,
@@
     )
-  old_z <- summary(mod)$coef["bioage10:bioage_catOld", "z value"] 
-  pnorm(old_z)
+  summary(mod)$coef["bioage10:bioage_catOld", "Pr(>|z|)"]
 })
-
-tibble(`Biological age threshold` = -12:-3,
-       `p value` = p_by_tau)
-
-# Model comparison
-breed_age_mod2 <- glmer(
-  observed_int ~ age10 : age_cat + (1 | animalID) + (1 | year_fct),
-  drop_na(sealdat2, lastrepro),
-  family = "binomial",
-  control = glmerControl(optimizer = "bobyqa")
-)
 
@@
   summarize(n_obs = n(), 
-            n_breed = sum(observed_int),
-            perc_breed = n_breed / n_obs, 
-            breed_lwr = binom.test(n_breed, n_obs)$conf.int[1], 
-            breed_upr = binom.test(n_breed, n_obs)$conf.int[2])
-
-# Create ANNUAL (i.e. rand effect of year only) predictions
-bioage_pred_years <- ggpredict(
+            perc_breed = mean(observed_int), 
+            se = sqrt(perc_breed * (1 - perc_breed) / n_obs))
+
+# Create POPULATION-LEVEL (i.e. without rand effects) predictions
+bioage_pred_pop <- ggpredict(breed_bioage_mod, 
+                             terms = c("bioage10 [all]", "bioage_cat")) %>% 
+  as_tibble() %>% 
+  mutate(bioage = x * 10 + bioage_senesce,
+         bioage_cat = factor(group, levels = c("Young", "Old"))) %>% 
+  filter((bioage_cat == "Young" & bioage < bioage_senesce) |
+           (bioage_cat == "Old" & bioage >= bioage_senesce))
+
+# Create WEIGHTED-AVERAGE predictions using years' random effects
+bioage_pred_wgt <- ggpredict(
   breed_bioage_mod,
-  terms = c("bioage10 [all]", "bioage_cat", "year_fct"),
-  type = "random"
+  terms = c("bioage10 [all]", "bioage_cat", "year_fct [all]"),
+  type = "random",
+  interval = "confidence"
 ) %>% 
@@
          bioage_cat = factor(group, levels = c("Young", "Old")),
-         year_fct = factor(facet)) %>% 
+         year_fct = facet) %>% 
   filter((bioage_cat == "Young" & bioage < bioage_senesce) |
-           (bioage_cat == "Old" & bioage >= bioage_senesce),
-         bioage <= 20)
-
-# Create WEIGHTED-AVERAGE predictions using years' random effects
-# ggpredict can't do that for confidence intervals
-# create prediction grid
-bioage_grid <- expand_grid(
-  bioage = -17:0,
-  year_fct = factor(1988:2019)
-) %>% 
-  mutate(bioage10 = (bioage - bioage_senesce) / 10,
-         bioage_cat = factor(ifelse(bioage < bioage_senesce, "Young", "Old"),
-                          levels = c("Young", "Old")))
-
-bioage_years <- count(sealdat2, year_fct)
-bioage_pred_wgt <- bioage_grid %>% 
-  mutate(predicted = predict(breed_bioage_mod, 
-                             newdata = bioage_grid,
-                             re.form = ~ (1 | year_fct)),
-         predicted_se = suppressWarnings(
-           # warning: "se.fit computation uses an approximation to estimate the sampling distribution of the parameters"
-           predict(breed_bioage_mod, 
-                   newdata = bioage_grid,
-                   re.form = NA,
-                   se.fit = TRUE)
-         )$se.fit,
-         predicted_pop = predict(breed_bioage_mod, 
-                                 newdata = bioage_grid,
-                                 re.form = NA),
-         conf_lo = predicted - 1.96 * predicted_se,
-         conf_hi = predicted + 1.96 * predicted_se) %>% 
-  left_join(bioage_years, by = "year_fct") %>% 
+           (bioage_cat == "Old" & bioage >= bioage_senesce)) %>% 
+  left_join(seal_years, by = "year_fct") %>% 
   group_by(bioage, bioage_cat) %>% 
-  summarize(across(c(predicted, predicted_pop, conf_lo, conf_hi), 
-                   \(x) weighted.mean(x, n)),
-            .groups = "drop") %>% 
-  # invert logit
-  mutate(across(c(predicted, predicted_pop, conf_lo, conf_hi), 
-                family(breed_bioage_mod)$linkinv)) %>% 
-  left_join(count(sealdat2, bioage), by = "bioage")
-
-ggplot(bioage_pred_wgt, aes(bioage, predicted)) +
-  # Predictions for individual years
-  geom_line(aes(group = interaction(bioage_cat, year_fct)),
-            bioage_pred_years,
-            alpha = 0.1) +
-  # Weighted model (CI and mean response)
-  geom_ribbon(aes(fill = bioage_cat, ymin = conf_lo, ymax = conf_hi), 
-              alpha = 0.5) +
+  summarize(predicted = weighted.mean(predicted, n),
+            .groups = "drop")
+
+ggplot(bioage_pred_pop, aes(bioage, predicted)) +
+  geom_ribbon(aes(fill = bioage_cat, ymin = conf.low, ymax = conf.high), 
+              alpha = 0.2) +
+  geom_line(aes(color = bioage_cat), linewidth = 1.2, linetype = "dashed") +
   geom_line(aes(color = bioage_cat),
-            linewidth = 1.2) +
-  # Population-level mean response
-  geom_line(aes(y = predicted_pop, group = bioage_cat),
-            linetype = "dotted",
+            bioage_pred_wgt, 
             linewidth = 1.2,
-            alpha = 0.5) +
-  # Summarized raw data
+            linetype = "dotted") +
   geom_pointrange(aes(y = perc_breed, 
-                      ymin = breed_lwr,
-                      ymax = breed_upr),
+                      ymin = perc_breed - se, 
+                      ymax = perc_breed + se),
                   bioage_summ) +
   geom_vline(xintercept = bioage_senesce - 0.5, linetype = "dashed") +
-  scale_y_continuous("Breeding probability", 
-                     labels = scales::percent,
-                     limits = c(NA, 1.02),
-                     breaks = seq(0.4, 1.0, by = 0.2)) +
+  scale_y_continuous("Breeding probability", labels = scales::percent, limits = c(.7, 1)) +
   scale_color_manual(values = c("#7fbc41", "#de77ae")) +
   scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
-  labs(x = "Biological age (years)") +
-  theme(legend.position = "none") + 
-  geom_text(aes(label = n), y = 1, vjust = -0.5)
+  scale_x_continuous(limits = c(-17, 0)) + 
+  labs(x = "Biological age (years-to-death)") +
+  theme(legend.position = "none")
 
@@
 Choice of breakpoint for senescence did not affect the significance of this coefficient.
-For breakpoints -12 to -3 years, *p* reached a minimum at biological age -7 years (see table above).
+For breakpoints -10 to -3 years, *p* reached a minimum at biological age -7 years.
 
@@
 More accurate measures of biological age, such as telomere length [@jylhävä2017], may be necessary to investigate senescence in elephant seals and other species with high extrinsic sources of mortality from a non-chronological perspective.
-
-::: callout-note
-Max's commentary: almost all seals bred in the last year they were observed.
-As a result, 96.6% (95% CI: 95.2% - 97.7%) of seals in the bioage 0 group bred.
-That proportion was greater than all bioages except -17 and -16, which had much smaller sample sizes (0: 903, -17: 1, -16: 4).
-With such a high probability in bioage 0, it's just about impossible to get a negative slope in the old age category, because that one data point is pulling the right side of the line up.
-FWIW, 53.7% of the bioage 0 animals have a chronological age 4-7, so that part of the "old" biological age category is mostly chronologically young animals.
-:::
 
@@
 
-missing_age_se <- summary(missing_mod)$coef["age10", "Std. Error"]
-missing_age_est <- summary(missing_mod)$coef["age10", "Estimate"]
-missing_age_coef <- sprintf("%0.2f (95%% CI: %0.2f - %0.2f)",
-                            missing_age_est,
-                            missing_age_est - 1.96 * missing_age_se,
-                            missing_age_est + 1.96 * missing_age_se)
-
 ```
@@
 missing_years <- count(missing_seals, year_fct)
-missing_grid <- expand_grid(age = 5:19, 
+missing_pred <- expand_grid(age = 5:19, 
                             year_fct = factor(1989:2022)) %>% 
-  mutate(age10 = (age - age_senesce) / 10)
-
-missing_pred <- missing_grid %>% 
+  mutate(age10 = (age - age_senesce) / 10) %>% 
   mutate(predicted = predict(missing_mod, 
@@
                              re.form = ~ (1 | year_fct)),
-         se = suppressWarnings(predict(missing_mod, 
+         se = predict(missing_mod, 
                       newdata = .,
                       se.fit = TRUE,
-                                       re.form = NA)$se.fit),
+                      re.form = NA)$se.fit,
          conf.low = predicted - 1.96 * se,
@@
 
-missing_pred_years <- missing_grid %>% 
+missing_pred_years <- expand_grid(age = 5:19, 
+                            year_fct = factor(1989:2022)) %>% 
+  mutate(age10 = (age - age_senesce) / 10) %>% 
   mutate(predicted = predict(missing_mod, 
@@
                              re.form = NA))
+
 missing_seals %>% 
   group_by(age) %>% 
-  summarize(n = n(),
-            n_unobserved = sum(unobserved),
-            unobserved_mean = n_unobserved / n,
-            unobserved_lwr = binom.test(n_unobserved, n)$conf.int[1],
-            unobserved_upr = binom.test(n_unobserved, n)$conf.int[2]) %>% 
+  summarize(unobserved_mean = mean(unobserved),
+            unobserved_se = sqrt(unobserved_mean * (1 - unobserved_mean) / n()),
+            unobserved_lwr = unobserved_mean - unobserved_se,
+            unobserved_upr = unobserved_mean + unobserved_se) %>% 
   ggplot(aes(age)) +
@@
 
-## Breakpoint vs quadratic
-
-### H1: Reproductive senescence
-
-```{r}
-#| label: quad-h1
-
-breed_age_quad <- glmer(
-  observed_int ~ poly(age10, 2, raw = TRUE) + (1 | animalID) + (1 | year_fct),
-  sealdat,
-  family = "binomial",
-  control = glmerControl(optimizer = "bobyqa")
-)
-
-h1_age10_peak <- -summary(breed_age_quad)$coefficients[2, 1] / 
-  (2 * summary(breed_age_quad)$coefficients[3, 1])
-h1_age_peak <- h1_age10_peak * 10 + age_senesce
-
-summary(breed_age_quad)$coefficients
-
-sprintf("Age peak (H1) = %0.1f", h1_age_peak)
-
-```
-
-### H2: Maternal effect senescence
-
-```{r}
-#| label: quad-h2
-
-# H2a
-surv_age_quad <- glmer(
-  pup_surv_int ~ poly(age10, 2, raw = TRUE) + (1 | animalID) + (1 | year_fct),
-  surv_recr_data,
-  family = "binomial",
-  control = glmerControl(optimizer = "bobyqa")
-)
-
-h2a_age10_peak <- -summary(surv_age_quad)$coefficients[2, 1] / 
-  (2 * summary(surv_age_quad)$coefficients[3, 1])
-h2a_age_peak <- h2a_age10_peak * 10 + age_senesce
-
-summary(surv_age_quad)$coefficients
-
-sprintf("Age peak (H2a) = %0.1f", h2a_age_peak)
-
-# H2b
-
-recr_age_quad <- glmer(
-  pup_recr_int ~ poly(age10, 2, raw = TRUE) + (1 | year_fct),
-  filter(surv_recr_data, pupsex == "F"),
-  family = "binomial",
-  control = glmerControl(optimizer = "bobyqa")
-)
-
-h2b_age10_peak <- -summary(recr_age_quad)$coefficients[2, 1] / 
-  (2 * summary(recr_age_quad)$coefficients[3, 1])
-h2b_age_peak <- h2b_age10_peak * 10 + age_senesce
-
-summary(recr_age_quad)$coefficients
-
-sprintf("Age peak (H2b) = %0.1f", h2b_age_peak)
-
-```
+## Threshold vs

### 2024-07-15 paper/paperv3.qmd

@@
   scale_y_continuous("Day of year", expand = expansion(mult = 0)) +
+    geom_text(aes(label = "Breeding haul out"), 
+            x = 8,
+            y = 30,
+            color = "white") +
+    geom_text(aes(label = "Molting haul out"), 
+            x = 8,
+            y = 135,
+            color = "white") + 
+      geom_text(aes(label = "Foraging"), 
+            x = 8,
+            y = 250,
+            color = "white") + 
   theme(panel.spacing = unit(1, "lines"), 
-        legend.position = "bottom",
-        legend.title = element_blank(),
-        legend.text = element_text(size = 8),
-        legend.key.size = unit(3, "mm")) 
-  # geom_text(aes(label = n, color = "white", y = 300))
+        legend.position = "none")
 

### 2024-07-16 paper/paperv3.qmd
model clarity and TODOs

@@
 
-resight <- read.csv(here::here("data/raw/fullresights.csv")) %>% 
+resight <- read_csv(here::here("data/raw/fullresights.csv"), 
+                    show_col_types = FALSE) %>% 
   select(animalID, resightID, season, date, timeofyear, calyear, yday) %>% 
@@
                      labels = scales::percent,
-                     limits = c(NA, 1.02),
-                     breaks = seq(0.4, 1.0, by = 0.2)) +
+                     breaks = seq(0.4, 1.0, by = 0.1)) +
   scale_color_manual(values = c("#7fbc41", "#de77ae")) +
   scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
+  coord_cartesian(ylim = c(0.68, 1.02)) +
   labs(x = "Age (years)") +
@@
   scale_y_continuous("Offspring survival", 
-                     labels = scales::percent,
-                     limits = c(0, 1)) +
+                     labels = scales::percent) +
   scale_color_manual(values = c("#7fbc41", "#de77ae")) +
@@
   scale_linetype_manual(values = c("dashed", "solid")) +
+  coord_cartesian(ylim = c(0, 0.6)) +
   labs(x = "Age (years)") +
   theme(legend.position = "none") + 
-  geom_text(aes(label = n), y = 1, vjust = 1)
+  geom_label(aes(label = n), y = 0.6, vjust = 1)
 
@@
   scale_y_continuous("Offspring recruitment", 
-                     labels = scales::percent,
-                     limits = c(0, 1)) +
+                     labels = scales::percent) +
   scale_color_manual(values = c("#7fbc41", "#de77ae")) +
   scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
+  coord_cartesian(ylim = c(0, 0.6)) +
   labs(x = "Age (years)") +
   theme(legend.position = "none") + 
-  geom_text(aes(label = n), y = 1, vjust = 1)
+  geom_label(aes(label = n), y = 0.6, vjust = 1)
 
@@
 d &\sim NB(\mu, k) \\
-\mu &= exp(\eta) \\
-\eta &= \beta_0 + \beta_p * a + Zy
+e^\mu &= \beta_0 + \beta_p a + Zy
 \end{align}
@@
 
-Where: $d$ is the duration of the phase in days, $NB$ is the negative binomial distribution with location $\mu$ and dispersion $k$, $\eta$ is the linear predictor, $\beta_0$ is the intercept, $\beta_p$ is the coefficient for age ($a$) by phase (i.e., for $p$ in "molting" or "breeding"), $y$ is the year, and $Z$ is the design matrix relating $y$ to $\eta$.
+Where: $d$ is the duration of the phase in days, $NB$ is the negative binomial distribution with location $\mu$ and dispersion $k$, $\beta_0$ is the intercept, $\beta_p$ is the coefficient for age ($a$) by phase (i.e., for $p$ in "molting" or "breeding"), $y$ is the year, and $Z$ is the design matrix relating $y$ to $e^\mu$.
 
@@
   geom_line(aes(age, predicted, group = interaction(phase, facet)), 
-            phen_pred_years, 
+            filter(phen_pred_years, phase == "breeddur"), 
             alpha = 0.1) + 
@@
   geom_line(aes(age, predicted, group = interaction(phase, facet)), 
-            phen_pred_years, 
+            filter(phen_pred_years, phase == "moltdur"), 
             alpha = 0.1) + 
@@
                 "H2b (recruitment)" = extract_p("h2_recr"),
-                "H3 (sex ratio)" = extract_p("h3"))
+                "H3 (sex ratio)" = extract_p("h3") * 2)
 
@@
   observed_int ~ bioage10 : bioage_cat + (1 | animalID) + (1 | year_fct),
-  sealdat2,
+  filter(sealdat2, bioage < 0),
   family = "binomial",
@@
                                labels = c("Old", "Young"))) %>% 
+    filter(bioage < 0) %>% 
     glmer(
@@
   observed_int ~ age10 : age_cat + (1 | animalID) + (1 | year_fct),
-  drop_na(sealdat2, lastobs),
+  filter(sealdat2, bioage < 0),
   family = "binomial",
@@
 bioage_summ <- sealdat2 %>%
+  # exclude n=1 age class
+  filter(bioage >= -15) %>% 
   group_by(bioage) %>% 
@@
            (bioage_cat == "Old" & bioage >= bioage_senesce),
-         bioage <= 20)
+         bioage >= -15)
 
@@
 bioage_grid <- expand_grid(
-  bioage = -17:0,
-  year_fct = factor(1988:2019)
+  bioage = -15:-1,
+  year_fct = factor(1988:2018)
 ) %>% 
@@
 
+TODO: fix sample size for bioage=0
+
 ## Observation frequency
@@
 Here, we assess those potential biases.
+
+TODO: include from age 4 on
 
@@
 
+TODO: AIC comparison
+
 ### H1: Reproductive senescence

### 2024-07-19 paper/paperv3.qmd
Updating phenology with the breakpoint model

--- 2024-07-15 paper/paperv3.qmd
@@
 seal_phenology <- sealdat %>% 
-  filter(year >= 2011) %>% 
+  filter(year >= 2011, 
+         moltdur >= 7) %>% 
   mutate(BA = firstobsbreeddoy, 
@@
 pheno_model <- seal_phenology %>% 
-  transmute(animalID, year, age10 = age / 10, moltdur, breeddur) %>% 
+  transmute(animalID, 
+            year_fct, 
+            age10 = (age - age_senesce) / 10, 
+            age_cat = factor(age >= age_senesce, 
+                          labels = c("Young", "Old")),
+            moltdur, 
+            breeddur) %>% 
   pivot_longer(moltdur:breeddur, 
@@
                values_to = "dur_days") %>% 
-  glmer.nb(dur_days ~ age10:phase + (1 | year), 
+  glmer.nb(dur_days ~ age10:phase:age_cat + (1 | year_fct), 
            data = .)
@@
 summary(pheno_model)
+
+```
+
+```{r}
+
+#Make sample size data frame
+n_phen_age <- count(seal_phenology, age)
+n_phen_year <- count(seal_phenology, year_fct)
+
+pheno_grid <- expand_grid(
+  age = 4:20,
+  year_fct = factor(2011:2023), 
+  phase = c("breeddur", "moltdur")
+) %>% 
+  mutate(age10 = (age - age_senesce) / 10,
+         age_cat = factor(ifelse(age < age_senesce, "Young", "Old"),
+                          levels = c("Young", "Old")))
+
+pheno_pred <- pheno_grid %>% 
+  mutate(predicted = predict(pheno_model, 
+                             newdata = pheno_grid,
+                             re.form = ~ (1 | year_fct)),
+         predicted_se = suppressWarnings(
+           # warning: "se.fit computation uses an approximation to estimate the sampling distribution of the parameters"
+           predict(pheno_model, 
+                   newdata = pheno_grid,
+                   re.form = NA,
+                   se.fit = TRUE),
+         )$se.fit,
+         predicted_pop = predict(pheno_model, 
+                                 newdata = pheno_grid,
+                                 re.form = NA),
+         conf_lo = predicted - 1.96 * predicted_se,
+         conf_hi = predicted + 1.96 * predicted_se) %>% 
+  left_join(n_phen_year, by = "year_fct") %>% 
+  group_by(age, age_cat) %>% 
+  summarize(across(c(predicted, predicted_pop, conf_lo, conf_hi), 
+                   \(x) weighted.mean(x, n)),
+            .groups = "drop") %>% 
+  # invert logit
+  mutate(across(c(predicted, predicted_pop, conf_lo, conf_hi), 
+                family(pheno_model)$linkinv)) %>% 
+  left_join(n_phen_age, by = "age")
+
+phen_pred_years <- ggpredict(
+  pheno_model,
+  terms = c("age10 [all]", "age_cat", "phase", "year_fct [all]"),
+  type = "random",
+  interval = "confidence"
+) %>% 
+  as_tibble() %>% 
+  mutate(age = x * 10 + age_senesce,
+         age_cat = group,
+         phase = facet,
+         year_fct = panel) %>% 
+  filter((age_cat == "Young" & age < age_senesce) |
+           (age_cat == "Old" & age >= age_senesce))
+
+pheno_raw <- seal_phenology %>% 
+  select(breeddur, moltdur, age) %>% 
+  pivot_longer(c(breeddur, moltdur), 
+               names_to = "phase", 
+               values_to = "dur_days") %>% 
+  group_by(age, phase) %>% 
+  summarize(n = n(),
+            dur_days_se = sd(dur_days) / sqrt(n),
+            dur_days = mean(dur_days),
+            dur_days_lwr = dur_days - dur_days_se,
+            dur_days_upr = dur_days + dur_days_se,
+            .groups = "drop")
+
+breed_pheno_raw <- pheno_raw[pheno_raw$phase == "breeddur",]
+molt_pheno_raw <- pheno_raw[pheno_raw$phase == "moltdur",]
+
+breed_duration_plot <- ggpredict(
+  pheno_model, 
+  terms = c("age10 [all]", "age_cat", "phase", "year_fct [all]")
+) %>% 
+  as_tibble() %>%
+  transmute(age = x * 10 + age_senesce,
+            age_cat = group,
+            dur_days = predicted,
+            dur_days_lwr = conf.low,
+            dur_days_upr = conf.high,
+            phase = facet,
+            year_fct = panel) %>% 
+  left_join(n_phen_year, by = "year_fct") %>% 
+  filter(phase == "breeddur",
+         (age_cat == "Young" & age < age_senesce) |
+           (age_cat == "Old" & age >= age_senesce)) %>% 
+  group_by(age, age_cat) %>% 
+  summarize(across(starts_with("dur_days"), \(x) weighted.mean(x, n)),
+            .groups = "drop") %>% 
+  ggplot(aes(age)) +
+  # Predictions by year
+  geom_line(aes(age, predicted, group = interaction(phase, age_cat, year_fct)), 
+            filter(phen_pred_years, phase == "breeddur"),
+            alpha = 0.1) + 
+  # Weighted model (CI + mean)
+  geom_ribbon(aes(ymin = dur_days_lwr, ymax = dur_days_upr, fill = age_cat),
+              alpha = 0.25) +
+  geom_line(aes(y = dur_days, color = age_cat, linetype = age_cat)) +
+  # Summarized data
+  geom_pointrange(aes(y = dur_days, 
+                      ymin = dur_days_lwr, 
+                      ymax = dur_days_upr),
+                  breed_pheno_raw,
+                  position = position_dodge(width = 0.5)) +
+  scale_linetype_manual(values = c("solid", "dashed")) +
+  labs(x = "Maternal age (years)",
+       y = "Breeding haul out (days)") +
+  theme(legend.position = "none",
+        axis.text.y = element_text(size = 8))
+
+molt_duration_plot <- ggpredict(
+  pheno_model, 
+  terms = c("age10 [all]", "age_cat", "phase", "year_fct [all]")
+) %>% 
+  as_tibble() %>%
+  transmute(age = x * 10 + age_senesce,
+            age_cat = group,
+            dur_days = predicted,
+            dur_days_lwr = conf.low,
+            dur_days_upr = conf.high,
+            phase = facet,
+            year_fct = panel) %>% 
+  left_join(n_phen_year, by = "year_fct") %>% 
+  filter(phase == "moltdur",
+         (age_cat == "Young" & age < age_senesce) |
+           (age_cat == "Old" & age >= age_senesce)) %>% 
+  group_by(age, age_cat) %>% 
+  summarize(across(starts_with("dur_days"), \(x) weighted.mean(x, n)),
+            .groups = "drop") %>% 
+  ggplot(aes(age)) +
+  # Predictions by year
+  geom_line(aes(age, predicted, group = interaction(phase, age_cat, year_fct)), 
+            filter(phen_pred_years, phase == "moltdur"),
+            alpha = 0.1) + 
+  # Weighted model (CI + mean)
+  geom_ribbon(aes(ymin = dur_days_lwr, ymax = dur_days_upr, fill = age_cat),
+              alpha = 0.25) +
+  geom_line(aes(y = dur_days, color = age_cat, linetype = age_cat)) +
+  # Summarized data
+  geom_pointrange(aes(y = dur_days, 
+                      ymin = dur_days_lwr, 
+                      ymax = dur_days_upr),
+                  molt_pheno_raw,
+                  position = position_dodge(width = 0.5)) +
+  scale_linetype_manual(values = c("solid", "dashed")) +
+  labs(x = "Maternal age (years)",
+       y = "Molt haul out (days)") +
+  theme(legend.position = "none",
+        axis.text.y = element_text(size = 8))
 
@@
 
-#Make sample size data frame
-n_phen <- seal_phenology %>% 
-  count(age)
-
 phen_dist <- seal_phenology %>% 
@@
   geom_col(fill = "grey50") + 
-  labs(x = "Age (Years)", y = "Number of individuals") + 
+  labs(x = "Maternal age (Years)", y = "Number of individuals") + 
   theme()
@@
                                           "foraging2"))) %>% 
-    left_join(n_phen, by = "age")
+    left_join(n_phen_age, by = "age")
 
@@
             .groups = "drop")
-
-phen_pred_years <- ggpredict(
-  pheno_model,
-  terms = c("age10 [all]", "phase", "year [all]"),
-  type = "random",
-  interval = "confidence"
-) %>% 
-  as_tibble() %>% 
-  mutate(age = x * 10,
-         phase = group)
-
-breed_pheno_raw <- pheno_raw[pheno_raw$phase == "breeddur",]
-molt_pheno_raw <- pheno_raw[pheno_raw$phase == "moltdur",]
-
-breed_duration_plot <- ggpredict(
-  pheno_model, 
-  terms = c("age10 [all]", "phase", "year [all]")
-) %>% 
-  as_tibble() %>% 
-  transmute(age = x * 10,
-            dur_days = predicted,
-            dur_days_lwr = conf.low,
-            dur_days_upr = conf.high,
-            phase = group,
-            year = as.numeric(as.character(facet))) %>% 
-  left_join(n_phenology, by = "year") %>% 
-  group_by(age) %>% 
-  filter(phase == "breeddur") %>% 
-  summarize(across(starts_with("dur_days"), \(x) weighted.mean(x, n)),
-            .groups = "drop") %>% 
-  ggplot(aes(age)) +
-  geom_line(aes(age, predicted, group = interaction(phase, facet)), 
-            phen_pred_years, 
-            alpha = 0.1) + 
-  geom_ribbon(aes(ymin = dur_days_lwr, ymax = dur_days_upr),
-              alpha = 0.25, 
-              fill = "#D05794") +
-  geom_pointrange(aes(y = dur_days, 
-                      ymin = dur_days_lwr, 
-                      ymax = dur_days_upr),
-                  breed_pheno_raw,
-                  position = position_dodge(width = 0.5), 
-                  color = "#D05794") +
-  geom_line(aes(y = dur_days), 
-            linetype = "dashed", 
-            color = "#D05794") +
-  labs(x = "Maternal age (years)",
-       y = "Breeding haul out (days)") +
-  theme(legend.position = "none",
-        axis.text.y = element_text(size = 8))
-
-molt_duration_plot <- ggpredict(
-  pheno_model,
-  terms = c("age10 [all]", "phase", "year [all]")
-) %>%
-  as_tibble() %>%
-  transmute(age = x * 10,
-            dur_days = predicted,
-            dur_days_lwr = conf.low,
-            dur_days_upr = conf.high,
-            phase = group,
-            year = as.numeric(as.character(facet))) %>%
-  left_join(n_phenology, by = "year") %>%
-  group_by(age, phase) %>%
-  filter(phase == "moltdur") %>%
-  summarize(across(starts_with("dur_days"), \(x) weighted.mean(x, n)),
-            .groups = "drop") %>%
-  ggplot(aes(age)) +
-  geom_line(aes(age, predicted, group = interaction(phase, facet)), 
-            phen_pred_years, 
-            alpha = 0.1) + 
-  geom_ribbon(aes(ymin = dur_days_lwr, ymax = dur_days_upr),
-              alpha = 0.25, 
-              fill = "#E09A2D") +
-  geom_pointrange(aes(y = dur_days, 
-                      ymin = dur_days_lwr, 
-                      ymax = dur_days_upr),
-                  molt_pheno_raw,
-                  position = position_dodge(width = 0.5), 
-                  color = "#E09A2D") +
-  geom_line(aes(y = dur_days), 
-            color = "#E09A2D") +
-  labs(x = "Maternal age (years)",
-       y = "Molting haul out (days)") +
-  theme(legend.position = "none", 
-        axis.text.y = element_text(size = 8))
 
@@
   scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
-  labs(x = "Biological age (years)") +
+  labs(x = "Biological age (years before death)") +
   theme(legend.position = "none") + 

### 2024-07-19 paper/paperv3.qmd
fix bioage 0 in bio v chrono fig

--- 2024-07-16 paper/paperv3.qmd
@@
                       ymax = breed_upr),
-                  bioage_summ) +
+                  filter(bioage_summ, bioage < 0)) +
+  geom_pointrange(aes(y = perc_breed,
+                      ymin = breed_lwr,
+                      ymax = breed_upr),
+                  color = "grey50",
+                  filter(bioage_summ, bioage == 0)) +
   geom_vline(xintercept = bioage_senesce - 0.5, linetype = "dashed") +
+  # Sample size labels
+  geom_text(aes(label = n_obs),
+            filter(bioage_summ, bioage < 0), 
+            y = 1, 
+            vjust = -0.5) +
+  geom_text(aes(label = n_obs),
+            filter(bioage_summ, bioage == 0), 
+            y = 1, 
+            vjust = -0.5,
+            color = "grey50") +
   scale_y_continuous("Breeding probability", 
@@
   labs(x = "Biological age (years)") +
-  theme(legend.position = "none") + 
-  geom_text(aes(label = n), y = 1, vjust = -0.5)
+  theme(legend.position = "none")
 

### 2024-07-19 paper/paperv3.qmd
other phen changes

--- 2024-07-19 paper/paperv3.qmd
@@
   filter(year >= 2011, 
-         moltdur >= 7) %>% 
+         moltdur >= 7, 
+         age <= 20) %>% 
   mutate(BA = firstobsbreeddoy, 
@@
 
+# #population level mean
+# phen_plot_pred <- ggpredict(
+#   pheno_model, 
+#   terms = c("age10 [all]", "age_cat", "phase", "year_fct [all]")
+# ) %>% 
+#   as_tibble() %>%
+#   transmute(age = x * 10 + age_senesce,
+#             age_cat = group,
+#             dur_days = predicted,
+#             dur_days_lwr = conf.low,
+#             dur_days_upr = conf.high,
+#             phase = facet,
+#             year_fct = panel) %>% 
+#   left_join(n_phen_year, by = "year_fct") %>% 
+#   filter((age_cat == "Young" & age < age_senesce) |
+#            (age_cat == "Old" & age >= age_senesce))
+
+#weighted model
 pheno_pred <- pheno_grid %>% 
@@
   left_join(n_phen_year, by = "year_fct") %>% 
-  group_by(age, age_cat) %>% 
+  group_by(age, age_cat, phase) %>% 
   summarize(across(c(predicted, predicted_pop, conf_lo, conf_hi), 
@@
 
-phen_pred_years <- ggpredict(
-  pheno_model,
-  terms = c("age10 [all]", "age_cat", "phase", "year_fct [all]"),
-  type = "random",
-  interval = "confidence"
-) %>% 
-  as_tibble() %>% 
-  mutate(age = x * 10 + age_senesce,
-         age_cat = group,
-         phase = facet,
-         year_fct = panel) %>% 
-  filter((age_cat == "Young" & age < age_senesce) |
-           (age_cat == "Old" & age >= age_senesce))
-
+#random effect of year
+phen_pred_years <- pheno_grid %>% 
+  mutate(predicted = predict(pheno_model, 
+                             newdata = pheno_grid,
+                             type = "response",
+                             re.form = ~ (1 | year_fct)))
+
+#raw data
 pheno_raw <- seal_phenology %>% 
@@
 
-breed_pheno_raw <- pheno_raw[pheno_raw$phase == "breeddur",]
-molt_pheno_raw <- pheno_raw[pheno_raw$phase == "moltdur",]
-
-breed_duration_plot <- ggpredict(
-  pheno_model, 
-  terms = c("age10 [all]", "age_cat", "phase", "year_fct [all]")
-) %>% 
-  as_tibble() %>%
-  transmute(age = x * 10 + age_senesce,
-            age_cat = group,
-            dur_days = predicted,
-            dur_days_lwr = conf.low,
-            dur_days_upr = conf.high,
-            phase = facet,
-            year_fct = panel) %>% 
-  left_join(n_phen_year, by = "year_fct") %>% 
-  filter(phase == "breeddur",
-         (age_cat == "Young" & age < age_senesce) |
-           (age_cat == "Old" & age >= age_senesce)) %>% 
-  group_by(age, age_cat) %>% 
-  summarize(across(starts_with("dur_days"), \(x) weighted.mean(x, n)),
-            .groups = "drop") %>% 
+breed_duration_plot <- pheno_pred %>% 
+  filter(phase == "breeddur") %>% 
   ggplot(aes(age)) +
@@
   # Weighted model (CI + mean)
-  geom_ribbon(aes(ymin = dur_days_lwr, ymax = dur_days_upr, fill = age_cat),
+  geom_ribbon(aes(ymin = conf_lo, ymax = conf_hi, fill = age_cat),
               alpha = 0.25) +
-  geom_line(aes(y = dur_days, color = age_cat, linetype = age_cat)) +
+  geom_line(aes(y = predicted, color = age_cat, linetype = age_cat)) +
   # Summarized data
@@
                       ymax = dur_days_upr),
-                  breed_pheno_raw,
+                  filter(pheno_raw, phase == "breeddur"),
                   position = position_dodge(width = 0.5)) +
+    # Population-level mean response
+  geom_line(aes(y = predicted_pop, group = age_cat),
+            linetype = "dotted", linewidth = 1.2, alpha = 0.5) +
+  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
+  geom_text(aes(label = n), y = 36.5, vjust = 0) +
   scale_linetype_manual(values = c("solid", "dashed")) +
+  scale_color_manual(values = c("#7fbc41", "#de77ae")) +
+  scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
+  scale_y_continuous(limits = c(15, 37)) +
   labs(x = "Maternal age (years)",
@@
 
-molt_duration_plot <- ggpredict(
-  pheno_model, 
-  terms = c("age10 [all]", "age_cat", "phase", "year_fct [all]")
-) %>% 
-  as_tibble() %>%
-  transmute(age = x * 10 + age_senesce,
-            age_cat = group,
-            dur_days = predicted,
-            dur_days_lwr = conf.low,
-            dur_days_upr = conf.high,
-            phase = facet,
-            year_fct = panel) %>% 
-  left_join(n_phen_year, by = "year_fct") %>% 
-  filter(phase == "moltdur",
-         (age_cat == "Young" & age < age_senesce) |
-           (age_cat == "Old" & age >= age_senesce)) %>% 
-  group_by(age, age_cat) %>% 
-  summarize(across(starts_with("dur_days"), \(x) weighted.mean(x, n)),
-            .groups = "drop") %>% 
+molt_duration_plot <- pheno_pred %>% 
+  filter(phase == "moltdur") %>% 
   ggplot(aes(age)) +
@@
                       ymax = dur_days_upr),
-                  molt_pheno_raw,
+                  filter(pheno_raw, phase == "moltdur"),
                   position = position_dodge(width = 0.5)) +
+    # Population-level mean response
+  geom_line(aes(y = predicted_pop, group = age_cat),
+            linetype = "dotted", linewidth = 1.2, alpha = 0.5) +
+  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
+  geom_text(aes(label = n), y = 36.5, vjust = 0) +
   scale_linetype_manual(values = c("solid", "dashed")) +
+  scale_color_manual(values = c("#7fbc41", "#de77ae")) +
+  scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
+  scale_y_continuous(limits = c(15, 37)) +
   labs(x = "Maternal age (years)",
@@
 B: Raw data for breeding and molting durations (mean ± se) and predictions (mean and 95% CI) from negative binomial weighted by number of seals obersved per year."
-#| fig-width: 9
+#| fig-width: 12
 #| fig-height: 9
@@
   geom_col(fill = "grey50") + 
-  labs(x = "Maternal age (Years)", y = "Number of individuals") + 
+  labs(x = "Maternal age (years)", y = "Number of individuals") + 
   theme()

### 2024-07-19 paper/paperv3.qmd
sex ratio 1, 2 tail p vals

--- 2024-07-19 paper/paperv3.qmd
@@
 # Are hypotheses supported by the model?
-extract_p <- \(h) map_dbl(mod_sens, \(m) pnorm(m[[h]]["age10:age_catOld", "z value"]))
+extract_p <- function(h, tails = 1) {
+  if (tails == 1) {
+    map_dbl(mod_sens, \(m) pnorm(m[[h]]["age10:age_catOld", "z value"]))
+  } else if (tails == 2) {
+    map_dbl(mod_sens, \(m) 2 * pnorm(m[[h]]["age10:age_catOld", "z value"],
+                                     lower.tail = FALSE))
+  } else {
+    stop("tails must be 1 or 2")
+  }
+}
 p_tbl <- tibble(age_cutoff = cutoff, 
@@
                 "H2b (recruitment)" = extract_p("h2_recr"),
-                "H3 (sex ratio)" = extract_p("h3") * 2)
+                "H3 (sex ratio)" = extract_p("h3"),
+                "H3 (sex ratio; 2 tail)" = extract_p("h3", tails = 2))
 
@@
   pivot_longer(-age_cutoff, names_to = "hypothesis", values_to = "p") %>%
+  mutate(hypgrp = ifelse(str_detect(hypothesis, "H3"), 
+                         "H3 (sex ratio)", 
+                         hypothesis)) %>% 
   ggplot(aes(age_cutoff, p)) +
-  geom_point() +
+  geom_point(aes(color = hypothesis == "H3 (sex ratio; 2 tail)")) +
   geom_hline(yintercept = 0.05, linetype = "dashed") +
   geom_text(aes(x, y, label = tag),
-            tibble(hypothesis = colnames(p_tbl)[-1],
+            tibble(hypgrp = colnames(p_tbl)[2:5],
                    tag = LETTERS[1:4],
@@
   scale_y_log10(labels = scales::number) +
+  scale_color_manual(values = c("black", "grey50")) +
   labs(x = "Senescence onset (years)") +
   expand_limits(y = 3) +
-  facet_wrap(~hypothesis)
+  facet_wrap(~hypgrp) +
+  theme(legend.position = "none")
 

### 2024-07-19 paper/paperv3.qmd
missing obs (age 4 on) and quad/break aic

@@
 
-TODO: fix sample size for bioage=0
-
 ## Observation frequency
@@
 
-TODO: include from age 4 on
-
 ```{r}
@@
 missing_seals <- sealdat %>% 
-  # Need to see a seal 3+ years for calculation to make sense
-  group_by(animalID) %>% 
-  mutate(n_years = n_distinct(year)) %>% 
-  ungroup() %>% 
-  filter(n_years >= 3,
-         age <= 20) %>% 
+  filter(age <= 20) %>% 
   # Fill in non-resighted years
@@
   arrange(year) %>% 
-  reframe(tibble(year2 = (min(year) + 1):(max(year) - 1),
+  reframe(tibble(year2 = (yearborn[1] + 4):max(year), # from age 4 to last obs
                  unobserved = as.integer(!year2 %in% year),
@@
   rename(year = year2) %>% 
+  filter(year >= 1988) %>% 
   mutate(age_cat = factor(age >= age_senesce, 
@@
 
-missing_mod <- glmer(unobserved ~ age10 + (1 | year_fct) + (1 | animalID), 
+# Summary: mean and 95% CI 
+missing_summ <- missing_seals %>% 
+  group_by(age, age_cat) %>% 
+  summarize(n = n(),
+            n_unobserved = sum(unobserved),
+            unobserved_mean = n_unobserved / n,
+            unobserved_lwr = binom.test(n_unobserved, n)$conf.int[1],
+            unobserved_upr = binom.test(n_unobserved, n)$conf.int[2],
+            .groups = "drop")
+
+missing_mod <- glmer(
+  unobserved ~ age10 + (1 | year_fct) + (1 | animalID), 
   missing_seals, 
-                     family = binomial)
+  family = binomial
+)
 
@@
 
-missing_age_se <- summary(missing_mod)$coef["age10", "Std. Error"]
-missing_age_est <- summary(missing_mod)$coef["age10", "Estimate"]
-missing_age_coef <- sprintf("%0.2f (95%% CI: %0.2f - %0.2f)",
-                            missing_age_est,
-                            missing_age_est - 1.96 * missing_age_se,
-                            missing_age_est + 1.96 * missing_age_se)
-
-missing_mod_old <- glmer(
-  unobserved ~ age10 + (1 | year_fct),
-  filter(missing_seals, age >= 11),
-  family = binomial
-)
-
-summary(missing_mod_old)
-
-missing_old_se <- summary(missing_mod_old)$coef["age10", "Std. Error"]
-missing_old_est <- summary(missing_mod_old)$coef["age10", "Estimate"]
-missing_old_coef <- sprintf("%0.2f (95%% CI: %0.2f - %0.2f)",
-                            missing_old_est,
-                            missing_old_est - 1.96 * missing_old_se,
-                            missing_old_est + 1.96 * missing_old_se)
-
 ```
@@
 First, we tested whether probability of detection was related to age.
-We found the proportion of unobserved animals increased with age.
-For this analysis, we excluded the first and last year of observation for each animal, because animals were necessarily observed in those years.
-For all years between, we determined whether animals were observed.
-Then we estimated the effect of age on probability of an animal being unobserved using a GLMM (binomial response, logit link) with random effects for year and animal ([@fig-unobserved]).
-The coefficient for age was `r missing_age_coef` (*p*\<0.0001).
-The old-animals-only model estimated the age coefficient to be `r missing_old_coef`.
+We found the proportion of unobserved animals decreased with age (i.e., older seals are more likely to be observed).
+For this analysis, we excluded the last year each animal was observed.
+Then, for all years between animals' age 4 year to the year prior to final observation, we determined whether each animal was observed.
+We estimated the effect of age on probability of an animal being unobserved using a GLMM (binomial response, logit link) with random effects for year and animal ([@fig-unobserved]).
+The coefficient for age was -0.2593 (95% CI: -0.4896 - -0.0290, *p* = 0.027).
+The change in mean unobserved rate from age 4 to 20 was 20.0% to 14.2%.
 
@@
 #| label: fig-unobserved
-#| fig-cap: The probability of an animal being unobserved increased with age.
-Points and error bars represent the proportion of unobserved animals within age classes plus/minus one standard error.
-The blue line and shaded area are the mean and 95% CI probability of an animal being unobserved estimated with a generalized linear mixed effect model.
+#| fig-cap: The probability of an animal being unobserved decreased with age.
+Points and error bars represent the proportion of unobserved animals with 95% CI.
+The blue line and shaded area are the mean and 95% CI probability of an animal being unobserved (GLMM, with random effect of year weighted by number of seals observed each year).
+The dotted line is the mean response of the unweighted predictions.
+Numbers at the top indicate number of seals in each age category.
 See text for model details.
 
-missing_grid <- expand_grid(age = 5:19, 
+# Prediction grid
+missing_grid <- expand_grid(age = 4:20, 
                             year_fct = factor(1989:2022)) %>% 
@@
 
+# Missing seal predictions (weighted model)
 missing_pred <- missing_grid %>% 
@@
                              re.form = ~ (1 | year_fct)),
+         pred_pop = predict(missing_mod, 
+                            newdata = .,
+                            type = "response",
+                            re.form = NA),
          se = suppressWarnings(predict(missing_mod, 
@@
   group_by(age) %>% 
-  summarize(across(c(predicted, conf.low, conf.high), mean)) %>% 
-  mutate(across(c(predicted, conf.low, conf.high), family(missing_mod)$linkinv))
-
+  summarize(across(c(predicted, pred_pop, conf.low, conf.high), mean)) %>% 
+  mutate(across(c(predicted, conf.low, conf.high), 
+                family(missing_mod)$linkinv))
+
+# Yearly predictions
 missing_pred_years <- missing_grid %>% 
@@
 
-missing_pred_old <- expand_grid(age = 11:19, 
-                                year_fct = factor(1989:2022)) %>% 
-  mutate(age10 = (age - age_senesce) / 10) %>% 
-  mutate(predicted = predict(missing_mod_old, 
-                             newdata = .,
-                             re.form = ~ (1 | year_fct)),
-         se = suppressWarnings(predict(missing_mod, 
-                                       newdata = .,
-                                       se.fit = TRUE,
-                                       re.form = NA)$se.fit),
-         conf.low = predicted - 1.96 * se,
-         conf.high = predicted + 1.96 * se) %>% 
-  left_join(count(filter(missing_seals, age >= 11), year_fct), 
-            by = "year_fct") %>% 
-  group_by(age) %>% 
-  summarize(across(c(predicted, conf.low, conf.high), mean)) %>% 
-  mutate(across(c(predicted, conf.low, conf.high), family(missing_mod)$linkinv))
-
-missing_pred_pop <- tibble(age = 5:19,
-                           age10 = (age - age_senesce) / 10) %>% 
-  mutate(predicted = predict(missing_mod, 
-                             newdata = .,
-                             type = "response",
-                             re.form = NA))
-
-missing_seals %>% 
-  group_by(age) %>% 
-  summarize(n = n(),
-            n_unobserved = sum(unobserved),
-            unobserved_mean = n_unobserved / n,
-            unobserved_lwr = binom.test(n_unobserved, n)$conf.int[1],
-            unobserved_upr = binom.test(n_unobserved, n)$conf.int[2]) %>% 
-  ggplot(aes(age)) +
+ggplot(missing_pred, aes(age)) +
+  # Line per year
   geom_line(aes(y = predicted, group = year_fct), 
@@
             alpha = 0.2) +
-  # Overall model
+  # Weighted model
   geom_ribbon(aes(ymin = conf.low, ymax = conf.high),
-              missing_pred,
-              alpha = 0.3,
-              fill = "firebrick") +
-  geom_line(aes(y = predicted),
-            missing_pred,
-            color = "firebrick",
-            linewidth = 2) +
-  # Old only
-  geom_ribbon(aes(ymin = conf.low, ymax = conf.high),
-              missing_pred_old,
               alpha = 0.3,
@@
   geom_line(aes(y = predicted),
-            missing_pred_old,
-            linetype = "dashed",
-            color = "cornflowerblue",
-            linewidth = 2) +
+            linewidth = 2,
+            color = "cornflowerblue") +
+  # Population-level model
+  geom_line(aes(y = pred_pop), 
+            linetype = "dotted",
+            linewidth = 1.2,
+            color = "cornflowerblue") +
   # Raw data
   geom_pointrange(aes(y = unobserved_mean, 
-                      ymin = unobserved_lwr, ymax = unobserved_upr)) +
-  scale_x_continuous("Age", breaks = seq(5, 19, by = 4)) +
+                      ymin = unobserved_lwr, 
+                      ymax = unobserved_upr),
+                  missing_summ) +
+  # Sample size
+  geom_text(aes(age, label = n), missing_summ, y = 0.8, vjust = 1.1) +
+  scale_x_continuous("Age", breaks = seq(4, 20, by = 4)) +
   scale_y_continuous("Unobserved proportion", 
                      labels = scales::label_percent(accuracy = 1),
-                     limits = c(0, NA))
+                     limits = c(0, NA)) +
+  theme(legend.position = "none")
 ```
@@
 ```
+
+This needs to be rethought now!
 
@@
 
-TODO: AIC comparison
-
 ### H1: Reproductive senescence
@@
 # H2b
-
 recr_age_quad <- glmer(
@@
 ```
+
+### AIC Comparison
+
+```{r}
+#| label: quad-aic
+
+aic <- function(break_mod, quad_mod, hyp) {
+  AIC(break_mod, quad_mod) %>% 
+    as_tibble() %>% 
+    mutate(dAIC = AIC - min(AIC),
+         rellik = exp(-0.5 * dAIC),
+         weight = rellik / sum(rellik),
+         hypothesis = hyp,
+         model = c("Breakpoint", "Quadratic")) %>% 
+  select(hypothesis, model, AIC, weight)
+}
+
+pmap(list(break_mod = list(breed_age_mod, pup_surv_mod, pup_recr_mod),
+          quad_mod = list(breed_age_quad, surv_age_quad, recr_age_quad),
+          hyp = c("H1", "H2a", "H2b")),
+     aic) %>% 
+  list_rbind() %>% 
+  knitr::kable()
+```

### 2024-07-19 paper/paperv3.qmd

@@
 seal_phenology <- sealdat %>% 
-  filter(year >= 2011) %>% 
+  filter(year >= 2011, 
+         moltdur >= 7, 
+         age <= 20) %>% 
   mutate(BA = firstobsbreeddoy, 
@@
 ```{r}
-#| label: fig-pheno
-#| fig-cap: "Molting gets shorter but breeding doesn't.
-A: Observed annual time allocation averaged over individuals.
-B: Raw data for breeding and molting durations (mean ± se) and predictions (mean and 95% CI) from negative binomial weighted by number of seals obersved per year."
-#| fig-width: 9
-#| fig-height: 9
+
+#Make sample size data frame
+n_phen_age <- count(seal_phenology, age)
+n_phen_year <- count(seal_phenology, year_fct)
+
+pheno_grid <- expand_grid(
+  age = 4:20,
+  year_fct = factor(2011:2023), 
+  phase = c("breeddur", "moltdur")
+) %>% 
+  mutate(age10 = (age - age_senesce) / 10,
+         age_cat = factor(ifelse(age < age_senesce, "Young", "Old"),
+                          levels = c("Young", "Old")))
+
+# #population level mean
+# phen_plot_pred <- ggpredict(
+#   pheno_model, 
+#   terms = c("age10 [all]", "age_cat", "phase", "year_fct [all]")
+# ) %>% 
+#   as_tibble() %>%
+#   transmute(age = x * 10 + age_senesce,
+#             age_cat = group,
+#             dur_days = predicted,
+#             dur_days_lwr = conf.low,
+#             dur_days_upr = conf.high,
+#             phase = facet,
+#             year_fct = panel) %>% 
+#   left_join(n_phen_year, by = "year_fct") %>% 
+#   filter((age_cat == "Young" & age < age_senesce) |
+#            (age_cat == "Old" & age >= age_senesce))
+
+#weighted model
+pheno_pred <- pheno_grid %>% 
+  mutate(predicted = predict(pheno_model, 
+                             newdata = pheno_grid,
+                             re.form = ~ (1 | year_fct)),
+         predicted_se = suppressWarnings(
+           # warning: "se.fit computation uses an approximation to estimate the sampling distribution of the parameters"
+           predict(pheno_model, 
+                   newdata = pheno_grid,
+                   re.form = NA,
+                   se.fit = TRUE),
+         )$se.fit,
+         predicted_pop = predict(pheno_model, 
+                                 newdata = pheno_grid,
+                                 re.form = NA),
+         conf_lo = predicted - 1.96 * predicted_se,
+         conf_hi = predicted + 1.96 * predicted_se) %>% 
+  left_join(n_phen_year, by = "year_fct") %>% 
+  group_by(age, age_cat, phase) %>% 
+  summarize(across(c(predicted, predicted_pop, conf_lo, conf_hi), 
+                   \(x) weighted.mean(x, n)),
+            .groups = "drop") %>% 
+  # invert logit
+  mutate(across(c(predicted, predicted_pop, conf_lo, conf_hi), 
+                family(pheno_model)$linkinv)) %>% 
+  left_join(n_phen_age, by = "age")
+
+#random effect of year
+phen_pred_years <- pheno_grid %>% 
+  mutate(predicted = predict(pheno_model, 
+                             newdata = pheno_grid,
+                             type = "response",
+                             re.form = ~ (1 | year_fct)))
+
+#raw data
+pheno_raw <- seal_phenology %>% 
+  select(breeddur, moltdur, age) %>% 
+  pivot_longer(c(breeddur, moltdur), 
+               names_to = "phase", 
+               values_to = "dur_days") %>% 
+  group_by(age, phase) %>% 
+  summarize(n = n(),
+            dur_days_se = sd(dur_days) / sqrt(n),
+            dur_days = mean(dur_days),
+            dur_days_lwr = dur_days - dur_days_se,
+            dur_days_upr = dur_days + dur_days_se,
+            .groups = "drop")
+
+breed_duration_plot <- pheno_pred %>% 
+  filter(phase == "breeddur") %>% 
+  ggplot(aes(age)) +
+  # Predictions by year
+  geom_line(aes(age, predicted, group = interaction(phase, age_cat, year_fct)), 
+            filter(phen_pred_years, phase == "breeddur"),
+            alpha = 0.1) + 
+  # Weighted model (CI + mean)
+  geom_ribbon(aes(ymin = conf_lo, ymax = conf_hi, fill = age_cat),
+              alpha = 0.25) +
+  geom_line(aes(y = predicted, color = age_cat, linetype = age_cat)) +
+  # Summarized data
+  geom_pointrange(aes(y = dur_days, 
+                      ymin = dur_days_lwr, 
+                      ymax = dur_days_upr),
+                  filter(pheno_raw, phase == "breeddur"),
+                  position = position_dodge(width = 0.5)) +
+    # Population-level mean response
+  geom_line(aes(y = predicted_pop, group = age_cat),
+            linetype = "dotted", linewidth = 1.2, alpha = 0.5) +
+  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
+  geom_text(aes(label = n), y = 36.5, vjust = 0) +
+  scale_linetype_manual(values = c("solid", "dashed")) +
+  scale_color_manual(values = c("#7fbc41", "#de77ae")) +
+  scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
+  scale_y_continuous(limits = c(15, 37)) +
+  labs(x = "Maternal age (years)",
+       y = "Breeding haul out (days)") +
+  theme(legend.position = "none",
+        axis.text.y = element_text(size = 8))
+
+molt_duration_plot <- pheno_pred %>% 
+  filter(phase == "moltdur") %>% 
+  ggplot(aes(age)) +
+  # Predictions by year
+  geom_line(aes(age, predicted, group = interaction(phase, age_cat, year_fct)), 
+            filter(phen_pred_years, phase == "moltdur"),
+            alpha = 0.1) + 
+  # Weighted model (CI + mean)
+  geom_ribbon(aes(ymin = conf_lo, ymax = conf_hi, fill = age_cat),
+              alpha = 0.25) +
+  geom_line(aes(y = predicted, color = age_cat, linetype = age_cat)) +
+  # Summarized data
+  geom_pointrange(aes(y = dur_days, 
+                      ymin = dur_days_lwr, 
+                      ymax = dur_days_upr),
+                  filter(pheno_raw, phase == "moltdur"),
+                  position = position_dodge(width = 0.5)) +
+    # Population-level mean response
+  geom_line(aes(y = predicted_pop, group = age_cat),
+            linetype = "dotted", linewidth = 1.2, alpha = 0.5) +
+  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
+  geom_text(aes(label = n), y = 36.5, vjust = 0) +
+  scale_linetype_manual(values = c("solid", "dashed")) +
+  scale_color_manual(values = c("#7fbc41", "#de77ae")) +
+  scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
+  scale_y_continuous(limits = c(15, 37)) +
+  labs(x = "Maternal age (years)",
+       y = "Molt haul out (days)") +
+  theme(legend.position = "none",
+        axis.text.y = element_text(size = 8))
+
+```
+
+```{r}
 
@@
   count(year)
-
-#Make sample size data frame
-n_phen <- seal_phenology %>% 
-  count(age)
 
@@
   geom_col(fill = "grey50") + 
-  labs(x = "Age (Years)", y = "Number of individuals") + 
+  labs(x = "Maternal age (years)", y = "Number of individuals") + 
   theme()
@@
   theme(legend.position = "none")
-
 ```

### 2024-07-22 paper/paperv3.qmd
Small figure changes

@@
   mutate(animalID = factor(animalID, levels = unique(animalID))) %>% 
+  filter(year > 1987) %>% 
   ggplot(aes(year, animalID)) +
@@
         legend.title = element_blank(),
-        # legend.position.inside = c(0.01, 0.99),
+        legend.position = c(0.1, 0.95),
         legend.justification = c(0, 1),
@@
 
-ggplot(breed_pred_wgt, aes(age, predicted)) +
+breed_plot <- ggplot(breed_pred_wgt, aes(age, predicted)) +
   # Predictions for individual years
@@
 
-ggsave("outputs/repro_senesce.svg", width = 13, height = 6, units = "in")
+ggsave("outputs/repro_senesce.svg", breed_plot, width = 13, height = 6, units = "in")
 ```
@@
 #| label: h1-pred-deets
+
+pnorm(summary(breed_age_mod)$coef["age10:age_catOld", "z value"])
 
@@
 summary(pup_recr_mod)$coefficients
+
+#survival 1 tailed
+pnorm(summary(pup_surv_mod)$coef["age10:age_catOld", "z value"])
+#recruit 1 tailed
+pnorm(summary(pup_recr_mod)$coef["age10:age_catOld", "z value"])
 
@@
   coord_cartesian(ylim = c(0, 0.6)) +
-  labs(x = "Age (years)") +
-  theme(legend.position = "none") + 
-  geom_label(aes(label = n), y = 0.6, vjust = 1)
+  theme(legend.position = "none", 
+        axis.title.x = element_blank()) + 
+  geom_text(aes(label = n), y = 0.6, vjust = 1, nudge_x = 0.2)
 
@@
   count(age)
+n_recruit_indiv <- surv_recr_data %>% 
+  filter(pupsex == "F") %>% 
+  group_by(animalID)
 
@@
   theme(legend.position = "none") + 
-  geom_label(aes(label = n), y = 0.6, vjust = 1)
+  geom_text(aes(label = n), y = 0.6, vjust = 1, nudge_x = 0.2)
 
@@
 Sample size for offspring survival was n = `r nrow(surv_recr_data)` observations of `r n_distinct(surv_recr_data$animalID)` individuals.
-Sample size for offspring recruitment was n = `r nrow(n_recruit)` observations of `r n_distinct(n_recruit$animalID)` female individuals.
+Sample size for offspring recruitment was n = `r n_distinct(n_recruit_indiv)` observations of `r n_distinct(n_recruit_indiv$animalID)` female individuals.
 
@@
 summary(sexrat_age_mod)$coefficients
+pnorm(summary(sexrat_age_mod)$coef["age10:age_catOld", "z value"])
 
@@
   scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
-  labs(x = "Age (years)") +
+  labs(x = "Maternal age (years)") +
   theme(legend.position = "none") +
@@
             alpha = 0.1) + 
+    # Population-level mean response
+  geom_line(aes(y = predicted_pop, group = age_cat),
+            linetype = "dotted", linewidth = 1.2, alpha = 0.5) +
   # Weighted model (CI + mean)
@@
                   filter(pheno_raw, phase == "breeddur"),
-                  position = position_dodge(width = 0.5)) +
-    # Population-level mean response
-  geom_line(aes(y = predicted_pop, group = age_cat),
-            linetype = "dotted", linewidth = 1.2, alpha = 0.5) +
+                  position = position_dodge(width = 0.5), 
+                  size = 0.2) +
   geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
-  geom_text(aes(label = n), y = 36.5, vjust = 0) +
+  geom_text(aes(label = n), y = 36.5, vjust = 0, size = 3) +
   scale_linetype_manual(values = c("solid", "dashed")) +
@@
   scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
-  scale_y_continuous(limits = c(15, 37)) +
-  labs(x = "Maternal age (years)",
-       y = "Breeding haul out (days)") +
+  scale_y_continuous(limits = c(15, 40)) +
+  labs(y = "Breeding haul out (days)") +
   theme(legend.position = "none",
-        axis.text.y = element_text(size = 8))
+        axis.text.y = element_text(size = 6), 
+        axis.title.x = element_blank())
 
@@
             alpha = 0.1) + 
+    # Population-level mean response
+   geom_line(aes(y = predicted_pop, group = age_cat),
+             linetype = "dotted", linewidth = 1.2, alpha = 0.2) +
   # Weighted model (CI + mean)
@@
                   filter(pheno_raw, phase == "moltdur"),
-                  position = position_dodge(width = 0.5)) +
-    # Population-level mean response
-  geom_line(aes(y = predicted_pop, group = age_cat),
-            linetype = "dotted", linewidth = 1.2, alpha = 0.5) +
+                  position = position_dodge(width = 0.5), 
+                  size = 0.2) +
   geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
-  geom_text(aes(label = n), y = 36.5, vjust = 0) +
+  geom_text(aes(label = n), y = 36.5, vjust = 0, size = 3) +
   scale_linetype_manual(values = c("solid", "dashed")) +
@@
   scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
-  scale_y_continuous(limits = c(15, 37)) +
+  scale_y_continuous(limits = c(15, 40)) +
   labs(x = "Maternal age (years)",
@@
   theme(legend.position = "none",
-        axis.text.y = element_text(size = 8))
-
-```
-
-```{r}
+        axis.text.y = element_text(size = 6))
+
+```
+
+```{r}
+
+mean_yb_phen <- pheno_raw %>% 
+  filter(age < age_senesce, phase == "breeddur") %>% 
+  summarize(mean_yb = mean(dur_days), 
+            sd_yb = sd(dur_days))
+mean_ob_phen <- pheno_raw %>% 
+  filter(age >= age_senesce, phase == "breeddur") %>% 
+  summarize(mean_ob = mean(dur_days), 
+            sd_ob = sd(dur_days))
+mean_ym_phen <- pheno_raw %>% 
+  filter(age < age_senesce, phase == "moltdur") %>% 
+  summarize(mean_ym = mean(dur_days), 
+            sd_ym = sd(dur_days))
+mean_om_phen <- pheno_raw %>% 
+  filter(age >= age_senesce, phase == "moltdur") %>% 
+  summarize(mean_om = mean(dur_days), 
+            sd_om = sd(dur_days))
+```
+
+Overall, the breeding haul out was `r mean_yb_phen$mean_yb` ± `r mean_yb_phen$sd_yb` days for animals younger than 11 years old and `r mean_ob_phen$mean_ob` ± `r mean_ob_phen$sd_ob` days for animals 11 years old and older.
+The molting haul out was `r mean_ym_phen$mean_ym` ± `r mean_ym_phen$sd_ym` days for animals younger than 11 years old and `r mean_om_phen$mean_om` ± `r mean_om_phen$sd_om` days for animals 11 years old and older.
+
+```{r}
+#| label: phen-fig
+#| fig-width: 8
 
@@
     geom_text(aes(label = "Breeding haul out"), 
-            x = 8,
+            x = 9,
             y = 30,
@@
     geom_text(aes(label = "Molting haul out"), 
-            x = 8,
+            x = 9,
             y = 135,
@@
 
-pheno_raw <- seal_phenology %>% 
-  select(breeddur, moltdur, age) %>% 
-  pivot_longer(c(breeddur, moltdur), 
-               names_to = "phase", 
-               values_to = "dur_days") %>% 
-  group_by(age, phase) %>% 
-  summarize(n = n(),
-            dur_days_se = sd(dur_days) / sqrt(n),
-            dur_days = mean(dur_days),
-            dur_days_lwr = dur_days - dur_days_se,
-            dur_days_upr = dur_days + dur_days_se,
-            .groups = "drop")
-
-phen_pred_years <- ggpredict(
-  pheno_model,
-  terms = c("age10 [all]", "phase", "year [all]"),
-  type = "random",
-  interval = "confidence"
-) %>% 
-  as_tibble() %>% 
-  mutate(age = x * 10,
-         phase = group)
-
-breed_pheno_raw <- pheno_raw[pheno_raw$phase == "breeddur",]
-molt_pheno_raw <- pheno_raw[pheno_raw$phase == "moltdur",]
-
-breed_duration_plot <- ggpredict(
-  pheno_model, 
-  terms = c("age10 [all]", "phase", "year [all]")
-) %>% 
-  as_tibble() %>% 
-  transmute(age = x * 10,
-            dur_days = predicted,
-            dur_days_lwr = conf.low,
-            dur_days_upr = conf.high,
-            phase = group,
-            year = as.numeric(as.character(facet))) %>% 
-  left_join(n_phenology, by = "year") %>% 
-  group_by(age) %>% 
-  filter(phase == "breeddur") %>% 
-  summarize(across(starts_with("dur_days"), \(x) weighted.mean(x, n)),
-            .groups = "drop") %>% 
-  ggplot(aes(age)) +
-  geom_line(aes(age, predicted, group = interaction(phase, facet)), 
-            filter(phen_pred_years, phase == "breeddur"), 
-            alpha = 0.1) + 
-  geom_ribbon(aes(ymin = dur_days_lwr, ymax = dur_days_upr),
-              alpha = 0.25, 
-              fill = "#D05794") +
-  geom_pointrange(aes(y = dur_days, 
-                      ymin = dur_days_lwr, 
-                      ymax = dur_days_upr),
-                  breed_pheno_raw,
-                  position = position_dodge(width = 0.5), 
-                  color = "#D05794") +
-  geom_line(aes(y = dur_days), 
-            linetype = "dashed", 
-            color = "#D05794") +
-  labs(x = "Maternal age (years)",
-       y = "Breeding haul out (days)") +
-  theme(legend.position = "none",
-        axis.text.y = element_text(size = 8))
-
-molt_duration_plot <- ggpredict(
-  pheno_model,
-  terms = c("age10 [all]", "phase", "year [all]")
-) %>%
-  as_tibble() %>%
-  transmute(age = x * 10,
-            dur_days = predicted,
-            dur_days_lwr = conf.low,
-            dur_days_upr = conf.high,
-            phase = group,
-            year = as.numeric(as.character(facet))) %>%
-  left_join(n_phenology, by = "year") %>%
-  group_by(age, phase) %>%
-  filter(phase == "moltdur") %>%
-  summarize(across(starts_with("dur_days"), \(x) weighted.mean(x, n)),
-            .groups = "drop") %>%
-  ggplot(aes(age)) +
-  geom_line(aes(age, predicted, group = interaction(phase, facet)), 
-            filter(phen_pred_years, phase == "moltdur"), 
-            alpha = 0.1) + 
-  geom_ribbon(aes(ymin = dur_days_lwr, ymax = dur_days_upr),
-              alpha = 0.25, 
-              fill = "#E09A2D") +
-  geom_pointrange(aes(y = dur_days, 
-                      ymin = dur_days_lwr, 
-                      ymax = dur_days_upr),
-                  molt_pheno_raw,
-                  position = position_dodge(width = 0.5), 
-                  color = "#E09A2D") +
-  geom_line(aes(y = dur_days), 
-            color = "#E09A2D") +
-  labs(x = "Maternal age (years)",
-       y = "Molting haul out (days)") +
-  theme(legend.position = "none", 
-        axis.text.y = element_text(size = 8))
-
-(annual_plot | phen_dist | breed_duration_plot | molt_duration_plot) + plot_annotation(tag_levels = "A") + plot_layout(ncol = 2, nrow = 2)
+design <- "
+  12
+  13
+"
+
+annual_plot + breed_duration_plot + molt_duration_plot + 
+  plot_annotation(tag_levels = "A") + 
+  plot_layout(design = design, 
+              widths = c(1, 2)) 
 ```
@@
 
-chrono_bio_aic <- AIC(breed_bioage_mod, breed_age_mod2)
-daic <- chrono_bio_aic$AIC - min(chrono_bio_aic$AIC)
-chrono_bio_rellik <- exp(-0.5 * daic)
-chrono_bio_aicw <- chrono_bio_rellik / sum(chrono_bio_rellik)
+```
+
+```{r}
+#| label: chrono-bio-aic
+
+AIC(breed_bioage_mod, breed_age_mod2) %>% 
+  as_tibble() %>% 
+  mutate(dAIC = AIC - min(AIC),
+         rellik = exp(-0.5 * dAIC),
+         weight = rellik / sum(rellik),
+         model = c("Biological age", "Chronological age")) %>% 
+  select(model, AIC, dAIC, weight) %>% 
+  knitr::kable()
 
@@
 The chronological age model outperformed the biological age model according to AIC.
-The difference in AIC between the two models was `r sprintf("%0.1f", -diff(chrono_bio_aic$AIC))` and the chronological age model had an AIC weight of `r sprintf("%0.3f", chrono_bio_aicw[2])`.
 

### 2024-07-29 paper/paperv3.qmd
Fixing figure captions

@@
 
-4.  Phenology (breeding haulout and molt haulout) is shifted later for older animals and breeders.
+4.  Annual cycle phenology - specifically, the duration of the molt haulout - will shift for older animals.
 
@@
 #| fig-width: 9
-#| fig-cap: "A: Raw longitudinal data for 1,202 known-age female elephant seals and their observations from birth, to recruitment, to death.
+#| fig-cap: "A large number of elephant seals from each cohort are tagged during their birth year (yellow) and observations take place for the rest of their lives so that breeding status (greens and blues) can be assigned and compared to age.
+A: Raw longitudinal data for 1,203 known-age female elephant seals and their observations from birth, to recruitment, to death.
 B: Histogram of the number of seals in each age class in the dataset."
@@
 #| label: fig-breeding-mod
-#| fig-cap: "Breeding rates increased until age 10, then decreased from age 11 on.
-Black points represent mean and standard error of the observed breeding rates.
-The solid lines and shaded areas represent the mean and 95% confidence interval of the population-level model predictions.
-The observed data are unequally distributed across years, so the dotted lines indicate the predicted breeding rates averaged over years (i.e., including the random effect of year), weighted by the number of seals observed in each year."
+#| fig-cap: "Breeding probability for adult female elephant seals increased up to age 11, and decreased above it, providing evidence for reproductive senescence.
+Black points and error bars show the mean and 95% CI of breeding rates.
+Sample sizes for each age class are included above the points.
+Thin gray lines show the mean response for each year of the study (i.e., including the random effect of year).
+The thick solid lines and shaded areas show the mean response and 95% confidence interval of the fitted model, weighted by the number of seals observed in each year.
+The unweighted fitted model is shown by the dotted gray line.
+The weighted model is emphasized because of the large impact of the random effect of year."
 
@@
 
-ggsave("outputs/repro_senesce.svg", breed_plot, width = 13, height = 6, units = "in")
+breed_plot
+
+#ggsave("outputs/repro_senesce.svg", breed_plot, width = 13, height = 6, units = "in")
 ```
@@
 #| fig-height: 8
+#| fig-cap: "Elephant seal pup survival (A) but not recruitment (B) decreased with maternal age above the threshold age of 11 years.
+Both males and females were included in the survival analysis, but only females were included for recruitment.
+This caused some values of recruitment to be higher than survival at the same maternal age.
+Black points and error bars show means and 95% CI of survival or recruitment for each maternal age.
+Sample sizes for each age class are included above the points.
+Thin gray lines show the mean response for each year of the study (i.e., including the random effect of year).
+Thick dashed and solid lines show the weighted mean response and 95% CI, with solid lines indicating significant trends.
+The unweighted fitted model is indicated by a dotted gray line.
+The weighted model is emphasized because of the large impact of the random effect of year."
 
@@
 
-We found that as maternal age increases, pup sex skews slightly towards males, but it is not significant (@fig-sex-ratio).
+We found that as maternal age increases, the pup sex ratio does not change (@fig-sex-ratio).
 
@@
 pup_sex_data <- sealdat %>% 
-  filter(pupsex %in% c("M", "F"), 
-         age < 19) %>% 
+  filter(pupsex %in% c("M", "F")) %>% 
   mutate(is_male = as.integer(pupsex == "M"))
@@
 #| label: fig-sex-ratio
-#| fig-cap: "Offspring sex ratio skews towards males with increasing maternal age, but the result is not significant (p = 0.6).
-Black points show the mean and solid black lines show the standard error of the observed sex ratio."
+#| fig-cap: "Offspring sex ratio trended towards more males with maternal age above the age threshold, but the trend was not significant.
+Black points and error bars show the mean and 95% CI.
+Sample sizes for each age class are included above the points.
+Thin gray lines show the mean response for each year of the study (i.e., including the random effect of year).
+Thick dashed lines show the weighted mean response and 95% CI; neither trend was significant.
+The unweighted fitted model is indicated by a dotted gray line.
+The weighted model is emphasized because of the large impact of the random effect of year."
 
@@
 sexrat_grid <- expand_grid(
-  age = 4:18,
+  age = 4:20,
   year_fct = factor(1988:2023)
@@
 
-**Note**: In the above figure, the population-level and weighted-average would lie directly on top of each other, so we've excluded the weighted average.
+**Note**: In the above figure, the population-level and weighted-average lie directly on top of each other.
 This is because the random effect of year has very little impact on offspring sex ratio, which makes sense biologically.
@@
 ```{r}
-#| label: phen-fig
+#| label: fig-phen
 #| fig-width: 8
+#| fig-cap: "Neither molting nor breeding haul out duration changed with old age.
+A: Raw data plotting the observed annual time allocation averaged over all n = 387 individuals.
+B & C: Points and error bars represent the mean and CI for haul out durations.
+Sample sizes for each age class are included above the points.
+Thin gray lines show the mean response for each year of the study (i.e., including the random effect of year).
+Thick solid and dashed lines represent the mean response and 95% CI of the fitted model, weighted by the number of seals observed per year, with solid lines indicating significant trends.
+The unweighted fitted model is indicated by a dotted gray line.
+The weighted model is emphasized because of the large impact of the random effect of year."
 
@@
 #| label: fig-mod-sens
-#| fig-cap: Model sensitivity to onset of senescence.
+#| fig-cap: "Model sensitivity to onset of senescence.
 Age at onset of senescence on the x-axis, p-value for the coefficient of the interaction (age category = post-senescence) x (response variable) on the y-axis.
@@
 Hypothesis H2a (maternal-effect senescence; offspring survival) was significantly supported for onset of senescence age 11+.
-At age 10, the p-vale for the coefficient of the interaction (age category = post-senescence) x (offspring survival) was 0.058.
+At age 10, the p-vale for the coefficient of the interaction (age category = post-senescence) x (offspring survival) was 0.058."
 
@@
 #| label: fig-selective
-#| fig-cap: The coefficient for age in the old age class in models accounting for selective appearance and disappearance.
+#| fig-cap: "The coefficient for age in the old age class in models accounting for selective appearance and disappearance.
 Red and blue indicate  reproductive senescence (H1) and maternal effect senescence (H2a), respectively.
-Compared to the base model, the appearance model includes a term for age at first reproduction, the disappearance model includes a term for age at last reproduction, and the combined model includes both.
+Compared to the base model, the appearance model includes a term for age at first reproduction, the disappearance model includes a term for age at last reproduction, and the combined model includes both."
 
@@
 #| label: fig-bio-age
-#| fig-cap: Breeding probability over biological age.
+#| fig-cap: "Breeding probability over biological age.
 Points and error bars are the mean ± SE of observed breeding proportions within age classes.
@@
 Solid lines and ribbons are the mean and 95% CI of the population-level breeding probability, estimated by a GLMM with random effects for individual and year.
-The dashed lines represent the mean of yearly breeding probabilities, weighted by the number of seals observed each year, which should match the observed data more closely than the population-level probabilities.
+The dashed lines represent the mean of yearly breeding probabilities, weighted by the number of seals observed each year, which should match the observed data more closely than the population-level probabilities."
 
@@
 
-This needs to be rethought now!
-
-The pattern of unobserved animals increasing with age means it is highly unlikely our inferences about reproductive senescence are due to partial detection.
-If there was no reproductive senescence in the population, then the apparent senescence we observed would have to result from older breeders being less detectable.
-However, the biology of northern elephant seals makes breeding a more detectable behavior than non-breeding.
-Animals must come to land for an extended period to give birth and wean their pups.
+The pattern of unobserved animals decreasing with age means it is highly unlikely our inferences about reproductive senescence are due to partial detection.
+
 Among the animals in this study, breeders hauled out for longer than non-breeders (1st - 3rd quartile: breeders `r breed_haulout` days; non-breeders `r non_breed_haulout` days).
@@
 Furthermore, northern elephant seals exhibit a high degree of site fidelity.
-Once they recruit to a breeding colony, temporary emigration is rare (cite me).
-The only plausible explanation for unobserved animals being disproportionately breeders would be crowding at the breeding colonies, which could obscure identifying marks.
-However, breeding animals were resighted `r breed_resight` times per year, compared to `r nonbreed_resight` for non-breeders (mean, 1st - 3rd quartile), which suggests that breeders are more readily observable than non-breeders.
-Therefore, it is biologically implausible that the unobserved seals were disproportionately breeders and our estimates for reproductive senescence are, if anything, conservative.
+Once they recruit to a breeding colony, temporary emigration is rare.
+Breeding animals were resighted `r breed_resight` times per year, compared to `r nonbreed_resight` for non-breeders (mean, 1st - 3rd quartile), which suggests that breeders are more readily observable than non-breeders.
 

### 2024-07-30 scratch/reviewer_asks.R
reviewer requests

@@ -1,0 +1,80 @@
+library(tidyverse)
+
+seals <- sealdat %>% 
+  select(animalID, year) %>% 
+  mutate(in_study = TRUE)
+
+foo <- read_csv(here::here("data/raw/fullresights.csv"), 
+                    show_col_types = FALSE) %>% 
+  filter(calyear > 1987, 
+         tagsex == "F", 
+         timeofyear == "Breeding") %>% 
+  mutate(animalID = as.factor(animalID), 
+         year = calyear) 
+
+longev <- foo %>% 
+  select(animalID, year, yearborn) %>%
+  filter(year < 2018) %>%
+  group_by(animalID) %>% 
+  mutate(longevity = max(year) - yearborn) %>% 
+  select(animalID, longevity) %>% 
+  unique() 
+
+longev %>% 
+  group_by(longevity) %>% 
+  count(longevity) %>% 
+  ggplot(aes(x = longevity, y = n)) + 
+  geom_col(fill = "grey50") + 
+  labs(x = "Longevity (Age in Years)", y = "Number of individuals") + 
+  theme_classic()
+
+mean(longev$longevity)
+
+foo2 <- left_join(foo, seals, by = c("animalID", "year")) %>% 
+  mutate(in_study = ifelse(is.na(in_study), yes = "FALSE", no = "TRUE")) %>% 
+  group_by(animalID, year) %>% 
+  mutate(n_obs = n_distinct(date))
+
+foo2 %>% filter(animalID == 10297) %>% 
+  # select(animalID, date, year, in_study, n_obs) %>% 
+    view()
+
+obs_by_breed_status <- foo2 %>% 
+  filter(!is.na(withpup)) %>% 
+  select(animalID, season, date, withpup, tagsex, yearborn, year, in_study, n_obs) %>% 
+  mutate(breed_status = if_else(withpup == 1, "Breeder", "Non-breeder")) %>% 
+  ungroup()
+
+n_distinct(obs_by_breed_status$animalID)
+#What proportion of seals were seen 4 times in a breeding season? 
+bred <- obs_by_breed_status %>% filter(breed_status == "Breeder")
+bred_used_df <- bred %>% filter(n_obs > 3) 
+#I think this part might be wrong? get max to check
+total_bred <- n_distinct(bred$animalID)
+bred_used <- n_distinct(bred_used$animalID)
+  
[…]

### 2024-08-12 paper/paperv3.qmd
updating with pop dynamics section

@@
 seal_years <- count(sealdat, year_fct)
+
 breed_pred_wgt <- breed_grid %>% 
@@
   left_join(n_breed, by = "age")
+
+breed_pred_wgt %>% filter(age > 9, 
+                          age <13) %>% 
+  summarize(mean_breed = mean(predicted))
+
+breed_pred_wgt %>% 
+  mutate(perc_pop = n/sum(n), 
+         pred_pups = n * predicted)
 
@@
   # In recruit model, only 17 individuals (4%) had >3 pups (female pups only)
-  pup_recr_int ~ age10 : age_cat + (1 | year_fct),
+  pup_recr_int ~ age10 : age_cat + (1 | year_fct) + (1 | animalID),
   filter(surv_recr_data, pupsex == "F"),
@@
                values_to = "dur_days") %>% 
-  glmer.nb(dur_days ~ age10:phase:age_cat + (1 | year_fct), 
-           data = .)
+  glmer(dur_days ~ age10:phase:age_cat + (1 | year_fct) + (1 | animalID), 
+        data = ., 
+        family = Gamma(link = "log"))
 
@@
 mean_yb_phen <- pheno_raw %>% 
-  filter(age < age_senesce, phase == "breeddur") %>% 
-  summarize(mean_yb = mean(dur_days), 
-            sd_yb = sd(dur_days))
+  filter(age == 4, phase == "breeddur") 
 mean_ob_phen <- pheno_raw %>% 
-  filter(age >= age_senesce, phase == "breeddur") %>% 
-  summarize(mean_ob = mean(dur_days), 
-            sd_ob = sd(dur_days))
+  filter(age == 11, phase == "breeddur")
 mean_ym_phen <- pheno_raw %>% 
-  filter(age < age_senesce, phase == "moltdur") %>% 
-  summarize(mean_ym = mean(dur_days), 
-            sd_ym = sd(dur_days))
+  filter(age == 4, phase == "moltdur")
 mean_om_phen <- pheno_raw %>% 
-  filter(age >= age_senesce, phase == "moltdur") %>% 
-  summarize(mean_om = mean(dur_days), 
-            sd_om = sd(dur_days))
-```
-
-Overall, the breeding haul out was `r mean_yb_phen$mean_yb` ± `r mean_yb_phen$sd_yb` days for animals younger than 11 years old and `r mean_ob_phen$mean_ob` ± `r mean_ob_phen$sd_ob` days for animals 11 years old and older.
-The molting haul out was `r mean_ym_phen$mean_ym` ± `r mean_ym_phen$sd_ym` days for animals younger than 11 years old and `r mean_om_phen$mean_om` ± `r mean_om_phen$sd_om` days for animals 11 years old and older.
+  filter(age == 11, phase == "moltdur") 
+```
+
+Overall, the breeding haul out was `r mean_yb_phen$dur_days` (`r mean_yb_phen$dur_days_lwr` - `r mean_yb_phen$dur_days_upr`) days for 4 year olds and `r mean_ob_phen$dur_days` (`r mean_ob_phen$dur_days_lwr` - `r mean_ob_phen$dur_days_upr`) days for 11 year old animals.
+The molting haul out was `r mean_ym_phen$dur_days` (`r mean_ym_phen$dur_days_lwr` - `r mean_ym_phen$dur_days_upr`) days for 4 year olds and `r mean_om_phen$dur_days` (`r mean_om_phen$dur_days_lwr` - `r mean_om_phen$dur_days_upr`) days for animals 11 years old and older.
 
@@
 ```
+
+## Population dynamics?
+
+What effect does senescence have on reproduction and offspring survival?
+First, calculate the total pup production of the age-structured population in one year.
+Then, let's pretend senescence *doesn't happen*.
+In this hypothetical population, all animals older than 11 years old reproduce like an 11 year old.
+How many more pups would that population produce?
+And how many more of those pups would survive to reproductive age?
+
+The observed pup production (in the age-structured population) is:
+
+$$
+\pi_{1a} = \frac{\sum_{a=4}^{20} b(a)n(a)}{\sum_{a=4}^{20} n(a)}
+$$
+
+Where $\pi_1$ is pup production per mother per year by the age-structured population. $a$ is age, $b$ is estimated breeding percentage, and $n$ is the number of mothers in that age class.
+
+Hypothetical pup production in a non-senescing population is:
+
+$$
+\pi_{2a} = \frac{\sum_{a=4}^{11} b(a) n(a) + \sum_{a=12}^{20} b(11) n(a)}{\sum_{a=4}^{20}n(a)}
+$$
+
+I.e., $b$ does not decline post-age 11.
+
+Similarly, the observed and hypothetical quantity of pups that survive to adulthood adds a term $s$ for offspring survival.
+
+$$
+\pi_{1b} = \frac{\sum_{a=4}^{20} b(a) n(a) s(a)}{\sum_{a=4}^{20} n(a)}
+$$
+
+$$
+\pi_{2b} = \frac{\sum_{a=4}^{11} b(a) n(a) s(a) + \sum_{a=12}^{20} b(11) n(a) (s11)}{\sum_{a=4}^{20}n(a)}
+$$
+
+```{r}
+#| label: popdy
+
+# a, b, n, and s 
+popdy_df <- count(sealdat, age) %>% 
+  filter(age <= 20) %>% 
+  mutate(age10 = (age - age_senesce) / 10,
+         age_cat = factor(age >= age_senesce, 
+                          labels = c("Young", "Old"))) %>% 
+  mutate(b = predict(breed_age_mod, newdata = ., re.form = NA, type = "response"),
+         s = predict(pup_surv_mod, newdata = ., re.form = NA, type = "response"))
+
+# pup production (pi1a and pi2a)
+pi1a <- with(popdy_df, sum(b * n) / sum(n))
+
+b11 <- popdy_df$b[popdy_df$age == 11]
+pi2a <- with(popdy_df, sum(ifelse(age > 11, b11, b) * n) / sum(n))
+
+# Survivng offspring (pi1b and pi2b)
+pi1b <- with(popdy_df, sum(b * n * s) / sum(n))
+
+s11 <- popdy_df$s[popdy_df$age == 11]
+pi2b <- with(popdy_df, sum(ifelse(age > 11, b11 * s11, b * s) * n) / sum(n))
+
+str_glue("
+pi1a: {round(pi1a, 5)}
+pi2a: {round(pi2a, 5)}
+pi1b: {round(pi1b, 5)}
+pi2b: {round(pi2b, 5)}")
+
+str_glue("
+delta pi a {round(100 * (pi1a - pi2a) / pi2a, 2)}%,
+delta pi b {round(100 * (pi1b - pi2b) / pi2b, 2)}%")
+
+```

### 2024-08-12 scratch/reviewer_asks.R

@@ -2,22 +2,24 @@ library(tidyverse)
 
 seals <- sealdat %>% 
   select(animalID, year) %>% 
-  mutate(in_study = TRUE)
+  mutate(in_study = TRUE,
+         animalID = factor(animalID))
 
 foo <- read_csv(here::here("data/raw/fullresights.csv"), 
-                    show_col_types = FALSE) %>% 
+                show_col_types = FALSE) %>% 
   filter(calyear > 1987, 
-         tagsex == "F", 
-         timeofyear == "Breeding") %>% 
-  mutate(animalID = as.factor(animalID), 
+         tagsex == "F") %>% 
+  mutate(animalID = factor(animalID), 
          year = calyear) 
 
 longev <- foo %>% 
   select(animalID, year, yearborn) %>%
   filter(year < 2018) %>%
   group_by(animalID) %>% 
-  mutate(longevity = max(year) - yearborn) %>% 
-  select(animalID, longevity) %>% 
+  mutate(lastyear = max(year), 
+         longevity = lastyear - yearborn) %>% 
+  filter(lastyear < 2018) %>% 
+  select(animalID, longevity, yearborn, lastyear) %>% 
   unique() 
 
 longev %>% 
@@ -29,21 +31,29 @@ longev %>%
   theme_classic()
 
 mean(longev$longevity)
-
+ 
 foo2 <- left_join(foo, seals, by = c("animalID", "year")) %>% 
-  mutate(in_study = ifelse(is.na(in_study), yes = "FALSE", no = "TRUE")) %>% 
-  group_by(animalID, year) %>% 
-  mutate(n_obs = n_distinct(date))
+  mutate(in_study = ifelse(is.na(in_study), yes = "FALSE", no = "TRUE"))
 
 foo2 %>% filter(animalID == 10297) %>% 
   # select(animalID, date, year, in_study, n_obs) %>% 
     view()
 
 obs_by_breed_status <- foo2 %>% 
-  filter(!is.na(withpup)) %>% 
-  select(animalID, season, date, withpup, tagsex, yearborn, year, in_study, n_obs) %>% 
-  mutate(breed_status = if_else(withpup == 1, "Breeder", "Non-breeder")) %>% 
-  ungroup()
+  filter(timeofyear == "Breeding") %>% 
+  group_by(animalID, year) %>% 
+  summarize(n_obs = n_distinct(yday),
[…]

### 2024-08-13 paper/paperv3.qmd
Nursing duration calculation

--- 2024-07-22 paper/paperv3.qmd
@@
   mutate(animalID = factor(animalID, levels = unique(animalID))) %>% 
-  filter(year > 1987) %>% 
   ggplot(aes(year, animalID)) +
@@
         legend.title = element_blank(),
-        legend.position = c(0.1, 0.95),
+        # legend.position.inside = c(0.01, 0.99),
         legend.justification = c(0, 1),
@@
 
-breed_plot <- ggplot(breed_pred_wgt, aes(age, predicted)) +
+ggplot(breed_pred_wgt, aes(age, predicted)) +
   # Predictions for individual years
@@
 
-ggsave("outputs/repro_senesce.svg", breed_plot, width = 13, height = 6, units = "in")
+ggsave("outputs/repro_senesce.svg", width = 13, height = 6, units = "in")
 ```
@@
 #| label: h1-pred-deets
-
-pnorm(summary(breed_age_mod)$coef["age10:age_catOld", "z value"])
 
@@
 summary(pup_recr_mod)$coefficients
-
-#survival 1 tailed
-pnorm(summary(pup_surv_mod)$coef["age10:age_catOld", "z value"])
-#recruit 1 tailed
-pnorm(summary(pup_recr_mod)$coef["age10:age_catOld", "z value"])
 
@@
   coord_cartesian(ylim = c(0, 0.6)) +
-  theme(legend.position = "none", 
-        axis.title.x = element_blank()) + 
-  geom_text(aes(label = n), y = 0.6, vjust = 1, nudge_x = 0.2)
+  labs(x = "Age (years)") +
+  theme(legend.position = "none") + 
+  geom_label(aes(label = n), y = 0.6, vjust = 1)
 
@@
   count(age)
-n_recruit_indiv <- surv_recr_data %>% 
-  filter(pupsex == "F") %>% 
-  group_by(animalID)
 
@@
   theme(legend.position = "none") + 
-  geom_text(aes(label = n), y = 0.6, vjust = 1, nudge_x = 0.2)
+  geom_label(aes(label = n), y = 0.6, vjust = 1)
 
@@
 Sample size for offspring survival was n = `r nrow(surv_recr_data)` observations of `r n_distinct(surv_recr_data$animalID)` individuals.
-Sample size for offspring recruitment was n = `r n_distinct(n_recruit_indiv)` observations of `r n_distinct(n_recruit_indiv$animalID)` female individuals.
+Sample size for offspring recruitment was n = `r nrow(n_recruit)` observations of `r n_distinct(n_recruit$animalID)` female individuals.
 
@@
 summary(sexrat_age_mod)$coefficients
-pnorm(summary(sexrat_age_mod)$coef["age10:age_catOld", "z value"])
 
@@
   scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
-  labs(x = "Maternal age (years)") +
+  labs(x = "Age (years)") +
   theme(legend.position = "none") +
@@
             alpha = 0.1) + 
-    # Population-level mean response
-  geom_line(aes(y = predicted_pop, group = age_cat),
-            linetype = "dotted", linewidth = 1.2, alpha = 0.5) +
   # Weighted model (CI + mean)
@@
                   filter(pheno_raw, phase == "breeddur"),
-                  position = position_dodge(width = 0.5), 
-                  size = 0.2) +
+                  position = position_dodge(width = 0.5)) +
+    # Population-level mean response
+  geom_line(aes(y = predicted_pop, group = age_cat),
+            linetype = "dotted", linewidth = 1.2, alpha = 0.5) +
   geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
-  geom_text(aes(label = n), y = 36.5, vjust = 0, size = 3) +
+  geom_text(aes(label = n), y = 36.5, vjust = 0) +
   scale_linetype_manual(values = c("solid", "dashed")) +
@@
   scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
-  scale_y_continuous(limits = c(15, 40)) +
-  labs(y = "Breeding haul out (days)") +
+  scale_y_continuous(limits = c(15, 37)) +
+  labs(x = "Maternal age (years)",
+       y = "Breeding haul out (days)") +
   theme(legend.position = "none",
-        axis.text.y = element_text(size = 6), 
-        axis.title.x = element_blank())
+        axis.text.y = element_text(size = 8))
 
@@
             alpha = 0.1) + 
-    # Population-level mean response
-   geom_line(aes(y = predicted_pop, group = age_cat),
-             linetype = "dotted", linewidth = 1.2, alpha = 0.2) +
   # Weighted model (CI + mean)
@@
                   filter(pheno_raw, phase == "moltdur"),
-                  position = position_dodge(width = 0.5), 
-                  size = 0.2) +
+                  position = position_dodge(width = 0.5)) +
+    # Population-level mean response
+  geom_line(aes(y = predicted_pop, group = age_cat),
+            linetype = "dotted", linewidth = 1.2, alpha = 0.5) +
   geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
-  geom_text(aes(label = n), y = 36.5, vjust = 0, size = 3) +
+  geom_text(aes(label = n), y = 36.5, vjust = 0) +
   scale_linetype_manual(values = c("solid", "dashed")) +
@@
   scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
-  scale_y_continuous(limits = c(15, 40)) +
+  scale_y_continuous(limits = c(15, 37)) +
   labs(x = "Maternal age (years)",
@@
   theme(legend.position = "none",
-        axis.text.y = element_text(size = 6))
-
-```
-
-```{r}
-
-mean_yb_phen <- pheno_raw %>% 
-  filter(age < age_senesce, phase == "breeddur") %>% 
-  summarize(mean_yb = mean(dur_days), 
-            sd_yb = sd(dur_days))
-mean_ob_phen <- pheno_raw %>% 
-  filter(age >= age_senesce, phase == "breeddur") %>% 
-  summarize(mean_ob = mean(dur_days), 
-            sd_ob = sd(dur_days))
-mean_ym_phen <- pheno_raw %>% 
-  filter(age < age_senesce, phase == "moltdur") %>% 
-  summarize(mean_ym = mean(dur_days), 
-            sd_ym = sd(dur_days))
-mean_om_phen <- pheno_raw %>% 
-  filter(age >= age_senesce, phase == "moltdur") %>% 
-  summarize(mean_om = mean(dur_days), 
-            sd_om = sd(dur_days))
-```
-
-Overall, the breeding haul out was `r mean_yb_phen$mean_yb` ± `r mean_yb_phen$sd_yb` days for animals younger than 11 years old and `r mean_ob_phen$mean_ob` ± `r mean_ob_phen$sd_ob` days for animals 11 years old and older.
-The molting haul out was `r mean_ym_phen$mean_ym` ± `r mean_ym_phen$sd_ym` days for animals younger than 11 years old and `r mean_om_phen$mean_om` ± `r mean_om_phen$sd_om` days for animals 11 years old and older.
-
-```{r}
-#| label: phen-fig
-#| fig-width: 8
+        axis.text.y = element_text(size = 8))
+
+```
+
+```{r}
 
@@
     geom_text(aes(label = "Breeding haul out"), 
-            x = 9,
+            x = 8,
             y = 30,
@@
     geom_text(aes(label = "Molting haul out"), 
-            x = 9,
+            x = 8,
             y = 135,
@@
 
-design <- "
-  12
-  13
-"
-
-annual_plot + breed_duration_plot + molt_duration_plot + 
-  plot_annotation(tag_levels = "A") + 
-  plot_layout(design = design, 
-              widths = c(1, 2)) 
+pheno_raw <- seal_phenology %>% 
+  select(breeddur, moltdur, age) %>% 
+  pivot_longer(c(breeddur, moltdur), 
+               names_to = "phase", 
+               values_to = "dur_days") %>% 
+  group_by(age, phase) %>% 
+  summarize(n = n(),
+            dur_days_se = sd(dur_days) / sqrt(n),
+            dur_days = mean(dur_days),
+            dur_days_lwr = dur_days - dur_days_se,
+            dur_days_upr = dur_days + dur_days_se,
+            .groups = "drop")
+
+phen_pred_years <- ggpredict(
+  pheno_model,
+  terms = c("age10 [all]", "phase", "year [all]"),
+  type = "random",
+  interval = "confidence"
+) %>% 
+  as_tibble() %>% 
+  mutate(age = x * 10,
+         phase = group)
+
+breed_pheno_raw <- pheno_raw[pheno_raw$phase == "breeddur",]
+molt_pheno_raw <- pheno_raw[pheno_raw$phase == "moltdur",]
+
+breed_duration_plot <- ggpredict(
+  pheno_model, 
+  terms = c("age10 [all]", "phase", "year [all]")
+) %>% 
+  as_tibble() %>% 
+  transmute(age = x * 10,
+            dur_days = predicted,
+            dur_days_lwr = conf.low,
+            dur_days_upr = conf.high,
+            phase = group,
+            year = as.numeric(as.character(facet))) %>% 
+  left_join(n_phenology, by = "year") %>% 
+  group_by(age) %>% 
+  filter(phase == "breeddur") %>% 
+  summarize(across(starts_with("dur_days"), \(x) weighted.mean(x, n)),
+            .groups = "drop") %>% 
+  ggplot(aes(age)) +
+  geom_line(aes(age, predicted, group = interaction(phase, facet)), 
+            filter(phen_pred_years, phase == "breeddur"), 
+            alpha = 0.1) + 
+  geom_ribbon(aes(ymin = dur_days_lwr, ymax = dur_days_upr),
+              alpha = 0.25, 
+              fill = "#D05794") +
+  geom_pointrange(aes(y = dur_days, 
+                      ymin = dur_days_lwr, 
+                      ymax = dur_days_upr),
+                  breed_pheno_raw,
+                  position = position_dodge(width = 0.5), 
+                  color = "#D05794") +
+  geom_line(aes(y = dur_days), 
+            linetype = "dashed", 
+            color = "#D05794") +
+  labs(x = "Maternal age (years)",
+       y = "Breeding haul out (days)") +
+  theme(legend.position = "none",
+        axis.text.y = element_text(size = 8))
+
+molt_duration_plot <- ggpredict(
+  pheno_model,
+  terms = c("age10 [all]", "phase", "year [all]")
+) %>%
+  as_tibble() %>%
+  transmute(age = x * 10,
+            dur_days = predicted,
+            dur_days_lwr = conf.low,
+            dur_days_upr = conf.high,
+            phase = group,
+            year = as.numeric(as.character(facet))) %>%
+  left_join(n_phenology, by = "year") %>%
+  group_by(age, phase) %>%
+  filter(phase == "moltdur") %>%
+  summarize(across(starts_with("dur_days"), \(x) weighted.mean(x, n)),
+            .groups = "drop") %>%
+  ggplot(aes(age)) +
+  geom_line(aes(age, predicted, group = interaction(phase, facet)), 
+            filter(phen_pred_years, phase == "moltdur"), 
+            alpha = 0.1) + 
+  geom_ribbon(aes(ymin = dur_days_lwr, ymax = dur_days_upr),
+              alpha = 0.25, 
+              fill = "#E09A2D") +
+  geom_pointrange(aes(y = dur_days, 
+                      ymin = dur_days_lwr, 
+                      ymax = dur_days_upr),
+                  molt_pheno_raw,
+                  position = position_dodge(width = 0.5), 
+                  color = "#E09A2D") +
+  geom_line(aes(y = dur_days), 
+            color = "#E09A2D") +
+  labs(x = "Maternal age (years)",
+       y = "Molting haul out (days)") +
+  theme(legend.position = "none", 
+        axis.text.y = element_text(size = 8))
+
+(annual_plot | phen_dist | breed_duration_plot | molt_duration_plot) + plot_annotation(tag_levels = "A") + plot_layout(ncol = 2, nrow = 2)
 ```
@@
 
-```
-
-```{r}
-#| label: chrono-bio-aic
-
-AIC(breed_bioage_mod, breed_age_mod2) %>% 
-  as_tibble() %>% 
-  mutate(dAIC = AIC - min(AIC),
-         rellik = exp(-0.5 * dAIC),
-         weight = rellik / sum(rellik),
-         model = c("Biological age", "Chronological age")) %>% 
-  select(model, AIC, dAIC, weight) %>% 
-  knitr::kable()
+chrono_bio_aic <- AIC(breed_bioage_mod, breed_age_mod2)
+daic <- chrono_bio_aic$AIC - min(chrono_bio_aic$AIC)
+chrono_bio_rellik <- exp(-0.5 * daic)
+chrono_bio_aicw <- chrono_bio_rellik / sum(chrono_bio_rellik)
 
@@
 The chronological age model outperformed the biological age model according to AIC.
+The difference in AIC between the two models was `r sprintf("%0.1f", -diff(chrono_bio_aic$AIC))` and the chronological age model had an AIC weight of `r sprintf("%0.3f", chrono_bio_aicw[2])`.
 
@@
 ```
+
+## Nursing duration
+
+What's the mean and 95% CI duration of nursing? [@reiter1981] provide the following information in Table III:
+
++-------+----------+----------+----------+----------+----------+----------+----------+
+| 2     | 3        | 4        | 5        | 6        | 7        | 8        | 9        |
++=======+==========+==========+==========+==========+==========+==========+==========+
+| 22±0  | 23.7±1.6 | 26.6±1.6 | 26.1±1.9 | 26.5±2.2 | 27.1±1.2 | 28.2±0.8 | 28.5±0.7 |
+|       |          |          |          |          |          |          |          |
+| \(1\) | \(9\)    | \(24\)   | \(12\)   | \(6\)    | \(8\)    | \(5\)    | \(2\)    |
++-------+----------+----------+----------+----------+----------+----------+----------+
+
+Where the column names are the age of the seal in years, and row values indicate mean±SD (n).
+To get the mean and 95% CI duration of nursing, we pooled the samples for mothers age 4 and up (i.e., excluded 2- and 3-year olds).
+
+The formulas we used for pooling the variances come from [@oneill2014].
+
+```{r}
+#| label: pooled-nursing
+
+# N, mean, and SD from Reiter et al. (1981) Table III for 4+ yo mothers
+n <- c(24, 12, 6, 8, 5, 2)
+m <- c(26.6, 26.1, 26.5, 27.1, 28.2, 28.5)
+s <- c(1.6, 1.9, 2.2, 1.2, 0.8, 0.7)
+
+# Function for pooling two samples
+pool_samples <- function(sample1, sample2) {
+  # sample1 and sample2 must be numeric vectors of length 3 with names v, m, and n
+  is_valid <- function(x) {
+    is.numeric(x) &&
+      is.vector(x) &&
+      (length(x) == 3) &&
+      all(c("v", "m", "n") %in% names(x))
+  }
+
+  if (!(is_valid(sample1) && is_valid(sample2))) 
+    stop("sample1 and sample2 must be numeric vectors of length 3 with names v, m, and n (variance, mean, and number of elements)")
+
+  v1 <- sample1["v"]
+  m1 <- sample1["m"]
+  n1 <- sample1["n"]
+  v2 <- sample2["v"]
+  m2 <- sample2["m"]
+  n2 <- sample2["n"]
+
+  n3 <- n1 + n2
+  m3 <- weighted.mean(c(m1, m2), c(n1, n2))
+  v3 <- 1 / (n1 + n2 - 1) * ((n1 - 1) * v1 + (n2 - 1) * v2 + n1 * n2 / (n1 + n2) * (m1 - m2)^2)
+  result <- c(v3, m3, n3)
+  names(result) <- c("v", "m", "n")
+  result
+}
+
+# Pool the samples iteratively using reduction
+samples <- pmap(list(n, m, s), \(.n, .m, .s) c(v = .s^2, m = .m, n = .n))
+pooled_samples <- reduce(samples, pool_samples)
+
+sprintf("%0.1f (95%%CI: %0.1f - %0.1f)", 
+        pooled_samples["m"],
+        pooled_samples["m"] - 1.96 * sqrt(pooled_samples["v"]),
+        pooled_samples["m"] + 1.96 * sqrt(pooled_samples["v"]))
+
+```

### 2024-08-13 paper/paperv3.qmd

--- 2024-08-12 paper/paperv3.qmd
@@
 
+## Nursing duration
+
+What's the mean and 95% CI duration of nursing? [@reiter1981] provide the following information in Table III:
+
++-------+----------+----------+----------+----------+----------+----------+----------+
+| 2     | 3        | 4        | 5        | 6        | 7        | 8        | 9        |
++=======+==========+==========+==========+==========+==========+==========+==========+
+| 22±0  | 23.7±1.6 | 26.6±1.6 | 26.1±1.9 | 26.5±2.2 | 27.1±1.2 | 28.2±0.8 | 28.5±0.7 |
+|       |          |          |          |          |          |          |          |
+| \(1\) | \(9\)    | \(24\)   | \(12\)   | \(6\)    | \(8\)    | \(5\)    | \(2\)    |
++-------+----------+----------+----------+----------+----------+----------+----------+
+
+Where the column names are the age of the seal in years, and row values indicate mean±SD (n).
+To get the mean and 95% CI duration of nursing, we pooled the samples for mothers age 4 and up (i.e., excluded 2- and 3-year olds).
+
+The formulas we used for pooling the variances come from [@oneill2014].
+
+```{r}
+#| label: pooled-nursing
+
+# N, mean, and SD from Reiter et al. (1981) Table III for 4+ yo mothers
+n <- c(24, 12, 6, 8, 5, 2)
+m <- c(26.6, 26.1, 26.5, 27.1, 28.2, 28.5)
+s <- c(1.6, 1.9, 2.2, 1.2, 0.8, 0.7)
+
+# Function for pooling two samples
+pool_samples <- function(sample1, sample2) {
+  # sample1 and sample2 must be numeric vectors of length 3 with names v, m, and n
+  is_valid <- function(x) {
+    is.numeric(x) &&
+      is.vector(x) &&
+      (length(x) == 3) &&
+      all(c("v", "m", "n") %in% names(x))
+  }
+
+  if (!(is_valid(sample1) && is_valid(sample2))) 
+    stop("sample1 and sample2 must be numeric vectors of length 3 with names v, m, and n (variance, mean, and number of elements)")
+
+  v1 <- sample1["v"]
+  m1 <- sample1["m"]
+  n1 <- sample1["n"]
+  v2 <- sample2["v"]
+  m2 <- sample2["m"]
+  n2 <- sample2["n"]
+
+  n3 <- n1 + n2
+  m3 <- weighted.mean(c(m1, m2), c(n1, n2))
+  v3 <- 1 / (n1 + n2 - 1) * ((n1 - 1) * v1 + (n2 - 1) * v2 + n1 * n2 / (n1 + n2) * (m1 - m2)^2)
+  result <- c(v3, m3, n3)
+  names(result) <- c("v", "m", "n")
+  result
+}
+
+# Pool the samples iteratively using reduction
+samples <- pmap(list(n, m, s), \(.n, .m, .s) c(v = .s^2, m = .m, n = .n))
+pooled_samples <- reduce(samples, pool_samples)
+
+sprintf("%0.1f (95%%CI: %0.1f - %0.1f)", 
+        pooled_samples["m"],
+        pooled_samples["m"] - 1.96 * sqrt(pooled_samples["v"]),
+        pooled_samples["m"] + 1.96 * sqrt(pooled_samples["v"]))
+```
+
 ## Population dynamics?
@@
 $$
-\pi_{1a} = \frac{\sum_{a=4}^{20} b(a)n(a)}{\sum_{a=4}^{20} n(a)}
+\pi_{1a} = \frac{\sum_{a=4}^{20} b_a n_a}{\sum_{a=4}^{20} n_a}
 $$
@@
 $$
-\pi_{2a} = \frac{\sum_{a=4}^{11} b(a) n(a) + \sum_{a=12}^{20} b(11) n(a)}{\sum_{a=4}^{20}n(a)}
+\pi_{2a} = \frac{\sum_{a=4}^{11} b_a n_a + \sum_{a=12}^{20} b_{11} n_a}{\sum_{a=4}^{20}n_a}
 $$
@@
 $$
-\pi_{1b} = \frac{\sum_{a=4}^{20} b(a) n(a) s(a)}{\sum_{a=4}^{20} n(a)}
+\pi_{1b} = \frac{\sum_{a=4}^{20} b_a n_a s_a}{\sum_{a=4}^{20} n_a}
 $$
@@
 $$
-\pi_{2b} = \frac{\sum_{a=4}^{11} b(a) n(a) s(a) + \sum_{a=12}^{20} b(11) n(a) (s11)}{\sum_{a=4}^{20}n(a)}
+\pi_{2b} = \frac{\sum_{a=4}^{11} b_a n_a s_a + \sum_{a=12}^{20} b_{11} n_a s_{11}}{\sum_{a=4}^{20} n_a}
 $$
@@
 delta pi b {round(100 * (pi1b - pi2b) / pi2b, 2)}%")
-
-```
+```

### 2024-08-16 paper/paperv3.qmd
Last few changes to supplemental plots

@@
 
-### Phenology
+### H4: Phenology
 
@@
 
-mean_yb_phen <- pheno_raw %>% 
-  filter(age == 4, phase == "breeddur") 
-mean_ob_phen <- pheno_raw %>% 
-  filter(age == 11, phase == "breeddur")
-mean_ym_phen <- pheno_raw %>% 
-  filter(age == 4, phase == "moltdur")
-mean_om_phen <- pheno_raw %>% 
-  filter(age == 11, phase == "moltdur") 
-```
-
-Overall, the breeding haul out was `r mean_yb_phen$dur_days` (`r mean_yb_phen$dur_days_lwr` - `r mean_yb_phen$dur_days_upr`) days for 4 year olds and `r mean_ob_phen$dur_days` (`r mean_ob_phen$dur_days_lwr` - `r mean_ob_phen$dur_days_upr`) days for 11 year old animals.
-The molting haul out was `r mean_ym_phen$dur_days` (`r mean_ym_phen$dur_days_lwr` - `r mean_ym_phen$dur_days_upr`) days for 4 year olds and `r mean_om_phen$dur_days` (`r mean_om_phen$dur_days_lwr` - `r mean_om_phen$dur_days_upr`) days for animals 11 years old and older.
+pheno_linkinv <- family(pheno_model)$linkinv
+expand_grid(
+  age = c(4, 10),
+  phase = c("moltdur", "breeddur")
+) %>% 
+  mutate(age_cat = factor(age >= age_senesce, 
+                          labels = c("Young")),
+         age10 = (age - age_senesce) / 10) %>% 
+  mutate(dur_days_eta = predict(pheno_model, newdata = ., re.form = NA),
+         dur_days_se = suppressWarnings(predict(pheno_model, 
+                                                newdata = ., 
+                                                re.form = NA, 
+                                                se.fit = TRUE)$se.fit),
+         dur_days = pheno_linkinv(dur_days_eta),
+         dur_days_lwr = pheno_linkinv(dur_days_eta - 1.96 * dur_days_se),
+         dur_days_upr = pheno_linkinv(dur_days_eta + 1.96 * dur_days_se)) %>% 
+  select(-age10, -dur_days_eta, -dur_days_se)
+
+```
 
@@
 
-ggplot(selective_coef, aes(model, mean, color = hypothesis)) + 
+ggplot(selective_coef, aes(hypothesis, mean, color = model)) + 
   geom_pointrange(aes(ymin = lwr, ymax = upr), position = position_dodge(width = 0.2)) +
   geom_hline(yintercept = 0, linetype = "dashed") +
-  scale_color_manual(values = c("firebrick", "cornflowerblue")) +
   expand_limits(y = 0) + 
   labs(y = "Coefficient", x = "Model") + 
-  theme(legend.position = "none")
+  theme(legend.title = element_blank())
 
@@
 
-## Observation frequency
+## Detection frequency
 
@@
 ```{r}
-#| label: missing-seals
-
-## Are we missing animals more or less with age?
-missing_seals <- sealdat %>% 
+#| label: seal-detections
+
+# Are seal detections changing with age?
+detection_seals <- sealdat %>% 
   filter(age <= 20) %>% 
@@
   group_by(animalID) %>% 
-  arrange(year) %>% 
-  reframe(tibble(year2 = (yearborn[1] + 4):max(year), # from age 4 to last obs
-                 unobserved = as.integer(!year2 %in% year),
+  # Keep only animals seen at least 2 years apart (i.e. if only seen in consecutive years remove them)
+  mutate(year_range = max(year) - min(year)) %>% 
+  filter(year_range > 2) %>% 
+  select(-year_range) %>% 
+  # Interpolate seen/not seen
+  reframe(tibble(year2 = (yearborn[1] + 4):(max(year) - 1), # from age 4 to year prior last obs
+                 observed = as.integer(year2 %in% year),
                  age = year2 - yearborn[1],
@@
 # Summary: mean and 95% CI 
-missing_summ <- missing_seals %>% 
+detection_summ <- detection_seals %>% 
   group_by(age, age_cat) %>% 
   summarize(n = n(),
-            n_unobserved = sum(unobserved),
-            unobserved_mean = n_unobserved / n,
-            unobserved_lwr = binom.test(n_unobserved, n)$conf.int[1],
-            unobserved_upr = binom.test(n_unobserved, n)$conf.int[2],
+            n_observed = sum(observed),
+            observed_mean = n_observed / n,
+            observed_lwr = binom.test(n_observed, n)$conf.int[1],
+            observed_upr = binom.test(n_observed, n)$conf.int[2],
             .groups = "drop")
 
-missing_mod <- glmer(
-  unobserved ~ age10 + (1 | year_fct) + (1 | animalID), 
-  missing_seals, 
+detection_mod <- glmer(
+  observed ~ age10:age_cat + (1 | year_fct) + (1 | animalID), 
+  detection_seals, 
   family = binomial
@@
 
-summary(missing_mod)
-
-```
-
-First, we tested whether probability of detection was related to age.
-We found the proportion of unobserved animals decreased with age (i.e., older seals are more likely to be observed).
+summary(detection_mod)
+
+```
+
+First, we tested whether detection probability was related to age.
+We found detection probalities decreased with age (i.e., older seals are less likely to be detected).
 For this analysis, we excluded the last year each animal was observed.
-Then, for all years between animals' age 4 year to the year prior to final observation, we determined whether each animal was observed.
-We estimated the effect of age on probability of an animal being unobserved using a GLMM (binomial response, logit link) with random effects for year and animal ([@fig-unobserved]).
-The coefficient for age was -0.2593 (95% CI: -0.4896 - -0.0290, *p* = 0.027).
-The change in mean unobserved rate from age 4 to 20 was 20.0% to 14.2%.
+Then, for all years between animals' age 4 year to the year prior to final observation, we determined whether each animal was detected.
+We estimated the effect of age on detection probability using a GLMM (binomial response, logit link) with an interaction between age and age category (young = age \< 11, old = age \> 11) and random effects for year and animal ([@fig-unobserved]).
+The coefficient for age was not significantly different than 0 for animals older than 11, but detection probability decreased significantly with age up to age 11.
+The change in detection probability from age 11 to 19 was 70.1% (61.7 - 77.3%) to 59.9% (44.4% - 73.7%).
 
@@
 #| label: fig-unobserved
-#| fig-cap: The probability of an animal being unobserved decreased with age.
-Points and error bars represent the proportion of unobserved animals with 95% CI.
-The blue line and shaded area are the mean and 95% CI probability of an animal being unobserved (GLMM, with random effect of year weighted by number of seals observed each year).
+#| fig-cap: Detection probabilities decreased with age for young, but not old, animals.
+Points and error bars represent the proportion of detected animals with 95% CI.
+The green and pink lines and shaded areas are the mean and 95% CI probability of an animal being detected (GLMM with random effect of year, weighted by number of seals observed each year), color-coded by age category.
 The dotted line is the mean response of the unweighted predictions.
-Numbers at the top indicate number of seals in each age category.
+Numbers at the top indicate number of seals of each age.
 See text for model details.
@@
 # Prediction grid
-missing_grid <- expand_grid(age = 4:20, 
+detection_grid <- expand_grid(age = 4:19, 
                               year_fct = factor(1989:2022)) %>% 
-  mutate(age10 = (age - age_senesce) / 10)
-
-# Missing seal predictions (weighted model)
-missing_pred <- missing_grid %>% 
-  mutate(predicted = predict(missing_mod, 
+  mutate(age10 = (age - age_senesce) / 10,
+         age_cat = factor(age >= age_senesce, 
+                          labels = c("Young", "Old")))
+
+# Seal detection predictions (weighted model)
+detection_pred <- detection_grid %>% 
+  mutate(predicted = predict(detection_mod, 
                              newdata = .,
                              re.form = ~ (1 | year_fct)),
-         pred_pop = predict(missing_mod, 
+         pred_pop = predict(detection_mod, 
                             newdata = .,
@@
                             re.form = NA),
-         se = suppressWarnings(predict(missing_mod, 
+         se = suppressWarnings(predict(detection_mod, 
                                        newdata = .,
@@
          conf.high = predicted + 1.96 * se) %>% 
-  left_join(count(missing_seals, year_fct), by = "year_fct") %>% 
-  group_by(age) %>% 
-  summarize(across(c(predicted, pred_pop, conf.low, conf.high), mean)) %>% 
+  left_join(count(detection_seals, year_fct), by = "year_fct") %>% 
+  group_by(age, age_cat) %>% 
+  summarize(across(c(predicted, pred_pop, conf.low, conf.high), mean), .groups = "drop") %>% 
   mutate(across(c(predicted, conf.low, conf.high), 
-                family(missing_mod)$linkinv))
+                family(detection_mod)$linkinv))
 
 # Yearly predictions
-missing_pred_years <- missing_grid %>% 
-  mutate(predicted = predict(missing_mod, 
+detection_pred_years <- detection_grid %>% 
+  mutate(predicted = predict(detection_mod, 
                              newdata = .,
@@
 
-ggplot(missing_pred, aes(age)) +
+ggplot(detection_pred, aes(age, color = age_cat, fill = age_cat)) +
   # Line per year
-  geom_line(aes(y = predicted, group = year_fct), 
-            missing_pred_years, 
+  geom_line(aes(y = predicted, group = interaction(age_cat, year_fct)), 
+            detection_pred_years, 
             alpha = 0.2) +
@@
               alpha = 0.3, 
-              fill = "cornflowerblue") +
-  geom_line(aes(y = predicted),
-            linewidth = 2,
-            color = "cornflowerblue") +
+              color = NA) +
+  geom_line(aes(y = predicted, linetype = age_cat),
+            linewidth = 1.5) +
   # Population-level model
@@
             linetype = "dotted",
-            linewidth = 1.2,
-            color = "cornflowerblue") +
+            linewidth = 1.1) +
   # Raw data
-  geom_pointrange(aes(y = unobserved_mean, 
-                      ymin = unobserved_lwr, 
-                      ymax = unobserved_upr),
-                  missing_summ) +
+  geom_pointrange(aes(y = observed_mean, 
+                      ymin = observed_lwr, 
+                      ymax = observed_upr),
+                  detection_summ) +
   # Sample size
-  geom_text(aes(age, label = n), missing_summ, y = 0.8, vjust = 1.1) +
-  scale_x_continuous("Age", breaks = seq(4, 20, by = 4)) +
-  scale_y_continuous("Unobserved proportion", 
+  geom_text(aes(age, label = n), detection_summ, y = 0, vjust = 0) +
+  scale_x_continuous("Age", breaks = seq(4, 19, by = 4)) +
+  scale_y_continuous("Detection rate", 
                      labels = scales::label_percent(accuracy = 1),
                      limits = c(0, NA)) +
+  scale_linetype_manual(values = c("solid", "dashed")) +
+  scale_color_manual(values = c("#7fbc41", "#de77ae")) +
+  scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
   theme(legend.position = "none")
@@
 
-## Population dynamics?
+## Population dynamics
 

### 2024-10-09 paper/paperv3.qmd
Updating model sensitivity chunk to all be 2-tailed tests instead of 1-tailed

@@
     h2_recr = summary(pup_recr_mod)$coefficients,
-    h3 = summary(sexrat_age_mod)$coefficients
+    h3 = summary(sexrat_age_mod)$coefficients,
+    h1_aic = AIC(breed_age_mod),
+    h2_surv_aic = AIC(pup_surv_mod),
+    h2_recr_aic = AIC(pup_recr_mod),
+    h3_aic = AIC(sexrat_age_mod)
   )
@@
 # Fit models with cut-off at ages 8-13
-cutoff <- 8:13
+cutoff <- 6:14
 mod_sens <- map(cutoff, fit_models)
@@
     map_dbl(mod_sens, \(m) 2 * pnorm(m[[h]]["age10:age_catOld", "z value"],
-                                     lower.tail = FALSE))
+                                     lower.tail = (m[[h]]["age10:age_catOld", "z value"] < 0)
+                                     ))
   } else {
@@
 }
+extract_aic <- function(h) {
+  map_dbl(mod_sens, paste(h, "aic", sep = "_"))
+}
 p_tbl <- tibble(age_cutoff = cutoff, 
-                "H1 (breeding)" = extract_p("h1"),
-                "H2a (survival)" = extract_p("h2_surv"),
-                "H2b (recruitment)" = extract_p("h2_recr"),
-                "H3 (sex ratio)" = extract_p("h3"),
+                "H1 (breeding)" = extract_p("h1", tails = 2),
+                "H2a (survival)" = extract_p("h2_surv", tails = 2),
+                "H2b (recruitment)" = extract_p("h2_recr", tails = 2),
                 "H3 (sex ratio; 2 tail)" = extract_p("h3", tails = 2))
+aic_weight <- function(x) {
+  rellik <- exp(-0.5 * (x - min(x)))
+  rellik / sum(rellik)
+}
+aic_tbl <- tibble(age_cutoff = cutoff,
+                  "H1 (breeding)" = extract_aic("h1"),
+                  "H2a (survival)" = extract_aic("h2_surv"),
+                  "H2b (recruitment)" = extract_aic("h2_recr"),
+                  "H3 (sex ratio)" = extract_aic("h3")) %>% 
+  mutate(across(-age_cutoff, aic_weight))
 
 knitr::kable(p_tbl)
+knitr::kable(aic_tbl)
 

### 2024-10-10 paper/paperv3.qmd
Fixing an error in one of the models (singular fit); adding phenology to the detailed model outputs.

@@
   geom_raster(aes(fill = observed)) +
-  scale_fill_viridis_d(direction = -1) +
+  scale_fill_brewer(palette = "YlGnBu", direction = 1) +
   guides(fill = guide_legend(position = "inside", 
@@
   # In recruit model, only 17 individuals (4%) had >3 pups (female pups only)
-  pup_recr_int ~ age10 : age_cat + (1 | year_fct) + (1 | animalID),
+  pup_recr_int ~ age10 : age_cat + (1 | year_fct),
   filter(surv_recr_data, pupsex == "F"),
@@
         family = Gamma(link = "log"))
-
-summary(pheno_model)
 
@@
 
+### H4: Phenology
+
+```{r}
+#| label: detailed-h4
+
+summary(pheno_model)
+
+```
+
 ## Models' sensitivity to age of senescence
@@
 # Are hypotheses supported by the model?
+note to future readers...this part could have been done way easier, and you will have to trust our process that this was the path of least resistence for us at the point of final revisions!
 extract_p <- function(h, tails = 1) {

### 2024-10-10 paper/paperv3.qmd
Adding a comparison to a linear model

@@
                 "H2b (recruitment)" = extract_p("h2_recr", tails = 2),
-                "H3 (sex ratio; 2 tail)" = extract_p("h3", tails = 2))
+                "H3 (sex ratio)" = extract_p("h3", tails = 2))
 aic_weight <- function(x) {
@@
 
-## Breakpoint vs quadratic
+## Breakpoint vs quadratic vs linear
 
 ### H1: Reproductive senescence
+
+```{r}
+#| label: linear-h1
+
+breed_age_line <- lmer(
+  observed_int ~ age10 + (1 | animalID) + (1 | year),
+  sealdat
+)
+
+summary(breed_age_line)$coefficients
+```
 
@@
 ### H2: Maternal effect senescence
+
+```{r}
+#| label: linear-h2
+
+#h2a
+surv_age_line <- lmer(
+  pup_surv_int ~ age10 + (1 | animalID) + (1 | year),
+  surv_recr_data
+)
+
+summary(surv_age_line)$coefficients
+
+#h2b
+recr_age_line <- lmer(
+  pup_recr_int ~ age10 + (1 | year),
+  filter(surv_recr_data, pupsex == "F")
+)
+
+summary(recr_age_line)$coefficients
+```
 
@@
 
-aic <- function(break_mod, quad_mod, hyp) {
-  AIC(break_mod, quad_mod) %>% 
+aic <- function(break_mod, quad_mod, line_mod, hyp) {
+  AIC(break_mod, quad_mod, line_mod) %>% 
     as_tibble() %>% 
@@
          hypothesis = hyp,
-         model = c("Breakpoint", "Quadratic")) %>% 
+         model = c("Breakpoint", "Quadratic", "Linear")) %>% 
   select(hypothesis, model, AIC, weight)
@@
           quad_mod = list(breed_age_quad, surv_age_quad, recr_age_quad),
+          line_mod = list(breed_age_line, surv_age_line, recr_age_line), 
           hyp = c("H1", "H2a", "H2b")),

### 2024-10-10 paper/paperv3.qmd
Adding a linear model for detection to compare to the breakpoint
adding library for akaike weights

@@
 library(patchwork)
+> library(qpcR)
 source(here::here("R/format.R"))
@@
 
+detection_mod_line <- lmer(
+  observed ~ age10 + (1 | year_fct) + (1 | animalID), 
+  detection_seals
+)
+
 summary(detection_mod)
+summary(detection_mod_line)
+AIC(detection_mod, detection_mod_line)
+akaike.weights(AIC(detection_mod, detection_mod_line))
 

### 2024-10-13 paper/paperv3.qmd
Fixing lmer -> glmer

@@
 library(patchwork)
-> library(qpcR)
+library(qpcR)
 source(here::here("R/format.R"))
@@
 
-detection_mod_line <- lmer(
+detection_mod_line <- glmer(
   observed ~ age10 + (1 | year_fct) + (1 | animalID), 
-  detection_seals
+  detection_seals, 
+  family = "binomial"
 )
@@
 
-breed_age_line <- lmer(
+breed_age_line <- glmer(
   observed_int ~ age10 + (1 | animalID) + (1 | year),
-  sealdat
+  sealdat, 
+  family = "binomial", 
+  control = glmerControl(optimizer = "bobyqa")
 )

### 2024-10-14 paper/paperv3.qmd
fixing more lmer > glmer errors

@@
 #h2a
-surv_age_line <- lmer(
+surv_age_line <- glmer(
   pup_surv_int ~ age10 + (1 | animalID) + (1 | year),
-  surv_recr_data
+  surv_recr_data, 
+  family = "binomial",
+  control = glmerControl(optimizer = "bobyqa")
 )
@@
 #h2b
-recr_age_line <- lmer(
+recr_age_line <- glmer(
   pup_recr_int ~ age10 + (1 | year),
-  filter(surv_recr_data, pupsex == "F")
+  filter(surv_recr_data, pupsex == "F"), 
+  family = "binomial",
+  control = glmerControl(optimizer = "bobyqa")
 )

### 2024-10-15 paper/paperv3.qmd
Updating the detection probability figure

@@
   geom_raster(aes(fill = observed)) +
-  scale_fill_brewer(palette = "YlGnBu", direction = 1) +
+  scale_fill_viridis_d(direction = 1) +
   guides(fill = guide_legend(position = "inside", 
@@
 
-ggplot(detection_pred, aes(age, color = age_cat, fill = age_cat)) +
+ggplot(detection_pred, aes(age)) +
   # Line per year
-  geom_line(aes(y = predicted, group = interaction(age_cat, year_fct)), 
+  geom_line(aes(y = predicted, group = interaction(age, year_fct)), 
             detection_pred_years, 
@@
               color = NA) +
-  geom_line(aes(y = predicted, linetype = age_cat),
+  geom_line(aes(y = predicted),
             linewidth = 1.5) +
@@
                      limits = c(0, NA)) +
-  scale_linetype_manual(values = c("solid", "dashed")) +
-  scale_color_manual(values = c("#7fbc41", "#de77ae")) +
-  scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
+  # scale_linetype_manual(values = c("solid", "dashed")) +
+  # scale_color_manual(values = c("#7fbc41", "#de77ae")) +
+  # scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
   theme(legend.position = "none")

### 2024-10-15 paper/paperv3.qmd
Updating pop dynamics equation notation

@@
 $$
-\pi_{1a} = \frac{\sum_{a=4}^{20} b_a n_a}{\sum_{a=4}^{20} n_a}
+F_{1} = \frac{\sum_{a=4}^{20} b_a n_a}{\sum_{a=4}^{20} n_a}
 $$
 
-Where $\pi_1$ is pup production per mother per year by the age-structured population. $a$ is age, $b$ is estimated breeding percentage, and $n$ is the number of mothers in that age class.
+Where $F$ represents fertility, described by pup production per mother per year by the age-structured population.
+$a$ is age, $b$ is estimated breeding percentage, and $n$ is the number of mothers in that age class.
 
@@
 $$
-\pi_{2a} = \frac{\sum_{a=4}^{11} b_a n_a + \sum_{a=12}^{20} b_{11} n_a}{\sum_{a=4}^{20}n_a}
+F_{2} = \frac{\sum_{a=4}^{11} b_a n_a + \sum_{a=12}^{20} b_{11} n_a}{\sum_{a=4}^{20}n_a}
 $$
@@
 
-Similarly, the observed and hypothetical quantity of pups that survive to adulthood adds a term $s$ for offspring survival.
+Similarly, to account for the influence of maternal effects ($M$), we add a term $s$ for offspring survival to estimate the observed and hypothetical quantity of pups that survive to adulthood.
 
 $$
-\pi_{1b} = \frac{\sum_{a=4}^{20} b_a n_a s_a}{\sum_{a=4}^{20} n_a}
+M_{1} = \frac{\sum_{a=4}^{20} b_a n_a s_a}{\sum_{a=4}^{20} n_a}
 $$
@@
 $$
-\pi_{2b} = \frac{\sum_{a=4}^{11} b_a n_a s_a + \sum_{a=12}^{20} b_{11} n_a s_{11}}{\sum_{a=4}^{20} n_a}
+M_{2} = \frac{\sum_{a=4}^{11} b_a n_a s_a + \sum_{a=12}^{20} b_{11} n_a s_{11}}{\sum_{a=4}^{20} n_a}
 $$

### 2024-10-16 paper/paperv3.qmd
comparison to linear for h3

@@
 
+### H3: Offspring Sex Ratio
+
+```{r}
+#| label: linear-h3
+
+#h2a
+sexrat_line <- glmer(
+  is_male ~ age10 + (1 | animalID) + (1 | year),
+  pup_sex_data, 
+  family = "binomial",
+  control = glmerControl(optimizer = "bobyqa")
+)
+
+summary(sexrat_line)$coefficients
+
+```
+
+```{r}
+#| label: quad-h3
+
+sexrat_quad <- glmer(
+  is_male ~ poly(age10, 2, raw = TRUE) + (1 | animalID) + (1 | year_fct),
+  pup_sex_data,
+  family = "binomial",
+  control = glmerControl(optimizer = "bobyqa")
+)
+
+h3_age10_peak <- -summary(sexrat_quad)$coefficients[2, 1] / 
+  (2 * summary(sexrat_quad)$coefficients[3, 1])
+h3_age_peak <- h3_age10_peak * 10 + age_senesce
+
+summary(surv_age_quad)$coefficients
+
+sprintf("Age peak (H2a) = %0.1f", h2a_age_peak)
+```
+
 ### AIC Comparison
@@
 str_glue("
-pi1a: {round(pi1a, 5)}
-pi2a: {round(pi2a, 5)}
-pi1b: {round(pi1b, 5)}
-pi2b: {round(pi2b, 5)}")
+P1: {round(pi1a, 5)}
+P2: {round(pi2a, 5)}
+M1: {round(pi1b, 5)}
+M2: {round(pi2b, 5)}")
 
 str_glue("
-delta pi a {round(100 * (pi1a - pi2a) / pi2a, 2)}%,
-delta pi b {round(100 * (pi1b - pi2b) / pi2b, 2)}%")
-```
+delta P {round(100 * (pi1a - pi2a) / pi2a, 2)}%,
+delta M {round(100 * (pi1b - pi2b) / pi2b, 2)}%")
+```

### 2024-10-17 paper/paperv3.qmd
Fixing figure S1

@@
 fit_models <- function(a) {
-  age_senesce <- a
   sealdat <- sealdat %>% 
-    mutate(age_cat = factor(age >= age_senesce, 
-                            labels = c("Young", "Old")))
+    mutate(age_cat = factor(age >= a, 
+                            labels = c("Young", "Old")),
+           age10 = (age - a) / 10)
 
@@
   )
+
+  effect_size <- function(m) {
+    probs <- tibble(age = c(15, 19),
+                    age_cat = c("Old", "Old")) %>% 
+      mutate(age10 = (age - a) / 10) %>% 
+      predict(m, re.form = NA, newdata = ., type = "response")
+    odds <- probs / (1 - probs)
+    odds[2] / odds[1]
+  }
 
@@
     h2_recr_aic = AIC(pup_recr_mod),
-    h3_aic = AIC(sexrat_age_mod)
+    h3_aic = AIC(sexrat_age_mod),
+    h1_effect = effect_size(breed_age_mod),
+    h2_surv_effect = effect_size(pup_surv_mod),
+    h2_recr_effect = effect_size(pup_recr_mod),
+    h3_effect = effect_size(sexrat_age_mod)
   )
@@
 
+extract_coef <- function(h) {
+  map_dbl(mod_sens, \(m) m[[h]]["age10:age_catOld", "Estimate"])
+}
+coef_tbl <- tibble(age_cutoff = cutoff, 
+                "H1 (breeding)" = extract_coef("h1"),
+                "H2a (survival)" = extract_coef("h2_surv"),
+                "H2b (recruitment)" = extract_coef("h2_recr"),
+                "H3 (sex ratio)" = extract_coef("h3"))
+
 knitr::kable(p_tbl)
 knitr::kable(aic_tbl)
+knitr::kable(coef_tbl)
 
@@
 
-p_tbl %>% 
-  pivot_longer(-age_cutoff, names_to = "hypothesis", values_to = "p") %>%
-  mutate(hypgrp = ifelse(str_detect(hypothesis, "H3"), 
-                         "H3 (sex ratio)", 
-                         hypothesis)) %>% 
-  ggplot(aes(age_cutoff, p)) +
-  geom_point(aes(color = hypothesis == "H3 (sex ratio; 2 tail)")) +
-  geom_hline(yintercept = 0.05, linetype = "dashed") +
-  geom_text(aes(x, y, label = tag),
-            tibble(hypgrp = colnames(p_tbl)[2:5],
-                   tag = LETTERS[1:4],
-                   x = 8, y = 3),
-            vjust = 1,
-            fontface = "bold") +
-  scale_y_log10(labels = scales::number) +
-  scale_color_manual(values = c("black", "grey50")) +
+left_join(
+  pivot_longer(p_tbl, -age_cutoff, names_to = "hypothesis", values_to = "p"),
+  pivot_longer(coef_tbl, -age_cutoff, names_to = "hypothesis", values_to = "coef"),
+  by = c("age_cutoff", "hypothesis")
+  ) %>% 
+  left_join(
+    pivot_longer(se_tbl, -age_cutoff, names_to = "hypothesis", values_to = "se"),
+    by = c("age_cutoff", "hypothesis")
+  ) %>% 
+  ggplot(aes(age_cutoff, coef)) +
+  geom_hline(yintercept = 0, linetype = "dashed") +
+  geom_pointrange(aes(size = p, 
+                      color = p < 0.05, 
+                      ymin = coef - 1.96 * se,
+                      ymax = coef + 1.96 * se)) +
   labs(x = "Senescence onset (years)") +
+  scale_size_continuous(range = c(0.25, 1)) +
   expand_limits(y = 3) +
-  facet_wrap(~hypgrp) +
+  facet_wrap(~hypothesis) +
   theme(legend.position = "none")

### 2024-10-17 paper/paperv3.qmd
updating model sensitivity to threshold analyses with h4 phenology

@@
 
-  effect_size <- function(m) {
-    probs <- tibble(age = c(15, 19),
-                    age_cat = c("Old", "Old")) %>% 
-      mutate(age10 = (age - a) / 10) %>% 
-      predict(m, re.form = NA, newdata = ., type = "response")
-    odds <- probs / (1 - probs)
-    odds[2] / odds[1]
-  }
+  # H4
+molt_data <- sealdat %>% 
+  filter(year >= 2011, 
+         moltdur >= 7, 
+         age <= 20) %>% 
+  mutate(BA = firstobsbreeddoy, 
+         BD = firstobsbreeddoy + breeddur, 
+         MA = lastobsmoltdoy - moltdur, 
+         MD = lastobsmoltdoy) %>% 
+  drop_na(age, BA, BD, MA, MD) %>% 
+  transmute(animalID, 
+            year_fct, 
+            age10 = (age - age_senesce) / 10, 
+            age_cat = factor(age >= age_senesce, 
+                          labels = c("Young", "Old")),
+            moltdur, 
+            breeddur) %>% 
+  pivot_longer(moltdur:breeddur, 
+               names_to = "phase", 
+               values_to = "dur_days") %>% 
+  filter(phase == "moltdur")
+
+molt_mod <- glmer(dur_days ~ age10 : age_cat + (1 | year_fct) + (1 | animalID), 
+        molt_data, 
+        family = Gamma(link = "log"))
+
+breed_data <- sealdat %>% 
+  filter(year >= 2011, 
+         moltdur >= 7, 
+         age <= 20) %>% 
+  mutate(BA = firstobsbreeddoy, 
+         BD = firstobsbreeddoy + breeddur, 
+         MA = lastobsmoltdoy - moltdur, 
+         MD = lastobsmoltdoy) %>% 
+  drop_na(age, BA, BD, MA, MD) %>% 
+  transmute(animalID, 
+            year_fct, 
+            age10 = (age - age_senesce) / 10, 
+            age_cat = factor(age >= age_senesce, 
+                          labels = c("Young", "Old")),
+            moltdur, 
+            breeddur) %>% 
+  pivot_longer(moltdur:breeddur, 
+               names_to = "phase", 
+               values_to = "dur_days") %>% 
+  filter(phase == "breeddur")
+
+breed_mod <- glmer(dur_days ~ age10 : age_cat + (1 | year_fct) + (1 | animalID), 
+        data = breed_data, 
+        family = Gamma(link = "log"), 
+        control = glmerControl(optimizer = "bobyqa")
+)
+
+#Making the tables
+#Not sure about this function
+  # effect_size <- function(m) {
+  #   probs <- tibble(age = c(15, 19),
+  #                   age_cat = c("Old", "Old")) %>% 
+  #     mutate(age10 = (age - a) / 10) %>% 
+  #     predict(m, re.form = NA, newdata = ., type = "response")
+  #   odds <- probs / (1 - probs)
+  #   odds[2] / odds[1]
+  # }
 
@@
     h3 = summary(sexrat_age_mod)$coefficients,
+    h4_molt = summary(molt_mod)$coefficients, 
+    h4_breed = summary(breed_mod)$coefficients,
     h1_aic = AIC(breed_age_mod),
@@
     h3_aic = AIC(sexrat_age_mod),
-    h1_effect = effect_size(breed_age_mod),
-    h2_surv_effect = effect_size(pup_surv_mod),
-    h2_recr_effect = effect_size(pup_recr_mod),
-    h3_effect = effect_size(sexrat_age_mod)
+    h4_molt_aic = AIC(molt_mod), 
+    h4_breed_aic = AIC(breed_mod)
+    # h1_effect = effect_size(breed_age_mod),
+    # h2_surv_effect = effect_size(pup_surv_mod),
+    # h2_recr_effect = effect_size(pup_recr_mod),
+    # h3_effect = effect_size(sexrat_age_mod)
   )
@@
 
+```
+
+```{r}
 # Are hypotheses supported by the model?
 note to future readers...this part could have been done way easier, and you will have to trust our process that this was the path of least resistence for us at the point of final revisions!
-extract_p <- function(h, tails = 1) {
-  if (tails == 1) {
-    map_dbl(mod_sens, \(m) pnorm(m[[h]]["age10:age_catOld", "z value"]))
-  } else if (tails == 2) {
-    map_dbl(mod_sens, \(m) 2 * pnorm(m[[h]]["age10:age_catOld", "z value"],
-                                     lower.tail = (m[[h]]["age10:age_catOld", "z value"] < 0)
-                                     ))
-  } else {
-    stop("tails must be 1 or 2")
-  }
+# extract_p <- function(h, tails = 1) {
+#   if (tails == 1) {
+#     map_dbl(mod_sens, \(m) pnorm(m[[h]]["age10:age_catOld", "z value"]))
+#   } else if (tails == 2) {
+#     map_dbl(mod_sens, \(m) 2 * pnorm(m[[h]]["age10:age_catOld", "z value"],
+#                                      lower.tail = (m[[h]]["age10:age_catOld", "z value"] < 0)
+#                                      ))
+#   } else {
+#     stop("tails must be 1 or 2")
+#   }
+# }
+extract_p <- function(h) {
+  map_dbl(mod_sens, \(m) m[[h]]["age10:age_catOld", "Pr(>|z|)"])
 }
@@
 }
-p_tbl <- tibble(age_cutoff = cutoff, 
-                "H1 (breeding)" = extract_p("h1", tails = 2),
-                "H2a (survival)" = extract_p("h2_surv", tails = 2),
-                "H2b (recruitment)" = extract_p("h2_recr", tails = 2),
-                "H3 (sex ratio)" = extract_p("h3", tails = 2))
 aic_weight <- function(x) {
@@
 }
+extract_coef <- function(h) {
+  map_dbl(mod_sens, \(m) m[[h]]["age10:age_catOld", "Estimate"])
+}
+extract_se <- function(h) {
+  map_dbl(mod_sens, \(m) m[[h]]["age10:age_catOld", "Std. Error"])
+}
+
+#make the tables
+p_tbl <- tibble(age_cutoff = cutoff, 
+                "H1 (breeding)" = extract_p("h1"),
+                "H2a (survival)" = extract_p("h2_surv"),
+                "H2b (recruitment)" = extract_p("h2_recr"),
+                "H3 (sex ratio)" = extract_p("h3"), 
+                "H4a (molt duration)" = extract_p("h4_molt"), 
+                "H4b (breed duration)" = extract_p("h4_breed"))
+
 aic_tbl <- tibble(age_cutoff = cutoff,
@@
                   "H2b (recruitment)" = extract_aic("h2_recr"),
-                  "H3 (sex ratio)" = extract_aic("h3")) %>% 
+                  "H3 (sex ratio)" = extract_aic("h3"), 
+                  "H4a (molt duration)" = extract_aic("h4_molt"), 
+                  "H4b (breed duration)" = extract_aic("h4_breed")) %>% 
   mutate(across(-age_cutoff, aic_weight))
 
-extract_coef <- function(h) {
-  map_dbl(mod_sens, \(m) m[[h]]["age10:age_catOld", "Estimate"])
-}
 coef_tbl <- tibble(age_cutoff = cutoff, 
@@
                 "H2b (recruitment)" = extract_coef("h2_recr"),
-                "H3 (sex ratio)" = extract_coef("h3"))
+                "H3 (sex ratio)" = extract_coef("h3"), 
+                "H4a (molt duration)" = extract_coef("h4_molt"), 
+                "H4b (breed duration)" = extract_coef("h4_breed"))
+
+se_tbl <- tibble(age_cutoff = cutoff, 
+                "H1 (breeding)" = extract_se("h1"),
+                "H2a (survival)" = extract_se("h2_surv"),
+                "H2b (recruitment)" = extract_se("h2_recr"),
+                "H3 (sex ratio)" = extract_se("h3"), 
+                "H4a (molt duration)" = extract_se("h4_molt"), 
+                "H4b (breed duration)" = extract_se("h4_breed"))
 

### 2024-10-17 paper/paperv3.qmd
Adding the random effect of year back into the detection probability plot

@@
                              type = "response",
-                             re.form = ~ (1 | year_fct)))
+                             re.form = ~ (1 | year_fct))) %>% 
+  group_by(age, year_fct)
+
+  geom_line(aes(age, predicted, group = interaction(phase, age_cat, year_fct)), 
+            filter(phen_pred_years, phase == "breeddur"),
+            alpha = 0.1) +
 
@@
   # Line per year
-  geom_line(aes(y = predicted, group = interaction(age, year_fct)), 
-            detection_pred_years, 
+  geom_line(aes(x = age, y = predicted, group = interaction(age_cat, year_fct)), 
+            mutate(detection_pred_years, age_cat = "all"), 
             alpha = 0.2) +
@@
                      limits = c(0, NA)) +
-  # scale_linetype_manual(values = c("solid", "dashed")) +
-  # scale_color_manual(values = c("#7fbc41", "#de77ae")) +
-  # scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
   theme(legend.position = "none")

### 2024-10-17 paper/paperv3.qmd
Fixing an error in the mapping for h4 sensitivity tests

@@
   # H4
-molt_data <- sealdat %>% 
+  pheno_data <- sealdat %>% 
     filter(year >= 2011, 
@@
     drop_na(age, BA, BD, MA, MD) %>% 
-  transmute(animalID, 
+    select(animalID,
+           age10,
+           age_cat,
            year_fct,
-            age10 = (age - age_senesce) / 10, 
-            age_cat = factor(age >= age_senesce, 
-                          labels = c("Young", "Old")),
            moltdur, 
@@
                  names_to = "phase", 
-               values_to = "dur_days") %>% 
-  filter(phase == "moltdur")
+                 values_to = "dur_days")
 
   molt_mod <- glmer(dur_days ~ age10 : age_cat + (1 | year_fct) + (1 | animalID), 
-        molt_data, 
+                    filter(pheno_data, phase == "moltdur"), 
                     family = Gamma(link = "log"))
 
-breed_data <- sealdat %>% 
-  filter(year >= 2011, 
-         moltdur >= 7, 
-         age <= 20) %>% 
-  mutate(BA = firstobsbreeddoy, 
-         BD = firstobsbreeddoy + breeddur, 
-         MA = lastobsmoltdoy - moltdur, 
-         MD = lastobsmoltdoy) %>% 
-  drop_na(age, BA, BD, MA, MD) %>% 
-  transmute(animalID, 
-            year_fct, 
-            age10 = (age - age_senesce) / 10, 
-            age_cat = factor(age >= age_senesce, 
-                          labels = c("Young", "Old")),
-            moltdur, 
-            breeddur) %>% 
-  pivot_longer(moltdur:breeddur, 
-               names_to = "phase", 
-               values_to = "dur_days") %>% 
-  filter(phase == "breeddur")
-
   breed_mod <- glmer(dur_days ~ age10 : age_cat + (1 | year_fct) + (1 | animalID), 
-        data = breed_data, 
+                     filter(pheno_data, phase == "breeddur"), 
                      family = Gamma(link = "log"), 
-        control = glmerControl(optimizer = "bobyqa")
-)
+                     control = glmerControl(optimizer = "bobyqa"))
 
@@
 
-# Fit models with cut-off at ages 8-13
+# Fit models with cut-off at ages 6-14
 cutoff <- 6:14

### 2024-10-21 paper/paperv3.qmd
Updated effect size tables

@@
 
-#Making the tables
-#Not sure about this function
-  # effect_size <- function(m) {
-  #   probs <- tibble(age = c(15, 19),
-  #                   age_cat = c("Old", "Old")) %>% 
-  #     mutate(age10 = (age - a) / 10) %>% 
-  #     predict(m, re.form = NA, newdata = ., type = "response")
-  #   odds <- probs / (1 - probs)
-  #   odds[2] / odds[1]
-  # }
-
   list(
@@
     h4_breed_aic = AIC(breed_mod)
-    # h1_effect = effect_size(breed_age_mod),
-    # h2_surv_effect = effect_size(pup_surv_mod),
-    # h2_recr_effect = effect_size(pup_recr_mod),
-    # h3_effect = effect_size(sexrat_age_mod)
   )
@@
 extract_coef <- function(h) {
-  map_dbl(mod_sens, \(m) m[[h]]["age10:age_catOld", "Estimate"])
+  #map_dbl(mod_sens, \(m) m[[h]]["age10:age_catOld", "Estimate"])
+  exp(map_dbl(mod_sens, \(m) m[[h]]["age10:age_catOld", "Estimate"]))
 }
@@
 
-knitr::kable(p_tbl)
-knitr::kable(aic_tbl)
-knitr::kable(coef_tbl)
+knitr::kable(p_tbl, 
+             caption = "P values for threshold ages 6-14")
+knitr::kable(aic_tbl, 
+             caption = "AIC weights for threshold ages 6-14")
+knitr::kable(coef_tbl, 
+             caption = "Effect size for threshold ages 6-14")
 
@@
                       ymax = coef + 1.96 * se)) +
-  labs(x = "Senescence onset (years)") +
+  labs(x = "Senescence onset (years)", 
+       y = "Effect size") +
   scale_size_continuous(range = c(0.25, 1)) +
@@
 
-  geom_line(aes(age, predicted, group = interaction(phase, age_cat, year_fct)), 
-            filter(phen_pred_years, phase == "breeddur"),
-            alpha = 0.1) +
-
 ggplot(detection_pred, aes(age)) +

### 2024-10-22 paper/paperv3.qmd
changing from effect size back to coefficient

@@
 extract_coef <- function(h) {
-  #map_dbl(mod_sens, \(m) m[[h]]["age10:age_catOld", "Estimate"])
-  exp(map_dbl(mod_sens, \(m) m[[h]]["age10:age_catOld", "Estimate"]))
+  map_dbl(mod_sens, \(m) m[[h]]["age10:age_catOld", "Estimate"])
 }
@@
   labs(x = "Senescence onset (years)", 
-       y = "Effect size") +
+       y = "Coefficient") +
   scale_size_continuous(range = c(0.25, 1)) +

### 2024-11-05 paper/paperv3.qmd
Adding a plot for the SMM talk on selective disappearance

@@
 
+selective_diss_coef <- selective_coef %>% filter(model == "Base" | model == "Disappearance")
+
+ggplot(selective_diss_coef, aes(hypothesis, mean, color = model)) + 
+  geom_pointrange(aes(ymin = lwr, ymax = upr), position = position_dodge(width = 0.2)) +
+  geom_hline(yintercept = 0, linetype = "dashed") +
+  expand_limits(y = 0) + 
+  labs(y = "Age Coefficient for Old Animals", x = "Model") + 
+  scale_color_manual(values = c("#de77ae", "#00b6b9")) +
+  scale_fill_manual(values = c("#de77ae", "#00b6b9")) +
+  theme(legend.title = element_blank())
+
 ```
